---
title: "Animal Water"
output: html_document
---

## Script Author
Paul-Eric Rayner

## Water consumption section of Supp.Mat

Water consumption per individual was estimated by summing service water and drinking water consumption. Service water values were extracted from Chapagain & Hoekstra (2003) for all livestock. For cows and goats, drinking water was estimated as a function of age and temperature based on values from Winchester (1956) and Dahlanuddin (1993) respectively. Water consumption data for different temperatures for sheep and pigs was not available. Model outputs are available in Table X. The model was applied to the WorldClim2 gridded average air temperature map (Fick and Hijmans, 2017) at a 10 km2 resolution. Total water consumption per raster cell was calculated by multiplying the number of individuals in each cell by the respective summed water consumption value (drinking + service water) per individual in each cell. We checked our final values against estimates from Chapagain & Hoekstra (2003) that we calculated by multiplying their drinking water.

## Objectives
Model water consumption of livestock, preferably based on temperature averages. 

## Inputs
Chapagain & Hoekstra (2003) drinking water and service water values.

## Outputs
Water consumption maps for each gleam file.

# Set-up
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("./_workflow/common.R")
source(here("_workflow/common_spatial.R"))
source(here("_spatial/template_raster.R"))

## load libraries, read in data
library(doParallel)
```

# Data import
```{r}
# Water data
model_results       <- vroom::vroom(here("animal_farm/water/data/model_results.csv"))
livestock_water_use <- vroom::vroom(here("animal_farm/water/data/livestock_water_use.csv"))

# Monthly temperature stack
avg_monthly_temp <- 
  list.files(file.path(raw,
                       "World_Clim_climate_data/avg_temp",
                       fsep = ""),
             pattern = ".tif", 
             full.names = TRUE) %>% 
  stack()
```

--------------------------------------------------------------------------------

# Setting-up functions

## Create function that applies monthly temperature data to the `poly` model
```{r}
water_temperature_model <- function(avg_monthly_temp_boundaries) {
  
  if (individual_model_results$polynomial[1] == 1) {
    
    individual_model_results$estimate[1] +
      individual_model_results$estimate[2] * avg_monthly_temp_boundaries
  
  } else if (individual_model_results$polynomial[1] == 2) {
    
    individual_model_results$estimate[1] +
      individual_model_results$estimate[2] * avg_monthly_temp_boundaries +
      individual_model_results$estimate[3] * (avg_monthly_temp_boundaries ^ 2)
    
  } else if (individual_model_results$polynomial[1] == 3) {
    
    individual_model_results$estimate[1] +
      individual_model_results$estimate[2] * avg_monthly_temp_boundaries +
      individual_model_results$estimate[3] * (avg_monthly_temp_boundaries ^ 2) +
      individual_model_results$estimate[4] * (avg_monthly_temp_boundaries ^ 3)
  
  } else {
    
    print(paste("Check", livestock_name))
    
  }
}
```

## Loop set-up
```{r}
# Set up a table with final values
total_water <- 
  tibble(gleam_prod_system = as.character(),
         gleam_product = as.character(),
         gleam_livestock = as.character(),
         total_water_consumption = as.double(),
         MekHoek_2012_comparison = as.double())

gleam_files <- 
  list.files("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/animal_farm/farm",
             pattern = "location_df", 
             full.names = TRUE)
```

# Run loop (`calc` won't let me use functions?)
```{r}
for (gleam_file in gleam_files) {

  head_count <- 
    map_df(gleam_file, 
           ~ vroom::vroom(.x, col_select = c(x, y, contains("current")))) %>%
    rasterFromXYZ(., crs = food_crs)
  
  names(head_count) <- "current_count"
  
  # Extract drinking and service water use
  gleam_livestock_name <- 
    str_extract(gleam_file, "buffaloes|cows|sheep|goats|pigs|chickens")
  
  gleam_prod_system_name <- 
    str_extract(gleam_file, "grassland|feedlot|industrial|mixed|backyard|intermediate")
  
  gleam_product_name <- 
    str_extract(gleam_file, "meat|milk|eggs&meat|eggs")
  
  gleam_full_name <- paste(gleam_livestock_name, 
                           gleam_prod_system_name, 
                           gleam_product_name, 
                           sep = "_")
  
  livestock_water_use_filtered <- 
    livestock_water_use %>% 
      filter(gleam_livestock %in% gleam_livestock_name &
             gleam_prod_system %in% gleam_prod_system_name & 
             gleam_product %in% gleam_product_name)
  
  # Extract gleam livestock name
  livestock_name <- unique(livestock_water_use_filtered$animal)
  
  # Extract service water use
  service_water <- 
    livestock_water_use_filtered %>% 
    filter(water_type %in% "service") %>% 
    .$tonnes_consumed

  # Run water temperature model
  
  if (livestock_name %in% c("sheep", "pigs")) {

    drinking_water_raw <-
      livestock_water_use_filtered %>%
      filter(water_type %in% "drinking") %>%
      .$tonnes_consumed

    # Rasterize
    water_consumption_raster <-
      (drinking_water_raw + service_water) * head_count * 365

  } else {

    # Filter by livestock name
    individual_model_results <-
      model_results %>%
        filter(livestock %in% livestock_name)

    avg_monthly_temp_boundaries <- avg_monthly_temp

    avg_monthly_temp_boundaries[avg_monthly_temp_boundaries < individual_model_results$min_temp[1]] <-
      individual_model_results$min_temp[1]

    avg_monthly_temp_boundaries[avg_monthly_temp_boundaries > individual_model_results$max_temp[1]] <-
      individual_model_results$max_temp[1]

    model_output <-
      calc(avg_monthly_temp_boundaries,
           water_temperature_model) %>%
      mean()

    # Save output for further inspection
    writeRaster(model_output,
                file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/animal_farm/water/",
                          gleam_livestock_name,
                          "_",
                          gleam_prod_system_name,
                          "_",
                          gleam_product_name,
                          "_model_output.tif",
                          fsep = ""),
                overwrite = TRUE)

    # Rasterize
    water_consumption_raster <-
      overlay(model_output,
              head_count,
              fun = function(drinking_water, head_count) {

                (drinking_water + service_water) * head_count * 365

              })

  }

  # Save
  writeRaster(water_consumption_raster,
              file.path(layers,
                        "land_",
                        gleam_full_name,
                        "_water",
                        ".tif",
                        fsep = ""),
              overwrite = TRUE,
              format = "GTiff")

  zonal_water <-
    zonal(water_consumption_raster,
          food_rgns_tif,
          fun = "sum",
          na.rm = TRUE) %>%
    as_tibble() %>%
    rename(ID_0 = zone) %>%
    left_join(food_rgns) %>%
    dplyr::select(-c(ID_0, Country)) %>%
    rename(water_consumption = sum)

  zonal_water <-
    raster_df(head_count) %>%
    left_join(food_rgns_xy) %>%
    dplyr::select(x, y, iso3c, current_count) %>%
    group_by(iso3c) %>%
    summarize(head_count = sum(current_count, na.rm = TRUE), .groups = 'drop') %>%
    left_join(zonal_water) %>%
    mutate(full_name = gleam_full_name) %>%
    mutate(gleam_prod_system = gleam_prod_system_name) %>%
    mutate(gleam_product = gleam_product_name) %>%
    mutate(gleam_livestock = gleam_livestock_name) %>%
    arrange(water_consumption)

  # Save tbl
  write_csv(zonal_water,
            here(file.path("animal_farm/water/error_check/",
                           gleam_full_name,
                           "_tbl.csv",
                           fsep = "")))

  ## Error checking plot
  png(file = here(file.path("animal_farm/water/error_check/",
                            gleam_full_name,
                            "_plot.png",
                            fsep = "")),
      width  = 1920,
      height = 1080)

  ggplot(zonal_water,
         aes(x = head_count,
             y = water_consumption)) +
    geom_point() +
    labs(title = Hmisc::capitalize(gsub(pattern = "_",
                                        replacement = " ",
                                        gleam_full_name))) +
    theme_classic()

  dev.off()

# Plots
  png(file = here(paste("animal_farm/water/plots/",
                        gleam_full_name,
                        ".png",
                        sep = "")),
      width = 1920,
      height = 1080)

  plot(water_consumption_raster,
       legend = TRUE,
       legend.args = list(
       text = "m³", # unit of measurement for the stressor
       cex  = 1,
       side = 3,
       line = 1),
       main = Hmisc::capitalize(gsub(pattern = "_",
                                     replacement = " ",
                                     gleam_full_name)))

  dev.off()

  ## Create a table with M&H sums using our population numbers for each animal group.

    drinking_water_raw <- 
      livestock_water_use_filtered %>% 
      filter(water_type %in% "drinking") %>% 
      .$tonnes_consumed
  
    # Rasterize
    water_consumption_raster_check <- (drinking_water_raw + service_water) * head_count * 365
  
    ## Error checking tbl
    total_water <- 
      total_water %>% 
        add_row(tibble_row(
          gleam_prod_system = gleam_prod_system_name,
          gleam_product = gleam_product_name,
          gleam_livestock = gleam_livestock_name,
          total_water_consumption = cellStats(water_consumption_raster, 
                                              stat = "sum", 
                                              na.rm = TRUE),
          MekHoek_2012_comparison = cellStats(water_consumption_raster_check, 
                                              stat = "sum",
                                              na.rm = TRUE)))
}

total_water <- 
  total_water %>% 
  mutate(difference = total_water_consumption / MekHoek_2012_comparison)

write_csv(total_water, here("animal_farm/water/error_check/total_water.csv"))
```

# Data checks
```{r}
# Compare different systems totals
p <- 
  vroom::vroom(here("animal_farm/water/error_check/total_water.csv")) %>% 
  ggplot(aes(x = gleam_livestock, y = total_water_consumption)) +
  geom_point() + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 90)) 
plotly::ggplotly(p)


p2 <- 
  list.files(here(file.path("animal_farm/water/error_check/",
                           fsep = "")),
             pattern = "tbl",
             full.names = TRUE) %>% 
  vroom::vroom() %>% 
  ggplot(aes(x   = head_count,
             y   = water_consumption,
             col = gleam_livestock,
             label = iso3c)) +
  geom_point() +
  theme_classic()

plotly::ggplotly(p2)

map(list.files(layers,
               pattern = "land_.*(meat|milk|eggs)_water.tif",
               full.names = TRUE),
    ~ paste(str_extract(.x, pattern = "(?<=land_).*(?=_water)"), 
            ": ",
            cellStats(raster(.x), stat = "sum"), 
            " tonnes of water",
            sep = ""))

vroom::vroom(here("animal_farm/water/error_check/total_water.csv")) %>%
  summarise(sum(total_water_consumption))

vroom::vroom(here("animal_farm/water/error_check/total_water.csv")) %>%
  filter(!gleam_livestock %in% "buffaloes") %>% 
  group_by(gleam_livestock) %>% 
  summarise(total_water_consumption = sum(total_water_consumption)) %>% 
  ungroup() %>% 
  mutate(percentage_proportion = (total_water_consumption / sum(total_water_consumption)) * 100) %>% 
  arrange(desc(.$percentage_proportion)) 
```

Compare animal totals
```{r}
vroom::vroom(here("animal_farm/water/error_check/total_water.csv")) %>% 
  mutate(gleam_livestock = if_else(gleam_livestock %in% "cows" & gleam_product %in% "milk", 
                                   "dairy cows",
                                   gleam_livestock),
         gleam_livestock = Hmisc::capitalize(gleam_livestock)) %>%
  group_by(gleam_livestock) %>% 
  summarise(Total_water_consumption = sum(total_water_consumption),
            MekHoek_2012_comparison = sum(MekHoek_2012_comparison)) %>% 
  rename(Animal = gleam_livestock) %>% 
  janitor::adorn_totals("row") %>% 
  mutate(Difference = Total_water_consumption / MekHoek_2012_comparison)
```

