---
title: "Model inputs"
author: "Paul-Eric Rayner"
date: "12/12/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Script objective
Calculate model outputs for livestock where we have water consumption ~ temperature data

# Data description

Cows data
Data source: NRC
Collected on: 12/12/19

Goat data
Data source: Dahlanuddin and Thwaites, 1993
Collected on: 12/16/19

Chickens data
Data source: Daghir, 2008

# Set-up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("./_workflow/common.R")
## load libraries, read in data
library(janitor)
```

# Data import and wrangling
```{r}
cows_water <- 
  vroom::vroom(here("animal_farm/water/data/winchester_1956_cattle_water.csv")) %>% 
  clean_names() %>% 
  rename(livestock = cattle_class) %>% 
  mutate(livestock = tolower(gsub(" ", 
                                  "_", 
                                  livestock, 
                                  fixed = TRUE))) %>% 
  pivot_longer(cols      = -c(wt_kg, livestock),
               names_to  = "temperature", 
               values_to = "water_intake") %>%
  mutate(temperature = as.numeric(str_remove(string = temperature,
                                             pattern = "[x]"))) %>% 
  mutate(water_intake = water_intake * 0.001) %>%  # metric cubes unit
  arrange() %>% 
  group_by(livestock, 
           temperature) %>%
  summarise(water_intake = mean(water_intake, na.rm = TRUE)) %>% 
  ungroup() %>% 
  na.omit()

goat_water <- 
  vroom::vroom(here("animal_farm/water/data/dahlanuddin_1993_goat_water.csv")) %>% 
  clean_names() %>% 
  rename(water_intake = mean) %>% 
  mutate(water_intake = water_intake * 0.001)

leghorn_pullets_water <- 
  tibble(temperature = c(10, 21.1, 32.2, 37.8),
         water_intake = c(0.00012, 0.000155, 0.000255, 0.00031))

broiler_chickens_water <- 
  vroom::vroom(here("animal_farm/water/data/broiler_chickens_water_temp.csv")) %>% 
  pivot_longer(cols      = - Age_Weeks,
               names_to  = "Temperature", 
               values_to = "Water_Consumption",
               names_transform = list(Temperature = as.double)) %>%
  clean_names() %>% 
  group_by(temperature) %>% 
  summarise(water_intake = mean(water_consumption) / 1000 / 1000, .groups = 'drop') 

backyard_chickens_water <- 
  bind_rows(leghorn_pullets_water,
            broiler_chickens_water) %>% 
  group_by(temperature) %>% 
  summarise(water_intake = mean(water_intake), .groups = 'drop')
```

# Visualize
```{r}
ggplot(cows_water, aes(x = temperature, 
                       y = water_intake, 
                       col = livestock)) +
       theme_classic() +
       geom_point() +
       geom_smooth(method = "lm", 
                   formula = y ~ poly(x, 2, raw = TRUE), se = FALSE) +
       labs(x   = "Temperature", 
            y   = "Water intake",
            col = "cows class") +
       theme(axis.title   = element_text(size = 14),
             axis.text    = element_text(size = 11, colour = "black"),
             legend.title = element_text(size = 14, colour = "black"),
             legend.text  = element_text(size = 11))

ggplot(goat_water, aes(x = temperature, y = water_intake)) +
        theme_classic() +
        geom_point() +
        geom_smooth(method = "lm", formula = y ~ poly(x, 2, raw = TRUE), se = FALSE) +
        labs(lab = "Temperature", y  = "Water Intake") +
        theme(axis.title   = element_text(size = 14),
              axis.text    = element_text(size = 11, colour = "black"),
              legend.title = element_text(size = 14, colour = "black"),
              legend.text  = element_text(size = 11))

ggplot(leghorn_pullets_water, aes(x = temperature, y = water_intake))+
        theme_classic()+
        geom_point()+
        geom_smooth(method = "lm", formula = y ~ poly(x, 2, raw = TRUE), se = FALSE) +
        labs(x = "Temperature", y  = "Water Intake") +
        theme(axis.title   = element_text(size = 14),
              axis.text    = element_text(size = 11, colour = "black"),
              legend.title = element_text(size = 14, colour = "black"),
              legend.text  = element_text(size = 11))

ggplot(backyard_chickens_water, aes(x = temperature, y = water_intake))+
        theme_classic()+
        geom_point()+
        geom_smooth(method = "lm", formula = y ~ poly(x, 2, raw = TRUE), se = FALSE) +
        labs(x = "Temperature", y  = "Water Intake") +
        theme(axis.title   = element_text(size = 14),
              axis.text    = element_text(size = 11, colour = "black"),
              legend.title = element_text(size = 14, colour = "black"),
              legend.text  = element_text(size = 11))
```

We can see that the line fits relatively well for all cows classes except
finishing cows, which performs better with a third polynomial.

# Extract model coefficients for each livestock
```{r}
# cows
model_cows_water <- function(class_name, polynomial) {

    cows_water %>% 
    filter(livestock %in% class_name) %>% 
    lm(water_intake ~ poly(temperature,
                           polynomial,
                           raw = TRUE),
       data = .) %>% 
    broom::tidy() %>% 
    add_column(livestock = class_name, 
               .before   = TRUE) %>% 
    vroom::vroom_write(here(file.path("animal_farm/water/data/", 
                             class_name, 
                             "_model.csv",
                             fsep = "")))
}

map2(c("finishing_cattle", "growing_cattle", "lactating_cows"),
     c(2, 2, 3),
     model_cows_water)

# Goat
lm(water_intake ~ poly(temperature, 
                       2, 
                       raw = TRUE), 
   data = goat_water) %>% 
  broom::tidy() %>%  
  add_column(livestock = "goats", 
             .before = TRUE) %>% 
  vroom::vroom_write(here("animal_farm/water/data/goat_model.csv"))

# Layer chickens
lm(water_intake ~ poly(temperature, 
                       2, 
                       raw = TRUE), 
   data = leghorn_pullets_water) %>% 
  broom::tidy() %>% 
  add_column(livestock = "layer_chickens", 
             .before = TRUE) %>% 
  vroom::vroom_write(here("animal_farm/water/data/layer_chickens_model.csv"))

# Backyard chickens
lm(water_intake ~ poly(temperature, 
                       2, 
                       raw = TRUE), 
   data = backyard_chickens_water) %>% 
  broom::tidy() %>% 
  add_column(livestock = "backyard_chickens", 
             .before = TRUE) %>% 
  vroom::vroom_write(here("animal_farm/water/data/backyard_chickens_model.csv"))

# Broiler chickens
lm(water_intake ~ poly(temperature, 
                       2, 
                       raw = TRUE), 
   data = broiler_chickens_water) %>% 
  broom::tidy() %>% 
  add_column(livestock = "broiler_chickens", 
             .before = TRUE) %>% 
  vroom::vroom_write(here("animal_farm/water/data/broiler_chickens_model.csv"))
```
