---
title: "MAPSPAM_tonnes_production.Rmd"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Determine total crop grown in each region
This should be run if there are changes to the MAPSPAM production data.

Calculates total crop production using the MAPSPAM data adjusted to 2017 production.  These data include all growing systems (intensive, irrigated, etc.) and are in units of metric tonnes.

This is used in subsequent script to estimate the proportion of each crop going to animal feed in each country.

```{r}


library(here)
library(tidyverse)
library(countrycode)

focus_year = 2017

## quick check to make sure categories are aligned
mapspam_crops <- read_csv(here("feed/data/MapSPAM_crop_info.csv")) %>%
  select(SPAM_short_name, SPAM_super) %>%
  unique()


```

Extract by region the crop production from the raster maps.
```{r}

regions <- read_csv(here("_spatial/_output/food_rgns.csv"))

crop_tonnes <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/crop/farm/scaled_maps_2017", pattern = "_A", full=TRUE)
crop_tonnes <- grep("rest", crop_tonnes, invert=TRUE, value=TRUE)

crop_stack <- raster::stack(crop_tonnes)
zones <- raster::raster("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/spatial/land_eez_rgns.tif")

crop_production <- raster::zonal(crop_stack, zones, fun="sum", na.rm=TRUE, progress="text")

crop_production_df <- data.frame(crop_production) %>%
  gather("SPAM_short_name", "tonnes_producing_crop", -1) %>%
  rename("ID_0" = zone) %>%
  left_join(regions, by="ID_0") %>%
  mutate(SPAM_short_name = gsub("crop_", "", SPAM_short_name)) %>%
  mutate(SPAM_short_name = gsub("_A_scaled", "", SPAM_short_name)) %>%
  dplyr::select(iso3c_producing=iso3c, SPAM_short_name, tonnes_producing_crop)

```

Combine the super categories

```{r}
crop_production_df <- left_join(crop_production_df, mapspam_crops, by="SPAM_short_name") 

# check that all crop categories add up: if not, adjust the crop code datasheet
crop_production_df <- crop_production_df %>%
  group_by(iso3c_producing, SPAM_super) %>%
  summarize(tonnes_producing_crop = sum(tonnes_producing_crop)) %>%
#  filter(!(SPAM_super %in% c("toba", "ofib", "rcof", "acof", "teas"))) %>%
  ungroup()

```

Check and save: should be about equal to 2017 FAO production

```{r}

check <- crop_production_df %>%
  group_by(SPAM_super) %>% summarize(our_tonnes = sum(tonnes_producing_crop, na.rm=TRUE)) %>% data.frame()
  
fao_translate <- vroom::vroom(here("crop/farm/data/crop_codes_updated.csv")) %>%
  select(SPAM_super, item_code) %>%
  unique()

fao <- read_csv(here("crop/farm/data/FAOSTAT_crop_production_2010_2017.csv")) %>% 
  filter(Year == 2017) %>%
  filter(Area != "China") %>%
  select(Area, item_code = `Item Code`, Value) %>%
  left_join(fao_translate, by="item_code") %>%
  group_by(SPAM_super) %>%
  summarize(total = sum(Value, na.rm=TRUE)) 
  
compare <- check %>% left_join(fao, by="SPAM_super") %>%
  mutate(prop = our_tonnes/total)
compare
sum(compare$our_tonnes, na.rm=TRUE)/sum(compare$total) # this all looks reasonable! Some small differences due to gapfilling and other things.

write_csv(crop_production_df, here("feed/data/MAPSPAMcrop_production.csv"))

```
