---
title: "Calculating proportion of each 'feedgroup' for ruminants globally and the prop feedlot out of feedlot+mixed"
author: "Juliette"
date: "10/19/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

For livestock: GLEAM provides feed composition data for different cohorts within a system (e.g., Cattle: Adult females, Meat animals, Replacement animals and adult males.

This script estimates the proportion of animals in each cohort to weight the feed composition data to better estimate the average feed composition for the entire system.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
source(here("_workflow/common.R"))

'%notin%' <- Negate('%in%')

ruminants_raw <- read_csv(here("animal_farm/farm/data/ruminants_GLEAMi_v2.csv"))

```


```{r}

ruminants <- ruminants_raw %>% 
  select(-Unit) %>% 
  mutate(Herd_type = ifelse(Production_system == "Feedlot operations", "Feedlot", Herd_type))%>% 
  filter(Production_system != "All systems",
         Herd_type != "Whole herd",
         Herd_type != "Feedlot") %>% ## these have their own specific diet composition so not included in this prop breakdown 
  filter(Variable %in% c("HERD: adult females", "HERD: adult males", "HERD: replacement females", "HERD: replacement males", "HERD: fattening females", "HERD: fattening males")) %>%
  mutate(feedgroup = case_when(Variable == "HERD: adult females" ~ "Adult females",
                               Variable %in% c("HERD: adult males", "HERD: replacement females", "HERD: replacement males") ~ "Replacement animals and adult males",
                               Variable %in% c("HERD: fattening females", "HERD: fattening males") ~ "Meat animals (non feedlot)")) %>% 
  group_by(iso3c, Species, Production_system, Herd_type, feedgroup) %>% 
  dplyr::summarise(feedgroup_headcount = sum(Value)) %>% 
  ungroup() %>% 
  group_by(iso3c, Species, Production_system, Herd_type)%>% 
  mutate(headcount = sum(feedgroup_headcount)) %>% 
  ungroup() %>% 
  rowwise() %>% 
  mutate(prop = feedgroup_headcount/headcount) %>% 
  mutate(prop = ifelse(feedgroup_headcount==0 & headcount == 0, 0, prop))
  
```

Need to gapfill missing countries

```{r}
length(unique(ruminants$iso3c)) ## 233

## figure out the global average for each species, production_system, herd_type, feedgroup 

ruminants_no_zero <- ruminants %>% 
  filter(prop != 0)

## goats
goats_plot <- ggplot(subset(ruminants_no_zero, Species == "Goats")) + 
geom_point(aes(y = prop, x = feedgroup, color = feedgroup), 
             position = position_jitterdodge(dodge.width = 0.8)) +
  facet_grid(cols = vars(Production_system), rows = vars(Herd_type)) +
  theme_light()  + 
  theme(legend.position = "none")+
  labs(title = "Goats") +
  stat_summary(mapping = aes(y = prop, x = feedgroup),
               fun = "mean", geom = "point", shape = 0, size = 4, color = "black",
               position = position_dodge(width = 0.8)) +
  stat_summary(mapping = aes(y = prop, x = feedgroup),
               fun = "median", geom = "point", shape = 2, size = 4, color = "black",
               position = position_dodge(width = 0.8)) + 
  theme(legend.position = "none")

## cows
cows_plot <- ggplot(subset(ruminants_no_zero, Species == "Cattle")) + 
geom_point(aes(y = prop, x = feedgroup, color = feedgroup), 
             position = position_jitterdodge(dodge.width = 0.8)) +
  facet_grid(cols = vars(Production_system), rows = vars(Herd_type)) +
  theme_light()  + 
  theme(legend.position = "none")+
  labs(title = "Cows") +
  stat_summary(mapping = aes(y = prop, x = feedgroup),
               fun = "mean", geom = "point", shape = 0, size = 4, color = "black",
               position = position_dodge(width = 0.8)) +
  stat_summary(mapping = aes(y = prop, x = feedgroup),
               fun = "median", geom = "point", shape = 2, size = 4, color = "black",
               position = position_dodge(width = 0.8)) + 
  theme(legend.position = "none")

## sheep
sheep_plot <- ggplot(subset(ruminants_no_zero, Species == "Sheep")) + 
geom_point(aes(y = prop, x = feedgroup, color = feedgroup), 
             position = position_jitterdodge(dodge.width = 0.8)) +
  facet_grid(cols = vars(Production_system), rows = vars(Herd_type)) +
  theme_light()  + 
  theme(legend.position = "none")+
  labs(title = "Sheep") +
  stat_summary(mapping = aes(y = prop, x = feedgroup),
               fun = "mean", geom = "point", shape = 0, size = 4, color = "black",
               position = position_dodge(width = 0.8)) +
  stat_summary(mapping = aes(y = prop, x = feedgroup),
               fun = "median", geom = "point", shape = 2, size = 4, color = "black",
               position = position_dodge(width = 0.8)) + 
  theme(legend.position = "none")

## buffaloes
buffaloes_plot <- ggplot(subset(ruminants_no_zero, Species == "Buffalo")) + 
geom_point(aes(y = prop, x = feedgroup, color = feedgroup), 
             position = position_jitterdodge(dodge.width = 0.8)) +
  facet_grid(cols = vars(Production_system), rows = vars(Herd_type)) +
  theme_light()  + 
  theme(legend.position = "none")+
  labs(title = "Buffaloes") +
  stat_summary(mapping = aes(y = prop, x = feedgroup),
               fun = "mean", geom = "point", shape = 0, size = 4, color = "black",
               position = position_dodge(width = 0.8)) +
  stat_summary(mapping = aes(y = prop, x = feedgroup),
               fun = "median", geom = "point", shape = 2, size = 4, color = "black",
               position = position_dodge(width = 0.8)) + 
  theme(legend.position = "none")


goats_plot
cows_plot
sheep_plot
buffaloes_plot

```

Calculate the mean and save

```{r}

proportions <- ruminants_no_zero %>% 
  group_by(Species, feedgroup) %>% 
  dplyr::summarise(mean_prop = mean(prop)) %>% 
  ungroup() %>% 
  mutate(mean_prop = round(mean_prop, 2)) %>% 
  group_by(Species) %>% 
  mutate(total_prop = sum(mean_prop)) %>% 
  ungroup() %>% 
  ## the prop total for cattle is 0.01. this probably wont be a big differnece later but we can just subtract the extra 
  mutate(mean_prop = ifelse(Species == "Cattle" & feedgroup == "Adult females", mean_prop - (0.01/3),
                            ifelse(Species == "Cattle" & feedgroup == "Meat animals (non feedlot)", mean_prop - (0.01/3),
                                   ifelse(Species == "Cattle" & feedgroup == "Replacement animals and adult males", mean_prop - (0.01/3), mean_prop))))%>% 
  group_by(Species) %>% 
  mutate(total_prop = sum(mean_prop)) %>% 
  ungroup() %>% 
  mutate(Species = ifelse(Species == "Goats", "Goat", Species))

prop_save <- proportions %>% 
  select(Species, feedgroup, mean_prop) %>% 
  write_csv(here("feed/data/ruminant_feedgroup_prop.csv"))

```
