---
title: "Aquaculture diet composition wrangling"
output: html_document
editor_options: 
  chunk_output_type: console
---

Wrangling and aggragating some aquaculture diet composition data.

```{r}
library(tidyverse)
library(readxl)

regions <- read_csv(here("_spatial/_output/food_rgns.csv")) %>%
  select(iso3c) %>%
  unique()

# Halley diet data
diet <- read_excel(here("feed/diet_composition_wrangling/Final_DietCrop_Numbers.xlsx"), sheet = "% type of crop input")

```

Shrimp: see scripts for "shrimp_mariculture_diet...". (regional data is the average shrimp feed, weighted by most common species)
Salmon: diet from Aas source, global values used for regions
Tuna: Halley data, average marine fish data

### shrimp
```{r}

shrimp <- read_csv(here("feed/diet_composition_wrangling/shrimp_aquaculture_diet_composition.csv"))

```

### marine-fish-general
```{r}

other_fish <- read_csv(here("feed/diet_composition_wrangling/marine-fish-general_aquaculture_diet_composition.csv"))

```

### salmon

```{r}

salmon <- read_csv(here("feed/diet_composition_wrangling/salmon_aquaculture_diet_composition.csv")) %>%
  mutate(prop_diet = per_diet * 0.01) %>%
  merge(regions) %>%
  select(iso3c, raw_name, prop_diet, production_system, species)

```

### tuna
```{r}

tuna_prop_fofm <- 0.85

tuna <- filter(diet, common_name=="Tunas") %>%
  filter(prop_crop_inputs > 0) %>%
  mutate(species = "tuna",
         production_system = "aquaculture") %>%
  group_by(species) %>%
  mutate(total_input_prop = sum(prop_crop_inputs)) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(prop_diet = prop_crop_inputs/total_input_prop * tuna_prop_fofm) %>%
  merge(regions) %>%
  select(iso3c, raw_name = crop, prop_diet, production_system, species)

# should be 0.85
tuna %>%  
group_by(iso3c) %>%
  summarize(sum(prop_diet))

tuna_fofm <- select(tuna, iso3c) %>%
  unique() %>%
  mutate(raw_name = "fofm",
         prop_diet = 1-tuna_prop_fofm,
         production_system = "aquaculture",
         species="tuna") 
  
tuna <- rbind(tuna, tuna_fofm)

# should now be one:
tuna %>%  
group_by(iso3c) %>%
  summarize(sum(prop_diet))

```


## crustaceans
```{r}

crusts_prop_fofm <- 0.885

crusts <- filter(diet, group=="Crustacean") %>%
  filter(!comment %in% "freshwater") %>%
  filter(common_name != "Chinese mitten crab") %>%
  group_by(crop) %>%
  summarize(prop_crop = sum(prop_crop_inputs)/3 * crusts_prop_fofm)  # three species, but check this in future iterations

sum(crusts$prop_crop)

crusts <- rbind(crusts,
                data.frame(crop="fofm", prop_crop = 1-crusts_prop_fofm))
sum(crusts$prop_crop) # should now be 1

crusts <- merge(crusts, regions) %>%
  mutate(production_system = "aquaculture",
         species="crustaceans") %>%
   select(iso3c, raw_name = crop, prop_diet = prop_crop, production_system, species)

```


## Combine everything
```{r}

aquaculture_diet <- rbind(shrimp, salmon, tuna, crusts, other_fish)


write_csv(aquaculture_diet, here("feed/data/aquaculture_diet_composition.csv"))