---
title: "Ruminant diet: feeditems rates"
author: "Juliette"
date: "10/23/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(here)
source(here("_workflow/common.R"))

'%notin%' <- Negate('%in%')

un_rgns <- read_csv(here("_spatial/_output/UNSD_Methodology.csv")) %>%
  filter(!is.na(iso3c)) %>%
  filter(iso3c %in% food_rgns$iso3c)

livestock_diet_composition <- read_csv(here("feed/data/livestock_diet_composition.csv")) %>% 
  mutate(orientation = ifelse(production_system == "Feedlot", "Meat", orientation),
         orientation = ifelse(orientation == "Dairy", "Milk", orientation))


tmp <- raster("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/animal_farm/farm/production_system2_tifs/mixed_prod_system.tif")
tmp <- raster::raster("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/animal_farm/farm/production_system2_tifs/grassland_prod_system.tif")
raster::plot(tmp)

```


## Fix the mixed and grassland diet compositions

For the mixed and feedlot categories (based on our description) we assume they do not graze on fresh grass, and instead, roughage is provided as hay or other fodder.
```{r}

roughage <- data.frame(raw_name = c( "Crop residues from maize",
                  "Crop residues from millet",
                  "Crop residues from other grains",
                  "Crop residues from rice",
                  "Crop residues from sorghum",
                  "Crop residues from sugarcane",
                  "Crop residues from wheat",
                  "Fodder beet",
                  "Fresh mixture of grass and legumes",
                  "Hay or silage from alfalfa",
                  "Hay or silage from cultivated grass",
                  "Hay or silage from grass and legumes",
                  "Silage from whole grain plants.",
                  "Silage from whole maize plant",
                  "Hay from adjacent areas", 
                  "Fresh grass"),
                  roughage_category = 1)
roughage <- roughage %>%
  mutate(roughage_category_no_grass = 1) %>%
  mutate(roughage_category_no_grass = ifelse(raw_name == "Fresh grass", 0, roughage_category_no_grass))

fixed_mixed_int <- livestock_diet_composition %>% 
  filter(production_system %in% c("Mixed", "Feedlot")) %>% 
  left_join(roughage, by="raw_name") %>%
  mutate(percent_roughage = percent * roughage_category) %>%
  mutate(percent_roughage_no_grass = percent * roughage_category_no_grass) %>%
  group_by(iso3c, species, orientation, production_system) %>%
  mutate(total_roughage = sum(percent_roughage, na.rm=TRUE)) %>%
  mutate(total_roughage_no_grass = sum(percent_roughage_no_grass, na.rm=TRUE)) %>%
  mutate(roughage_correct = total_roughage/total_roughage_no_grass) %>%
  mutate(percent_roughage_corrected = percent_roughage_no_grass * roughage_correct) %>%
  mutate(new_percent = ifelse(is.na(percent_roughage_corrected), percent, percent_roughage_corrected)) %>%
  group_by(iso3c, species, orientation, production_system) %>% 
  mutate(test = sum(new_percent))  ## yay it looks good, all test should be 100

tmp <- filter(fixed_mixed_int, iso3c=="USA" & species=="Cattle" & orientation=="Milk") %>% data.frame()
sum(tmp$new_percent)

fixed_mixed <- fixed_mixed_int %>% 
  select(iso3c, species, orientation, production_system, raw_name, percent = new_percent)

```


For grassland/grazed animals we assume that all roughage comes from grazing.
```{r}

fixed_grass_int <- livestock_diet_composition %>% 
  filter(production_system == "Grassland Based") 

## pull out a list of everything that is considered fodder or crop residues
unique(fixed_grass_int$raw_name)
remove_list <- c( "Crop residues from maize",
                  "Crop residues from millet",
                  "Crop residues from other grains",
                  "Crop residues from rice",
                  "Crop residues from sorghum",
                  "Crop residues from sugarcane",
                  "Crop residues from wheat",
                  "Fodder beet",
                  "Fresh mixture of grass and legumes",
                  "Hay or silage from alfalfa",
                  "Hay or silage from cultivated grass",
                  "Hay or silage from grass and legumes",
                  "Silage from whole grain plants.",
                  "Silage from whole maize plant",
                  "Hay from adjacent areas")

fixed_grass <- fixed_grass_int %>% 
  mutate(raw_name = ifelse(raw_name %in% remove_list, "Fresh grass", raw_name)) %>% 
  group_by(iso3c, species, orientation, production_system, raw_name) %>% 
  dplyr::summarise(percent = sum(percent)) %>% 
  ungroup()
  
```

Add the fixed compositions to the main diet comp df

```{r}

livestock_diet_composition <-  rbind(fixed_grass, fixed_mixed) %>% 
  mutate(species = ifelse(species == "Goat", "Goats", species))

```


Gapfill regions with no data with regional data:
```{r}

livestock_diet_composition_all <- livestock_diet_composition %>%
  select(species, orientation, production_system, raw_name) %>%
  unique() %>%
  merge(food_rgns$iso3c) %>%
  select(iso3c=y, species, orientation, production_system, raw_name)

# quick check
filter(livestock_diet_composition_all, iso3c=="USA" & species == "Cattle", production_system=="Grassland Based" & orientation == "Meat")
filter(livestock_diet_composition_all, iso3c=="USA" & species == "Cattle", production_system=="Mixed" & orientation == "Meat")



livestock_diet_composition_gf <- left_join(livestock_diet_composition_all, livestock_diet_composition, by=c("iso3c", "species", "orientation", "production_system", "raw_name")) %>%
  left_join(un_rgns) %>%
  group_by(species, orientation, production_system, raw_name, Intermediate_Region_Name) %>%
  mutate(percent = ifelse(is.na(percent), mean(percent, na.rm=TRUE), percent)) %>%
  group_by(species, orientation, production_system, raw_name, Sub_region_Name) %>%
    mutate(percent = ifelse(is.na(percent), mean(percent, na.rm=TRUE), percent)) %>%
group_by(species, orientation, production_system, raw_name, Region_Name) %>%
    mutate(percent = ifelse(is.na(percent), mean(percent, na.rm=TRUE), percent)) %>%
group_by(species, orientation, production_system, raw_name) %>%
    mutate(percent = ifelse(is.na(percent), mean(percent, na.rm=TRUE), percent)) %>%
ungroup() %>%
  select(iso3c, species, orientation, production_system, raw_name, percent)

# check that all gapfilled sum to 100
tmp <- livestock_diet_composition_gf %>%
 group_by(species, orientation, production_system, iso3c) %>%
 summarize(total = sum(percent))
summary(tmp)


write_csv(livestock_diet_composition_gf, here("feed/data/ruminant_diet_composition_update1.csv"))

```



Let's do the total amount of each item needed for each ruminant and category 
```{r}

total_feed <- read_csv(here("animal_farm/farm/data/ruminants_GLEAMi_v2.csv"))

rum_list <- c("Sheep", "Buffalo", "Goats", "Cattle")

ruminant_total_feed <- total_feed %>% 
  filter(Species %in% rum_list,
         Variable %in% c("INTAKE: Total intake", "HERD: total number of animals"),
         Production_system != "All systems") %>% 
  mutate(Herd_type = ifelse(Production_system == "Feedlot operations", "Non-dairy", Herd_type)) %>% 
  filter(Herd_type != "Whole herd") %>% 
  select(-Country, - Unit) %>%  
  pivot_wider(names_from = Variable, values_from = Value) %>% 
  rename(kg_dm_feed_intake = 'INTAKE: Total intake',
         count = 'HERD: total number of animals',
         species = Species,
         production_system = Production_system) %>% 
  mutate(production_system = case_when(production_system == "Grassland systems" ~ "Grassland Based",
                                      production_system == "Mixed systems" ~ "Mixed",
                                      production_system == "Feedlot operations" ~ "Feedlot"),
         Herd_type = case_when(Herd_type == "Dairy" ~ "Milk",
                               Herd_type == "Non-dairy" ~ "Meat")) %>% 
  rename(orientation = Herd_type) %>%
  mutate(consumption_rate = kg_dm_feed_intake/count,
         consumption_rate = ifelse(kg_dm_feed_intake == 0 & count == 0, 0, consumption_rate),
         consumption_rate = ifelse(is.infinite(kg_dm_feed_intake), 0, consumption_rate)) %>%
   filter(consumption_rate > 0) 


summary(ruminant_total_feed)

## some checks
sheep_grass_milk <- filter(ruminant_total_feed, species=="Sheep" & production_system=="Grassland Based" & orientation == "Milk")
summary(sheep_grass_milk)
sheep_grass_meat <- filter(ruminant_total_feed, species=="Sheep" & production_system=="Grassland Based" & orientation == "Meat")
summary(sheep_grass_meat)
sheep_mixed_milk <- filter(ruminant_total_feed, species=="Sheep" & production_system=="Mixed" & orientation == "Milk")
summary(sheep_mixed_milk)
sheep_mixed_meat <- filter(ruminant_total_feed, species=="Sheep" & production_system=="Mixed" & orientation == "Meat")
summary(sheep_mixed_meat)
filter(ruminant_total_feed, iso3c=="USA" & species=="Sheep")

goat_grass_milk <- filter(ruminant_total_feed, species=="Goats" & production_system=="Grassland Based" & orientation == "Milk")
summary(goat_grass_milk)
goat_grass_meat <- filter(ruminant_total_feed, species=="Goats" & production_system=="Grassland Based" & orientation == "Meat")
summary(goat_grass_meat)
goat_mixed_milk <- filter(ruminant_total_feed, species=="Goats" & production_system=="Mixed" & orientation == "Milk")
summary(goat_mixed_milk)
goat_mixed_meat <- filter(ruminant_total_feed, species=="Goats" & production_system=="Mixed" & orientation == "Meat")
summary(goat_mixed_meat)

cows_grass_milk <- filter(ruminant_total_feed, species=="Cattle" & production_system=="Grassland Based" & orientation == "Milk")
summary(cows_grass_milk)
cows_grass_meat <- filter(ruminant_total_feed, species=="Cattle" & production_system=="Grassland Based" & orientation == "Meat")
summary(cows_grass_meat)
cows_mixed_milk <- filter(ruminant_total_feed, species=="Cattle" & production_system=="Mixed" & orientation == "Milk")
summary(cows_mixed_milk)
cows_mixed_meat <- filter(ruminant_total_feed, species=="Cattle" & production_system=="Mixed" & orientation == "Meat")
summary(cows_mixed_meat)
cows_feedlot_meat <- filter(ruminant_total_feed, species=="Cattle" & production_system=="Feedlot" & orientation == "Meat")
summary(cows_feedlot_meat)

buffalo_grass_milk <- filter(ruminant_total_feed, species=="Buffalo" & production_system=="Grassland Based" & orientation == "Milk")
summary(buffalo_grass_milk)
buffalo_grass_meat <- filter(ruminant_total_feed, species=="Buffalo" & production_system=="Grassland Based" & orientation == "Meat")
summary(buffalo_grass_meat)
buffalo_mixed_milk <- filter(ruminant_total_feed, species=="Buffalo" & production_system=="Mixed" & orientation == "Milk")
summary(buffalo_mixed_milk)
buffalo_mixed_meat <- filter(ruminant_total_feed, species=="Buffalo" & production_system=="Mixed" & orientation == "Meat")
summary(buffalo_mixed_meat)


``` 
https://pir.sa.gov.au/__data/assets/pdf_file/0007/272869/Calculating_dry_matter_intakes.pdf : sheep and cattle are 2-3% of liveweight in DM
https://beeflambnz.com/knowledge-hub/PDF/guide-feed-planning-sheep-farmers.pdf


Gapfill regions with no feed rates.
```{r}

# get all possible combinations of animals and regions

systems <- select(ruminant_total_feed, species, production_system, orientation) %>%
  unique()
all_sites <- merge(un_rgns$iso3c, systems) %>%
  rename(iso3c=x)

ruminant_total_feed_gf <- left_join(all_sites, ruminant_total_feed, by=c("iso3c", "species", "production_system", "orientation")) %>%
  left_join(un_rgns) %>%
  group_by(species, production_system, orientation, Intermediate_Region_Name) %>%
  mutate(consumption_rate = ifelse(is.na(consumption_rate), median(consumption_rate, na.rm=TRUE), consumption_rate)) %>%
  group_by(species, production_system, orientation, Sub_region_Name) %>%
    mutate(consumption_rate = ifelse(is.na(consumption_rate), median(consumption_rate, na.rm=TRUE), consumption_rate)) %>%
group_by(species, production_system, orientation, Region_Name) %>%
    mutate(consumption_rate = ifelse(is.na(consumption_rate), median(consumption_rate, na.rm=TRUE), consumption_rate)) %>%
group_by(species, production_system, orientation) %>%
    mutate(consumption_rate = ifelse(is.na(consumption_rate), median(consumption_rate, na.rm=TRUE), consumption_rate)) %>%
ungroup() %>%
  select(iso3c, species, production_system, orientation, consumption_rate)



```
 
 
Calculate feed amounts and rates for each feed item
```{r}


  combo <- left_join(livestock_diet_composition_gf, ruminant_total_feed_gf, by = c("iso3c", "species", "orientation", "production_system")) %>% 
    rowwise() %>% 
    mutate(percent = percent/100,
         feeditem_kg_per_head = percent*consumption_rate) %>% 
  select(iso3c, animal = species, product = orientation, system = production_system, feedstuff = raw_name, feeditem_kg_per_head) %>% 
    mutate(animal = tolower(animal),
           product = case_when(product == "Milk" ~ "milk",
                               product == "Meat" ~ "meat"),
           system = case_when(system == "Grassland Based" ~ "grassland",
                              system == "Mixed" ~ "mixed",
                              system == "Feedlot" ~"feedlot"),
           animal = ifelse(animal == "cattle", "cows",
                           ifelse(animal == "goat", "goats",
                                  ifelse(animal == "buffalo", "buffaloes", animal))))
    
  

  write_csv(combo, here("feed/data/ruminant_feed_consumption_rates.csv"))
  


```


```{r}

ruminant_feed_consumption_rates <- read_csv(here("feed/data/ruminant_feed_consumption_rates.csv"))

tmp<- ruminant_feed_consumption_rates %>%
  group_by(iso3c, animal, product, system) %>%
  summarize(total_feed_per_animal = sum(feeditem_kg_per_head))
filter(tmp, iso3c=="USA")
test <- ruminant_feed_consumption_rates %>% 
  filter(system == "mixed") %>% 
  arrange(desc(feeditem_kg_per_head)) %>% 
  mutate(feedstuff = as_factor(feedstuff))

#should be around 5000 for milk c ows and feedlot, and 3000-4000 for meat cows.  This is less than a grown cow eats, but this seems
# about right (although maybe a little low) given not all animals in the herd are eating the adult rate.

ggplot(test) +
  geom_point(aes(y = feedstuff, x = feeditem_kg_per_head)) +
  facet_grid(cols = vars(animal), rows = vars(product))



```
 

 