---
title: "Animal_consumption"
output: html_document
editor_options: 
  chunk_output_type: console
---

Get proportion of each MAPSPAM crop that is consumed by animals in each country.

Outputs: a raster for each crop describing the proportion of each crop consumed in each country by animal feed systems. This is a final output.

Be sure to run previous scripts if changes to data.

```{r}

library(here)
library(tidyverse)
library(countrycode)

focus_year = 2017

```


# STEP 1: determine amount of crop produced in each country for each food system.
Trade proportion data:
```{r}

trade_data <- read_csv(here("feed/data/FAO_MAPSPAMcrop_trade_data.csv"))

```


Total consumption by animals per country in MAPSPAM crops
```{r}

consumption <- read_csv(here("feed/data/system_country_mapspam_tonnes_consumption.csv")) %>%
  rename(iso3c_consuming = iso3c, consuming_crop_consumed_tonnes =tonnes_product)

```


Join consumption by country and crop to trade data to get likely country of origin regardless of county of consumption.

```{r}
# check that all feed products match up: should be nothing showing up here
setdiff(consumption$SPAM_super, trade_data$SPAM_super)

production_location <- left_join(trade_data, consumption, by=c("iso3c_consuming", "SPAM_super")) %>%
  filter(!is.na(animal_system)) %>%  #crops not fed to animal system in given country
  mutate(consuming_crop_consumed_tonnes = ifelse(is.na(consuming_crop_consumed_tonnes), 0, consuming_crop_consumed_tonnes)) %>%
  mutate(producing_crop_consumed_tonnes = consuming_crop_consumed_tonnes * prop_of_supply)

# should all be equal tonnes_crop_consumed. prop_of_supply should sum to 1.
tmp <- filter(production_location, iso3c_consuming=="USA" & SPAM_super=="maiz" & animal_system == "cows_feedlot_meat")
tmp
summary(tmp)
sum(tmp$prop_of_supply)
sum(tmp$producing_crop_consumed_tonnes) # should equal consuming_crop_consumed_tonnes
head(production_location)

```


Summarize by producing country to get the total production of feed for animal systems across all countries.
(This is used for efficiency calculations: tonnes of production in each country that is consumed by each food system)

```{r}

feed_source <- production_location %>%
  group_by(iso3c_producing, animal_system, SPAM_super) %>%
  summarize(consumed_tonnes = sum(producing_crop_consumed_tonnes)) 
write_csv(feed_source, here("feed/data/tonnes_feedproduced_per_country_system.csv"))

#tmp <- read_csv(here("feed/data/tonnes_feedproduced_per_country_system.csv"))
```

# STEP 2 Calculate proportion of production for animal system for each country

Crop production according to MAPSPAM data adjusted to 2017 production.  These data include all growing systems (intensive, irrigated, etc.) and are in units of metric tonnes.

```{r}

crop_production_df <- read_csv( here("feed/data/MAPSPAMcrop_production.csv"))
  
crop_production_global <- crop_production_df %>%
  group_by(SPAM_super) %>%
  summarize(tonnes_global = sum(tonnes_producing_crop, na.rm=TRUE))

```

Divide tonnes of consumption for each country/food system by total production to get the proportion of production produced for each food system:

```{r}

# if the data were perfect, we could leave it at this...but there are some corrections to make
# when consumption overshoots production
proportion <- left_join(feed_source, crop_production_df, by=c("iso3c_producing", "SPAM_super")) %>%
  mutate(tonnes_producing_crop = ifelse(is.na(tonnes_producing_crop), 0, tonnes_producing_crop)) %>%
  mutate(prop_produced_for_system = ifelse(consumed_tonnes == 0, 0, consumed_tonnes/tonnes_producing_crop))

```

# STEP 3: Distribute excess, and adjust proportion produced for system values
NOTE: if our consumption and production data were perfect these complicated steps would be unnecessary and the above calculation for prop_produced_for_system would be good enough. But due to mismatches, we spread excess consumption based on relative global production.


This is only for interest, I tried to approach this category of exceedence differently, but it didn't work out.
```{r}
## which crops exceed global production (NOTE: even if <1 a country/crop consumption can still exceed production):
global_exceed <- proportion %>%
  group_by(SPAM_super) %>%
  summarize(global_consumed = sum(consumed_tonnes)) %>%
  ungroup() %>%
  left_join(crop_production_global) %>%
  mutate(prop_consumed = global_consumed/tonnes_global) %>%
  select(SPAM_super, prop_consumed)

crop_exceed <- global_exceed$SPAM_super[global_exceed$prop_consumed>1]

```


Starting again:
Identify the countries/crops with total consumption rates (across all systems) that exceed production
```{r}
## step 1 ID regions where consumption exceeds production
consumption_exceed <- proportion %>%
  group_by(iso3c_producing, SPAM_super) %>%
  mutate(tonnes_consumed_country = sum(consumed_tonnes)) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(extra_crop_tonnes = tonnes_producing_crop - tonnes_consumed_country) %>%
  mutate(percent_exceed = ifelse(tonnes_producing_crop == 0, 0, tonnes_consumed_country/tonnes_producing_crop)) %>%
    data.frame()

```


Adjust proportion of consumption to not exceed production by more than 5%,
```{r}

corrected_prop <- consumption_exceed %>%
  mutate(prop_of_consumption_within_iso = ifelse(tonnes_consumed_country == 0, 0, consumed_tonnes/tonnes_consumed_country)) %>%  # calculate proportion of crop each system eats within a country
  mutate(prop_produced_for_system_adjust = ifelse(percent_exceed >1 , 
                                        ifelse(percent_exceed > 1.05, prop_of_consumption_within_iso*1.05, prop_of_consumption_within_iso*percent_exceed),
                                        prop_produced_for_system))

summary(corrected_prop)
filter(corrected_prop, iso3c_producing=="USA", animal_system=="chickens_industrial_meat", SPAM_super=="maiz")

corrected_prop %>%
  group_by(iso3c_producing, SPAM_super) %>%
  summarize(total_adj_prop = sum(prop_produced_for_system_adjust),
            total_prop = sum(prop_produced_for_system))

tmp <- filter(corrected_prop, iso3c_producing=="AFG" & SPAM_super=="maiz")
sum(tmp$prop_produced_for_system)
sum(tmp$prop_produced_for_system_adjust)


tmp <- filter(corrected_prop, iso3c_producing=="USA" & SPAM_super=="maiz")
sum(tmp$prop_produced_for_system)
sum(tmp$prop_produced_for_system_adjust)
#
tmp <- filter(corrected_prop, iso3c_producing=="USA" & SPAM_super=="sorg")
sum(tmp$prop_produced_for_system)
sum(tmp$prop_produced_for_system_adjust)

tmp <- filter(corrected_prop, iso3c_producing=="AGO" & SPAM_super=="whea")
sum(tmp$prop_of_consumption_within_iso)
sum(tmp$prop_produced_for_system)
sum(tmp$prop_produced_for_system_adjust)

```

For each animal system determine tonnes that was included and excluded.  The excluded will  be disbursed.
```{r}

tonnes_included_system <- corrected_prop %>%
  mutate(included_tonnes = prop_produced_for_system_adjust * tonnes_producing_crop) %>%
  mutate(lost_crop_tonnes = consumed_tonnes - included_tonnes) %>%
  mutate(included_prop = ifelse(tonnes_producing_crop == 0 , 0, included_tonnes/tonnes_producing_crop))
summary(tonnes_included_system)

# check, sum of included and excluded should equal sum of all consumption
tmp <- tonnes_included_system %>%
  group_by(SPAM_super) %>%
  summarize(included_crop = sum(included_tonnes, na.rm=TRUE),
            excluded_crop = sum(lost_crop_tonnes, na.rm=TRUE)) %>% 
  mutate(prop_redistributed = excluded_crop/(excluded_crop + included_crop)) %>%
           data.frame() 
sum(tmp$included_crop)/(sum(tmp$included_crop) + sum(tmp$excluded_crop)) #84% included and does not need to be distributed

tonnes_lost_system <- tonnes_included_system %>%
  group_by(SPAM_super, animal_system) %>%
  summarize(lost_crop_tonnes = sum(lost_crop_tonnes, na.rm=TRUE))


#calculate extra crop tonnes in each country, and determine what proportion of extra each country accounts for
tonnes_extra_crop <- corrected_prop %>%
  dplyr::select(iso3c_producing, SPAM_super, extra_crop_tonnes, tonnes_producing_crop) %>%
  unique() %>%
  mutate(extra_crop_tonnes = ifelse(extra_crop_tonnes<0, 0, extra_crop_tonnes)) %>%
  group_by(SPAM_super) %>%
  mutate(excess_crop_global = sum(extra_crop_tonnes)) %>%
  rowwise() %>%
  mutate(prop_excess_crop = ifelse(excess_crop_global == 0, 0, extra_crop_tonnes/excess_crop_global))

# distribute the tonnes lost for each crop/system based on each countries extra production
excess_prop <- merge(tonnes_extra_crop, tonnes_lost_system) %>%
  mutate(tonnes_dispersed = prop_excess_crop * lost_crop_tonnes) %>%
  mutate(prop_dispersed = ifelse(tonnes_producing_crop == 0, 0, tonnes_dispersed/tonnes_producing_crop))


### NOTE: you can get higher total props for countries with very small proportions. This happens only for soybeans and for small production areas.
## this explores this:
excess_excess_prop <- excess_prop %>%
  mutate(extra_dispersed_tonnes = tonnes_dispersed - extra_crop_tonnes) %>%
  mutate(extra_dispersed_tonnes = ifelse(prop_dispersed > 1, extra_dispersed_tonnes, 0)) %>%
  left_join(crop_production_global, by = "SPAM_super") %>%
  mutate(prop_global_crop_production = tonnes_producing_crop/tonnes_global) %>%
  group_by(SPAM_super, animal_system) %>%
  mutate(extra_dispersed_tonnes_by_system_crop = sum(extra_dispersed_tonnes))
  

filter(excess_excess_prop, prop_dispersed>1) %>% data.frame()

```


Combine the included and lost proportions for each system and country for plotting below

```{r}

assigned_prop <- tonnes_included_system  %>%
  dplyr::select(iso3c_producing, animal_system, SPAM_super, prop_produced_for_system = included_prop)
excess_prop <- excess_prop %>%
  dplyr::select(iso3c_producing, animal_system, SPAM_super, prop_extra_for_system = prop_dispersed)

total_prop <- left_join(assigned_prop, excess_prop, by=c("iso3c_producing", "animal_system", "SPAM_super")) %>%
  rowwise() %>%
  mutate(total_prop = prop_produced_for_system + prop_extra_for_system)
summary(total_prop)
filter(total_prop, iso3c_producing=="AFG" & SPAM_super=="soyb") %>% data.frame()

```

Checking
```{r}

summary(total_prop) # ok for total prop to be a bit above 1

filter(total_prop, prop_produced_for_system>1)
tmp <- filter(total_prop, iso3c_producing == "BHS", SPAM_super=="maiz") %>% data.frame()
sum(tmp$prop_produced_for_system)
sum(tmp$total_prop)

## some go as high as 1.22x production, this is mostly allowing the sugarbeets to go over.
tmp <- total_prop %>%
  dplyr::group_by(iso3c_producing, SPAM_super) %>%
  dplyr::summarize(total_prop = sum(total_prop))
summary(tmp)
filter(tmp, total_prop>1.05)
            
## check: should match original consumption:
consumption_global <- feed_source %>%
  group_by(animal_system, SPAM_super) %>%
  summarize(obs_tonnes = sum(consumed_tonnes, na.rm=TRUE))

check <- left_join(total_prop, crop_production_df) %>%
  mutate(total_tonnes = total_prop*tonnes_producing_crop) %>%
  group_by(animal_system, SPAM_super) %>%
  summarize(est_tonnes = sum(total_tonnes, na.rm=TRUE)) %>%
  left_join(consumption_global) %>%
  rowwise() %>%
  mutate(difference = obs_tonnes - est_tonnes )

filter(check, SPAM_super=="maiz" & animal_system == "chickens_industrial_meat")
# these should be the same
check %>%
  group_by(SPAM_super) %>%
  summarize(est_tonnes =sum(est_tonnes),
            obs_tonnes = sum(obs_tonnes)) %>% data.frame()

plot(log(check$est_tonnes+1), log(check$obs_tonnes+1))
abline(0,1, col="red") # these should look about the same!

```


Fill in missing regions/crops with 0 values
```{r}

regions <- read_csv(here("_spatial/_output/food_rgns.csv")) %>%
  dplyr::select(iso3c_producing = iso3c)

spam <- read_csv(here("_analysis/food_system_categories.csv")) %>%
  filter(subcategory == "crops") %>%
  pull(organism) %>%
  unique()

spam <- grep("fodd", spam, invert=TRUE, value=TRUE)  # fodder taken care of elsewhere

all_combos <- expand.grid(iso3c_producing = regions$iso3c_producing, 
                           animal_system = unique(total_prop$animal_system),
                           SPAM_super = spam) %>%
  data.frame() %>%
  mutate(iso3c_producing = as.character(iso3c_producing)) %>%
  mutate(animal_system = as.character(animal_system)) %>%
      mutate(SPAM_super = as.character(SPAM_super)) 



proportion_distribute <- left_join(all_combos, total_prop) %>%
  mutate(total_prop = ifelse(is.na(total_prop), 0, total_prop)) %>%
  dplyr::select(iso3c_producing, animal_system, SPAM_super, total_prop) 

summary(proportion_distribute)
check <- filter(proportion_distribute, animal_system=="buffaloes_grassland_meat" & SPAM_super=="barl") %>% data.frame()
check <- filter(proportion_distribute, iso3c_producing=="ZMB" & SPAM_super=="barl") %>% data.frame()

write_csv(proportion_distribute, here("feed/data/proportion_feed_per_country_system.csv"))

```



For each crop determine the proportion of each countries production dedicated to the animal system.

```{r}

library(raster)

## organize region data
rgns <- read_csv("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/spatial/food_rgns_xy.csv", col_types = "ddddcc")

####################
## start loop here
#################3

todo <- unique(dplyr::select(proportion_distribute, SPAM_super, animal_system))


for(i in 1:dim(todo)[1]){
#i=182
crop <- todo$SPAM_super[i]
system <- todo$animal_system[i]

# total country feed production
feed_crop <- filter(proportion_distribute, SPAM_super==crop, animal_system==system) %>%
  dplyr::select(iso3c = iso3c_producing, prop_feed = total_prop)

# proportion crop grown for feed per country
prop_feed_crop <- left_join(rgns, feed_crop, by="iso3c") %>%
  dplyr::select(x, y, prop_feed)

prop_feed_crop_raster <- rasterFromXYZ(prop_feed_crop)
# plot(prop_feed_crop_raster)
crs(prop_feed_crop_raster) <- "+proj=longlat +datum=WGS84"
writeRaster(prop_feed_crop_raster, filename= sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/datalayers/land_%s_x_land_%s_crop_produce.tif", system, crop), overwrite=TRUE)

#old <- raster(sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/datalayers/land_%s_x_land_%s_crop_produce.tif", system, crop))
#plot(log(old+1))
}


```
 
 Edit error introduced above in filename, fix so the marine animals have a "marine" prefix (vs. default land)
```{r}


shrimp_files <- list.files('/home/shares/food-systems/Food_footprint/all_food_systems/datalayers', full=TRUE, pattern = "land_shrimp")
file.rename(from= shrimp_files, to = 
                gsub("land_shrimp", "marine_shrimp", shrimp_files))

salmon_files <- list.files('/home/shares/food-systems/Food_footprint/all_food_systems/datalayers', full=TRUE, pattern = "land_salmon")
file.rename(from= salmon_files, to = 
                gsub("land_salmon", "marine_salmon", salmon_files))

tuna_files <- list.files('/home/shares/food-systems/Food_footprint/all_food_systems/datalayers', full=TRUE, pattern = "land_tuna")
file.rename(from= tuna_files, to = 
                gsub("land_tuna", "marine_tuna", tuna_files))

fish_files <- list.files('/home/shares/food-systems/Food_footprint/all_food_systems/datalayers', full=TRUE, pattern = "land_marine-fish-general")
file.rename(from= fish_files, to = 
                gsub("land_marine-fish-general", "marine_marine-fish-general", fish_files))

crusts_files <- list.files('/home/shares/food-systems/Food_footprint/all_food_systems/datalayers', full=TRUE, pattern = "land_crustaceans")
file.rename(from= crusts_files, to = 
                gsub("land_crustaceans", "marine_crustaceans", crusts_files))

aux_files <- list.files('/home/shares/food-systems/Food_footprint/all_food_systems/datalayers', full=TRUE, pattern = ".aux.xml")
file.remove(aux_files)

```


## Data check
```{r}

all_props <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/datalayers", pattern="_x_", full=TRUE)
all_props <- grep("xpul", all_props, value=TRUE)
all_props <- grep("human_human_human", all_props, invert=TRUE, value=TRUE)
all_crop_stack <- stack(all_props)
all_crop_sum <- calc(all_crop_stack, sum, na.rm=TRUE)

## look at all crops
crop_list <- c("bana", "barl", "cass", "maiz", "ocer", "oilp", "rice",
               "sorg", "soyb", "sugb", "whea")

# do something special for these
xmil <- c("smil", "pmil")


total_feed_by_crop <- data.frame(SPAM_super = crop_list, calc_tonnes_from_rast = NA)

for(crop in crop_list){ #crop="soyb"
all_props <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/datalayers", pattern="_x_", full=TRUE)
all_props <- grep(crop, all_props, value=TRUE)
all_props <- grep("human_human_human", all_props, invert=TRUE, value=TRUE)
all_crop_stack <- raster::stack(all_props)
all_crop_sum <- raster::calc(all_crop_stack, sum, na.rm=TRUE)

#plot(all_crop_sum)
#click(all_crop_sum)
prod <- raster(sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/crop/farm/scaled_maps_2017/crop_%s_A_scaled.tif", crop))

tonnes_consumed <- prod*all_crop_sum
total_tonnes_consumed <- cellStats(tonnes_consumed, "sum", na.rm=TRUE)
total_feed_by_crop$calc_tonnes_from_rast[total_feed_by_crop$SPAM_super==crop] <- total_tonnes_consumed
}

old_consumption <- read_csv(here("feed/data/system_country_mapspam_tonnes_consumption.csv")) %>%
  group_by(SPAM_super) %>%
  summarize(start_tonnes_feed = sum(tonnes_product))

left_join(total_feed_by_crop, old_consumption) %>%
  mutate(prop = calc_tonnes_from_rast/start_tonnes_feed)


## combine groups
xpul <- c("bean", "chic", "cowp", "pige", "lent", "opul")

all_props <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/datalayers", pattern="_x_", full=TRUE)
all_props <- grep("xpul", all_props, value=TRUE)
all_props <- grep("human_human_human", all_props, invert=TRUE, value=TRUE)
all_crop_stack <- stack(all_props)
all_crop_sum <- calc(all_crop_stack, sum, na.rm=TRUE)

#plot(all_crop_sum)
#click(all_crop_sum)
pul_list <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/crop/farm/scaled_maps_2017", pattern = "_A_", full=TRUE)
pul_list <- grep(paste(xpul, collapse="|"), pul_list, value=TRUE)
pul_stack <- stack(pul_list)
prod <- calc(pul_stack, sum, na.rm=TRUE)

tonnes_consumed <- prod*all_crop_sum
total_tonnes_consumed <- cellStats(tonnes_consumed, "sum", na.rm=TRUE)
total_tonnes_consumed
8860114/8909291

xoil <- c("grou", "sunf", "rape", "sesa", "ooil", "cott")

all_props <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/datalayers", pattern="_x_", full=TRUE)
all_props <- grep("xoil", all_props, value=TRUE)
all_props <- grep("human_human_human", all_props, invert=TRUE, value=TRUE)
all_crop_stack <- stack(all_props)
all_crop_sum <- calc(all_crop_stack, sum, na.rm=TRUE)

#plot(all_crop_sum)
#click(all_crop_sum)
pul_list <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/crop/farm/scaled_maps_2017", pattern = "_A_", full=TRUE)
pul_list <- grep(paste(xoil, collapse="|"), pul_list, value=TRUE)
pul_stack <- stack(pul_list)
prod <- calc(pul_stack, sum, na.rm=TRUE)

tonnes_consumed <- prod*all_crop_sum
total_tonnes_consumed <- cellStats(tonnes_consumed, "sum", na.rm=TRUE)
total_tonnes_consumed
79593545/79755715
```