---
title: "Human and other consumption"
output: html_document
editor_options: 
  chunk_output_type: console
---

Human feed is calculated as the proportion of crop after animal feed consumption is accounted for.

Creates proportion of each MAPSPAM crop that is consumed by humans in each country and outputs a raster map for each crop.

Be sure to run previous scripts if changes to data.

```{r}

library(here)
library(tidyverse)
library(countrycode)

focus_year = 2017

spam_cats <- read_csv(here("feed/data/MapSPAM_crop_info.csv")) %>%
  select(SPAM_super) %>%
  unique()

```

# Calculate "other (non food/feed) consumption

First estimate other consumption for crops with large amounts of non-food/non-feed uses (e.g., biofuel, etc.):
maize (~20%), sugarcane (~18%), coconut (~12%), cassava (~9%)

These data come from FAO Food Balance Sheets, and particularly, the "Element = Other" category of use.
```{r}

fao <- read_csv(here("_analysis/checking_data/data/FAOSTAT_FBS_5-5-2022.csv")) %>%
  select(Area, Element, item_code = "Item Code", Item, Year, Unit, Value) %>%
  filter(Element == "Other uses (non-food)" & Item %in% c("Cassava and products", "Coconuts - Incl Copra", "Maize and products", "Sugar cane")) %>%
  filter(Value > 0) %>%
  mutate(Value = Value*1000)

crop_translate <- read_csv(here("feed/data/MapSPAM_to_FAO_v2.csv")) %>%
  filter(!is.na(FAO_item_code_NFB)) %>%
    filter(is.na(non_food)) %>%
  select(SPAM_super, item_code = FAO_item_code_NFB) %>%
  unique()

fao_spam <- fao %>%
  left_join(crop_translate, by="item_code") %>%
  group_by(Area, SPAM_super, Element, Year, Unit) %>%
  dplyr::summarize(other_tonnes = sum(Value, na.rm=TRUE)) %>%
  ungroup()


fao_iso <- fao_spam %>%
    mutate(Area = iconv(Area, "UTF-8", "UTF-8",sub=''),
         Area = ifelse(Area == "Eswatini", "Swaziland", Area),
         Area = ifelse(Area == "Netherlands Antilles (former)", "Bonaire, Sint Eustatius and Saba", Area),
         Area = ifelse(Area == "French Guyana", "French Guiana", Area),
         Area = ifelse(Area == "China, mainland", "China", Area))

## add standardized country names
fao_iso$iso3c_consuming <- countrycode(as.character(fao_iso$Area), origin="country.name", destination = "iso3c")

## make sure all countries have an iso3c code
check <- filter(fao_iso, is.na(iso3c_consuming))

fao_other <- fao_iso %>%
  select(iso3c_consuming, SPAM_super, other_tonnes)

```

Merge with the trade data to get a better estimate of where it comes from.
```{r}

trade_data <- read_csv(here("feed/data/FAO_MAPSPAMcrop_trade_data.csv"))

production_location <- left_join(trade_data, fao_other, by=c("iso3c_consuming", "SPAM_super")) %>%
  filter(!is.na(other_tonnes)) %>%  #crops not fed to animal system in given country
  mutate(producing_crop_other_tonnes = other_tonnes * prop_of_supply)

# should all be equal tonnes_crop_consumed. prop_of_supply should sum to 1.
tmp <- filter(production_location, iso3c_consuming=="USA" & SPAM_super=="maiz")
tmp
summary(tmp)
sum(tmp$prop_of_supply)
sum(tmp$producing_crop_other_tonnes) # should equal consuming_crop_consumed_tonnes
head(production_location)

```

Summarize by producing country to get the total production of "other" across all countries.

```{r}

other_source <- production_location %>%
  group_by(iso3c_producing, SPAM_super) %>%
  summarize(other_tonnes = sum(producing_crop_other_tonnes)) %>%
  filter(other_tonnes > 0)

# check that we end up with the same amount! 
# these should match pretty closely
other_source %>%
  group_by(SPAM_super) %>%
  summarize(other_tonnes = sum(other_tonnes))

fao_spam %>%
  group_by(SPAM_super) %>%
  summarize(start_tonnes = sum(other_tonnes))

```

Join with total production. This will help us determine whether we exceed production with other and feed. And, also how to distribute the excess.

```{r}

crop_production_df <- read_csv( here("feed/data/MAPSPAMcrop_production.csv"))

other_source <- left_join(crop_production_df, other_source, by=c("iso3c_producing", "SPAM_super")) %>%
  filter(SPAM_super %in% c("maiz", "sugc", "cass", "cnut"))

```


# Data for livestock aquaculture feed consumption
(calculated in previous scripts in feed folder)
```{r}

prop_feed <- read_csv(here("feed/data/proportion_feed_per_country_system.csv"))


prop_feed_total <- prop_feed %>%
  group_by(SPAM_super, iso3c_producing) %>%
  summarize(animal_feed_prop = sum(total_prop)) %>%
    mutate(animal_feed_prop = ifelse(animal_feed_prop > 1, 1, animal_feed_prop)) %>% # allowed to float 5% above food production
  ungroup()

sum(duplicated(paste(prop_feed_total$iso3c_producing, prop_feed_total$SPAM_super)))

```

## Join feed and other

This step is necessary because we take into account the amount of production within a country used for feed and other.  Anything that results in more other/feed consumption than what is produced is distributed among all countries with production that excceds consumptions, proportional to the remaining supply.
```{r}

feed_other_tonnes <- left_join(other_source, prop_feed_total, by=c("iso3c_producing", "SPAM_super")) %>%
  mutate(feed_tonnes = animal_feed_prop * tonnes_producing_crop) %>%
  mutate(other_tonnes = ifelse(is.na(other_tonnes), 0, other_tonnes)) %>%
  mutate(feed_other_tonnes = other_tonnes + feed_tonnes) %>%
  mutate(total_prop = ifelse(tonnes_producing_crop == 0, 0, feed_other_tonnes/tonnes_producing_crop)) %>% 
  mutate(deficit_tonnes = ifelse(total_prop < 1, 0, (feed_other_tonnes - tonnes_producing_crop))) %>%
  mutate(excess_tonnes = ifelse(total_prop > 1 , 0, (tonnes_producing_crop - feed_other_tonnes)))


sum(feed_other_tonnes$deficit_tonnes)/sum(feed_other_tonnes$excess_tonnes) ## this is good! Accounting for nearly everything.  Not even sure it is worth messing with this further. But I will, just in case this changes in future analyses.

feed_other_tonnes <- feed_other_tonnes %>%
  group_by(SPAM_super) %>%
  mutate(global_spam_tonnes = sum(tonnes_producing_crop),
         global_spam_deficit_tonnes = sum(deficit_tonnes),
         global_spam_excess_tonnes = sum(excess_tonnes)) %>%
  ungroup() %>%
  rowwise() %>%
mutate(prop_global_deficit_dispersal = global_spam_deficit_tonnes/global_spam_excess_tonnes) %>%
  mutate(other_tonnes_dispersed = prop_global_deficit_dispersal*excess_tonnes) %>%
  mutate(other_tonnes_capped = (other_tonnes - deficit_tonnes)) %>%
  mutate(other_tonnes_total = other_tonnes_capped + other_tonnes_dispersed)

## check: the other_tonnnes_total should sum to the other_tonnes
sum(feed_other_tonnes$other_tonnes)
sum(feed_other_tonnes$other_tonnes_total)

# some more checks: something should be in each of these categories:
filter(feed_other_tonnes, other_tonnes_total != other_tonnes) # something should pop up here!
filter(feed_other_tonnes, other_tonnes_total > other_tonnes)
filter(feed_other_tonnes, other_tonnes_total < other_tonnes) %>% data.frame() # should all have total_prop >1

feed_other_tonnes %>% group_by(SPAM_super) %>%
  summarize(global = mean(global_spam_tonnes),
            other = sum(other_tonnes_total)) %>%
  rowwise() %>%
  mutate(prop = other/global) 
# prop should be about equal to these: maize (~20%), sugarcane (~18%), coconut (~12%), cassava (~9%)

# check that we got all production
read_csv( here("feed/data/MAPSPAMcrop_production.csv")) %>%
  group_by(SPAM_super) %>%
  summarize(global_tonnes = sum(tonnes_producing_crop)) %>%
  filter(SPAM_super %in% c("cnut", "maiz", "sugc", "cass"))

```  

Clean.
```{r}

feed_other_tonnes_clean <- feed_other_tonnes %>% 
  rowwise() %>%
  mutate(other_prop = ifelse((other_tonnes_total == 0 & tonnes_producing_crop == 0), 0, other_tonnes_total/tonnes_producing_crop)) %>%
  mutate(total_non_human_prop = other_prop + animal_feed_prop) %>%
  select(iso3c_producing, SPAM_super, tonnes_producing_crop, animal_feed_prop, other_prop, total_non_human_prop)

summary(feed_other_tonnes_clean)


# rowbind with other data (above only incudes the data for the 4 crops that have a large portion going to other.)
prop_feed_total_cut <- filter(prop_feed_total, !(SPAM_super %in% unique(feed_other_tonnes_clean$SPAM_super))) %>%
  bind_rows(feed_other_tonnes_clean) %>%
  select(-tonnes_producing_crop) %>%
  mutate(other_prop = ifelse(is.na(other_prop), 0, other_prop),
         total_non_human_prop = animal_feed_prop + other_prop)


# these should be the same
dim(prop_feed_total_cut)
dim(prop_feed_total)

#total_non_human_prop should be 1 or less 
summary(prop_feed_total_cut)

write_csv(prop_feed_total_cut, here("feed/data/proportion_feed_and_other_per_country_system.csv"))
```


```{r}

prop_feed_human <- read_csv(here("feed/data/proportion_feed_and_other_per_country_system.csv"))

prop_feed_human <- prop_feed_human %>%
  mutate(human_feed_prop = 1 - total_non_human_prop) %>%
  select(SPAM_super, iso3c_producing, human_feed_prop)

unique(prop_feed_human$SPAM_super)

setdiff(spam_cats$SPAM_super, unique(prop_feed_human$SPAM_super))

write_csv(prop_feed_human, here("feed/data/human_crop_prop_consume.csv"))
filter(prop_feed_human, SPAM_super=="spis")
```
For each crop determine the proportion of each countries production dedicated to direct human consumption.

```{r}
library(raster)

## organize region data
rgns <- read_csv("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/spatial/food_rgns_xy.csv", col_types = "ddddcc")

####################
## start loop here
#################3

crop_list <-  unique(prop_feed_human$SPAM_super)

for(crop in crop_list){
#crop="spis"

# total country feed production
feed_crop <- filter(prop_feed_human, SPAM_super==crop) %>%
  dplyr::select(iso3c = iso3c_producing, prop_feed = human_feed_prop)

# proportion crop grown for feed per country
prop_feed_crop <- left_join(rgns, feed_crop, by="iso3c") %>%
  dplyr::select(x, y, prop_feed)

prop_feed_crop_raster <- rasterFromXYZ(prop_feed_crop)
# plot(prop_feed_crop_raster)
crs(prop_feed_crop_raster) <- "+proj=longlat +datum=WGS84"
writeRaster(prop_feed_crop_raster, filename= sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/datalayers/land_human_human_human_x_land_%s_crop_produce.tif", crop), overwrite=TRUE)

}

aux_files <- list.files('/home/shares/food-systems/Food_footprint/all_food_systems/datalayers', full=TRUE, pattern = ".aux.xml")
file.remove(aux_files)
```


Check to see that everything basically adds up to the original amounts of production.  Looks good!

```{r}

total_crop_consumption <- data.frame(SPAM_super = unique(prop_feed_human$SPAM_super), calc_tonnes_from_rast = NA) %>%
  filter(!(SPAM_super %in% c('xoil', 'xfru', 'xpul', 'xmil')))

for(crop in crop_list){ #crop="sugc"
all_props <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/datalayers", pattern="_x_", full=TRUE)
all_props <- grep(crop, all_props, value=TRUE)
all_crop_stack <- raster::stack(all_props)
all_crop_sum <- raster::calc(all_crop_stack, sum, na.rm=TRUE)

#plot(all_crop_sum)
#click(all_crop_sum)
prod <- raster(sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/crop/farm/scaled_maps_2017/crop_%s_A_scaled.tif", crop))

tonnes_consumed <- prod*all_crop_sum
total_tonnes_consumed <- cellStats(tonnes_consumed, "sum", na.rm=TRUE)
total_crop_consumption$calc_tonnes_from_rast[total_crop_consumption$SPAM_super==crop] <- total_tonnes_consumed
}

#prop_consumed should be close to 1
crop_production_df <- read_csv( here("feed/data/MAPSPAMcrop_production.csv")) %>%
  group_by(SPAM_super) %>%
  summarize(starting_tonnes_produced_spam = sum(tonnes_producing_crop)) %>%
  left_join(total_crop_consumption, by="SPAM_super") %>%
  mutate(prop_consumed = calc_tonnes_from_rast/starting_tonnes_produced_spam)
crop_production_df %>% data.frame()

xpul <- c("bean", "chic", "cowp", "pige", "lent", "opul")

all_props <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/datalayers", pattern="_x_", full=TRUE)
all_props <- grep("xpul", all_props, value=TRUE)
all_crop_stack <- stack(all_props)
all_crop_sum <- calc(all_crop_stack, sum, na.rm=TRUE)

#plot(all_crop_sum)
#click(all_crop_sum)
pul_list <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/crop/farm/scaled_maps_2017", pattern = "_A_", full=TRUE)
pul_list <- grep(paste(xpul, collapse="|"), pul_list, value=TRUE)
pul_stack <- stack(pul_list)
prod <- calc(pul_stack, sum, na.rm=TRUE)

tonnes_consumed <- prod*all_crop_sum
total_tonnes_consumed <- cellStats(tonnes_consumed, "sum", na.rm=TRUE)
total_tonnes_consumed
96130485/96296387

xoil <- c("grou", "sunf", "rape", "sesa", "ooil", "cott")
all_props <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/datalayers", pattern="_x_", full=TRUE)
all_props <- grep("xoil", all_props, value=TRUE)
all_crop_stack <- stack(all_props)
all_crop_sum <- calc(all_crop_stack, sum, na.rm=TRUE)

#plot(all_crop_sum)
#click(all_crop_sum)
oil_list <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/crop/farm/scaled_maps_2017", pattern = "_A_", full=TRUE)
oil_list <- grep(paste(xoil, collapse="|"), oil_list, value=TRUE)
oil_stack <- stack(oil_list)
prod <- calc(oil_stack, sum, na.rm=TRUE)

tonnes_consumed <- prod*all_crop_sum
total_tonnes_consumed <- cellStats(tonnes_consumed, "sum", na.rm=TRUE)
total_tonnes_consumed
291380837/292094582
```

