---
title: 'Figure 5: Ranking Foods'
author: "Juliette"
date: "11/17/2020"
output: html_document
---

the total crop pressures (counting everything) are here: /home/shares/food-systems/Food_footprint/all_food_systems/datalayers
and_maize_crop_produce_nutrient.tif includes all the nutrient pressures associated with maize regardless of where it is going.

The human consumed part would be in here; /home/shares/food-systems/Food_footprint/all_food_systems/analysis/raw_feed_combos

The animal consumed part of each crop is also in here: /home/shares/food-systems/Food_footprint/all_food_systems/analysis/raw_feed_combos

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(hrbrthemes)
library(RColorBrewer)
library(colorspace)
library(scales)
library(colorspace)
library(strex)
library(patchwork)
source(here("_workflow/common.R"))

'%!in%' <- function(x,y)!('%in%'(x,y))

summary_df_raw <- read_csv(file.path("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/raw_df_summary/rgn_raw_summary.csv")) 

rescaling_values <- read_csv(here("_analysis/rescale_values.csv"))

food_groupings <- read_csv(here("_analysis/figures/paper/data/grouping_naming_structure.csv")) 

# mapspam_names <- read_csv(here("crop/farm/data/MapSPAM_names.csv"))%>% 
#   select(sub_tier = SPAM_short_name, SPAM_full_name) %>% unique()

mapspam_names <- vroom::vroom(here("crop/farm/data/crop_codes_updated.csv")) %>%
  select(sub_tier = SPAM_super, SPAM_full_name_raw = SPAM_full_name) %>%
  mutate(SPAM_full_name = case_when(sub_tier == "xmil" ~ "millet",
                                    sub_tier == "xpul" ~ "pulses",
                                    sub_tier == "xcof" ~ "coffee",
                                    sub_tier == "xfru" ~ "other fruits",
                                    sub_tier == "xoil" ~ "other oil crops",
                                    T ~ SPAM_full_name_raw))  %>% 
  select(sub_tier, SPAM_full_name) %>% 
  unique() 


```

Groupings that we will want (this is by tier 2)
(1) livestock  and their feed
(2) crops (human consumption and feed)
(3) mariculture and their feed
(3) fisheries marine and fresh


## need to change salmon to salmonoide and fish meat to river fish meat
```{r}
summary_df_int <-summary_df_raw %>% 
  mutate(feed_item = case_when(category == "feedcrop"|category =="feedfodd"|category =="feedfofm" ~ "feed",
                               T ~ " ")) %>% 
  unite(product, c("product", "feed_item"), sep = " ") %>% 
  mutate(product = str_trim(product, side = c("right"))) %>% 
  select(organism, product, pressure, sum) %>% 
  ## fresh water fisheries is listed as fish and there is still salmon instead of salmonids so we will change that here 
  mutate(organism = ifelse(organism == "fish", "freshwater fish", 
                           ifelse(organism == "salmon", "salmonids", organism))) %>% 
  group_by(organism, product, pressure) %>% 
  dplyr::summarise(sum = sum(sum)) %>% 
  unite(tier_5, c("organism", "product"), sep = " ") %>% 
  ungroup() 

## need to split the back yard chickens eggs&meat
fix_chickens_1 <- summary_df_int %>% 
  filter(tier_5 == "chickens eggs&meat") 
fix_chickens_1 <- rbind(fix_chickens_1, fix_chickens_1) %>% 
  arrange(pressure) %>% 
  mutate(tier_5 = rep(c("chickens meat", "chickens eggs"), times = 4, each = 1))

fix_chickens_2 <- summary_df_int %>% 
  filter(tier_5 == "chickens eggs&meat feed") 
fix_chickens_2 <- rbind(fix_chickens_2, fix_chickens_2) %>% 
  arrange(pressure) %>% 
  mutate(tier_5 = rep(c("chickens meat", "chickens eggs"), times = 4, each = 1))

## add back
summary_df <- summary_df_int %>% 
  filter(tier_5 != "chickens eggs&meat",
         tier_5 != "chickens eggs&meat feed") %>% 
  rbind(fix_chickens_1, fix_chickens_2) %>% 
  group_by(tier_5, pressure) %>% 
  dplyr::summarise(sum = sum(sum)) %>% 
  left_join(rescaling_values, by = "pressure") %>% 
  ungroup() %>% 
  rowwise() %>% 
  mutate(prop_of_global = sum/global_total,
         prop_of_global = ifelse(tier_5 == "fofm meat", prop_of_global*0.1, prop_of_global)) %>% 
  left_join(food_groupings, by = "tier_5")

## quick check
sum(summary_df$prop_of_global) #  3.968501 looks good!
```

Now extract the feed from crops and assign to the crop it comes from. This takes A LONG TIME. NOTE: This only needs to be run once, and it was run on May 25th, 2022. Maybe there is a way to speed this up? There are ~2700 files that we loop through. 


```{r}

crop_feed_files <- grep(list.files(file.path("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/raw_feed_combos"), full.names = T, pattern = "crop_produce"), pattern= 'human', invert=TRUE, value=TRUE)




empty_df <- data.frame(tier_5 = NA, pressure = NA, sum = NA)

for(file in crop_feed_files){
  
  # file <- crop_feed_files[[1]]
  

  index <- match(file, crop_feed_files)
  
  file_name <- str_after_last(file, "\\/")
  
  crop_long <- str_after_first(file_name, "_x_")
  
  crop_name <- str_after_first(crop_long, "_")
  
    
  tier_5_name <- paste(str_before_first(crop_name, "_"), "produce feed")
  
  tier_2_name <- str_before_first(crop_name, "_")
  
  pressure_name <- str_after_last(file_name, "_")
  
  pressure_name <- str_before_first(pressure_name, "\\.")
  
  
  
  rast <- raster(file)
  
  sum_value <- cellStats(rast, "sum", na.rm = TRUE)

  new_df <- data.frame(tier_5 = tier_5_name, pressure = pressure_name, sum = sum_value)
  
  empty_df <- rbind(empty_df, new_df)
  
  cat(paste("done", index, "out of 2704"))
  
  }


crop_feed_df <- empty_df %>%
  group_by(tier_5, pressure) %>%
  summarise(sum = sum(sum)) %>%
  left_join(rescaling_values) %>%
  filter(!is.na(global_total)) %>%
  mutate(prop_of_global = sum/global_total) %>%
  mutate(tier_2 = "crop feed")

sum(crop_feed_df$prop_of_global) # 0.5413687

write.csv(crop_feed_df, here("_analysis/figures/extended_data/prep_Rmd/data/crop_feed_bar_fig.csv"), row.names = FALSE)


```

Combine with the derivative dataset of rgn_raw_summary

```{r}
crop_feed_df <- read.csv(here("_analysis/figures/extended_data/prep_Rmd/data/crop_feed_bar_fig.csv"))

sum(crop_feed_df$prop_of_global) # 0.5413687

check <- summary_df %>%
  filter(tier_1 %in% c("crop human consumption", "feed"))

sum(check$prop_of_global) # 2.283848; this number should be a higher than the sum above, since this includes FOFM feed. And it is. 

crop_human_df <- summary_df %>%
  filter(tier_2 == "crop human consumption") %>%
  dplyr::select(tier_5, pressure, sum, global_total, prop_of_global, tier_2)


all_crop_human_feed_df <- rbind(crop_feed_df, crop_human_df)

```


Colors **might want to darken the colors a bit 
```{r}
dis <- "#D5BD75"
ghg <- "#DA6098"
nut <- colorspace::lighten("#795B48", 0.4)
wat <- "#77C1D7"

dis_f <- "#93875F"
ghg_f <- "#965372"
nut_f <- "#5A4D44"
wat_f <- "#5B7A84"

pal <- c(dis, dis_f, ghg, ghg_f, nut, nut_f, wat, wat_f)
scales::show_col(pal, n=2)



```


Create DF that'll work for the plots
```{r}
bar_df_int <- all_crop_human_feed_df %>% 
  select(tier_2, tier_5, pressure, prop_of_global) %>% 
  mutate(sub_tier = str_remove(tier_5, " feed"),
         pressure_bar = ifelse(str_detect(tier_5, "feed"), paste(pressure, "_feed", sep = ""), pressure)) %>%
  group_by(tier_2, sub_tier) %>% 
  mutate(food_cat_sum = sum(prop_of_global)) %>% 
  ungroup()  %>% 
  mutate(fill_value = case_when(pressure_bar == "disturbance" ~ dis,
                                pressure_bar == "ghg" ~ ghg,
                                pressure_bar == "nutrient" ~ nut,
                                pressure_bar == "water" ~ wat,
                                pressure_bar == "disturbance_feed" ~ dis_f,
                                pressure_bar == "ghg_feed" ~ ghg_f,
                                pressure_bar == "nutrient_feed" ~ nut_f,
                                pressure_bar == "water_feed" ~ wat_f)) %>% 
  mutate(prop_of_global = prop_of_global/4)

## make naming changes to make it look better
bar_df <- bar_df_int%>% 
# mutate(sub_tier = ifelse(facet_tier ==  "Crops", str_remove(tier_5, " produce"), sub_tier)) %>% 
   mutate(sub_tier =  str_remove(sub_tier, " produce")) %>% 
  left_join(mapspam_names, by = "sub_tier") %>% 
  mutate(SPAM_full_name = ifelse(sub_tier == "xfru", "other fruits",
                                 ifelse(sub_tier == "xmil", "millet",
                                        ifelse(sub_tier == "xoil", "other oil crops",
                                               ifelse(sub_tier == "xpul", "pulses", 
                                                      ifelse(sub_tier == "tnut", "tree nuts", 
                                                             ifelse(sub_tier == "spis", "spices", SPAM_full_name))))))) %>% 
  mutate(sub_tier = ifelse(tier_2 %in% c("crop human consumption", "crop feed"), SPAM_full_name, sub_tier)) %>% 
  select(-SPAM_full_name)


  


plot_order <- bar_df %>% 
  select(sub_tier, food_cat_sum) %>% 
  unique() %>% 
  group_by(sub_tier) %>%
  summarise(food_cat_sum = sum(food_cat_sum)) %>%
  arrange(food_cat_sum) %>% 
  pull(sub_tier) 

bar_df$sub_tier <- factor(bar_df$sub_tier, levels = plot_order)

bar_df$fill_value <- factor(bar_df$fill_value, levels = c(dis, dis_f, ghg, ghg_f, nut, nut_f, wat, wat_f))


```


make the plots individually so can manage the size of the bars
```{r}


crops_all <- ggplot(bar_df, aes(y=sub_tier, x = prop_of_global, fill = fill_value)) + 
#  geom_vline(data = filter(bar_df, facet_tier %!in% c("Marine and Freshwater Fisheries", "Mariculture")), aes(xintercept = 0.05), color = "grey33", linetype = "F1") +
#  geom_vline(data = filter(bar_df, facet_tier !=  "Mariculture"), aes(xintercept = 0.025), color = "grey33", linetype = "twodash")+
#  geom_vline(xintercept = 0.002, color = "grey33", linetype = "dotted", size = 1) +
  geom_bar(stat="identity", width = 0.5) +
  theme_minimal() +
   scale_fill_identity() +
  labs(y = "Crops",
       x = "Proportion of global environmental pressures") +
  theme(axis.title.y = element_text(size = 12),
        axis.text.y = element_text(size = 10),
        axis.text.x = element_text(size = 10),
        axis.title.x = element_text(size = 12))


crops_all

ggsave(here("_analysis/figures/extended_data/output/ed_fig_7_crops_feed.png"), units=c("in"), width=9)
```


















