---
title: 'SI circle plots: ranking countries'
author: "Juliette"
date: "11/17/2020"
output: html_document
---

Ranking countries.

a) Radial plot of each country, with stacked bars for amount of each of 4 stressors (from all foods). Outside ring is land and freshwater; inner ring is country EEZs (asterix for land-locked countries);


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(hrbrthemes)
library(RColorBrewer)
library(colorspace)
source(here("_workflow/common.R"))

summary_df_raw <- read_csv(file.path("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/raw_df_summary/rgn_raw_summary.csv"))

rescaling_values <- read_csv(here("_analysis/rescale_values.csv"))

un_geopolitical <- read_csv(here("_spatial/_output/UNSD_Methodology.csv")) %>%
  dplyr::select(iso3c, georegion=Region_Name)

```


Create DF that works for the circle plot
```{r}
summary_df <- summary_df_raw %>% 
  mutate(origin = ifelse(category == "feedfofm", "marine", origin)) %>% 
  mutate(land_ocean = case_when(origin %in% c("land", "freshwater") ~ "land",
                                origin == "marine" ~ "ocean")) %>% 
  mutate(sum = ifelse(iso3c == "HSX" & category == "feedfodd", 0, sum)) %>% 
  group_by(country, iso3c, land_ocean, pressure) %>% 
  dplyr::summarise(pressure_sum = sum(sum)) %>% 
  left_join(rescaling_values, by = "pressure") %>% 
  ungroup() %>% 
  rowwise() %>% 
  mutate(prop_of_global = pressure_sum/global_total)  %>% 
  unite(source, c("land_ocean", "pressure"), sep = "-", remove = FALSE) %>% 
  left_join(un_geopolitical, by = "iso3c") %>% 
  mutate(georegion = ifelse(country == "High Seas", "High Seas", georegion))

## get georegion ranks
chi <- summary_df %>%
  group_by(georegion, country, iso3c) %>%
  summarize(chi = sum(prop_of_global, na.rm=TRUE)) %>%
  arrange(-chi)

rank_georgn <- chi %>%
  group_by(georegion) %>%
  summarize(sum_chi = sum(chi), ## we can also do it by mean
            count = length(chi)) %>%
  ungroup() %>%
  data.frame() %>%
  arrange(sum_chi)

rank_georgn$georegion <- factor(rank_georgn$georegion, 
                                         levels=unique(rank_georgn$georegion))
chi$georegion <- factor(chi$georegion, 
                                         levels=unique(rank_georgn$georegion))

### Organize the chi data and add a few extra variables to help plotting
# Including empty spaces to add a y-axes
rank_rgn <- chi %>%
  arrange(georegion, chi) %>%
  data.frame()
 
```

```{r}
 ## modfiy region names to be shorter
 rank_rgn <- rank_rgn %>% 
   mutate(rgn_name_short = country, 
         rgn_name_short = gsub("Islands", "Isl", rgn_name_short),
         rgn_name_short = gsub("Island", "Isl", rgn_name_short),
         rgn_name_short = gsub("Democratic", "Dem", rgn_name_short),
         rgn_name_short = gsub("Republic", "Rep", rgn_name_short),
         rgn_name_short = gsub("South", "S", rgn_name_short),
         rgn_name_short = gsub("American", "Am", rgn_name_short),
         rgn_name_short = gsub("the United States", "US", rgn_name_short),
         rgn_name_short = gsub("Territory", "Terr", rgn_name_short),
         rgn_name_short = gsub("Saint", "St", rgn_name_short),
         rgn_name_short = gsub(" and ", " & ", rgn_name_short),
         rgn_name_short = gsub("Republique", "Rep", rgn_name_short),
         rgn_name_short = gsub("Dem Rep of the", "Dem Rep of", rgn_name_short),
         rgn_name_short = gsub("Georgia and the", "Georgia and", rgn_name_short),
         rgn_name_short = gsub("St Vincent and the", "St Vincent and", rgn_name_short),
         rgn_name_short = gsub("Howland Isl & Baker Isl", "Howland & Baker Isl", rgn_name_short),
         rgn_name_short = gsub("Northern", "N", rgn_name_short), 
         rgn_name_short = gsub("Reunion", "Reunion", rgn_name_short))
```


## Stuff for both plots
i think that this would look better with each stressor having its own color that is paired with ocean/land

pick 4 distinct colors and then create a pair with one being more brown or green and the other being more blue
```{r}

source_name <- data.frame(source = c("land-disturbance", "land-ghg", "land-water", "land-nutrient",
                                     "ocean-disturbance", "ocean-ghg", "ocean-water", "ocean-nutrient"),
                          ## this may be where i remove the part before the source later on
                            source_name = c("Land: Disturbance", "Land: GHG", "Land: Water", "Land: Nutrient",
                                     "Ocean: Disturbance", "Ocean: GHG", "Ocean: Water", "Ocean: Nutrient"))

dist_1 <- "#D5BD75"
dist_1 <- lighten(dist_1, amount = -0.05)
dist_2 <- "#FFEAB1"
dist_2 <- lighten(dist_2, amount = -0.05)

ghg_1 <- "#DA6098"
ghg_1 <- lighten(ghg_1, amount = -0.05)
ghg_2 <- "#FFCEE0"
ghg_2 <- lighten(ghg_2, amount = -0.05)

nut_1 <- "#795B48"
nut_2 <- "#E7C7B5"

wat_1 <- "#3E638E"
wat_2 <- "#C5F0FF"

jv_palette <- c(dist_1, ghg_1, nut_1, wat_1,
                dist_2, ghg_2, nut_2, wat_2)
scales::show_col(jv_palette, n = 4)
#dist, ghg, water, nut

pal_df <- as_tibble(jv_palette) %>% 
  rename(fill_value = value) %>% 
  mutate(source = case_when(fill_value == dist_1 ~ "land-disturbance",
                            fill_value == dist_2 ~ "ocean-disturbance",
                            fill_value == wat_1 ~ "land-water",
                            fill_value == wat_2 ~ "ocean-water",
                            fill_value == nut_1 ~ "land-nutrient",
                            fill_value == nut_2 ~ "ocean-nutrient",
                            fill_value == ghg_1 ~ "land-ghg",
                            fill_value == ghg_2 ~ "ocean-ghg")) 


```


## circle part

```{r}

#filter to include only countries in circle

# cutoff_value <-  0.04653524 

cutoff_value <- 0.03531988 # ukraine chi value; this is the value that corresponds with the top 23 regions (i.e. the circle plot includes all regions but the top 23)

pressures_adj_circle <- summary_df %>%
  group_by(country, iso3c) %>%
  mutate(chi = sum(prop_of_global, na.rm=TRUE)) %>%
  ungroup() %>%
  arrange(chi) %>% 
 filter(chi< cutoff_value)


pressures_adj_circle_test <- summary_df %>%
  group_by(country, iso3c) %>%
  mutate(chi = sum(prop_of_global, na.rm=TRUE)) %>%
  ungroup() %>%
  arrange(chi) %>%
# filter(chi>cutoff_value) %>%
  mutate(prop_of_global_test = prop_of_global/4) %>%
  mutate(chi = chi/4)

sum(pressures_adj_circle_test$prop_of_global_test)
# [1] 0.9999999

sum(pressures_adj_circle_test$prop_of_global)
# 3.999999


# temp <- summary_df %>%
#   group_by(country, iso3c) %>%
#   mutate(chi = sum(prop_of_global, na.rm=TRUE)) %>%
#   ungroup() %>%
#   dplyr::select(country, chi) %>%
#   unique()
# hist(temp$chi)


##grab the highest country to highlight red and id where the circle ends and then side bar chart begins
highest_country <- pressures_adj_circle[dim(pressures_adj_circle)[1], "iso3c"]
pressures_highest_country <- filter(pressures_adj_circle, iso3c %in% highest_country$iso3c) #used below

# adjust the rank_rgn-adjust dataframe to match above
rank_rgn_adj <- rank_rgn %>%
  filter(iso3c %in% pressures_adj_circle$iso3c) 

# add empty space to create space in the circleplot for axes
 empty_bar <- 5
 to_add <- data.frame(matrix(NA, empty_bar, ncol(rank_rgn_adj)) )
 colnames(to_add) = colnames(rank_rgn_adj)
 to_add$country <- as.character(1:empty_bar)
 rank_rgn_adj  <- rbind(to_add, rank_rgn_adj)
 
 # some code to orient the country labels
sequence_length = length(unique(rank_rgn_adj$country))
first_sequence = c(1:(sequence_length%/%2)) 
second_sequence = c((sequence_length%/%2+1):sequence_length) 
first_angles = c(90 - 180/length(first_sequence) * first_sequence)
second_angles = c(-90 - 180/length(second_sequence) * second_sequence)
rank_rgn_adj$angle <- c(first_angles, second_angles)
rank_rgn_adj$hjust <- c(rep(0, length(first_sequence)), 
                    rep(1, length(second_sequence)))

# color for region labels
rank_rgn_adj <- rank_rgn_adj %>%
  mutate(country = factor(country, unique(country))) %>%
  mutate(georegion = factor(georegion, unique(georegion))) %>%
  mutate(color = "black") %>%
  mutate(color = ifelse(is.na(chi), "white", color)) %>%
  mutate(color = ifelse(iso3c %in% highest_country$iso3c, "red", color))

# get dataframe for georegion shift locations
rgn_shift <- rank_rgn_adj %>%
  mutate(georegion = ifelse(is.na(georegion), "tmp", georegion)) %>%
  mutate(georegion = as.factor(georegion)) %>%
  mutate(region_shift = as.numeric(georegion) - lag(as.numeric(georegion)), default=first(as.numeric(georegion)))

rgn_shift <- which(rgn_shift$region_shift > 0)
rgn_shift <- c(1, rgn_shift) -0.5
rgn_shift <- data.frame(rgn_shift_x=rgn_shift,
                         georegion = rank_georgn$georegion,
                         name_x= c(2, 15, 60, 115, 150, 200), # 140
                         name_y=c(55, 55, 55, 55, 55, 55))
rgn_shift <- rgn_shift %>%
   mutate(georegion = as.character(georegion))

# # add some blanks rows to match the master region list
empty_bar2 <-10
to_add <-  data.frame( matrix(NA, empty_bar2*nlevels(as.factor(pressures_adj_circle$source)), ncol(pressures_adj_circle)) )
  colnames(to_add) <- colnames(pressures_adj_circle)
  to_add$source <- rep(levels(as.factor(pressures_adj_circle$source)), each=empty_bar2)
  to_add$chi<-  0
  to_add$country <- as.character(rep(1:empty_bar2, nlevels(as.factor(pressures_adj_circle$source))))
  pressures_adj_circle <- rbind(to_add, pressures_adj_circle, to_add)
  
  pressures_adj_circle %>%
  group_by(source) %>%
  summarize(mean = mean(prop_of_global, na.rm=TRUE)) %>%
  arrange(mean)
  
pressures_adj_circle <- pressures_adj_circle %>%
  left_join(source_name, by = "source")

pressures_adj_circle$source_name <- factor(pressures_adj_circle$source_name, levels=rev(source_name$source_name))
pressures_adj_circle$country <- factor(pressures_adj_circle$country, levels=unique(rank_rgn_adj$country))
pressures_adj_circle$georegion <- factor(pressures_adj_circle$georegion, 
                                         levels=unique(rank_georgn$georegion))

```

make the plot
```{r}

circle_theme <- theme(axis.line=element_blank(),
                      axis.text.y=element_blank(),
                      axis.ticks=element_blank(),
                      axis.title.x=element_blank(),
                      axis.title.y=element_blank(),
                      legend.position="none",
                      # legend.position="left",
                      panel.background=element_blank(),
                      panel.border=element_blank(),
                      panel.grid.major=element_blank(),
                      panel.grid.minor=element_blank(),
                      plot.background=element_blank(),
                      axis.text.x = element_blank())

color_order <- c(dist_2, ghg_2, wat_2, nut_2, dist_1, ghg_1, wat_1, nut_1)

## WE DIVIDE EVERYTHING BY 4 HERE SO THAT WE CAN GET PROPORTION OUT OF 1 INSTEAD OF 4

pressures_adj_circle_plot <- pressures_adj_circle %>% 
  mutate(prop_of_global_stacked = abs(prop_of_global)/4) %>% 
  left_join(pal_df) %>% 
  mutate(fill_value = factor(fill_value, levels = color_order))

sum(pressures_adj_circle_plot$prop_of_global, na.rm = TRUE) # 1.179597

test <- pressures_adj_circle_plot %>%
  filter(!is.na(prop_of_global)) %>%
  mutate(testing = prop_of_global/4) 

sum(test$testing) # 0.2948992; this plot accounts for about 29% of all food here

rgn_shift <- rgn_shift %>% 
  mutate(georegion = ifelse(georegion == "High Seas", " ", georegion))

  ggplot(data=pressures_adj_circle_plot, aes(x=country, y=prop_of_global_stacked, fill = fill_value)) + 
  
  geom_bar(stat="identity") +
  
  # ## this adjust the inner circle size (ymin) and outer circle size (ymax), which can be expanded to include names
   geom_errorbar(aes(
       x     = 1, 
        ymin = (-0.01/4),   # ymin  = -0.03, this is for when ocean bars stack inwards -0.01 for when not
       ymax  = 0.00875),  # 0.035
       alpha = 0) +
      # Country name text placement
     geom_text(
       data = rank_rgn_adj,
       aes(x     = country,
           y     = max(pressures_adj_circle$chi/4, na.rm=TRUE),
           label = rgn_name_short,
           angle = angle,
           hjust = hjust),
       inherit.aes = FALSE,
       fontface = "bold",
       alpha = 0.9,
       size = 5,
       color = rank_rgn_adj$color) +
    # 
   # georegion names
    geom_text(
      data=rgn_shift,
      aes(y = 0.00375, x = name_x, label = georegion), # y=0.02
      # aes(x=name_x,
      #     y=name_y,
      #     label=georegion),
      inherit.aes=FALSE, size=7) +
  #  
  # # # Scale bar circles
  # ## This adds the 0 line
      geom_segment(
        x = 0, 
         y = 0, 
         xend = dim(rank_rgn_adj)[1]+1, 
         yend = 0, 
         fill = NULL,
        alpha = 1, color = "white", size = 0.1) + #color = "black for internal
  #    
  #  # Scale bar annotation, this adds the y-axis numbers
      # annotate("text",
      #   x     = c(3, 3, 3),
      #   y     = c(0, 0.0175, 0.035),
      #   label = c(0, 0.0175, 0.035),
      #   color = "darkgray", angle = -4, size = 4) +
    
          annotate("text",
        x     = c(3, 3, 3),
        y     = c(0, 0.004375, 0.00875),
        label = c(0, 0.004375, 0.00875),
        color = "darkgray", angle = -4, size = 5) +
    
    
  #     
 scale_color_identity() +
 scale_fill_identity() +
   # scale_fill_manual(values= rev(myPalette), c("Land: Disturbance", "Land: GHG", "Land: Water", "Land: Nutrient",
   #                                    "Ocean: Disturbance", "Ocean: GHG", "Ocean: Water", "Ocean: Nutrient")) +
  # scale_fill_manual(values= rev(jv_palette), c("Land: Disturbance", "Land: GHG", "Land: Water", "Land: Nutrient",
  #                                     "Ocean: Disturbance", "Ocean: GHG", "Ocean: Water", "Ocean: Nutrient")) +
  coord_polar() +
 circle_theme 
  
  ggsave(here("_analysis/figures/extended_data/output/ed_fig_1_circle_country.png"), height=18, width=18, units=c("in"))
# ggsave(here("_analysis/figures/extended_data/output/circle_plots_no_legends/ed_fig_4_circle_country.png"), height=18, width=18, units=c("in"))
# # bg="transparent"

```

