---
title: "shrimp_nutrient_pollution"
author: "Gage Clawson (NCEAS, UCSB, OHI)"
date: "12/22/2020"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    includes: 
toc: true
---

This markdown uses models and species specific parameters from the paper *Aquaculture Production is a Large, Spatially Concentrated Source of Nutrients in Chinese Freshwater and Coastal Seas* (2020 Wang et al)

## Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(sf)
library(raster)
library(readr)
library(data.table)

source(here("_spatial/template_raster.R"))

path_prep <- "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/shrimp"
path_final <-"/home/shares/food-systems/Food_footprint/all_food_systems/datalayers"
path_pre <- "/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers"

production <- read_csv(file.path(path_prep, "updated/shrimp_farm_updated.csv"))
food_raster <- raster(nrows=2160, ncols=4320, xmn=-180, xmx=180, ymn=-90, ymx=90)
```

shrimp N/P pollution values

N: 57 kg/tonne of growth
P: 9.2 kg/tonne of growth

Steps: 
1. Multiply tonnes production by 0.95 (this corrects for the growth that occurs outside the seapen on land), this is the value Lex uses
2. Then multiply growth by the N value and the P value
3. We report in total excess nutrients: Sum the N and P values.
4. We report in tonnes, so divide by 1000. 

```{r}

shrimp_pollution <- production %>%
  dplyr::mutate(prod_growth = tonnes_production*0.95) %>%
  dplyr::mutate(prod_growth_N_poll = prod_growth*57,
                prod_growth_P_poll = prod_growth*9.2) %>% ## in kg pollution now
  # dplyr::mutate(PO4eq_kg = prod_growth_N_poll*0.42 + prod_growth_P_poll*3.07) %>% cut (01/14/2021)
  dplyr::mutate(N_P_tonnes = (prod_growth_N_poll + prod_growth_P_poll)/1000,
                N_tonnes = prod_growth_N_poll/1000,
                P_tonnes = prod_growth_P_poll/1000) %>%
  dplyr::select(x, y, N_P_tonnes, N_tonnes, P_tonnes)

sum(shrimp_pollution$N_P_tonnes) # 320602.1
sum(production$tonnes_production) # 5097823

## save to dataprep folder: 
write.csv(shrimp_pollution, file.path(path_prep, "shrimp_nutrient.csv"))


## save N + P layer
N_P_df <- shrimp_pollution %>%
  dplyr::select(x,y, N_P_tonnes)

final_rast <- rasterFromXYZ(N_P_df, crs = food_crs)
crs(final_rast)
plot(final_rast)
final_rast
cellStats(final_rast, "sum") # 320602.1

final_area_fix <- final_rast/raster::area(final_rast)

final_rast_proj <- projectRaster(final_area_fix, food_raster, method = "ngb")

final_rast <- final_rast_proj*raster::area(final_rast_proj)
cellStats(final_rast, "sum") # 320602.1
final_rast

## Save to datalayers 
writeRaster(final_rast, file.path(path_final, "marine_shrimp_aquaculture_meat_nutrient.tif"), overwrite = T)


## save N layer
N_df <- shrimp_pollution %>%
  dplyr::select(x,y, N_tonnes)

final_rast <- rasterFromXYZ(N_df, crs = food_crs)
crs(final_rast)
plot(final_rast)
final_rast
cellStats(final_rast, "sum") # 276047.1

final_area_fix <- final_rast/raster::area(final_rast)

final_rast_proj <- projectRaster(final_area_fix, food_raster, method = "ngb")

final_rast <- final_rast_proj*raster::area(final_rast_proj)
cellStats(final_rast, "sum") # 276047.1
final_rast

## Save to datalayers 
writeRaster(final_rast, file.path(path_pre, "marine_shrimp_aquaculture_meat_nutrientN.tif"), overwrite = T)



## save P layer
P_df <- shrimp_pollution %>%
  dplyr::select(x,y, P_tonnes)

final_rast <- rasterFromXYZ(P_df, crs = food_crs)
crs(final_rast)
plot(final_rast)
final_rast
cellStats(final_rast, "sum") # 44554.97 

final_area_fix <- final_rast/raster::area(final_rast)

final_rast_proj <- projectRaster(final_area_fix, food_raster, method = "ngb")

final_rast <- final_rast_proj*raster::area(final_rast_proj)
cellStats(final_rast, "sum") # 44554.97
final_rast

## Save to datalayers 
writeRaster(final_rast, file.path(path_pre, "marine_shrimp_aquaculture_meat_nutrientP.tif"), overwrite = T)

```


Data check: 

Values from this paper: https://pubs.acs.org/doi/pdf/10.1021/acs.est.9b03340


   - 57 kg N / tonne
   - 9.2 kg P / tonne 

I believe the estimates from their paper for "crustaceans" lumps shrimp in with that, so we just used the same values for shrimp and crustaceans. There are no issues with the code. 

When just looking at China (and lumping shrimp/crustacean production together for 2017) and comparing to the paper, we see this: 

Our results:
 - 88328.67 tonnes N
 - 14256.56 tonnes P
 
Their results: 
 - 102000 tonnes N
 - 16600 tonnes P
 
This seems pretty reasonable to me... 


```{r}
shrimp_nutrient_N <- raster(file.path("/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers/marine_shrimp_aquaculture_meat_nutrientN.tif"))
cellStats(shrimp_nutrient_N, "sum") # 275847.3 N tonnes

shrimp_nutrient_P <- raster(file.path("/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers/marine_shrimp_aquaculture_meat_nutrientP.tif"))
cellStats(shrimp_nutrient_P, "sum") # 44522.72 P tonnes

sum(production$tonnes_production) # 5094133

## Total production is 5094133 tonnes for 2017

## Used 57 N kg and  9.2 P kg for shrimp

5094133*0.95*57/1000 # 275847.3 N tonnes
5094133*0.95*9.2/1000 # 44522.72 P tonnes

## both match our final layers 

final_rast <- raster(file.path(path_final, "marine_shrimp_aquaculture_meat_nutrient.tif"))
cellStats(final_rast, "sum") # 320370


## add crustaceans to this to check against paper
275847.3 + 21753.09 # 297600.4
44522.72 + 3511.025 # 48033.75

## now look at china production and see if it matches up 
china_shrimp_prod <- read.csv(file.path(here(),"aquaculture/marine/shrimp/shrimp_farms/data/fao_mariculture_shrimp.csv")) %>%
  filter(iso3c == "CHN",
         year == 2017) # 1345154

china_crust_prod <-  read.csv(file.path(here(),"aquaculture/marine/crustaceans/crustacean_farms/data/fao_mariculture_crustaceans.csv")) %>%
  filter(iso3c == "CHN",
         year == 2017) #286031

((1345154+286031)*0.95*57)/1000 # 88328.67
((1345154+286031)*0.95*9.2)/1000 # 14256.56


```

