---
title: "New disturbance"
output: html_document
editor_options: 
  chunk_output_type: console
---

Here we:
1. Combine the two metrics (habitat destruction and biomass removal) of marine fisheries disturbance by adding the two rescaled rasters and dividing by 2.  
2. Resample the data to match the higher resolution food rasters
3. Multiply by cell area to get km2eq and save to datalayers.

```{r setup, include=FALSE}
#load relevant packages, etc.

library(here)
library(raster)
library(tidyverse)

# raster template
food_raster_template <- raster(nrows=2160, ncols=4320, xmn=-180, xmx=180, ymn=-90, ymx=90)
fishery_raster_template <- raster(list.files("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/rasters/destructive_proportions", full=TRUE)[1])
```

Combine the biomass removal and habitat destruction rasters for each food group (sum and divide by 2).
```{r}

food_cats <- data.frame(filename = c("Benthic", "Demersal", "fofm", "Large pelagic", "Medium pelagic", "Reef-associated", "Small pelagic"),
                        savename = c("benthic", "demersal", "fofm", "large-pelagic", "medium-pelagic", "reef", "small-pelagic"))

biomass <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/rasters/biomass_removal", full=TRUE)
habitat <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/rasters/benthic_habitat_destruction", full=TRUE)

for(food_type in food_cats$filename){  #food_type = food_cats$filename[4]
  cat(food_type, "\n")
  biomass_file <- raster(grep(food_type, biomass, value=TRUE))
  biomass_file[is.na(biomass_file)] <- 0
  habitat_file <- raster(grep(food_type, habitat, value=TRUE))
  habitat_file[is.na(habitat_file)] <- 0
  cat("this should be 2 =", length(biomass_file) + length(habitat_file), "\n")
#plot(biomass_file)
#plot(habitat_file)
  pressure <- (biomass_file + habitat_file)/2
  pressure_food_scale <- resample(pressure, food_raster_template, method="ngb")
  cat(minValue(pressure_food_scale), " to ", maxValue(pressure_food_scale), "\n", "\n") 
  pressure_km2eq <- pressure_food_scale * area(pressure_food_scale)
  
#cellStats(pressure_km2eq, "sum", na.rm=TRUE)  
  savename = food_cats$savename[food_cats$filename==food_type]
  writeRaster(pressure_km2eq, sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/datalayers/marine_%s_fisheries_meat_disturbance.tif", savename), overwrite=TRUE)
  }

```

### some checks
```{r}
# make sure total of all rasters doesn't exceed 1

marine_disturb <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/datalayers", pattern="fisheries_meat_disturbance.tif", full=TRUE)

marine_disturb <- grep("freshwater", marine_disturb, invert=TRUE, value=TRUE)

total_marine_disturb <- calc(stack(marine_disturb), sum, na.rm=TRUE)
prop_total_marine_disturb <- total_marine_disturb/area(total_marine_disturb)
prop_total_marine_disturb
plot(prop_total_marine_disturb)

```


Compare to previous rasters
```{r}

old_files <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/archive_marine_fish_disturbance", full=TRUE)

old_files_stack <- stack(old_files)
old_area <- cellStats(old_files_stack, "sum", na.rm=TRUE)
sum(old_area)

marine_disturb <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/datalayers", pattern="fisheries_meat_disturbance.tif", full=TRUE)

marine_disturb <- grep("freshwater", marine_disturb, invert=TRUE, value=TRUE)
new_files <- stack(marine_disturb)
new_area <- cellStats(new_files, "sum", na.rm=TRUE)
sum(new_area)

(new_area-old_area)/old_area

(5047749 - 4817619)/4817619

```


### Scenario analysis

Here we are combining the data for fisheries using a 99.99th rescaling value (vs. 99.9th) to determine our level of uncertainty.

Combine the biomass removal and habitat destruction rasters for each food group (sum and divide by 2).
```{r}

food_cats <- data.frame(filename = c("Benthic", "Demersal", "fofm", "Large pelagic", "Medium pelagic", "Reef-associated", "Small pelagic"),
                        savename = c("benthic", "demersal", "fofm", "large-pelagic", "medium-pelagic", "reef", "small-pelagic"))

biomass <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/scenarios/scenarios_dataprep/fis_biomass_removal", full=TRUE)
habitat <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/scenarios/scenarios_dataprep/fis_benthic_destructive", full=TRUE)

for(food_type in food_cats$filename){  #food_type = food_cats$filename[4]
  cat(food_type, "\n")
  biomass_file <- raster(grep(food_type, biomass, value=TRUE))
  biomass_file[is.na(biomass_file)] <- 0
  habitat_file <- raster(grep(food_type, habitat, value=TRUE))
  habitat_file[is.na(habitat_file)] <- 0
  cat("this should be 2 =", length(biomass_file) + length(habitat_file), "\n")
#plot(biomass_file)
#plot(habitat_file)
  pressure <- (biomass_file + habitat_file)/2
  pressure_food_scale <- resample(pressure, food_raster_template, method="ngb")
  cat(minValue(pressure_food_scale), " to ", maxValue(pressure_food_scale), "\n", "\n") 
  pressure_km2eq <- pressure_food_scale * area(pressure_food_scale)
  
#cellStats(pressure_km2eq, "sum", na.rm=TRUE)  
  savename = food_cats$savename[food_cats$filename==food_type]
  writeRaster(pressure_km2eq, sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/scenarios/scenarios_datalayers/marine_%s_fisheries_meat_disturbance.tif", savename), overwrite=TRUE)
  }

```

