---
title: "New disturbance"
output: html_document
editor_options: 
  chunk_output_type: console
---

Exploring and preparing Global Fishing Watch data.
1. Combine daily effort data into annual effort data
2. Explore how big an issue unidentified geartypes are in the data (total proportion of effort and spatial distriubtion of unidentified geartypes)
3. Create maps of effort for destructive geartypes: trawl and dredge, which will be used in subsequent scripts.


## Data description

Data download: https://globalfishingwatch.org/data-download/datasets/public-fishing-effort
May 18 2021, MRF

Description: Fleet daily v2
Resolution: daily, 100th degree
Year: 2017

Geartypes:
- fishing: a combination of vessels of unknown fishing gear
 - drifting_longlines: drifting longlines
 - seiners: vessels using seine nets, including potential purse seine vessels
   targeting tuna and other species, as well as danish and other seines
     - purse_seines: purse seines, both pelagic and demersal
        - tuna_purse_seines: large purse seines primarily fishing for tuna.
        - other_purse_seines: purse seiners fishing for mackerel, anchovies, etc, often smaller and operating nearer the coast than tuna purse seines.
    - other_seines: danish seines and other seiners not using purse seines.
 - trawlers: trawlers, all types
 - pole_and_line: vessel from which people fish with pole and line.
 - trollers: vessel that tows multiple fishing lines.
 - fixed_gear: a category that includes potential set longlines, set gillnets,  and pots and traps
     - pots_and_traps: vessel that deploys pots (small, portable traps) or traps to
       catch fish
     - set_longlines: vessel that fishes by setting longlines anchored to the
       seafloor. These lines have shorter hooked, typically baited, lines hanging
       from them
     - set_gillnets: vessel that fishes by setting gillnets anchored to the seafloor.
 - dredge_fishing: vessel that tows a dredge the scrapes up edible bottom
   dwellers such as scallops or oysters.
 - squid_jigger: squid jiggers, mostly large industrial pelagic operating vessels

```{r setup, include=FALSE}
#load relevant packages, etc.

library(here)
library(raster)
library(tidyverse)

# raster template
r_template <- raster::raster(ncol=720, nrow=360, vals=c(1:259200))

  
```

Step 1: Summarize effort data
The effort data are provided as separate csvs for each day. Each row provides the following data:
date: 1/1/17
cell_ll_lat: 2.25
cell_ll_lon: -138.81
flag: USA
geartype: series of gear types
hours: 0.2575
fishing_hours: 0
mmsi_present: 1


I am going to open each of these files and sum fishing_hours by: lat, lon, geartype 
And save the resulting files.

```{r}

fish_effort <- list.files("/home/shares/food-systems/Food_footprint/_raw_data/GlobalFishingWatch_effort/d2020", pattern=".csv", full=TRUE)

fish_lat_long_gear <- function(filepath){
   #filepath <- fish_effort[1]
  saveName <- basename(filepath)
  day_file <- read_csv(filepath, col_types = "Dddccddn")
  day_file <- day_file %>%
    group_by(cell_ll_lat, cell_ll_lon, geartype) %>%
   summarize(fishing_hours = sum(fishing_hours)) %>%
    data.frame()

  write_csv(day_file, sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/effort_summarized_by_cell_gear/summary_%s", saveName))
  }

for(filepath in fish_effort){
  fish_lat_long_gear(filepath)
}


```


Now combine the daily csv files.
```{r}

fish_effort_df <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/effort_summarized_by_cell_gear/", pattern = "summary", full=TRUE) %>%
  lapply(data.table::fread) %>%
  bind_rows

fish_effort_annual <- fish_effort_df %>%
  group_by(cell_ll_lat, cell_ll_lon, geartype) %>%
   summarize(fishing_hours = sum(fishing_hours)) %>%
    data.frame()

write_csv(fish_effort_annual, "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/global_fishing_watch/FishingWatch_annual_effort_2017.csv")

```


# Determine proportion of fisheries with unidentified geartypes
```{r}

fishing_effort <- read_csv("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/global_fishing_watch/FishingWatch_annual_effort_2017.csv")

## quick peak
totals <- fishing_effort %>%
  group_by(geartype) %>%
  summarize(fishing_hours = sum(fishing_hours))
totals$fishing_hours[totals$geartype == "fishing"]/sum(totals$fishing_hours)

#Map all fishing effort
total_map <- fishing_effort %>%
  group_by(cell_ll_lat, cell_ll_lon) %>%
  summarize(fishing_hours = sum(fishing_hours))


total_map$lat <- total_map$cell_ll_lat
total_map$lon <- total_map$cell_ll_lon
total_map <- total_map[,c("lon", "lat","fishing_hours")]
coordinates(total_map) <- ~lon+lat
proj4string(total_map) = CRS("+init=epsg:4326")
total_map_raster <- rasterFromXYZ(total_map, crs = "+init=epsg:4326")
proj4string(total_map_raster) = CRS("+init=epsg:4326")
plot(log(total_map_raster + 1))

#Map only unknowns
unk_map <- fishing_effort %>%
  filter(geartype %in% c("fishing")) %>%
  group_by(cell_ll_lat, cell_ll_lon) %>%
  summarize(fishing_hours = sum(fishing_hours, na.rm=TRUE))


unk_map$lat <- unk_map$cell_ll_lat
unk_map$lon <- unk_map$cell_ll_lon
unk_map <- unk_map[,c("lon", "lat","fishing_hours")]
coordinates(unk_map) <- ~lon+lat
proj4string(unk_map)=CRS("+init=epsg:4326")
unk_map_raster<-rasterFromXYZ(unk_map, crs = "+init=epsg:4326")
proj4string(unk_map_raster)=CRS("+init=epsg:4326")
plot(unk_map_raster)
unk_map_raster[is.na(unk_map_raster)] <- 0

prop_unk <- unk_map_raster/total_map_raster
colr <- rev(RColorBrewer::brewer.pal(4, 'RdYlBu'))
plot(prop_unk, col=colr, breaks=c(-0.01, 0.10, 0.30, 0.66, 1.01))

library(sf)
pts = rasterToPoints(prop_unk, spatial = TRUE)
pts_sf <- st_as_sf(pts)
plot(pts_sf,  pal = colr, pch=16, cex=0.005, breaks=c(-0.01, 0.10, 0.30, 0.66, 1.01))

```

## Create maps of destructive fishing
Create a map including only trawl effort and a map of dredge effort:
```{r}

fishing_effort <- read_csv("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/global_fishing_watch/FishingWatch_annual_effort_2017.csv")


## trawl
destructive_map <- fishing_effort %>%
  filter(geartype %in% c("trawlers")) %>%
  group_by(cell_ll_lat, cell_ll_lon) %>%
  summarize(fishing_hours = sum(fishing_hours))


destructive_map$lat <- destructive_map$cell_ll_lat
destructive_map$lon <- destructive_map$cell_ll_lon
destructive_map <- destructive_map[,c("lon", "lat","fishing_hours")]
coordinates(destructive_map) <- ~lon+lat
proj4string(destructive_map)=CRS("+init=epsg:4326")
destructive_map_raster<-rasterFromXYZ(destructive_map, crs = "+init=epsg:4326")
proj4string(destructive_map_raster)=CRS("+init=epsg:4326")
plot(destructive_map_raster)

writeRaster(destructive_map_raster, "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/global_fishing_watch/FishingWatch_annual_effort_trawlers_2017.tif", overwrite=TRUE)


## dredge
destructive_map <- fishing_effort %>%
  filter(geartype %in% c("dredge_fishing")) %>%
  group_by(cell_ll_lat, cell_ll_lon) %>%
  summarize(fishing_hours = sum(fishing_hours))


destructive_map$lat <- destructive_map$cell_ll_lat
destructive_map$lon <- destructive_map$cell_ll_lon
destructive_map <- destructive_map[,c("lon", "lat","fishing_hours")]
coordinates(destructive_map) <- ~lon+lat
proj4string(destructive_map)=CRS("+init=epsg:4326")
destructive_map_raster<-rasterFromXYZ(destructive_map, crs = "+init=epsg:4326")
proj4string(destructive_map_raster)=CRS("+init=epsg:4326")
plot(destructive_map_raster)

writeRaster(destructive_map_raster, "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/global_fishing_watch/FishingWatch_annual_effort_dredge_2017.tif", overwrite=TRUE)


```

