---
title: "New disturbance"
output: html_document
editor_options: 
  chunk_output_type: console
---

Here we:
1. rescale the habitat destruction data 
2. partition between marine fisheries food groups.

```{r setup, include=FALSE}
#load relevant packages, etc.

library(here)
library(raster)
library(tidyverse)

# raster template
# raster template
r <- raster::raster(ncol=720, nrow=360, vals=c(1:259200))
food_raster_template <- raster(nrows=2160, ncols=4320, xmn=-180, xmx=180, ymn=-90, ymx=90)

fishery_raster_template <- raster(list.files("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/rasters/destructive_proportions", full=TRUE)[1])
```

Rescale the hours of habitat destructive fisheries to values between 0-1 using the 99.9th quantile.
```{r}

destructive_hours <- raster("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/global_fishing_watch/FishingWatch_annual_effort_destructive_hours_2017.tif")

destructive_hours_low_res <- aggregate(destructive_hours, fact=50, fun=sum, progress="text")
# both have 259200 cells now!
plot(log(destructive_hours_low_res+1))
cellStats(destructive_hours_low_res, "sum")
cellStats(destructive_hours, "sum", na.rm=TRUE)
# both equal 16009799

## divide by area to get hours effort per km2 to control for differences in raster area in lat/long projection
destructive_hours_per_km2 <- destructive_hours_low_res/area(destructive_hours_low_res)

## get the rescaling raster by removing land zero values:
land_remove_template <- raster("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/spatial/food_rgns.tif") %>%
  resample(destructive_hours_per_km2) %>%
  raster::calc(fun=function(x){ifelse(is.na(x), 1, NA)})

land_remove_template
plot(land_remove_template)

destructive_hours_per_km2_rescale_data <- destructive_hours_per_km2 * land_remove_template
destructive_hours_per_km2_rescale_data
plot(destructive_hours_per_km2_rescale_data)
rescale_value <- quantile(destructive_hours_per_km2_rescale_data, c(0.999))
#rescale_value <- quantile(destructive_hours_per_km2[destructive_hours_per_km2>0], c(0.999))

destructive_hours_per_km2_rescaled <- destructive_hours_per_km2 %>%
  raster::calc(fun=function(x){ifelse(x<0, 0,
                                      ifelse(x>rescale_value, 1, x/rescale_value))})

plot(destructive_hours_per_km2_rescaled)

```

Partition the rescaled pressures among the marine fisheries food categories.

```{r}

food_cat_props_rasters <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/rasters/destructive_proportions", full=TRUE)

cellStats(stack(food_cat_props_rasters), sum, na.rm=TRUE)

for(food_cat in food_cat_props_rasters){ # food_cat = food_cat_props_rasters[1]
  saveName <- basename(food_cat)
  saveName <- gsub("ben_hab_dest_prop_", "", saveName)
  food_cat_hab_dest <- raster(food_cat) * destructive_hours_per_km2_rescaled 
  writeRaster(food_cat_hab_dest, sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/rasters/benthic_habitat_destruction/benthic_habitat_destruction_%s", saveName), overwrite=TRUE)
  
  #plot(food_cat_hab_dest)
}


# change name to match Gage names
list.files("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/rasters/benthic_habitat_destruction", full=TRUE)
file.rename(from=
"/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/rasters/benthic_habitat_destruction/benthic_habitat_destruction_forage_fish.tif",
to="/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/rasters/benthic_habitat_destruction/benthic_habitat_destruction_fofm.tif")

```

Quick check that things are ok
```{r}

all_hab_dest <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/rasters/benthic_habitat_destruction", full=TRUE)

all_hab_dest_stack <- stack(all_hab_dest)
all_hab_total <- calc(all_hab_dest_stack, sum, na.rm=TRUE)

diff <- all_hab_total - destructive_hours_per_km2_rescaled
plot(diff)

```


## Uncertainty analysis
Rescale the hours of habitat destructive fisheries to values between 0-1 using the 99.99th quantile.  These data will be saved elsewhere and ran as a different scenario.
```{r}

destructive_hours <- raster("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/global_fishing_watch/FishingWatch_annual_effort_destructive_hours_2017.tif")

destructive_hours_low_res <- aggregate(destructive_hours, fact=50, fun=sum, progress="text")
# both have 259200 cells now!
plot(log(destructive_hours_low_res+1))
cellStats(destructive_hours_low_res, "sum")
cellStats(destructive_hours, "sum", na.rm=TRUE)
# both equal 16009799

## divide by area to get hours effort per km2 to control for differences in raster area in lat/long projection
destructive_hours_per_km2 <- destructive_hours_low_res/area(destructive_hours_low_res)

## get the rescaling raster by removing land zero values:
land_remove_template <- raster("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/spatial/food_rgns.tif") %>%
  resample(destructive_hours_per_km2) %>%
  raster::calc(fun=function(x){ifelse(is.na(x), 1, NA)})

land_remove_template
plot(land_remove_template)

destructive_hours_per_km2_rescale_data <- destructive_hours_per_km2 * land_remove_template
destructive_hours_per_km2_rescale_data
plot(destructive_hours_per_km2_rescale_data)
rescale_value <- quantile(destructive_hours_per_km2_rescale_data, c(0.9999))

destructive_hours_per_km2_rescaled <- destructive_hours_per_km2 %>%
  raster::calc(fun=function(x){ifelse(x<0, 0,
                                      ifelse(x>rescale_value, 1, x/rescale_value))})

plot(destructive_hours_per_km2_rescaled)

```

Partition the rescaled pressures among the marine fisheries food categories.

```{r}

food_cat_props_rasters <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/rasters/destructive_proportions", full=TRUE)

cellStats(stack(food_cat_props_rasters), sum, na.rm=TRUE)

for(food_cat in food_cat_props_rasters){ # food_cat = food_cat_props_rasters[1]
  saveName <- basename(food_cat)
  saveName <- gsub("ben_hab_dest_prop_", "", saveName)
  food_cat_hab_dest <- raster(food_cat) * destructive_hours_per_km2_rescaled 
  writeRaster(food_cat_hab_dest, sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/scenarios/scenarios_dataprep/fis_benthic_destructive/benthic_habitat_destruction_%s", saveName), overwrite=TRUE)
  
  #plot(food_cat_hab_dest)
}


# change name to match Gage names
list.files("/home/shares/food-systems/Food_footprint/all_food_systems/scenarios/scenarios_dataprep/fis_benthic_destructive", full=TRUE)
file.rename(from=
"/home/shares/food-systems/Food_footprint/all_food_systems/scenarios/scenarios_dataprep/fis_benthic_destructive/benthic_habitat_destruction_forage_fish.tif",
to="/home/shares/food-systems/Food_footprint/all_food_systems/scenarios/scenarios_dataprep/fis_benthic_destructive/benthic_habitat_destruction_fofm.tif")

```
