---
title: "Create rasters for new disturbance calculations"
author: "Gage Clawson (UCSB, NCEAS, OHI)"
date: "May 19, 2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r, eval = FALSE}
#load relevant packages, etc.

library(here)
library(raster)
library(tidyverse)
library(doParallel)
library(foreach)
library(parallel)

# raster template
r <- raster::raster(ncol=720, nrow=360, vals=c(1:259200))
food_raster <- raster(nrows=2160, ncols=4320, xmn=-180, xmx=180, ymn=-90, ymx=90)
hab_raster <- raster("/home/shares/food-systems/Food_footprint/_raw_data/Halpern_chi/habitats/coral_reef.tif")


cat_msg <- function(x, ...) {
  if(is.null(knitr:::.knitEnv$input.dir)) {
    ### not in knitr environment, so use cat()
    cat(x, ..., '\n')
  } else {
    ### in knitr env, so use message()
    message(x, ...)
  }
  return(invisible(NULL))
}

```

Here we will make rasters describing: 

1. Proportion of bottom trawling to all trawling: tonnes(bottom_trawls)/tonnes(bottom_trawl + midwater_trawl) for all marine fisheries catch
2. Proportion of destructive catch per species class to all destructive catch: n=7 species classes: raster_tonnes(raking_device + dredge + hand_dredges + bottom_trawl)_spp_i / raster_tonnes(raking_device + dredge + hand_dredges + bottom_trawl)_spp_all
3. Proportion of all catch per species class: n=7 species classes: all_catch_spp_i / all_catch_spp_all

We will include IUU, landings, discards. 

The catch file (`catch_spp_class_2017.csv`) we are reading in is made in `STEP2c_prep_watson_catch.Rmd`. 

## Raster 1: All catch trawling rasters 
 
```{r}
# tonnes(bottom_trawls)/tonnes(bottom_trawl + midwater_trawl) for all marine fisheries catch

## read in the catch data 
# catch_raw <- read.csv("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/catch_with_gear_cats.csv") 

catch_raw <- read_csv("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/catch_spp_class_2017.csv") 


catch_all <- catch_raw %>%
  mutate(catch = ReportedIND + IUUIND + DiscardsIND + ReportedNIND + IUUNIND + DiscardsNIND)


sum(catch_all$catch) # 126256514

check <- catch_all %>%
  group_by(species_class_fin) %>%
  summarise(catch_sum = sum(catch))

bottom_trawls_tonnes <- catch_all %>%
  filter(GearName == "bottom trawls") %>%
  group_by(Cell) %>%
  summarise(catch = sum(catch, na.rm = TRUE)) %>%
  ungroup()
sum(bottom_trawls_tonnes$catch) # 37332269
37332269/126256514 # around 30% of total catch is bottom trawl

## make bottom trawl raster 
bottom_trawls_tonnes_raster <- raster::subs(r, bottom_trawls_tonnes, by = "Cell", which = "catch", subsWithNA=TRUE)

plot(bottom_trawls_tonnes_raster)
cellStats(bottom_trawls_tonnes_raster, "sum") # 37332269 - perfect 

## make bottom and midwater trawl raster

all_trawls_tonnes <- catch_all %>%
  filter(str_detect(GearName, "trawl")) %>%
  group_by(Cell) %>%
  summarise(catch = sum(catch, na.rm = TRUE)) %>%
  ungroup()
sum(all_trawls_tonnes$catch) # 55119183
55119183/126256514 # any type of trawling account for ~44% of total catch

all_trawls_tonnes_raster <- raster::subs(r, all_trawls_tonnes, by = "Cell", which = "catch", subsWithNA=TRUE)
cellStats(all_trawls_tonnes_raster, "sum") # 55119183

trawl_prop_raster <- bottom_trawls_tonnes_raster/all_trawls_tonnes_raster
plot(trawl_prop_raster)

## save all three rasters 
writeRaster(bottom_trawls_tonnes_raster, file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/rasters/trawling/bottom_trawl_tonnes_raster.tif"))

writeRaster(all_trawls_tonnes_raster, file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/rasters/trawling/bottom_midwater_trawl_tonnes_raster.tif"))

writeRaster(trawl_prop_raster, file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/rasters/trawling/trawl_proportion_raster.tif"))

```

## Raster series 2

n=7 species classes: raster_tonnes(raking_device + dredge + hand_dredges + bottom_trawl)_spp_i / raster_tonnes(raking_device + dredge + hand_dredges + bottom_trawl)_spp_all

```{r}

check <- catch_all %>%
  group_by(category, GearName) %>%
  summarise(catch = sum(catch))

all_destructive_gears <- catch_all %>%
  filter(GearName %in% c("bottom trawls", "dredges", "hand dredges", "raking devices"))
sum(all_destructive_gears$catch) # 41601963

all_dest_gears_cell <- all_destructive_gears %>%
  group_by(Cell) %>%
  summarise(catch = sum(catch, na.rm = TRUE)) %>%
  ungroup()
sum(all_dest_gears_cell$catch) # 41601963

## make all destructive gears raster 
all_destructive_gears_rast <- raster::subs(r, all_dest_gears_cell, by = "Cell", which = "catch", subsWithNA=TRUE)
cellStats(all_destructive_gears_rast, "sum") # 41601963
plot(all_destructive_gears_rast)


## now write a for loop to make the rasters for each species class, and divide by the total raster
## get species, description, taxon ids
catch_cats <- all_destructive_gears %>%
  dplyr::select(species_class_fin) %>%
  distinct()

reload <- FALSE ## change this to true if you want to rerun all species

## rasterize catch
for(i in 1:nrow(catch_cats)){
  
  # i = 1

  ## get the ids for this i  
  spp_class_id <- catch_cats$species_class_fin[i]
  
  ## this will be the final file name
  spp_final_file <- file.path(sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/rasters/destructive_proportions/ben_hab_dest_prop_%s.tif", spp_class_id))
    
if(!file.exists(spp_final_file) | reload) { ## if the file already exists, skip over it (unless otherwise specified)
  
## filter for correct observations  
catch_summary_i <- all_destructive_gears %>%
  filter(species_class_fin == spp_class_id) %>%
  group_by(Cell) %>%
  summarise(catch = sum(catch, na.rm = TRUE)) %>%
  ungroup()

print(sum(catch_summary_i$catch)) # 29229250

## rasterize
tonnes_raster <- raster::subs(r, catch_summary_i, by = "Cell", which = "catch", subsWithNA=TRUE)
# plot(log(tonnes_raster+1))

print(cellStats(tonnes_raster, "sum")) # 29229250
print(cellStats(all_destructive_gears_rast, "sum")) # 41601963

## divide the species group tonnes by the total tonnes 
final_rast_i <- tonnes_raster/all_destructive_gears_rast
# plot(final_rast_i)

writeRaster(final_rast_i, file.path(spp_final_file))

cat_msg('Finished ', i, ' out of ', nrow(catch_cats), 'species')

} else {
      
      cat_msg('species file ', spp_final_file, ' already exists; skipping these spp')
      
    }
}

```

## Raster series 3

n=7 species classes: all_catch_spp_i / all_catch_spp_all

```{r}

sum(catch_all$catch) # 126256514

all_catch_cell <- catch_all %>%
  group_by(Cell) %>%
  summarise(catch = sum(catch, na.rm = TRUE)) %>%
  ungroup()
sum(all_catch_cell$catch) # 126256514

## make all catch raster
all_catch_rast <- raster::subs(r, all_catch_cell, by = "Cell", which = "catch", subsWithNA=TRUE)
cellStats(all_catch_rast, "sum") # 126256514
plot(log(all_catch_rast + 1))


writeRaster(all_catch_rast, file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/rasters/catch_rasters/catch_all_spp.tif"))


## now write a for loop to make the rasters for each species class, and divide by the total raster
## get species, description, taxon ids
catch_cats <- catch_all %>%
  dplyr::select(species_class_fin) %>%
  distinct()

reload <- FALSE ## change this to true if you want to rerun all rasters

## rasterize
for(i in 1:nrow(catch_cats)){
  
  # i = 1

  ## get the ids for this i  
  spp_class_id <- catch_cats$species_class_fin[i]
  
  ## this will be the final file name
  spp_final_file <- file.path(sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/rasters/all_catch_proportions/catch_prop_%s.tif", spp_class_id))
    
if(!file.exists(spp_final_file) | reload) { ## if the file already exists, skip over it (unless otherwise specified)
  
## filter for correct observations  
catch_summary_i <- catch_all %>%
  filter(species_class_fin == spp_class_id) %>%
  group_by(Cell) %>%
  summarise(catch = sum(catch, na.rm = TRUE)) %>%
  ungroup()

print(sum(catch_summary_i$catch)) # 8752618

## rasterize
tonnes_raster <- raster::subs(r, catch_summary_i, by = "Cell", which = "catch", subsWithNA=TRUE)
# plot(log(tonnes_raster+1))

print(cellStats(tonnes_raster, "sum")) # 8752618
print(cellStats(all_catch_rast, "sum")) # 126256514

writeRaster(tonnes_raster, file.path(sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/rasters/catch_rasters/catch_%s_spp.tif", spp_class_id)), overwrite=TRUE)

## divide the species group tonnes by the total tonnes 
final_rast_i <- tonnes_raster/all_catch_rast
# plot(final_rast_i)

writeRaster(final_rast_i, file.path(spp_final_file), overwrite = TRUE)

cat_msg('Finished ', i, ' out of ', nrow(catch_cats), 'species')

} else {
      
      cat_msg('species file ', spp_final_file, ' already exists; skipping these spp')
      
    }
}
```

