---
title: "New disturbance"
output: html_document
editor_options: 
  chunk_output_type: console
---

Here we:
1. correct trawl to include only the demersal destructive portion of trawling (removing midwater trawls based on proportion of catch from Watson data). Using data created in STEP3_create_prop_rasters.Rmd. 
2. combine the hours effort for the corrected trawl data to hours effort for dredge data to get total effort.



```{r setup, include=FALSE}
#load relevant packages, etc.

library(here)
library(raster)
library(tidyverse)

# raster template
r_template <- raster::raster(ncol=720, nrow=360, vals=c(1:259200))

  
```

## correct trawl data

Global Fishing Watch combines midwater and bottom trawls.  We are only concerned here with the bottom trawls, so we attempt to dissentangle this using watson data to generate a raster with these values:

bottom_trawl_tonnes / all_trawl_tonnes

We multiply that raster by the Global Fishing Watch raster for all trawl data.
```{r}

trawl_correction <- raster("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/rasters/trawling/trawl_proportion_raster.tif")
plot(trawl_correction)
avg_trawl_correction <- cellStats(trawl_correction, "mean", na.rm=TRUE)
trawl_correction[is.na(trawl_correction)] <- avg_trawl_correction
plot(trawl_correction)

GFW_all_trawl <- raster("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/global_fishing_watch/FishingWatch_annual_effort_trawlers_2017.tif")
plot(GFW_all_trawl)

resample(trawl_correction, GFW_all_trawl, method="ngb", filename = "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/rasters/trawling/trawl_proportion_raster_resampled.tif", progress="text", overwrite=TRUE)

trawl_correction_raster_resampled <- raster("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/rasters/trawling/trawl_proportion_raster_resampled.tif")
plot(trawl_correction_raster_resampled)

trawl_stack <- stack(trawl_correction_raster_resampled, GFW_all_trawl)
  
overlay(trawl_stack, fun=function(x,y){return(x*y)},
 filename = "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/global_fishing_watch/FishingWatch_annual_effort_destructive_trawlers_2017.tif", overwrite=TRUE, progress="text")


```


Sum the corrected trawl with the dredge effort to get the total demersal destructive fishing effort.
```{r}
corrected_trawl <- raster("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/global_fishing_watch/FishingWatch_annual_effort_destructive_trawlers_2017.tif")

dredge <- raster("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/global_fishing_watch/FishingWatch_annual_effort_dredge_2017.tif")

newextent=c(-180.005, 179.995, -89.995, 90.005)

dredge_extend <- extend(dredge, newextent, value=0)
corrected_trawl_extend <- extend(corrected_trawl, newextent, value=0)

damage_stack <- stack(dredge_extend, corrected_trawl_extend)

calc(damage_stack, sum, na.rm=TRUE,
 filename = "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/global_fishing_watch/FishingWatch_annual_effort_destructive_hours_2017.tif", overwrite=TRUE, progress="text")

```
