---
title: "Create raster of disturbance for all marine fisheries"
author: "Gage Clawson (UCSB, NCEAS, OHI)"
date: "October 2, 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

Here we prep the Watson fisheries data to be classified into the correct species classes, and save as a csv. 

- Final output is a dataset, `/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/catch_spp_class_2017.csv` that assigns each species to a specific species grouping class for 2017. 

```{r, eval = FALSE}
#load relevant packages, etc.

library(here)
library(raster)
library(tidyverse)
library(doParallel)
library(foreach)
library(parallel)

# raster template
r <- raster::raster(ncol=720, nrow=360, vals=c(1:259200))
food_raster <- raster(nrows=2160, ncols=4320, xmn=-180, xmx=180, ymn=-90, ymx=90)
hab_raster <- raster("/home/shares/food-systems/Food_footprint/_raw_data/Halpern_chi/habitats/coral_reef.tif")
```


## Step 1: Classify catch into demersal destructive/nondestructive
Get all the data together...
```{r, eval = FALSE}
## read in catch data
catch <- read_csv("/home/shares/food-systems/Food_footprint/_raw_data/watson_data/v5.0/Catch2015_2019.csv")

## read in taxa index 
taxa_index <- read_csv("/home/shares/food-systems/Food_footprint/_raw_data/watson_data/v5.0/Codes_taxa.csv")

## read in gear index 
gear_index <- read_csv("/home/shares/food-systems/Food_footprint/_raw_data/watson_data/v5.0/Codes_gear.csv") %>%
  dplyr::select(Gear, GearName = VBDesc) %>%
  unique()

taxa_gear_types <- read_csv(here("fisheries/marine/disturbance/int/taxa_gear_types.csv"))

head(taxa_gear_types)

## join datasets together
catch_full <- left_join(catch, taxa_index, by="Taxonkey")
dim(catch_full)
catch_full <- left_join(catch_full, gear_index, by="Gear")
dim(catch_full)
catch_full <- left_join(catch_full, taxa_gear_types, by=c("TaxonName", "Taxonkey", "CommonName", "Descript","GearName", "Gear")) 
dim(catch_full)


forage_fish_list <- read_csv(here("fisheries/marine/ghg/data/master_taxa_list.csv"))


catch_full <- catch_full %>%
  mutate(forage_fish = ifelse(TaxonName %in% forage_fish_list$forage_fish, 1, 0)) %>%
  mutate(forage_fish = ifelse(Descript == "krill", 1, forage_fish)) %>%
  dplyr::select(-foragefish)

```

Check to see if there are any NAs to gapfill in our gear type data.

```{r, eval = FALSE}

check <- catch_full %>%
  mutate(category = paste(destruction, type, sep="-"))

# explore ones with no category
check %>%
  mutate(total_tonnes = ReportedIND + IUUIND + DiscardsIND + ReportedNIND + IUUNIND + DiscardsNIND) %>%
  group_by(category) %>%
  summarize(total_tonnes = sum(total_tonnes))


# # A tibble: 4 x 2
#   category                 total_tonnes
#   <chr>                           <dbl>
# 1 destructive-demersal       102286775.
# 2 destructive-pelagic         21463146.
# 3 non-destructive-demersal    96149597.
# 4 non-destructive-pelagic    153890005.

check2 <- catch_full %>%
  mutate(total_tonnes = ReportedIND + IUUIND + DiscardsIND + ReportedNIND + IUUNIND + DiscardsNIND) %>%
  filter(forage_fish == 0, IYear == 2017)
sum(check2$total_tonnes) # 84562955 tonnes of non-forage fish (IUU, Reported, Discards)

check3 <- catch_full %>%
  mutate(total_tonnes = ReportedIND + IUUIND + ReportedNIND + IUUNIND) %>%
  filter(forage_fish == 0, IYear == 2017)
sum(check3$total_tonnes) # 72172118 tonnes of non-forage fish (IUU, Reported)



check4 <- catch_full %>%
  mutate(total_tonnes = ReportedIND + IUUIND + DiscardsIND + ReportedNIND + IUUNIND + DiscardsNIND) %>%
  filter(forage_fish == 1, IYear == 2017)
sum(check4$total_tonnes) # 41693559 tonnes of forage fish (IUU, Reported, Discards)

check5 <- catch_full %>%
  mutate(total_tonnes = ReportedIND + IUUIND + ReportedNIND + IUUNIND) %>%
  filter(forage_fish == 1, IYear == 2017)
sum(check5$total_tonnes) # 39054068 tonnes of forage fish (IUU, Reported, Discards)
```

For intermediate and final rasters, we need to split our species into coarser groups: 

Intermediate categories
 - Benthic
 - Small Demersal or Small bentho-pelagic
 - Medium Demersal or Medium bentho-pelagic
 - Large Demersal or Large bentho-pelagic
 - Large Pelagics (>90cm)
 - Medium Pelagic (30-60cm)
 - Small pelagics (<30cm)
 - Small Reef-associated
 - Medium Reef-associated
 - Large Reef-associated
 - FOFM - includes krill
 
 Conversions: 
 - pelagic >=90 cm : Large pelagics **
 - demersal <30cm : Small Demersal **
 - shark >=90cm : Large Demersal **
 - pelagic <30cm : Small Pelagics **
 - demersal 30 - 90 cm : Medium Demersal **
 - bathydemersal <30 cm : Small Demersal **
 - bathydemersal >= 90cm : Large Demersal **
 - pelagic 30 - 90cm : Medium pelagics **
 - reef-associated 30 - 90 cm : Medium Reef associated **
 - flatfish <90 cm : Medium Demersal **
 - cephalopods : See below
 - demersal >=90 cm : Large Demersal **
 - benthopelagic >=90 cm : Large Demersal **
 - reef-associated >=90 cm : Large Reef associated **
 - flatfish >=90 cm : Large Demersal **
 - shrimp : Small demersal **
 - demersal mollusc : Small Demersal **
 - rays <90 cm : Medium Demersal **
 - lobsters crab : Benthic **
 - shark <90 cm : Medium Demersal **
 - benthopelagic 30 - 90 cm : Medium Demersal **
 - benthopelagic <30 cm : Small Demersal **
 - reef-associated <30 cm : Small Reef associated **
 - rays >=90 cm : Large Demersal **
 - bathypelagic <30 cm : Small Pelagics **
 - bathypelagic 30 - 90 cm : Medium Pelagics **
 - bathypelagic >=90 cm : Large Pelagic **
 - bathydemersal 30 - 90 cm : Medium Demersal **
 - krill : FOFM 

Final raster categories:
 - Benthic
 - Demersal or bentho-pelagic
 - Large Pelagics (>90cm)
 - Medium Pelagic (30-60cm) 
 - Small pelagics (<30cm)
 - Reef-associated
 - FOFM

Conversions: 
 - pelagic >=90 cm : Large pelagics
 - demersal <30cm : Demersal
 - shark >=90cm : Demersal
 - pelagic <30cm : Small Pelagics
 - demersal 30 - 90 cm : Demersal
 - bathydemersal <30 cm : Demersal
 - bathydemersal >= 90cm : Demersal
 - pelagic 30 - 90cm : Medium pelagics
 - reef-associated 30 - 90 cm : Reef associated
 - flatfish <90 cm : Demersal
 - cephalopods : species dependent, see below 
 - demersal >=90 cm : Demersal
 - benthopelagic >=90 cm : Demersal
 - reef-associated >=90 cm : Reef associated
 - flatfish >=90 cm : Demersal
 - shrimp : Demersal
 - demersal mollusc : Demersal
 - rays <90 cm : Demersal
 - lobsters crab : Benthic
 - shark <90 cm : Demersal
 - benthopelagic 30 - 90 cm : Demersal
 - benthopelagic <30 cm : Demersal
 - reef-associated <30 cm : Reef associated
 - rays >=90 cm : Demersal
 - bathypelagic <30 cm : Small Pelagics
 - bathypelagic 30 - 90 cm : Medium Pelagics
 - bathypelagic >=90 cm : Large Pelagic 
 - bathydemersal 30 - 90 cm : Demersal
 - krill : FOFM 


Reclassify the Watson species classes to intermediate supplementary classes:
```{r, eval = FALSE}
catch_species_class_int <- catch_full %>%
  dplyr::mutate(species_class_int =
                  case_when(
                    str_detect(Descript, "demersal mollusc|demersal <30 cm|bathydemersal <30 cm|shrimp|benthopelagic <30 cm") ~ "Small demersal",
                    str_detect(Descript, "shark >=90 cm|bathydemersal >= 90cm|demersal >=90 cm|benthopelagic >=90 cm|flatfish >=90 cm|rays >=90 cm") ~ "Large demersal",
                    str_detect(Descript, "demersal 30 - 90 cm|flatfish <90 cm|rays <90 cm|shark <90 cm|benthopelagic 30 - 90 cm|bathydemersal 30 - 90 cm") ~ "Medium demersal",
                    str_detect(Descript, "lobsters crab") ~ "Benthic",
                    str_detect(Descript, "pelagic >=90 cm|bathypelagic >=90 cm") ~ "Large pelagic",
                    str_detect(Descript, "pelagic 30 - 90 cm|bathypelagic 30 - 90 cm") ~ "Medium pelagic",
                    str_detect(Descript, "pelagic <30 cm|bathypelagic <30 cm|krill") ~ "Small pelagic",
                    str_detect(Descript, "reef-associated >=90 cm") ~ "Large reef-associated",
                    str_detect(Descript, "reef-associated 30 - 90 cm") ~ "Medium reef-associated",
                    str_detect(Descript, "reef-associated <30 cm") ~ "Small reef-associated",
                    str_detect(CommonName, "uttlefish") & str_detect(Descript, "cephalopods") ~ "Small reef-associated",
                    str_detect(CommonName, "Horned") & str_detect(Descript, "cephalopods") ~ "Medium reef-associated",
                    str_detect(CommonName, "Common octopus|Octopuses") & str_detect(Descript, "cephalopods") ~ "Large reef-associated", 
                    str_detect(CommonName, "Argentine shortfin squid|Patagonian squid|Northern shortfin squid|California market squid|Cape Hope squid|Longfin squid|Sevenstar flying squid") & str_detect(Descript, "cephalopods") ~ "Small pelagic", 
                    str_detect(CommonName, "Jumbo flying squid|Broadtail shortfin squid|European flying squid|Japanese flying squid|Wellington flying squid|Neon flying squid|Veined Squid|Squids|Common squids|Cephalopods") & str_detect(Descript, "cephalopods") ~ "Medium pelagic", 
                    str_detect(CommonName, "Arrow squids") & str_detect(Descript, "cephalopods") ~ "Large pelagic"
                  )) 

## take a look to see if we caught all the cases... we did
unique(catch_species_class_int$species_class_int)

test <- catch_species_class_int %>%
  filter(Descript == "krill")

unique(test$forage_fish)
```


Reclassify the Watson species classes to final raster classes:

```{r, eval = FALSE}
catch_species_class_fin <- catch_species_class_int %>%
  dplyr::mutate(species_class_fin =
                  case_when(
                    str_detect(species_class_int, "demersal") ~ "Demersal",
                    str_detect(species_class_int, "reef-associated") ~ "Reef-associated",
                    TRUE  ~ species_class_int
                  ))

## take a look to see if we caught all the cases... we did
catch_species_class_summary <- catch_species_class_fin %>%
  filter(forage_fish == 0, IYear == 2017) %>%
  mutate(catch_total = ReportedIND + IUUIND + DiscardsIND + ReportedNIND + IUUNIND + DiscardsNIND) %>%
  group_by(species_class_fin) %>%
  summarise(catch_sum = sum(catch_total))
sum(catch_species_class_summary$catch_sum) # 84562955

catch_species_class_summary_2 <- catch_species_class_fin %>%
  filter(forage_fish == 1, IYear == 2017) %>%
   mutate(catch = ReportedIND + IUUIND + DiscardsIND + ReportedNIND + IUUNIND + DiscardsNIND) %>%
  group_by(species_class_fin) %>%
  summarise(catch_sum = sum(catch))
sum(catch_species_class_summary_2$catch_sum) # 41693559

```

Save the dataset
```{r, eval = FALSE}

catch_full_final <- catch_species_class_fin %>%
  mutate(category = paste(destruction, type, sep="-"))

## check no NAs
table(catch_full_final$category)

## This file will be used in a subsequent Rmd to isolate the pelagic fisheries
save_file <- catch_full_final %>%
  dplyr::select(Cell, ReportedIND, IUUIND, DiscardsIND, ReportedNIND, IUUNIND, DiscardsNIND, IYear, CNumber, Taxonkey, Gear, TaxonName, CommonName, Descript, species_class_int, species_class_fin, TaxLevel, ISSCAAP, ISSCAAPName, forage_fish, GearName, destruction, fish_type = type, category)

write_csv(save_file, "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/catch_with_gear_cats.csv")


catch_spp_class <- read_csv("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/catch_with_gear_cats.csv")

catch_spp_class_fofm_2017 <- catch_spp_class %>%
  mutate(species_class_fin = ifelse(forage_fish == 1, "forage_fish", species_class_fin)) %>%
  filter(IYear == 2017) ## filter for correct year

## save file 
write_csv(catch_spp_class_fofm_2017, "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/catch_spp_class_2017.csv")
  


# catch_summary <- read_csv(file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/catch_with_gear_cats.csv"))
# 
# catch_summary_table <- catch_summary %>%
#   dplyr::filter(IYear == 2017) %>%
#   dplyr::mutate(tonnes_catch = ReportedIND + IUUIND + DiscardsIND + ReportedNIND + IUUNIND + DiscardsNIND) %>%
#   mutate(species_class_fin = ifelse(forage_fish == 1, "fofm", species_class_fin)) %>%
#   dplyr::group_by(species_class_fin) %>%
#   dplyr::summarise(total_tonnes = sum(tonnes_catch)) %>%
#   dplyr::ungroup()

```


