---
title: "SI Tables prep"
output: html_document
---

This script contains code and data to fill out the relevant supplemental results tables. 

```{r setup, include=FALSE, eval = FALSE, eval = FALSE}
knitr::opts_chunk$set(eval = TRUE)

library(here)
library(tidyverse)
library(kableExtra)

source(here("_workflow/common.R"))
source(here("_workflow/common_spatial.R"))

### save the gall peters projection
gall_peters <- "+proj=cea +lon_0=0 +x_0=0 +y_0=0 +lat_ts=45 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

rgn_raw <- read_csv(file.path("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/raw_df_summary/rgn_raw_summary.csv"))

```

#### Supplementary Table XX: Total global pressures
Total global pressures measured across all included food types, used to rescale raster pressure data.

```{r total pressures, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, asis=TRUE}

units_df <- data.frame(pressure = c("ghg", "nutrient", "disturbance", "water"),
                       units = c("tonnes CO2eq", "tonnes N and P", "km2eq", "m3"))

pressure <- read.csv(here("_analysis/rescale_values.csv")) %>%
  left_join(units_df) %>%
  select(pressure, "global total" = global_total, units)

kable(pressure) %>%
 kableExtra::kable_styling("striped", full_width = FALSE, position = "left")
#kable_styling(x_html, "striped", position = "left", font_size = 7)

## save this 
write.csv(pressure, here::here("_analysis/SI_data/output/pressure_rescale_values.csv"), row.names = FALSE)
```
 
 
## Supplemental Table XX





```{r, eval = FALSE}
rescale <- read_csv("../rescale_values.csv")

### do overall, including feed and fodder
percent_pressure_rescale <- rgn_raw %>%
  left_join(rescale, by = "pressure") %>%
  mutate(global_pressure_prop = sum/global_total) %>%
  group_by(origin) %>%
  summarise(cum_pressure_prop_rescaled = sum(global_pressure_prop),
            cum_pressure = sum(sum)) %>%
  ungroup()

percent_pressure_origin <- percent_pressure_rescale %>%
  mutate(total_cum_prop = sum(cum_pressure_prop_rescaled)) %>%
  mutate(cum_pressure_prop_final = cum_pressure_prop_rescaled/total_cum_prop) %>%
  mutate(percent_final = round(cum_pressure_prop_final*100,2)) %>%
  dplyr::select(origin, cum_pressure_prop_final, percent_final)


write.csv(percent_pressure_origin, "output/percent_cum_pressure_origin.csv", row.names = FALSE)


```
 
 
## Supplemental table XX: Distribution of cumulative pressures by country: marine vs. land 


  
```{r, eval = FALSE}

rescale <- read_csv("../rescale_values.csv")


percent_pressure_ocean <- rgn_raw %>%
  left_join(rescale, by = "pressure") %>%
  mutate(global_pressure_prop = sum/global_total) %>%
  group_by(iso3c, country, origin) %>%
  summarise(cum_pressure_prop_rescaled = sum(global_pressure_prop),
            cum_pressure = sum(sum)) %>%
  ungroup()

percent_pressure_origin_countries <- percent_pressure_ocean %>%
  select(-cum_pressure) %>%
  group_by(iso3c, country) %>%
  mutate(total_cum_prop = sum(cum_pressure_prop_rescaled)) %>%
  ungroup() %>%
  mutate(cum_pressure_prop_final = cum_pressure_prop_rescaled/total_cum_prop) %>%
  mutate(cum_pressure_percent_final = round(cum_pressure_prop_final*100,2)) %>%
  arrange(iso3c) %>%
  dplyr::select(iso3c, country, origin, cum_pressure_prop_final, cum_pressure_percent_final)


write.csv(percent_pressure_origin_countries, "output/percent_cum_pressure_country_origin_fw.csv", row.names = FALSE)


``` 

## Supplemental table XX: Distribution of cumulative pressures by product: marine vs. land 

Data S5


```{r}
rescale <- read_csv("../rescale_values.csv")

### rescale pressures 
percent_pressure_rescale <- rgn_raw %>%
  left_join(rescale, by = "pressure") %>%
  mutate(global_pressure_prop = sum/global_total) %>%
  group_by(organism, origin) %>%
  summarise(cum_pressure = sum(sum)) %>%
  ungroup()

### group by food group
percent_pressure_organisms_origin <- percent_pressure_rescale %>%
  group_by(organism) %>%
  mutate(total_organism_pressure = sum(cum_pressure)) %>%
  ungroup() %>%
   mutate(proportion_final = cum_pressure/total_organism_pressure) %>%
  filter(organism %in% c("chickens", "pigs", "crustaceans", "marine-fish-general", "bivalve", "salmon", "shrimp", "tuna"))
  

write.csv(percent_pressure_organisms_origin, "output/percent_cum_pressure_product_origin.csv")
```
