---
title: "Calculating GHG emissions from crop burning"
author: "Juliette"
date: "3/16/2020"
output: html_document
---

In this markdown we are calculating the total carbon emitted, converted to CH4, CO, NO2, NOx, and the carbon equivalant of these 4 emission types.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Read in libraries and paths
```{r}
#Load libraries
library(readr)
library(dplyr)
library(tidyverse)
library(doParallel)
library(raster)
library(rgdal)
library(here)

source(here("_spatial/template_raster.R"))
source(here("_workflow/common.R"))
source(here("_workflow/common_spatial.R"))

#Files we will need:
burn_parameters <- read_csv("crop/ghg/data/burning_parameters_gf.csv") %>% 
  dplyr::select(-gapfilled)

developed <- read_csv("crop/ghg/data/developed_regions.csv")

```

Total carbon released = annual production (tonnes of biomass per year) * ratio of residue to crop product (fraction) * average dry matter fraction of residue (tonnes dry matter/tonnes of biomass) * fraction burned in field * fraction oxidised * carbon fraction (tonnes carbon/tonnes of dry matter)

tc = prod * ratio_res_prod * dm * frac_ox * carbon_res * fraction_burned

```{r}
crop_production <- list.files(file.path(prep, "crop/farm/scaled_maps_2017"), pattern = "A_scaled.tif", full = TRUE)

crop_production <- 
  crop_production[!grepl(pattern = c("rest|toba|cof|teas"), 
                         x = crop_production)]

registerDoParallel(6)   
foreach(file = crop_production) %dopar% {
  
  #file = crop_production[40]
  crop <- str_sub(file, -17, -14)
  
  prod <- raster(file, crs = food_crs) %>% 
    raster_df() %>% 
    mutate(split_rest_names = crop) %>% 
    rename(production = paste0("crop_",crop, "_A_scaled")) %>% 
    mutate(split_rest_names = ifelse(is.na(production), NA , split_rest_names)) %>% 
    dplyr::select(-x,-y)

prod_country <- left_join(prod, food_rgns_xy, by = c("cellindex")) %>% 
  mutate(ID_0 = as.numeric(ID_0))
    
## Add the parameters for the mapspam crops
  prod_burn <- left_join(prod_country, burn_parameters, by =  "split_rest_names") %>% 
    left_join(developed)

## Calculate GHG emissions
   c02_eq <- prod_burn %>% 
     mutate(tc = production*ratio_res_prod*dm*frac_ox*carb_frac*fraction_burned) %>% 
     dplyr::select(x, y, tc)
   
   write_csv(c02_eq, file.path(prep, paste("crop/ghg/crop_residue_burning_ghg/total_carbon_emitted/", crop, ".csv", sep = "")))
}
```


Now we need to convert the total carbon released from each crop to the emission from 4 different GHG
CH4 Emission = Carbon Released * Emission Ratio * 16/12
CO Emission = Carbon Released * Emission Ratio * 28/12
N2O Emission = Carbon Releaed * Emission Ratio * N/C Ratio * 44/28
NOx Emission = Carbon Released * Emission Ratio * N/C Ratio * 46/14

Calculating CH4
```{r}
em_ratios <- read_csv(here("crop/ghg/data/ghg_em_ratio.csv"))
burn_ghg <- list.files(file.path(prep, "crop/ghg/crop_residue_burning_ghg/total_carbon_emitted/"), full = TRUE)

registerDoParallel(6)   
foreach(file = burn_ghg) %dopar% {
  
  #file = burn_ghg[39]
  crop <- str_sub(file, -8, -5)
  
  em_ratio <- as.numeric(em_ratios$ratio[em_ratios$compound == "ch4"])
  
  ch4_em <- read_csv(file.path(file), col_types = "ddc") %>% 
    mutate(tc = as.numeric(tc),
           ch4_emitted = tc*em_ratio*16/12) %>% 
    dplyr::select(x,y,ch4_emitted)
  
   write_csv(ch4_em, file.path(prep, paste("crop/ghg/crop_residue_burning_ghg/ch4_emitted/", crop, ".csv", sep = "")))
}
```

Calculating CO
```{r}
em_ratios <- read_csv(here("crop/ghg/data/ghg_em_ratio.csv"))
burn_ghg <- list.files(file.path(prep, "crop/ghg/crop_residue_burning_ghg/total_carbon_emitted/"), full = TRUE)

foreach(file = burn_ghg) %dopar% {
  
 # file = burn_ghg[41]
  crop <- str_sub(file, -8, -5)
  
  em_ratio <- as.numeric(em_ratios$ratio[em_ratios$compound == "co"])
  
  co_em <- read_csv(file.path(file), col_types = "ddc") %>% 
    mutate(tc = as.numeric(tc),
           ch4_emitted = tc*em_ratio*28/12) %>% 
    dplyr::select(x,y,ch4_emitted)
  
   write_csv(co_em, file.path(prep, paste("crop/ghg/crop_residue_burning_ghg/co_emitted/", crop, ".csv", sep = "")))
}
```

Calculating N2O
```{r}
em_ratios <- read_csv(here("crop/ghg/data/ghg_em_ratio.csv"))
burn_ghg <- list.files(file.path(prep, "crop/ghg/crop_residue_burning_ghg/total_carbon_emitted/"), full = TRUE)

registerDoParallel(6)   
foreach(file = burn_ghg) %dopar% {
  
# file = burn_ghg[1]
  crop <- str_sub(file, -8, -5)
  
  nc_ratio <- burn_parameters %>% 
    select(split_rest_names, n_c_ratio) %>% 
    rename(crop = split_rest_names)
  
  em_ratio <- as.numeric(em_ratios$ratio[em_ratios$compound == "n2o"])
  
  n2o_em <- read_csv(file.path(file), col_types = "ddc") %>% 
    mutate(crop = crop) %>% 
    left_join(nc_ratio) %>% 
    mutate(tc = as.numeric(tc),
           em_ratio = em_ratio,
           n2o_emitted = tc*em_ratio*n_c_ratio*44/28) %>% 
    dplyr::select(x,y,n2o_emitted)

   write_csv(n2o_em, file.path(prep, paste("crop/ghg/crop_residue_burning_ghg/n2o_emitted/", crop, ".csv", sep = "")))
}
```

Calculating Nox
```{r}
em_ratios <- read_csv(here("crop/ghg/data/ghg_em_ratio.csv"))
burn_ghg <- list.files(file.path(prep, "crop/ghg/crop_residue_burning_ghg/total_carbon_emitted/"), full = TRUE)

registerDoParallel(6)   
foreach(file = burn_ghg) %dopar% {
  
  # file = burn_ghg[1]
 crop <- str_sub(file, -8, -5)
  
  nc_ratio <- burn_parameters %>% 
    dplyr::select(split_rest_names, n_c_ratio) %>% 
    rename(crop = split_rest_names)
  
  em_ratio <- as.numeric(em_ratios$ratio[em_ratios$compound == "nox"])
  
  nox_em <- read_csv(file.path(file), col_types = "ddc") %>% 
    mutate(crop = crop) %>% 
    left_join(nc_ratio) %>% 
    mutate(tc = as.numeric(tc),
           em_ratio = em_ratio,
           nox_emitted = tc*em_ratio*n_c_ratio*44/28) %>% 
    dplyr::select(x,y,nox_emitted)
  
   write_csv(nox_em, file.path(prep, paste("crop/ghg/crop_residue_burning_ghg/nox_emitted/", crop, ".csv", sep = "")))
}
```

Now we want to convert CH4 nd NO2 values to carbon equivilant and sum those up in one raster per crop

CH4
```{r}
ch4 <- list.files(file.path(prep, "crop/ghg/crop_residue_burning_ghg/ch4_emitted/"), full = TRUE)
eq_factors <- read_csv(here("crop/ghg/data/co2_eq_factors.csv"))

registerDoParallel(6)   
foreach(file = ch4) %dopar% {
  
 # file = ch4[41]
 crop <- str_sub(file, -8, -5)
 
 eq_fac <- as.numeric(eq_factors$gwp[eq_factors$source == "ch4"])
  
  ch4_eq <- read_csv(file.path(file), col_types = "ddc") %>% 
    mutate(eq_fac = eq_fac,
           ch4_emitted = as.numeric(ch4_emitted),
           co2eq = ch4_emitted*eq_fac) %>% 
    dplyr::select(x, y, co2eq)
  
  ch4_eq_rast <- rasterFromXYZ(ch4_eq, crs = food_crs)
  
   writeRaster(ch4_eq_rast, file.path(prep, paste("crop/ghg/crop_residue_burning_ghg/co2eq_ch4/", crop, ".tif", sep = "")),  format="GTiff", overwrite=TRUE)
}
```

```{r}
n2o <- list.files(file.path(prep, "crop/ghg/crop_residue_burning_ghg/n2o_emitted/"), full = TRUE)
eq_factors <- read_csv(here("crop/ghg/data/co2_eq_factors.csv"))

registerDoParallel(6)   
foreach(file = n2o) %dopar% {
  
 #  file = n2o[38]
 crop <- str_sub(file, -8, -5)
 
 eq_fac <- as.numeric(eq_factors$gwp[eq_factors$source == "n2o"])
  
  n2o_eq <- read_csv(file.path(file), col_types = "ddc") %>% 
    mutate(eq_fac = eq_fac,
           n2o_emitted = as.numeric(n2o_emitted),
           co2eq = n2o_emitted*eq_fac) %>% 
    dplyr::select(x, y, co2eq)
   
     n2o_eq_rast <- rasterFromXYZ(n2o_eq, crs = food_crs)
  
   writeRaster(n2o_eq_rast, file.path(prep, paste("crop/ghg/crop_residue_burning_ghg/co2eq_n2o/", crop, ".tif", sep = "")),  format="GTiff", overwrite=TRUE)
}

list.files(file.path(prep, 
                     "crop/ghg/crop_residue_burning_ghg/", 
                     sep = ""))

```

```{r}
ch4_list <- list.files(file.path(prep, "crop/ghg/crop_residue_burning_ghg/co2eq_ch4"), full = TRUE)
  
registerDoParallel(6)           
foreach(ch4 = ch4_list) %dopar% {
  
# ch4 = ch4_list[38]
 crop <- str_sub(ch4, -8, -5)
 
 ch4 <- raster(ch4)
 n2o <-  raster(paste(file.path(prep, "crop/ghg/crop_residue_burning_ghg/co2eq_n2o/"), crop, ".tif", sep = ""))
 
  crop_stack <- raster::stack(ch4,n2o)
  
  crop_add <- sum(crop_stack, na.rm= TRUE)
 
   writeRaster(crop_add, filename = paste(prep, "crop/ghg/crop_residue_burning_ghg/co2eq_emitted/",crop,".tif", sep = ""), format="GTiff", overwrite=TRUE)
}
```

Lets look at it!
```{r}
plot(raster(file.path(prep, "crop/ghg/crop_residue_burning_ghg/co2eq_ch4/whea.tif")), main ="Wheat burning: co2-eq from ch4")

plot(raster(file.path(prep, "crop/ghg/crop_residue_burning_ghg/co2eq_ch4/maiz.tif")), main ="Corn burning: co2-eq from ch4")

plot(raster(file.path(prep, "crop/ghg/crop_residue_burning_ghg/co2eq_ch4/othr.tif")), main ="tnut burning: co2-eq from ch4")

```

# list files
```{r}
list.files(file.path(prep, 
                     "crop/ghg/crop_residue_burning_ghg/ch4_emitted", 
                     sep = ""),
           pattern = "tnut|othr|spis")

list.files(file.path(prep, 
                     "crop/ghg/crop_residue_burning_ghg/co_emitted", 
                     sep = ""),
           pattern = "tnut|othr|spis")

list.files(file.path(prep, 
                     "crop/ghg/crop_residue_burning_ghg/co2eq_ch4", 
                     sep = ""),
           pattern = "tnut|othr|spis")

list.files(file.path(prep, 
                     "crop/ghg/crop_residue_burning_ghg/co2eq_emitted", 
                     sep = ""),
           pattern = "tnut|othr|spis")

list.files(file.path(prep, 
                     "crop/ghg/crop_residue_burning_ghg/co2eq_n2o", 
                     sep = ""),
           pattern = "tnut|othr|spis")

list.files(file.path(prep, 
                     "crop/ghg/crop_residue_burning_ghg/nox_emitted", 
                     sep = ""),
           pattern = "tnut|othr|spis")

list.files(file.path(prep, 
                     "crop/ghg/crop_residue_burning_ghg/total_carbon_emitted", 
                     sep = ""),
           pattern = "tnut|othr|spis")

list.files(file.path(prep, 
                     "crop/ghg/crop_residue_burning_ghg/n2o_emitted", 
                     sep = ""),
           pattern = "tnut|othr|spis")
```

