---
title: "GHG emissions from N2O volatilization of synthetic fertilizers"
author: "Juliette"
date: "4/30/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

Calculate CO2-eq emissions from synthetic fertilizers left on fields. We used methods from Global Livestock Environmental Assessment Model Version 2.0

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
library(raster)
library(dplyr)
library(tidyverse)
library(here)
library(foreach)
library(doParallel)

source(here("_workflow/common.R"))
source(here("_workflow/common_spatial.R"))

conversion <- read_csv(here("crop/ghg/data/co2_eq_factors.csv"))
```


Using equation 6.6 in the GLEAM document

Equation: 

N2O = app_rate* (ef_dir + frac_vol* ef_vol + frac_leach * ef_leach) * 44/28

**N2O** = total nitrous oxide emissions from application of nitrogen fertilizer per hectare of feed material (kg N2O / ha)

**app_rate** = application rate of nitrogen fertilizer for feed material kg N/ha

**ef_dir** = emission factor for direct nitrous oxide emissions, kg N2O-N·kg N-1. Default value of 0.01. (0.003 is reported for flooded rice feed materials but we used the default value because of uncertainty about the location of flooded rice). 

**frac_vol** = fraction of nitrogen from synthetic fertilizers or crop residues that volatilizes as NH3 and NOx, fraction. Default value of 0.1 is used.

**ef_vol** = emission factor for N2O emissions from N volatilized as NH3 and NOx, kg N2O-N·kg N volatilized-1. Default value of 0.01 is used.

**frac_leach** = fraction of nitrogen lost due to leaching and runoff, fraction. Default value of 0.3 is used.

**ef_leach** = emission factor for N2O emissions from leaching and runoff, kg N2O-N·kg N leached-1. Default value of 0.0075 is used.

Emission factors
```{r}
ef_dir <- 0.01
ef_dir_rice <- 0.01  # gleam documents report 0.03, but orginal source is 0.003, https://www.ipcc-nggip.iges.or.jp/public/2006gl/pdf/4_Volume4/V4_11_Ch11_N2O&CO2.pdf, Table 11.1
frac_vol <- 0.1
ef_vol <- 0.01
frac_leach <- 0.3
ef_leach <- 0.0075
```

Need to convert the tonnnes of fertilizer applied to kg for these calcualtions and then back to tonnes for the final layer

Read in fertilizer application rasters
```{r}
fertilizer_list <- list.files(paste0(prep, "crop/nutrient", sep = ""), pattern = "application_N", full = TRUE)
fertilizer_list <- fertilizer_list[!grepl(pattern = "toba|cof|teas", fertilizer_list)]
registerDoParallel(6)   

foreach(fertilizer = fertilizer_list) %dopar% {

   #fertilizer <- fertilizer_list[38]
    crop <- str_sub(fertilizer, -22, -19)
    
    app_rate <- raster(fertilizer, crs = food_crs)*1000 #convert from tonnes to kg
    app_rate_df <- as.data.frame(app_rate, xy= TRUE)
    
    n2o <- app_rate_df %>% 
      mutate(crop = crop) %>% 
      mutate(ef_dir = ifelse(crop == "rice", ef_dir_rice, ef_dir),
           frac_vol = frac_vol,
           ef_vol = ef_vol,
           frac_leach =frac_leach,
           ef_leach = ef_leach) %>%
      rename(n_app = paste0("crop_", crop, "_application_N")) %>% 
      mutate(n2o_tonnes = n_app* (ef_dir + frac_vol*ef_vol + frac_leach*ef_leach)*(44/28)/1000) %>% 
      dplyr::select(x,y,n2o_tonnes)
      
  n2o_raster <- rasterFromXYZ(n2o, crs= food_crs)

    writeRaster(n2o_raster, filename =
                  paste(prep, "crop/ghg/crop_fertilizer_N2O/n2o/fertilizer_n2o_",crop,".tif", sep = ""), format="GTiff", overwrite=TRUE)
    
}

delete_aux <- list.files(file.path(prep, "crop/ghg/crop_fertilizer_N2O/n2o"), pattern="aux", full=TRUE)
file.remove(delete_aux)

```


Calculate the CO2 equivalent
```{r}

n2o_list <- list.files(file.path(prep, "crop/ghg/crop_fertilizer_N2O/n2o"), full = TRUE)
n2o_list <- n2o_list[!grepl(pattern = "toba|cof|teas", n2o_list)]

foreach(file = n2o_list) %dopar% {

  #  file <- n2o_list[38]
    crop <- str_sub(file, -8, -5)
    
    co2_eq <- as.numeric(conversion$gwp[conversion$source == "n2o"])
  
    co2_eq_raster <- raster(file)*co2_eq

    writeRaster(co2_eq_raster, filename =
                  paste(prep, "crop/ghg/crop_fertilizer_N2O/co2_eq/fert_n2o_co2eq_",crop,".tif", sep = ""), format="GTiff", overwrite=TRUE)
    
}

delete_aux <- list.files(file.path(prep, "crop/ghg/crop_fertilizer_N2O/co2_eq"), pattern="aux", full=TRUE)
file.remove(delete_aux)

```










