---
title: "ghg_pesticides"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Script Description
Map pesticide emissions

# Data information
## Pesticide use for agricultural use.
Accessed: 11/24/2020 from FAOSTAT database.
Year: 2016
Unit: tonnes.

## Pesticide emission factors data:
We used Lal 2004 values for herbicide, insecticide and fungicide emission factors. All other pesticide types in FAOSTAT were assumed to have an emission factor of 25.5 kg CO2-eq based on LEAP (2017) (whom quote Audley et al 2009).

We assume that proportional pesticide application perfectly correlates with synthetic nitrogen fertilizer application.

# Preamble
```{r setup, include = FALSE}
# getting packages we want
library(here)
source(here("_workflow/common.R"))
source(here("_workflow/common_spatial.R"))
source(here("_spatial/template_raster.R"))
library(janitor)
library(furrr)

food_rgns_xy <- select(food_rgns_xy, c(x, y, iso3c))
``` 

# Wrangle

Getting the units right:
Emission factor is from kg of CO2eq to kg of a.i (ie a one-to-one conversion)
FAOSTAT reports a.i in tonnes.
Thus, we don't need to convert units.

```{r}
FAO_pesticide <- 
  read_csv("crop/ghg/data/FAOSTAT_pesticide_use.csv") %>% 
  clean_names() %>% 
  select(area_code, item, value) %>% 
  rename(tonnes = value, pesticide_type = item) %>% 
  left_join(read_csv(here("_spatial/_output/FAO_rgn_codes.csv"))) %>% 
  filter(!area_code %in% 351) %>% # remove China w/ territories
  filter(pesticide_type %in% "Pesticides (total)") %>%  
  left_join(clean_names(food_rgns)) %>%
  filter(!iso3c %in% c("XMI", "GGY"))%>% 
  mutate(CO2eq = case_when(
    pesticide_type %in% "Herbicides" ~   6.3 * 44/12 * tonnes,
    pesticide_type %in% "Insecticides" ~ 5.1 * 44/12 * tonnes,
    pesticide_type %in% "Fungicides" ~   3.9 * 44/12 * tonnes,
    TRUE ~ 25.5 * tonnes)) %>% # from LEAP, 2017/Audley et al 2009; could also do an average of the three values.
  group_by(iso3c) %>%
  summarize(CO2eq = sum(CO2eq)) %>%
  ungroup() %>%
  mutate(CO2eq = ifelse(is.na(CO2eq),
                        0,
                        CO2eq))
```

# FAO_pesticide check
```{r}
FAO_pesticide %>% 
  filter(iso3c %in% c("USA", "BRA", "SRB", "SDN", 
                      "TWN", "ALA", "XMI", "XKO", 
                      "PRT", "GBR", "GGY")) %>%
  group_by(iso3c) %>% 
  summarize(total_CO2eq_tonnes = sum(CO2eq)) %>% 
  ungroup()
```

```{r}

pesticide_prop <- 
  read_csv(here("crop/nutrient/data/nutrient_df_full.csv")) %>% 
  filter(nutrient == "N") %>%
  group_by(iso3c) %>% 
  mutate(sum_harvested_area = sum(harvested_area)) %>% 
  ungroup() %>%
  mutate(prop_pesticide = prop_nutrient) %>%
  select(iso3c, nutrient, prod_system, split_rest_names, sum_harvested_area, prop_pesticide)

tmp <- pesticide_prop %>%
  group_by(iso3c) %>%
  summarize(total_prop = sum(prop_pesticide))
summary(tmp) # should all be one or zero

sum(duplicated(pesticide_prop)) # should be zero

pesticide_prop <- 
  pesticide_prop %>% 
  left_join(., FAO_pesticide)

## gapfill pesticide application when no FAO data
## based on 
rgn_median <- pesticide_prop %>%
  mutate(CO2eq_per_area = CO2eq/sum_harvested_area) %>%
  select(iso3c, CO2eq_per_area) %>%
  unique()
rgn_gf <- median(rgn_median$CO2eq_per_area, na.rm=TRUE)

pesticide_prop <- 
  pesticide_prop %>% 
  mutate(CO2eq_gf = case_when(
    is.na(CO2eq) & sum_harvested_area > 0 ~ rgn_gf*sum_harvested_area,
    is.na(CO2eq) & sum_harvested_area == 0 ~ 0,
    TRUE ~ CO2eq))

# Allocate

pesticide_prop <- 
  pesticide_prop %>% 
  mutate(allocated_pesticide = CO2eq_gf * prop_pesticide)


#should be about equal
sum(unique(pesticide_prop$CO2eq_gf))
sum(pesticide_prop$allocated_pesticide, na.rm = TRUE)

```

# Trim
```{r}
pesticide_prop <- 
  pesticide_prop %>% 
  select(iso3c, nutrient, prod_system, split_rest_names, allocated_pesticide)

# these should be about the same: pesticide_prop can be a bit higher due to gapfilling
sum(pesticide_prop$allocated_pesticide)
sum(FAO_pesticide$CO2eq)

```

# Save
```{r}
write_csv(pesticide_prop, here("crop/ghg/data/pesticide_df.csv"))
```


