---
title: "Caculating GHG emissions values from farming machinery for different production systems and types"
author: "Juliette"
date: "3/23/2020"
output: html_document
---

In this markdown we create a df of carbon emission factors from on farm maintaince for each of our crops.

All values taken from https://www.sciencedirect.com/science/article/pii/S0160412004000832#TBL4
The emission factors are in kg CE/ha

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read in libraries and paths
```{r}
#Load libraries
library(readr)
library(dplyr)
library(tidyverse)

#Files we will need:
crops <- vroom::vroom(here("crop/farm/data/crop_codes_updated.csv")) %>% 
  left_join(., vroom::vroom("crop/ghg/data/mapspam_crops.csv")[c("SPAM_short_name","GROUP")])
act_em <- read_csv(here("crop/ghg/data/farm_mach_activites_em.csv"))
```

Create a df with all the ghg emission averages for each category of 

Set up the df to fill in with the em values
```{r}
setup <- crops %>% 
  dplyr::select(split_rest_names) %>% 
  slice(rep(1:n(), each = 4)) %>% 
  group_by(split_rest_names) %>% 
  mutate(id = row_number()) %>% 
  ungroup() %>% 
  mutate(input = case_when(id == 1 ~ "high irrigated",
                           id == 2 ~ "high rain",
                           id == 3 ~ "low",
                           id == 4 ~ "subsistence")) %>% 
  left_join(crops, by = c("split_rest_names")) %>% 
   dplyr::select(split_rest_names, input, GROUP) 
```

Filling in the GHG values

We will start with the first three since those are the most straight forward and not crop specific (except one):

Fertilizer anad Herbicide/pesticide
```{r}
### Fertilizer
fertilizer_ir <- (as.numeric(act_em$value[act_em$action == "fertilizer spraying"]))*(as.numeric(act_em$irrigated[act_em$action == "fertilizer spraying"])) + (as.numeric(act_em$value[act_em$action == "fertilizer spreading"]))*(as.numeric(act_em$irrigated[act_em$action == "fertilizer spreading"]))

fertilizer_rain_hi <- (as.numeric(act_em$value[act_em$action == "fertilizer spraying"]))*(as.numeric(act_em$rain[act_em$action == "fertilizer spraying"])) + (as.numeric(act_em$value[act_em$action == "fertilizer spreading"]))*(as.numeric(act_em$rain[act_em$action == "fertilizer spreading"]))  

### Herbicide/pesticide
herbicide_ir <- as.numeric(act_em$value[act_em$action == "spray herbicide"])*(as.numeric(act_em$irrigated[act_em$action == "spray herbicide"]))

herbicide_rain <- as.numeric(act_em$value[act_em$action == "spray herbicide"])*(as.numeric(act_em$rain[act_em$action == "spray herbicide"]))

### Add it to the df
setup_fert_herb <- setup %>% 
  mutate(fertilizer_em = case_when(input == "high irrigated" ~ fertilizer_ir,
                                  input == "high rain" ~ fertilizer_rain_hi,
                                  input == "low" | input == "subsistence" ~ 0)) %>% 
  mutate(herbicide_em = case_when(input == "high irrigated" ~ herbicide_ir,
                                  input == "high rain" ~ herbicide_rain,
                                  input == "low" | input == "subsistence" ~ 0))
```  
  
Tilling/seed prep
Pathways are the same for high irrigation and rain, but I will separate out in case we change that in the future. So there is a little code repetition here
```{r}
## Tilling
till_ir <- (as.numeric(act_em$value[act_em$action == "moldboard plowing"]))*(as.numeric(act_em$irrigated[act_em$action == "moldboard plowing"])) + (as.numeric(act_em$value[act_em$action == "average of reduced tilling"]))*(as.numeric(act_em$irrigated[act_em$action == "average of reduced tilling"]))

till_rain <- (as.numeric(act_em$value[act_em$action == "moldboard plowing"]))*(as.numeric(act_em$rain[act_em$action == "moldboard plowing"])) + (as.numeric(act_em$value[act_em$action == "average of reduced tilling"]))*(as.numeric(act_em$rain[act_em$action == "average of reduced tilling"]))

## Seeding
seed_ir <- (as.numeric(act_em$value[act_em$action == "plant/sow/drill"]))*(as.numeric(act_em$irrigated[act_em$action == "plant/sow/drill"])) + (as.numeric(act_em$value[act_em$action == "no-till planting"]))*(as.numeric(act_em$irrigated[act_em$action == "no-till planting"]))

seed_rain <- (as.numeric(act_em$value[act_em$action == "plant/sow/drill"]))*(as.numeric(act_em$rain[act_em$action == "plant/sow/drill"])) + (as.numeric(act_em$value[act_em$action == "no-till planting"]))*(as.numeric(act_em$rain[act_em$action == "no-till planting"]))

## potatoes have their own seeding pathway
seed_ir_pota <- (as.numeric(act_em$value[act_em$action == "potato planter"]))*(as.numeric(act_em$irrigated[act_em$action == "potato planter"]))

seed_rain_pota <- (as.numeric(act_em$value[act_em$action == "potato planter"]))*(as.numeric(act_em$rain[act_em$action == "potato planter"]))

### combine
## irrigated high
till_seed_ir <- till_ir + seed_ir
## rain high
till_seed_rain <- till_rain + seed_rain

## irrigated high potatoes
till_seed_ir_pota <- till_ir + seed_ir_pota
## rain high potaotes
till_seed_rain_pota <- till_rain + seed_rain_pota

### Add it to the df
setup_till_seed <- setup_fert_herb %>% 
  mutate(till_seed_em = case_when(input == "high irrigated" ~ till_seed_ir,
                                  input == "high rain" ~ till_seed_rain,
                                  input == "low" | input == "subsistence" ~ 0)) %>% 
  mutate(till_seed_em = ifelse(input == "high irrigated" & split_rest_names == "pota", till_seed_ir_pota,
                               ifelse(input == "high rain" & split_rest_names == "pota", till_seed_rain_pota,
                                      till_seed_em)))
```

The harvesting portion of this is the more complicated one. We identified 7 different combinations of machinery usage, some are crop specific and some not

1) windrower + rake + baler 
2) forage harvesting + baler
3) forage harvesting 
4) windrower + rake
5) corn silage
6) corn harvesting combine + shred corn stalk
7) soybean harvesting combine

We're going to read in all the values here so it's easier to follow
```{r}
windrower <- as.numeric(act_em$value[act_em$action == "windrower"])
rake <- as.numeric(act_em$value[act_em$action == "rake"])
baler <- as.numeric(act_em$value[act_em$action == "baler (rectangle and large round, averaged)"])
forage_harvesting <- as.numeric(act_em$value[act_em$action == "forage harvesting"])
baler <- as.numeric(act_em$value[act_em$action == "baler (rectangle and large round, averaged)"])
corn_silage <- as.numeric(act_em$value[act_em$action == "corn silage"])
corn_combine <- as.numeric(act_em$value[act_em$action == "corn harvesting combine"])
shed_corn <- as.numeric(act_em$value[act_em$action == "shred corn stalk"])
soybean_combine <- as.numeric(act_em$value[act_em$action == "soybean harvesting combine"])
```

calculate for high irrigation and high rain
```{r}
## windrower + rake + baler
harv_1_ir <- windrower*(as.numeric(act_em$irrigated[act_em$action == "windrower"])) + rake*(as.numeric(act_em$irrigated[act_em$action == "rake"])) + baler*(as.numeric(act_em$irrigated[act_em$action == "baler (rectangle and large round, averaged)"]))

harv_1_rain <- windrower*(as.numeric(act_em$rain[act_em$action == "windrower"])) + rake*(as.numeric(act_em$rain[act_em$action == "rake"])) + baler*(as.numeric(act_em$rain[act_em$action == "baler (rectangle and large round, averaged)"]))

## forage harvesting + baler
harv_2_ir <- forage_harvesting*(as.numeric(act_em$irrigated[act_em$action == "forage harvesting"])) + baler*(as.numeric(act_em$irrigated[act_em$action == "baler (rectangle and large round, averaged)"]))

harv_2_rain <- harv_2_ir <- forage_harvesting*(as.numeric(act_em$rain[act_em$action == "forage harvesting"])) + baler*(as.numeric(act_em$rain[act_em$action == "baler (rectangle and large round, averaged)"]))

## forage harvesting
harv_3_ir <- forage_harvesting*(as.numeric(act_em$irrigated[act_em$action == "forage harvesting"]))

harv_3_rain <- forage_harvesting*(as.numeric(act_em$rain[act_em$action == "forage harvesting"]))

## windrower + rake
harv_4_ir <-  windrower*(as.numeric(act_em$irrigated[act_em$action == "windrower"])) + rake*(as.numeric(act_em$irrigated[act_em$action == "rake"])) 
harv_4_rain <-  windrower*(as.numeric(act_em$irrigated[act_em$action == "windrower"])) + rake*(as.numeric(act_em$rain[act_em$action == "rake"])) 

## corn silage
harv_5_ir <- corn_silage*(as.numeric(act_em$irrigated[act_em$action == "corn silage"]))

harv_5_rain <- corn_silage*(as.numeric(act_em$rain[act_em$action == "corn silage"]))

##corn combine and shred corn stalk
harv_6_ir <- corn_combine*(as.numeric(act_em$irrigated[act_em$action == "corn harvesting combine"])) + shed_corn*(as.numeric(act_em$irrigated[act_em$action == "shred corn stalk"]))

harv_6_rain <- corn_combine*(as.numeric(act_em$rain[act_em$action == "corn harvesting combine"])) + shed_corn*(as.numeric(act_em$rain[act_em$action == "shred corn stalk"]))

##soybean combine
harv_7_ir <- soybean_combine*(as.numeric(act_em$irrigated[act_em$action == "soybean harvesting combine"]))

harv_7_rain <- soybean_combine*(as.numeric(act_em$rain[act_em$action == "soybean harvesting combine"]))
```

Some of the path ways are crop specific:
harv_1 (windrower + rake + baler): cereals 
harv_2 (forage harvesting + baler): cereals
harv_3 (forage harvesting): cereals, corn, soybeans
harv_4 (windrower + rake): cereals, soybean
harv_5 (corn silage): corn
harv_6 (corn harvesting combine + shred corn stalk): corn
harv_7 (soybean harvesting combine): soybean

Calculate averages for each crop and input type, and then use an average off all the non-specialized equipment pathways as a value for the rest of crops
```{r}
harv_cereals_ir <- mean(c(harv_1_ir, harv_2_ir, harv_3_ir, harv_4_ir))
harv_cereals_rain <- mean(c(harv_1_rain, harv_2_rain, harv_3_rain, harv_4_rain))

harv_corn_ir <- mean(c(harv_3_ir, harv_5_ir, harv_6_ir))
harv_corn_rain <- mean(c(harv_3_rain, harv_5_rain, harv_6_rain))

harv_soy_ir <- mean(c(harv_3_ir, harv_4_ir, harv_7_ir))
harv_soy_rain <- mean(c(harv_3_rain, harv_4_rain, harv_7_rain))

harv_rest_ir <- mean(c(harv_1_ir, harv_2_ir, harv_3_ir, harv_4_ir, harv_5_ir))
harv_rest_rain <- mean(c(harv_1_rain, harv_2_rain, harv_3_rain, harv_4_rain, harv_5_rain))
```


Add a harvesting emissions column for each specific input and crop (where applicable)
```{r}
machinery_em <- setup_till_seed %>% 
  mutate(harvesting_em = case_when(input == "high irrigated" & split_rest_names == "maiz" ~ harv_corn_ir,
                                   input == "high rain" & split_rest_names == "maiz" ~ harv_corn_rain,
                                   
                                   input == "high irrigated" & split_rest_names == "soyb" ~ harv_soy_ir,
                                   input == "high rain" & split_rest_names == "soyb" ~ harv_soy_rain,

                                   input == "high irrigated" & GROUP == "cereals" ~ harv_cereals_ir,
                                   input == "high rain" & GROUP == "cereals" ~ harv_cereals_rain,
                                   
                                   input == "low" | input == "subsistence" ~ 0
                                   )) %>% 
  
  mutate(gapfill_desc = case_when(is.na(harvesting_em) ~ "harvesting_em gapfilled with the average of all pathways except those with soybean or corn combines")) %>% 
  
  mutate(harvesting_em = ifelse(input == "high irrigated" & is.na(harvesting_em), harv_rest_ir,
                                ifelse(input == "high rain" & is.na(harvesting_em), harv_rest_rain,
                                       harvesting_em))) %>% 
  rowwise() %>% 
  mutate(all_em  = sum(c(fertilizer_em, herbicide_em, till_seed_em, harvesting_em))) %>% 
  mutate(units = "kg_co2eq/hectare")

write_csv(machinery_em, here("crop/ghg/data/machinery_em.csv"))
```
