---
title: "crop_ghg_fertilizer_production_and_transport"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Script Description
Map GHG emissions stressor for the production and transport of fertilizer.

# Data information
Source: FAO 2017 LEAP report http://www.fao.org/3/a-i8275e.pdf [Accessed: 04/13/2020]

FAO_LEAP programme used Kool et al 2012 and Blanca(?) 2011 to calculate emission factors for the production and transport of fertilizer NPK and lime.
While we don't use potash (K2O) to calculate leaching, we are calculating emissions.

# Preamble
```{r setup, include = FALSE}
# getting packages we want
library(here)
source(here("_workflow/common.R"))
source(here("_workflow/common_spatial.R"))
source(here("_spatial/template_raster.R"))
library(janitor)
library(furrr)

# Import food_rgns xy df
food_rgns_xy <- 
  food_rgns_xy %>% 
  select(x, y, iso3c)
```

-------------------------------------------------------------------------------

# Tidy and transform

## Import and wrangling
```{r}
nutrient_EF <- 
  vroom::vroom(here("crop/ghg/data/FAO_2017_LEAP_EF.csv")) %>% 
  clean_names() %>% 
  rename(LEAP_region = region, 
         N           = n,
         P2O5        = p2o5,
         K2O         = k2o)

# Match LEAP regions with equivalent UNSD to get iso3c
match_region <- 
  vroom::vroom(here("_spatial/_output/UNSD_Methodology.csv")) %>% 
  mutate(LEAP_region = case_when(
    Sub_region_Name %in% "Western Europe" ~ "W. Europe",
    Sub_region_Name %in% "Eastern Europe" ~ "E.Europe inclu Russian federation",
    Intermediate_Region_Name %in% "Central America" ~ "Central & S.America",
    Intermediate_Region_Name %in% "South America" ~ "Central & S.America",
    Region_Name %in% "Asia" ~ "Asia",
    iso3c %in% "AUS" ~ "Australia",
    iso3c %in% "NZL" ~ "New Zealand",
    TRUE ~ "Global Average")) %>% 
  select(iso3c, LEAP_region) 

# Join iso3c codes to respective global region.
nutrient_EF <- 
  left_join(nutrient_EF, match_region) %>% 
  select(-LEAP_region)

# Gapfill
missing_rgns <- unique(nutrient_EF$iso3c)[!unique(nutrient_EF$iso3c) %in% food_rgns$iso3c]
nutrient_EF <- nutrient_EF %>% filter(!iso3c %in% missing_rgns)

# Change units
nutrient_EF <- 
  nutrient_EF %>% 
  gather(key = "nutrient", 
         value = "nutrient_CO2eq", 
         -iso3c, 
         na.rm = TRUE)

# Save
write_csv(nutrient_EF, here("crop/ghg/data/nutrient_EF.csv"))
```

Combine with nutrient dataframe (created in crop/nutrient folder)
```{r}
nutrient_EF <- read_csv(here("crop/ghg/data/nutrient_EF.csv"))
nutrient_df <- vroom::vroom(here("crop/nutrient/data/nutrient_df.csv"))
nutrient_EF <- left_join(nutrient_df, nutrient_EF)

```

# Data check
```{r}
# Must be the same #244
n_distinct(unique(nutrient_EF$iso3c));+ 
n_distinct(unique(nutrient_df$iso3c));+ 
n_distinct(food_rgns$iso3c) 

# Expecting reasonable values, XMI should be zeros due to absence of fertilizer data.
nutrient_EF %>% 
  filter(iso3c %in% c("USA", "BRA", "SRB", "SDN", "TWN", "ALA", "XMI", "XKO", "PRT", "GBR", "GGY")) %>%
  group_by(iso3c) %>% 
  summarize(total_tonnes = sum(allocated_nutrient))
```


-------------------------------------------------------------------------------

# Mapping
```{r}
# Distinct variables for loops
distinct_vars <- 
  nutrient_df %>% 
  distinct(split_rest_names, nutrient) 

```

## Rasterize
```{r}
# Read nutrient_EF_raster function 
nutrient_EF_raster <- function(this_crop, this_nutrient) { #this_nutrient = "N"; this_crop="maiz"
  
    # Split by nutrient and crop
  nutrient_EF %>% 
    filter(split_rest_names %in% this_crop & 
           nutrient %in% this_nutrient) %>%
    select(iso3c, nutrient_CO2eq) %>% 
    left_join(food_rgns_xy, .) %>% # add iso3c
    select(-iso3c) %>% 
    rasterFromXYZ(crs = food_crs) %>% # rasterize
    
    # Multiply emission factor by the tonnes of nutrients in each raster cell.
    overlay(.,
            raster(file.path(prep,
                         "crop/nutrient/crop_",
                         this_crop,
                         "_application_",
                         this_nutrient,
                         ".tif", 
                         fsep = "")), 
            fun = function(r1, r2) {
              
              return(r1 * r2)
              
            }) %>% 
    
    writeRaster(.,
                file.path(prep,
                          "crop/ghg/crop_nutrient_ghg/crop_",
                          this_crop,
                          "_",
                          this_nutrient,
                          "_nutrient_CO2eq.tif",
                          fsep = ""), 
                format = "GTiff", 
                overwrite = TRUE)
  }

plan(multisession, workers = 12)

options <- furrr_options(seed = 54232)

future_map2(distinct_vars$split_rest_names,
            distinct_vars$nutrient,
            nutrient_EF_raster,
            .options = options)

delete_aux <- list.files(file.path(prep, "crop/ghg/crop_nutrient_ghg"), pattern="aux", full=TRUE)
file.remove(delete_aux)
```                                      

# Sum together NPK values
```{r}
total_nutrient_CO2eq <- function(crop){
  
  CO2eq_raster <- overlay(
    raster(file.path(prep,
                     "crop/ghg/crop_nutrient_ghg/crop_",
                     crop,
                     "_N_nutrient_CO2eq.tif", 
                     fsep = "")), 
    raster(file.path(prep,
                     "crop/ghg/crop_nutrient_ghg/crop_",
                     crop,
                     "_P2O5_nutrient_CO2eq.tif", 
                     fsep = "")),
    raster(file.path(prep,
                     "crop/ghg/crop_nutrient_ghg/crop_",
                     crop,
                     "_K2O_nutrient_CO2eq.tif", 
                     fsep = "")),
    
    fun = function(r1, r2, r3) {
       
        r1[is.na(r1)] <- 0
        r2[is.na(r2)] <- 0
        r3[is.na(r3)] <- 0
        return(r1 + r2 + r3)
        
       })
  
  names(CO2eq_raster) <- paste(crop, "_all_nutrient_CO2eq", sep = "")

  writeRaster(CO2eq_raster, 
              file.path(prep,
                        "crop/ghg/crop_nutrient_ghg/crop_",
                        crop,
                        "_all_nutrient_CO2eq.tif", 
                        fsep = ""), 
              format = "GTiff", 
              overwrite = TRUE)
}

future_map(unique(distinct_vars$split_rest_names),
           total_nutrient_CO2eq,
           .options = options)
 
delete_aux <- list.files(file.path(prep, "crop/ghg/crop_nutrient_ghg"), pattern="aux", full=TRUE)
file.remove(delete_aux) 
```

# Check output
```{r}
test <- 
  raster(file.path(prep,
                   "crop/ghg/crop_nutrient_ghg/crop_",
                   "maiz",
                   "_all_nutrient_CO2eq.tif", 
                   fsep = ""))
plot(test)

final <- list.files(file.path(prep,
                     "crop/ghg/crop_nutrient_ghg/",
                     fsep = ""),
           pattern = "all_nutrient_CO2eq",
           full.names = TRUE) %>% 
  stack() %>% 
  stackApply(., 1, "sum") 

plot(final) 

# these should be approximately the same  
cellStats(final, stat = "sum")


start <- nutrient_EF %>% 
  mutate(co2 = allocated_nutrient*nutrient_CO2eq)
sum(start$co2)
```


-------------------------------------------------------------------------------

# grassland 
```{r}
grassland_df <- vroom(here("crop/nutrient/data/fao_nutrient_grassland_2.csv"))

nutrient_EF <- vroom(here("crop/ghg/data/nutrient_EF.csv"))

grassland_df <- left_join(grassland_df, nutrient_EF)
```

```{r}

grassland_EF_raster <- function(this_grazing_animal) {
  
  # N
  grazing_N <- nutrient_EF %>% 
      filter(nutrient %in% "N") %>%
      select(iso3c, nutrient_CO2eq) %>% 
      left_join(food_rgns_xy, .) %>% # add iso3c
      select(-iso3c) %>% 
      rasterFromXYZ(crs = food_crs) %>% # rasterize
      
      # Multiply emission factor by the tonnes of nutrients in each raster cell.
      overlay(.,
              raster(file.path(prep,
                           "animal_farm/nutrient/grassland_fertilizer_application/",
                           this_grazing_animal,
                           "_N.tif", 
                           fsep = "")), 
              fun = function(r1, r2) {
                
                return(r1 * r2)
                
              })
  
  # P2O5
  grazing_P2O5 <- nutrient_EF %>% 
      filter(nutrient %in% "P2O5") %>%
      select(iso3c, nutrient_CO2eq) %>% 
      left_join(food_rgns_xy, .) %>% # add iso3c
      select(-iso3c) %>% 
      rasterFromXYZ(crs = food_crs) %>% # rasterize
      
      # Multiply emission factor by the tonnes of nutrients in each raster cell.
      overlay(.,
              raster(file.path(prep,
                           "animal_farm/nutrient/grassland_fertilizer_application/",
                           this_grazing_animal,
                           "_P2O5.tif", 
                           fsep = "")), 
              fun = function(r1, r2) {
                
                return(r1 * r2)
                
              })
  
  # K2O
  grazing_K2O <- nutrient_EF %>% 
      filter(nutrient %in% "P2O5") %>%
      select(iso3c, nutrient_CO2eq) %>% 
      left_join(food_rgns_xy, .) %>% # add iso3c
      select(-iso3c) %>% 
      rasterFromXYZ(crs = food_crs) %>% # rasterize
      
      # Multiply emission factor by the tonnes of nutrients in each raster cell.
      overlay(.,
              raster(file.path(prep,
                           "animal_farm/nutrient/grassland_fertilizer_application/",
                           this_grazing_animal,
                           "_K2O.tif", 
                           fsep = "")), 
              fun = function(r1, r2) {
                
                return(r1 * r2)
                
              })
  
  CO2eq_raster <- overlay(grazing_N, grazing_P2O5, grazing_K2O,
    
    fun = function(r1, r2, r3) {
       
        r1[is.na(r1)] <- 0
        r2[is.na(r2)] <- 0
        r3[is.na(r3)] <- 0
        return(r1 + r2 + r3)
        
       })
  
  names(CO2eq_raster) <- paste("grassland_all_nutrient_CO2eq", sep = "")

  writeRaster(CO2eq_raster, 
              file.path(prep,
                        "animal_farm/ghg/",
                        str_extract(this_grazing_animal, ".+?(?=_)"),
                        "_grassland_",
                        str_extract(this_grazing_animal, "(?<=_).*$"),
                        "_grss_fertilizer_CO2eq.tif", 
                        fsep = ""), 
              format = "GTiff", 
              overwrite = TRUE)
  }

grazer_list <- 
  list.files(file.path(prep, "animal_farm/nutrient/grassland_fertilizer_application"),
           pattern = ".(N|P2O5|K2O)\\.tif$") %>% 
  str_extract(., pattern = ".+?((meat|milk))") %>% 
  unique()

plan(multisession, workers = 12)

options <- furrr_options(seed = 54232)

future_map(grazer_list,
            grassland_EF_raster,
            .options = options)
```

```{r}
tmp <- list.files(file.path(prep,
                        "animal_farm/ghg/",
                     fsep = ""),
           pattern = "grss_fertilizer_CO2eq",
           full.names = TRUE) %>% 
  stack() %>% 
  stackApply(., 1, "sum") 
plot(tmp)

```

