---
title: "Create production rasters and extract per food raster"
author: "Gage Clawson (UCSB, NCEAS, OHI)"
date: "May 13, 2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

Here we need to take the mariculture production rasters and extract per each region to make a csv, so that we avoid any weird mismatches between our pressure rasters and production reporting. (The output of this will really only be used for the efficiency calculations). We will also re-save the production rasters to predatalayers folder on Aurora.

```{r, eval = FALSE}
#load relevant packages, etc.

library(here)
library(raster)
library(tidyverse)
library(doParallel)
library(foreach)
library(parallel)

# raster template
r <- raster::raster(ncol=720, nrow=360, vals=c(1:259200))
food_raster <- raster(nrows=2160, ncols=4320, xmn=-180, xmx=180, ymn=-90, ymx=90)

prep <- file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep")


source(here("_workflow/common_spatial.R"))

food_rgns <- food_rgns %>% 
  add_row(iso3c = "HSX", ID_0 = 9999, Country = "High Seas")


cat_msg <- function(x, ...) {
  if(is.null(knitr:::.knitEnv$input.dir)) {
    ### not in knitr environment, so use cat()
    cat(x, ..., '\n')
  } else {
    ### in knitr env, so use message()
    message(x, ...)
  }
  return(invisible(NULL))
}

```



Mariculture code 

Read in the data
```{r, eval = FALSE}
aq_group = c("marine_fish_general", "shrimp", "bivalve", "tuna", "salmon", "crustaceans")
species_classes <- as.data.frame(aq_group) %>%
  mutate(aq_group_fix = ifelse(aq_group == "marine_fish_general", "marine-fish-general", aq_group))


for(i in 1:nrow(species_classes)){
  
 # i = 1
  spp_class <- species_classes$aq_group[i]
  
  spp_class_fix <- species_classes$aq_group_fix[i]

production_raster <- rasterFromXYZ(read_csv(sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/%s/%s_farm_updated.csv", spp_class, spp_class)), crs = food_crs) 

# production_raster
print(cellStats(production_raster, "sum")) # 3901214
# plot(production_raster)

## save this production raster to the predatalayers folder

writeRaster(production_raster, sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers/marine_%s_aquaculture_meat_tonnes_production.tif", spp_class_fix), overwrite = TRUE) ## save to predatalayers

land_eez_rgns <- raster(file.path(prep, "spatial/land_eez_rgns.tif"))

land_eez_rgns_hsx <- calc(land_eez_rgns, fun = function(x){ifelse(is.na(x), 9999, x)})


tonnes_df <- as.data.frame(raster::zonal(production_raster, land_eez_rgns_hsx, fun = "sum"))

print(sum(tonnes_df$sum)) # 3901214


tonnes_df_species <- tonnes_df %>%
  mutate(aq_group = spp_class) %>%
  left_join(food_rgns, by = c("zone" = "ID_0")) %>%
  rename("total_tonnes" = "sum", rgn_id = "zone")

print(sum(tonnes_df_species$total_tonnes)) # 3901214


write.csv(tonnes_df_species, sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/production/csvs_rgn/tonnes_%s.csv", spp_class), row.names = FALSE)
}


## compile all the files together
prod_files <- list.files(file.path(prep, 'aquaculture/marine/production/csvs_rgn'), 
                                    pattern = 'tonnes_', 
                                    full.names = TRUE)

all_prod <- lapply(prod_files, FUN = function(x) {
    read.csv(x)}) %>%
    bind_rows()

sum(all_prod$total_tonnes) # 29623515

401719.1 +2753752 + 37072.56 + 17435623 + 5094133 + 3901214  # 29623514 nearly perfect 

write.csv(all_prod, file.path(prep, "aquaculture/marine/production_extracted_final.csv"), row.names = FALSE)
```



