---
title: "Tuna: Aquaculture GHG"
author: "Gage Clawson (NCEAS, UCSB, OHI)"
date: "12/21/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Script Description

This markdown calculates the GHG emissions from general tuna aquaculture. 

# Data information
We used [Seafood Carbon Emissions Tool data](http://seafoodco2.dal.ca/) to extract ghg emissions associated with carcass weight of tuna from tuna farming.

One factor that we added was aquatic N<sub>2</sub>O emissions as discussed in [Hu *et al*., 2012](https://pubs.acs.org/doi/full/10.1021/es300110x) and [MacLoid *et al*., 2019](http://www.fao.org/3/ca7130en/ca7130en.pdf).


# Methods 
 - Take midpoint of ghg low and ghg high from [seafood carbon emissions tool data](http://seafoodco2.dal.ca/) for all tuna species
 - Multiply by % energy (to cut out the feed emissions)
 - Add the N20 value from the MacLeod paper to the emissions intensity estimate we got for tuna
 - Multiply emissions intensity estimate by tonnes of production in each cell
 
 
# Preamble
```{r setup, include = FALSE}
# getting packages we want
library(tidyverse)
library(here)
library(raster)
library(sf)
library(janitor)

# Raster templates
source(here("_spatial/template_raster.R"))

# File paths
raw   <- "/home/shares/food-systems/Food_footprint/_raw_data/"
prep_tuna <- "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/tuna"
prep <- "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/" 
path_final <- "/home/shares/food-systems/Food_footprint/all_food_systems/datalayers" 

# Import master_rgns xy df
master_rgns <- read_csv(here("_spatial/_output/food_rgns.csv"))
```  
 
## Import
```{r}
## Read in farmed emissions data 

farmed_emissions <- read_csv(file.path("/home/shares/food-systems/Food_footprint/_raw_data/seafood_carbon_emmisions_tool/d2020/Specie_List_-_11_24_2020,_11_39_21_AM_Farmed.csv")) %>%
  clean_names()

tuna_production <- read_csv(file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/tuna/updated/tuna_farm_updated.csv"))

fao_tuna_spp <- read_csv(here("aquaculture/marine/tuna/data/fao_mariculture_tuna_taxa_production.csv"))
sum(fao_tuna_spp$fao_tonnes_production) # 37072.56
```
 
 
# Wrangle

```{r}
unique(sort(fao_tuna_spp$species))
# [1] "Atlantic bluefin tuna" "Pacific bluefin tuna"  "Southern bluefin tuna"

unique(sort(farmed_emissions$common_name))
## we only want tunas that are in fao
 #  Emissions spp   -   FAO spp
 ###############################################
# "Atlantic bluefin tuna" - "Atlantic bluefin tuna"
# "Pacific bluefin tuna" - Pacific bluefin tuna
# "Southern bluefin tuna" - Southern bluefin tuna

## Filter for the appropriate species... also filter for "Marine net pen"... since we don't want any inland aquaculture
tuna_emissions_spp <- farmed_emissions %>%
  dplyr::filter(common_name %in% c("Atlantic bluefin tuna", "Pacific bluefin tuna", "Southern bluefin tuna")) 


## multiply the ghg low and ghg high by the percent energy
## take the mean of new ghg low and high (the 25th and 75th quantiles)
tuna_emissions <- tuna_emissions_spp %>%
  dplyr::mutate(ghg_low_energy = ghg_low*(percent_energy/100),
                ghg_high_energy = ghg_high*(percent_energy/100)) %>%
  rowwise() %>%
  dplyr::mutate(ghg_kgCO2_kg_fish_mean = mean(c(ghg_low_energy, ghg_high_energy))) 

ghg_kgCO2_kg_fish <- mean(tuna_emissions$ghg_kgCO2_kg_fish_mean)

## now multiply by the percent of that which is energy emissions (100%)
ghg_energy_kgCO2_kg_fish <- ghg_kgCO2_kg_fish # *1 ##   0.172 kg CO2eq/ kg fish (live/carcass weight, ratioed)

# Adding Aquatic N2O
# From MacLeod et al. 
# "The amount of N2O per species-group was determined by multiplying the
# production by the N2O emission factor per kg of production (Hu et al., 2012), i.e.
# 1.69 gN2O-N per kg of production, or 0.791 kgCO2e/kgLW production. This equates
# to a conversion rate of N to N2O-N of 1.8 percent, which is higher that the 0.71
# percent used in Henriksson et al. (2014a). "

# 0.791 kgCO2e/kgLW production 
ghg_total_kgCO2_kg_fish <- 0.791 + ghg_energy_kgCO2_kg_fish ## 0.963

## convert to kgco2/tfish
GHG_kgCO2e_tfish <- ghg_total_kgCO2_kg_fish/0.001

## convert to MTco2e/tfish
GHG_MTCO2e_tfish <- GHG_kgCO2e_tfish/1000


tuna_ghg <- tuna_production %>%
  dplyr::mutate(tuna_ghg_tonnes = GHG_MTCO2e_tfish*tonnes_production) %>%
  dplyr::select(x,y, tuna_ghg_tonnes)

sum(tuna_ghg$tuna_ghg_tonnes) # 35700.88

write.csv(tuna_ghg, file.path(prep_tuna, "tuna_ghg.csv"))


## Now multiply this by our production raster
tuna_production_raster <- rasterFromXYZ(tuna_production)
plot(log(tuna_production_raster + 1))
cellStats(tuna_production_raster, "sum") # 37072.56

ghg_em <- tuna_production_raster*GHG_MTCO2e_tfish
plot(log(ghg_em + 1))
cellStats(ghg_em, "sum") # 35700.88
ghg_em

#check
37072.56*GHG_MTCO2e_tfish # 35700.88
# it worked

## Transform CRS to be appropriate
food_raster <- raster(nrows=2160, ncols=4320, xmn=-180, xmx=180, ymn=-90, ymx=90)

crs(ghg_em) <- crs(food_raster)
ghg_em

ghg_em_final <- projectRaster(ghg_em, food_raster, method = "ngb")
plot(log(ghg_em_final+1))
cellStats(ghg_em_final, "sum") # 35700.88 
ghg_em_final


## Save the raster to datalayers folder
writeRaster(ghg_em_final, file.path(path_final, "marine_tuna_aquaculture_meat_ghg.tif"), overwrite = TRUE)


```

Datacheck: 

```{r}
tuna_production <- read_csv(file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/tuna/tuna_farm.csv"))

## sum production
sum(tuna_production$tonnes_production) # 37072.56

# now multiply that production by the ghg intensity calculated above: 
 
37072.56*0.963 # 35700.88


## now check against the final ghg raster for bivalves 
tuna_ghg <- raster(file.path(path_final, "marine_tuna_aquaculture_meat_ghg.tif"))
cellStats(tuna_ghg, "sum") # 35700.88 matches perfectly


## check our emissions against that of macleod et al... 

## tuna arent specifically included in macleod et al... so we will just compare to marine fish general. And will only look at the emissions intensity

# Their emissions intensity appears to be a little less than 1 kgco2/LW for on farm energy and N20, which is the only ones we include. Our EI is ~0.963, so it it actually really close, and seems reasonable. 
```