---
title: "STEP 1: Wrangling FAO livestock count"
author: "Juliette Verstaen"
date: "7/1/2020"
output: html_document
---

# Objectives
This markdown wrangles the FAO headcount data from the Emissions-Manure Management section of FAOSTAT into an easy format for the analysis later down the line.

# Inputs
**Reference** http://www.fao.org/faostat/en/#data/GM
**Description** Head counts/stock numbers for dairy cows, nondairy cows, buffaloes, broiler chickens, layer chickens, goats, and sheep
**Resolution** Country level
**Time Range** Avaliable 1961-2018; we downloaded 2010 and 2017
**Format** CSV file

## Outputs: 'fao_livestock_headcount.csv' (iso3c, year[2017, 2010], animal, product, fao_headcount)

**

## Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(countrycode)
source(here("_workflow/common.R"))

# un gapfilling regions
un <- read_csv(here("_spatial/_output/UNSD_Methodology.csv")) %>% 
  dplyr::select(iso3c, Global_Name, Region_Name, Sub_region_Name, Intermediate_Region_Name, Region_Name)
```

## Methods

Read in and clean up the FAO livestock head count numbers
```{r}
livestock_headcount_raw <- read_csv(file.path(raw, "FAO_data/v2020/FAO_livestock_head_count_2020_09_25.csv")) 

fao_livestock_headcount <- livestock_headcount_raw %>% 
  
  filter(Area != "China") %>% 
  mutate(Area = ifelse(Area == "Eswatini", "Swaziland", Area),
         Area = ifelse(Area == "Netherlands Antilles (former)", "Bonaire, Sint Eustatius and Saba", Area),
         Area = ifelse(Area == "French Guyana", "French Guiana", Area),
         Area = ifelse(Area == "China, mainland", "China", Area),
         iso3c = countrycode(Area, origin="country.name", destination = "iso3c")) %>% 
  rename(country = Area,
         fao_headcount = Value,
         fao_class = Item, 
         year = Year) %>% 
  select(country, iso3c, year, fao_class, fao_headcount) %>% 
  
  mutate(animal = case_when(fao_class == "Cattle, dairy" ~ "cows",
                            fao_class == "Cattle, non-dairy" ~ "cows",
                            fao_class == "Chickens, broilers" ~ "chickens",
                            fao_class == "Chickens, layers" ~ "chickens",
                            fao_class == "Goats" ~ "goats",
                            fao_class == "Sheep" ~ "sheep",
                            fao_class == "Swine, breeding" ~ "pigs",
                            fao_class == "Swine, market" ~ "pigs",
                            fao_class == "Buffaloes" ~ "buffaloes"),
  
         product = case_when(fao_class == "Cattle, dairy" ~ "milk",
                            fao_class == "Cattle, non-dairy" ~ "meat",
                            fao_class == "Chickens, broilers" ~ "meat",
                            fao_class == "Chickens, layers" ~ "eggs",
                            fao_class == "Goats" ~ "meat and milk",
                            fao_class == "Sheep" ~ "meat and milk",
                            fao_class == "Swine, breeding" ~ "meat",
                            fao_class == "Swine, market" ~ "meat",
                            fao_class == "Buffaloes" ~ "meat and milk")) %>% 
  
  group_by(iso3c, year, animal, product) %>% 
  mutate(fao_headcount = sum(fao_headcount)) %>% 
  select(iso3c,year, animal, product, fao_headcount) %>% 
  unique() ## we need to do this because the pigs are currently double counted the way the sum calculation is done above

write_csv(fao_livestock_headcount, here("animal_farm/farm/data/fao_livestock_headcount.csv"))

```
