---
title: "STEP 14: Creating country summary dfs by livestock"
author: "Juliette"
date: "9/15/2020"
output: html_document
---

# Objectives
This markdown creates a summary df for each of our animals.

# Inputs
All the maps we created of our animals, systems, products

# Output 

csv files for each animal with the regional totals


## Set up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(countrycode)
source(here("_workflow/common.R"))
source(here("_workflow/common_spatial.R"))

```

## Methods
Grab and consolidate all our map counts into country df
```{r}
library(doParallel)
file_list <- list.files(file.path(prep, "animal_farm/farm"), pattern = "location_df", full = TRUE)

registerDoParallel(4)
foreach(file = file_list) %dopar%  {
  
      ## grab the names of all the parts from the file
 # file <- file_list[5]
  
    animal_name <- str_split(str_split(file, pattern = "/")[[1]][11], pattern = "_")[[1]][1]
    system_name <- str_split(str_split(file, pattern = "/")[[1]][11], pattern = "_")[[1]][2]
    product_name <- str_split(str_split(file, pattern = "/")[[1]][11], pattern = "_")[[1]][3]
    
  if(animal_name == "chickens" ) {    
      col_type = c("ddnnnccnnccncnnc")
      
      }else{
      if(animal_name %in% c("goats", "sheep")) {    
      col_type = c("ddccccn")
      
      }else{
      if(animal_name == "cows") {    
      col_type = c("ddccccn")
      
      }else{
        if(animal_name == "buffaloes") {
          col_type = c("ddccccn")
      }else{
        ## this is pigs
        (col_type = c("ddcccnnnnncc")) } } } } 
 
  df <- read_csv(file, col_types = col_type)
  
  df <- df %>% 
    select(iso3c, current_count = contains("current")) %>% 
    group_by(iso3c) %>% 
    dplyr::summarise(rgn_count = sum(current_count, na.rm = TRUE)) %>% 
    ungroup() %>% 
    mutate(animal = animal_name, system = system_name, product = product_name)
    
write_csv(df, paste0(prep, "animal_farm/farm/rgn_df/region_counts_", animal_name, "_", system_name, "_", product_name, ".csv"))  
   
}

```

combine all similar animals into one df
```{r}
chicken_counts <- rbind(read_csv(file.path(prep, "animal_farm/farm/rgn_df/region_counts_chickens_industrial_meat.csv")),
                        read_csv(file.path(prep, "animal_farm/farm/rgn_df/region_counts_chickens_industrial_eggs.csv")),
                        read_csv(file.path(prep, "animal_farm/farm/rgn_df/region_counts_chickens_backyard_eggs&meat.csv")))

cow_counts <- rbind(read_csv(file.path(prep, "animal_farm/farm/rgn_df/region_counts_cows_grassland_meat.csv")),
                        read_csv(file.path(prep, "animal_farm/farm/rgn_df/region_counts_cows_grassland_milk.csv")),
                        read_csv(file.path(prep, "animal_farm/farm/rgn_df/region_counts_cows_mixed_meat.csv")),
                        read_csv(file.path(prep, "animal_farm/farm/rgn_df/region_counts_cows_mixed_milk.csv")))

goats_counts <- rbind(read_csv(file.path(prep, "animal_farm/farm/rgn_df/region_counts_goats_grassland_meat.csv")),
                        read_csv(file.path(prep, "animal_farm/farm/rgn_df/region_counts_goats_grassland_milk.csv")),
                        read_csv(file.path(prep, "animal_farm/farm/rgn_df/region_counts_goats_mixed_meat.csv")),
                        read_csv(file.path(prep, "animal_farm/farm/rgn_df/region_counts_goats_mixed_milk.csv")))

sheep_counts <- rbind(read_csv(file.path(prep, "animal_farm/farm/rgn_df/region_counts_sheep_grassland_meat.csv")),
                        read_csv(file.path(prep, "animal_farm/farm/rgn_df/region_counts_sheep_grassland_milk.csv")),
                        read_csv(file.path(prep, "animal_farm/farm/rgn_df/region_counts_sheep_mixed_meat.csv")),
                        read_csv(file.path(prep, "animal_farm/farm/rgn_df/region_counts_sheep_mixed_milk.csv")))

pigs_counts <- rbind(read_csv(file.path(prep, "animal_farm/farm/rgn_df/region_counts_pigs_backyard_meat.csv")),
                        read_csv(file.path(prep, "animal_farm/farm/rgn_df/region_counts_pigs_intermediate_meat.csv")),
                        read_csv(file.path(prep, "animal_farm/farm/rgn_df/region_counts_pigs_industrial_meat.csv")))

buffaloes_count <- rbind(read_csv(file.path(prep, "animal_farm/farm/rgn_df/region_counts_buffaloes_grassland_milk.csv")),
                        read_csv(file.path(prep, "animal_farm/farm/rgn_df/region_counts_buffaloes_grassland_milk.csv")))


write_csv(chicken_counts, file.path(prep, "animal_farm/farm/rgn_df/chicken_counts.csv"))
write_csv(cow_counts, file.path(prep, "animal_farm/farm/rgn_df/cow_counts.csv"))
write_csv(goats_counts, file.path(prep, "animal_farm/farm/rgn_df/goats_counts.csv"))
write_csv(sheep_counts, file.path(prep, "animal_farm/farm/rgn_df/sheep_counts.csv"))
write_csv(pigs_counts, file.path(prep, "animal_farm/farm/rgn_df/pigs_counts.csv"))
write_csv(buffaloes_count, file.path(prep, "animal_farm/farm/rgn_df/buffaloes_counts.csv"))
```


We will also create tif files of chickens and pigs here, since we already have it for ruminants


Grab all the livestock head count rasters and ID where areas have <10 heads per cell

```{r}
source(here("_workflow/common_spatial.R"))
file_list <- list.files(file.path(paste(prep, "animal_farm/farm/", sep = "")), pattern = "df.csv", full.names = TRUE)
chicken_list <-  Filter(function(x) any(grepl("chickens", x)), file_list)  
pigs_list <- Filter(function(x) !any(grepl("pigs", x)), file_list)  


#backyard
df_chick_back <- read_csv(file.path(prep,"animal_farm/farm/chickens_backyard_eggs&meat_location_df.csv"), col_type = c("ddnnnccnnccncnnc")) %>% 
  select(x,y, current_count) %>% 
  rasterFromXYZ(crs = food_crs)
writeRaster(df_chick_back, file.path(paste(prep, "animal_farm/farm/location_tifs/", "chickens_backyard_eggs&meat_location_df", ".tif", sep = "")), format = "GTiff", overwrite = TRUE)   

## ind eggs  
df_chick_ind_eggs<- read_csv(file.path(prep,"animal_farm/farm/chickens_industrial_eggs_location_df.csv"), col_type = c("ddnnnccnnccncnnc"))%>% 
  select(x,y, current_count) %>% 
  rasterFromXYZ(crs = food_crs)
writeRaster(df_chick_ind_eggs, file.path(paste(prep, "animal_farm/farm/location_tifs/", "chickens_industrial_eggs_location_df", ".tif", sep = "")), format = "GTiff", overwrite = TRUE)   

## ind meat
df_chick_ind_meat <- read_csv(file.path(prep,"animal_farm/farm/chickens_industrial_meat_location_df.csv"), col_type = c("ddnnnccnnccncnnc"))%>% 
  select(x,y, current_count) %>% 
  rasterFromXYZ(crs = food_crs)
writeRaster(df_chick_ind_eggs, file.path(paste(prep, "animal_farm/farm/location_tifs/", "chickens_industrial_meat_location_df", ".tif", sep = "")), format = "GTiff", overwrite = TRUE)   

#pigs back
pigs_back <- read_csv(file.path(prep,"animal_farm/farm/pigs_backyard_meat_location_df.csv"), col_type = c("ddcccnnnnncc")) %>% 
  select(x,y, current_count) %>% 
  rasterFromXYZ(crs = food_crs)
writeRaster(df_chick_back, file.path(paste(prep, "animal_farm/farm/location_tifs/", "pigs_backyard_meat_location_df", ".tif", sep = "")), format = "GTiff", overwrite = TRUE)   
#pigs inter
pigs_back <- read_csv(file.path(prep,"animal_farm/farm/pigs_intermediate_meat_location_df.csv"), col_type = c("ddcccnnnnncc")) %>% 
  select(x,y, current_count) %>% 
  rasterFromXYZ(crs = food_crs)
writeRaster(df_chick_back, file.path(paste(prep, "animal_farm/farm/location_tifs/", "pigs_intermediate_meat_location_df", ".tif", sep = "")), format = "GTiff", overwrite = TRUE)   
#pigs indu
pigs_back <- read_csv(file.path(prep,"animal_farm/farm/pigs_industrial_meat_location_df.csv"), col_type = c("ddcccnnnnncc")) %>% 
  select(x,y, current_count) %>% 
  rasterFromXYZ(crs = food_crs)
writeRaster(df_chick_back, file.path(paste(prep, "animal_farm/farm/location_tifs/", "pigs_industrial_meat_location_df", ".tif", sep = "")), format = "GTiff", overwrite = TRUE)   


```


