---
title: "STEP 9: Mapping cows, buffaloes, goats, and sheep locations"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Objectives
Create 4 maps each for all our ruminants by grassland/mixed and dairy/nondairy.

# Inputs:
animal_all_counts.tif files calculated in STEP 7
ruminant_dairy.tif and ruminant_nondairy.tif files calculated in STEP 6
ruminant_system_prod_system.tif files calculated in STEP 5

# Outputs: ruminant_system_product.csv and tif files for every combination of cows/buffaloes/goats/sheep, grassland/mixed, and milk/meat. located in the prep folder on the server

## Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(countrycode)
library(readxl)

source(here("_workflow/common.R"))
source(here("_workflow/common_spatial.R"))
source(here("_spatial/template_raster.R"))

## load libraries, read in data
food_rgns_area <- area(food_rgns_tif)

```

## Map them!

```{r}

ruminant_list <- c("sheep", "goats", "buffaloes")


for(ruminant in ruminant_list) {
  
   # ruminant <- ruminant_list[1]
  
  map_all <- raster(file.path(paste0(prep, "animal_farm/farm/updated_gridded_map_tifs/", ruminant, "_all_counts.tif", sep = "")))
  
  dairy <- raster(file.path(paste0(prep, "animal_farm/farm/production_system1_tifs/", ruminant, "_dairy.tif", sep = "")))
  nondairy <- raster(file.path(paste0(prep, "animal_farm/farm/production_system1_tifs/", ruminant, "_nondairy.tif", sep = "")))
  
  grass <- raster(file.path(prep, "animal_farm/farm/production_system2_tifs/grassland_prod_system.tif"))
  mixed <- raster(file.path(prep, "animal_farm/farm/production_system2_tifs/mixed_prod_system.tif"))
  
  
  dairy_grass <- map_all*dairy*grass
    names(dairy_grass) <- "current_count"
  dairy_mixed <- map_all*dairy*mixed
    names(dairy_mixed) <- "current_count"
  
  nondairy_grass <- map_all*nondairy*grass
    names(nondairy_grass) <- "current_count"
  nondairy_mixed <- map_all*nondairy*mixed
    names(nondairy_mixed) <- "current_count"
    
    ## save them as rasters
    writeRaster(dairy_grass, file.path(paste0(prep, "animal_farm/farm/location_tifs/", ruminant, "_grassland_milk.tif", sep = "")), format = "GTiff", overwrite = TRUE )
    writeRaster(dairy_mixed, file.path(paste0(prep, "animal_farm/farm/location_tifs/", ruminant, "_mixed_milk.tif", sep = "")), format = "GTiff", overwrite = TRUE )
    writeRaster(nondairy_grass, file.path(paste0(prep, "animal_farm/farm/location_tifs/", ruminant, "_grassland_meat.tif", sep = "")), format = "GTiff", overwrite = TRUE )
    writeRaster(nondairy_mixed, file.path(paste0(prep, "animal_farm/farm/location_tifs/", ruminant, "_mixed_meat.tif", sep = "")), format = "GTiff", overwrite = TRUE )

  
## convert to csvs so it remains consistent with the chickens and pigs. don't make the map for meat buffalo 

    ## grassland milk
    dairy_grass_df <- dairy_grass %>% 
        raster_df() %>% 
      left_join(food_rgns_xy, by = c("x", "y", "cellindex")) %>% 
      mutate(system = "grassland",
             product = "milk") %>% 
      select(x,y, iso3c, country = Country, system, product, current_count)
    
    write_csv(dairy_grass_df, file.path(paste0(prep, "animal_farm/farm/", ruminant, "_grassland_milk_location_df.csv", sep = "")))
    
    ## grassland meat   
    nondairy_grass_df <- nondairy_grass %>% 
        raster_df() %>% 
      left_join(food_rgns_xy, by = c("x", "y", "cellindex")) %>% 
      mutate(system = "grassland",
             product = "meat") %>% 
      select(x,y, iso3c, country = Country, system, product, current_count)
    
    write_csv(nondairy_grass_df, file.path(paste0(prep, "animal_farm/farm/", ruminant, "_grassland_meat_location_df.csv", sep = "")))
    
    
        ## mixed milk
    dairy_mixed_df <- dairy_mixed %>% 
        raster_df() %>% 
      left_join(food_rgns_xy, by = c("x", "y", "cellindex")) %>% 
      mutate(system = "mixed",
             product = "milk") %>% 
      select(x,y, iso3c, country = Country, system, product, current_count)
    
    write_csv(dairy_mixed_df, file.path(paste0(prep, "animal_farm/farm/", ruminant, "_mixed_milk_location_df.csv", sep = "")))
    
    ## mixed meat   
    nondairy_mixed_df <- nondairy_mixed %>% 
        raster_df() %>% 
      left_join(food_rgns_xy, by = c("x", "y", "cellindex")) %>% 
      mutate(system = "mixed",
             product = "meat") %>% 
      select(x,y, iso3c, country = Country, system, product, current_count)
    
    write_csv(nondairy_mixed_df, file.path(paste0(prep, "animal_farm/farm/", ruminant, "_mixed_meat_location_df.csv", sep = "")))
    
}

```


Cows are a little different
```{r}

product_list <- c("milk", "meat")


for(product_OI in product_list) {
  
   # product_OI <- product_list[2]
  
  map_all <- raster(file.path(paste0(prep, "animal_farm/farm/updated_gridded_map_tifs/cows_", product_OI, "_all_counts.tif", sep = "")))
  
  grass <- raster(file.path(prep, "animal_farm/farm/production_system2_tifs/grassland_prod_system.tif"))
  mixed <- raster(file.path(prep, "animal_farm/farm/production_system2_tifs/mixed_prod_system.tif"))
  
  grass_map <- map_all*grass
    names(grass_map) <- "current_count"
  mixed_map <- map_all*mixed
    names(mixed_map) <- "current_count"
  

    ## save them as rasters
    writeRaster(grass_map, file.path(paste0(prep, "animal_farm/farm/location_tifs/cows_grassland_", product_OI, ".tif", sep = "")), format = "GTiff", overwrite = TRUE )
    writeRaster(mixed_map, file.path(paste0(prep, "animal_farm/farm/location_tifs/cows_mixed_", product_OI, ".tif", sep = "")), format = "GTiff", overwrite = TRUE )

  
## convert to csvs so it remains consistent with the chickens and pigs. don't make the map for meat buffalo 

    ## grassland
    grass_df <- grass_map %>% 
        raster_df() %>% 
      left_join(food_rgns_xy, by = c("x", "y", "cellindex")) %>% 
      mutate(animal = "cows", 
             system = "grassland",
             product = product_OI) %>% 
      select(x,y, iso3c, country = Country, system, product, current_count)
    
    write_csv(grass_df, file.path(paste0(prep, "animal_farm/farm/cows_grassland_", product_OI, "_location_df.csv", sep = "")))
    
    ## mixed   
    mixed_map_df <- mixed_map %>% 
        raster_df() %>% 
      left_join(food_rgns_xy, by = c("x", "y", "cellindex")) %>% 
      mutate(animal = "cows", 
             system = "mixed",
             product = product_OI) %>% 
      select(x,y, iso3c, country = Country, system, product, current_count)
    
    write_csv(mixed_map_df, file.path(paste0(prep, "animal_farm/farm/cows_mixed_", product_OI, "_location_df.csv", sep = "")))
  
}

```


The mixed cow maps do not include feedlot cows. We want to have a separate map representing feedlot cow numbers, but in the locations where it is identified as mixed. This will replace the mixed meat cow raster created above after adjusting for feedlot counts.

```{r}
mixed_meat_cattle <- read_csv("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/animal_farm/farm/cows_mixed_meat_location_df.csv", col_types = "ddccccd")
total_count_mixed <- mixed_meat_cattle %>%
  group_by(iso3c, country, system, product) %>%
  summarize(current_count = sum(current_count, na.rm=TRUE))

feedlot_counts <- read_csv(here("animal_farm/farm/data/mixed_to_feedlot_factor.csv"))  

feedlot_factor <- left_join(feedlot_counts, total_count_mixed) %>%
  mutate(prop_feedlot = feedlot_count/current_count) %>%
  select(iso3c, prop_feedlot) 

feedlot_factor_raster <- feedlot_factor %>%
  right_join(food_rgns) %>% 
  mutate(prop_feedlot = ifelse(is.na(prop_feedlot), 0, prop_feedlot)) %>% 
  right_join(food_rgns_xy) %>% 
  select(x,y,prop_feedlot) %>% 
  rasterFromXYZ(crs = food_crs)

plot(feedlot_factor_raster)

cows_mixed_meat <- raster(file.path(prep, "animal_farm/farm/location_tifs/cows_mixed_meat.tif"))

cows_feedlots <-  cows_mixed_meat*feedlot_factor_raster
plot(cows_feedlots)
writeRaster(cows_feedlots, file.path(paste0(prep, "animal_farm/farm/location_tifs/cows_feedlot_meat.tif")), format = "GTiff", overwrite = TRUE) 

## save the location_df.csv

## feedlot 
df_feedlot_rast <- raster(file.path(prep, "animal_farm/farm/location_tifs/cows_feedlot_meat.tif"))

df_feedlot <- df_feedlot_rast %>% 
  raster_df() %>%
  left_join(food_rgns_xy, by = c("x", "y", "cellindex")) %>% 
  mutate(animal = "cows", 
         system = "feedlot",
         product = "meat") %>% 
  select(x,y, iso3c, country = Country, system, product, current_count = cows_feedlot_meat)
    
write_csv(df_feedlot, file.path(prep, "animal_farm/farm/cows_feedlot_meat_location_df.csv"))
    
#####
## now we have to remove the feedlot cows from the mixed cows
cows_mixed_meat <- raster(file.path(prep, "animal_farm/farm/location_tifs/cows_mixed_meat.tif"))

cows_mixed_meat <-  cows_mixed_meat*(1 - feedlot_factor_raster)
plot(cows_mixed_meat)
writeRaster(cows_mixed_meat, file.path(paste0(prep, "animal_farm/farm/location_tifs/cows_mixed_meat.tif")), format = "GTiff", overwrite = TRUE) 

## save the location_df.csv
df_mixed_rast <- raster(file.path(prep, "animal_farm/farm/location_tifs/cows_mixed_meat.tif"))

df_mixed <- df_mixed_rast %>% 
  raster_df() %>%
  left_join(food_rgns_xy, by = c("x", "y", "cellindex")) %>% 
  mutate(animal = "cows", 
         system = "mixed",
         product = "meat") %>% 
  select(x,y, iso3c, country = Country, system, product, current_count = cows_mixed_meat)
    
write_csv(df_mixed, file.path(prep, "animal_farm/farm/cows_mixed_meat_location_df.csv"))

```





Data checks and plots

Cows 
```{r}
cw_gr_mk <- raster(file.path(prep, "animal_farm/farm/location_tifs/cows_grassland_milk.tif"))
cw_gr_mt <- raster(file.path(prep, "animal_farm/farm/location_tifs/cows_grassland_meat.tif"))
cw_mx_mk <- raster(file.path(prep, "animal_farm/farm/location_tifs/cows_mixed_milk.tif"))
cw_mx_mt <- raster(file.path(prep, "animal_farm/farm/location_tifs/cows_mixed_meat.tif"))
```

Plot
```{r}

plot(log(cw_gr_mk+1),
     main= "Log Cows Grassland Milk")

plot(log(cw_gr_mt+1),
     main= "Log Cows Grassland Meat")

plot(log(cw_mx_mk+1),
     main= "Log Cows Mixed Milk")

plot(log(cw_mx_mt+1),
     main= "Log Cows Mixed Meat")

```


Data check

```{r}
## milk cows
fao_dairy <- read_csv(here("animal_farm/farm/data/fao_livestock_headcount.csv")) %>% 
  filter(animal == "cows" &
         product == "milk",
         year == 2017) %>% 
  dplyr::summarise(total_fao = sum(fao_headcount))
# 275,405,738			

our_dairy <- stack(cw_gr_mk,cw_mx_mk)
our_dairy <- sum(our_dairy)
raster::cellStats(our_dairy, stat = 'sum')
# 273,165,632

## meat cows
fao_meat <- read_csv(here("animal_farm/farm/data/fao_livestock_headcount.csv")) %>% 
  filter(animal == "cows" &
         product == "meat",
         year == 2017) %>% 
  dplyr::summarise(total_fao = sum(fao_headcount, na.rm = TRUE))
# 1,197,881,390			

our_meat <- stack(cw_gr_mt,cw_mx_mt)
our_meat <- sum(our_meat)
raster::cellStats(our_meat, stat = 'sum')

# 1,186,951,806

summary(our_meat)

```

Buffaloes
** note: while we have meat rasters we are not including it in our analysis. I am just including it here to be able to check that the numbers add up and nothing wonky happened.
```{r}
bf_gr_mk <- raster(file.path(prep, "animal_farm/farm/location_tifs/buffaloes_grassland_milk.tif"))
bf_gr_mt <- raster(file.path(prep, "animal_farm/farm/location_tifs/buffaloes_grassland_meat.tif"))
bf_mx_mk <- raster(file.path(prep, "animal_farm/farm/location_tifs/buffaloes_mixed_milk.tif"))
bf_mx_mt <- raster(file.path(prep, "animal_farm/farm/location_tifs/buffaloes_mixed_meat.tif"))
```

Plot
```{r}

plot(log(bf_gr_mk+1),
     main= "Log Buffaloes Grassland Milk")

plot(log(bf_mx_mk+1),
     main= "Log Buffaloes Mixed Milk")


```


Data check

```{r}
## buff
fao_buffalo <- read_csv(here("animal_farm/farm/data/fao_livestock_headcount.csv")) %>% 
  filter(animal == "buffaloes" ) %>% 
  dplyr::summarise(total_fao = sum(fao_headcount))
# 398,298,997	

our_buff <- bf_gr_mk + bf_mx_mk 
sum <- raster::cellStats(our_buff, 'sum')
# 175,767,486

```



Sheep
```{r}
sh_gr_mk <- raster(file.path(prep, "animal_farm/farm/location_tifs/sheep_grassland_milk.tif"))
sh_gr_mt <- raster(file.path(prep, "animal_farm/farm/location_tifs/sheep_grassland_meat.tif"))
sh_mx_mk <- raster(file.path(prep, "animal_farm/farm/location_tifs/sheep_mixed_milk.tif"))
sh_mx_mt <- raster(file.path(prep, "animal_farm/farm/location_tifs/sheep_mixed_meat.tif"))
```

Plot
```{r}

plot(log(sh_gr_mk+1),
     main= "Log Sheep Grassland Milk")

plot(log(sh_gr_mt+1),
     main= "Log Sheep Grassland Meat")

plot(log(sh_mx_mk+1),
     main= "Log Sheep Mixed Milk")

plot(log(sh_mx_mt+1),
     main= "Log Sheep Mixed Meat")

```


Data check

```{r}
## sheep
fao_sheep <- read_csv(here("animal_farm/farm/data/fao_livestock_headcount.csv")) %>% 
  filter(animal == "sheep",
         year == 2017) %>% 
  dplyr::summarise(total_fao = sum(fao_headcount))
# 1,201,076,128		

our_sheep <- sh_gr_mk + sh_mx_mk + sh_gr_mt + sh_mx_mt
sum <- raster::cellStats(our_sheep, 'sum')
# 1,201,639,086

```


Goats
```{r}
gt_gr_mk <- raster(file.path(prep, "animal_farm/farm/location_tifs/goats_grassland_milk.tif"))
gt_gr_mt <- raster(file.path(prep, "animal_farm/farm/location_tifs/goats_grassland_meat.tif"))
gt_mx_mk <- raster(file.path(prep, "animal_farm/farm/location_tifs/goats_mixed_milk.tif"))
gt_mx_mt <- raster(file.path(prep, "animal_farm/farm/location_tifs/goats_mixed_meat.tif"))
```

Plot
```{r}

plot(log(sh_gr_mk+1),
     main= "Log Goats Grassland Milk")

plot(log(sh_gr_mt+1),
     main= "Log Goats Grassland Meat")

plot(log(sh_mx_mk+1),
     main= "Log Goats Mixed Milk")

plot(log(sh_mx_mt+1),
     main= "Log Goats Mixed Meat")

```


Data check

```{r}
## buff
fao_goats <- read_csv(here("animal_farm/farm/data/fao_livestock_headcount.csv")) %>% 
  filter(animal == "goats" ) %>% 
  dplyr::summarise(total_fao = sum(fao_headcount))
# 1,021,738,193

our_goats <- gt_gr_mk + gt_mx_mk + gt_gr_mt + gt_mx_mt
sum <- raster::cellStats(our_goats, 'sum')
# 1,022,516,691

```
