---
title: "STEP 8: Chicken and Pig Farm Location"
output: html_document
---

## Objective
Map the distribution of total current living chickens and pigs globally.

## Inputs
We use 
1. [FAO gridded livestock maps](http://www.fao.org/livestock-systems/global-distributions/en/) to establish locations of livestock
2. [GLEAM](http://www.fao.org/gleam/results/en/) data on total herd numbers, production system type and product for proportion of animals going to each category
3. [FAO head count data](http://www.fao.org/faostat/en/#data/GM) for total living livestock values

The FAO head count data is wrangled in fao_livestock_count.Rmd
The other two data sets are used in the animal specific scripts, and wrangled accordingly

## Outputs
6 maps of each livestock system of interest

## Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(countrycode)
library(readxl)

source(here("_workflow/common.R"))
source(here("_workflow/common_spatial.R"))

## load libraries, read in data
food_rgns_area <- area(food_rgns_tif)

```

## Map all livestock distributions
```{r}
## chickens (cluckers)
source(here("animal_farm/farm/scripts/mapping_chickens.R"))
chicken_system_list <- c("industrial", "backyard")
chicken_product_list <- c("meat", "eggs", "eggs&meat")
mapping_chickens("chickens", "backyard")
mapping_chickens("chickens", "industrial")

## pigs (oinkies)
source(here("animal_farm/farm/scripts/mapping_pigs.R"))
pigs_system_list <- c("backyard", "intermediate", "industrial")
mapping_pigs("pigs")

```

## Error checking and initial maps

I'm thinking maybe we want this in another markdown for cleaner? maybe not, might be nice to have it all in one when we rerun and check

### Checking out Chickens
Looking at the maps (make this a loop in the end and save all plots as pngs)
```{r}
chick_ind_egg <- read_csv(file.path(prep, "animal_farm/farm/chickens_industrial_eggs_location_df.csv")) %>% 
  select(x,y, current_count) %>% 
  rasterFromXYZ(crs = food_crs)
writeRaster(chick_ind_egg, file.path(prep, "animal_farm/farm/check/chickens_industrial_eggs_raster.tif"), overwrite=TRUE)

chick_ind_meat <- read_csv(file.path(prep, "animal_farm/farm/chickens_industrial_meat_location_df.csv")) %>% 
  select(x,y, current_count) %>% 
  rasterFromXYZ(crs = food_crs)
writeRaster(chick_ind_meat, file.path(prep, "animal_farm/farm/check/chickens_industrial_meat_raster.tif"), overwrite=TRUE)

chick_back <- read_csv(file.path(prep, "animal_farm/farm/chickens_backyard_eggs&meat_location_df.csv"), col_type = "nnnnnccnn") %>% 
  select(x,y, current_count) %>% 
  rasterFromXYZ(crs = food_crs)
writeRaster(chick_back, file.path(prep, "animal_farm/farm/check/chickens_backyard_eggs$meat_raster.tif"), overwrite=TRUE)


```


Plot
```{r}

plot(log(chick_ind_egg+1),
     main= "Log Chicken Industrial Egg")

plot(chick_ind_egg,
     main= "Chicken Industrial Egg")

plot(log(chick_ind_meat+1),
     main= "Log Chicken Industrial Meat")

##backyard
plot(chick_back,
     main= "Chicken Backyard Eggs & Meat")

plot(log(1 + chick_back),
     main= "Log Chicken BackyardEggs & Meat")

##I think the logged versions look so weird because there jsut aren't many places with chcikens, so when we log+1 all the zero areas show up prominent and bright green

```


Data check

```{r}
## broilers
fao_broiler <- read_csv(here("animal_farm/farm/data/fao_livestock_headcount.csv")) %>% 
  filter(animal == "chickens" &
         product == "meat",
         year == 2017) %>% 
  dplyr::summarise(total_fao = sum(fao_headcount))
# 15,724,676,722		

our_broiler <- read_csv(file.path(prep, "animal_farm/farm/chickens_industrial_meat_location_df.csv")) %>% 
  dplyr::summarise(total_broilers_us = sum(current_count, na.rm = TRUE))
# 15,803,802,499		

## layers
fao_layer <- read_csv(here("animal_farm/farm/data/fao_livestock_headcount.csv")) %>% 
  filter(animal == "chickens", 
         product == "eggs",
         year == 2017) %>% 
  dplyr::summarise(total_fao = sum(fao_headcount))
# 7,511,742,112	

our_layer <- read_csv(file.path(prep, "animal_farm/farm/chickens_industrial_eggs_location_df.csv")) %>% 
  dplyr::summarise(total_broilers_us = sum(current_count, na.rm = TRUE))  

# 7,511,232,930	

##Backyard; we are comparing to 2010 gleam value so it will be off already because it is not adjusted for 2016 numbers
gleam_raw <- read_csv(here(paste0("animal_farm/farm/data/", animal, "_GLEAMi_v2.csv")))
        
gleam_backyard <- gleam_raw %>% 
      filter(Production_system =="Backyard",
      Variable == "HERD: total number of animals") %>% 
      dplyr::summarise(total = sum(Value))
# 3,725,587,649	

our_backyard <- read_csv(file.path(prep, "animal_farm/farm/chickens_backyard_eggs&meat_location_df.csv"), col_type = "nnnnnccnn") %>% 
  dplyr::summarise(total_broilers_us = sum(current_count, na.rm = TRUE))  
# 4,095,779,664		


```

### Checking out pigs

Compare to FAO counts
Pigs 
```{r}
fao_data <- read_csv(here("animal_farm/farm/data/fao_livestock_headcount.csv"))

pigs_fao <- fao_data %>% 
  filter(animal == "pigs",
         year == 2017) %>% 
  dplyr::summarise(total_pigs = sum(fao_headcount, na.rm = TRUE))
# 978,466,146		

## using the fao_count which includes both milk and meat so only need to read in one of them
pigs1 <- read_csv(file.path(prep,  "animal_farm/farm/pigs_backyard_meat_location_df.csv"), col_type = "ddcccnnnnncc") %>% 
  select(x,y, current_count) %>% 
  rasterFromXYZ()

pigs2 <- read_csv(file.path(prep,  "animal_farm/farm/pigs_intermediate_meat_location_df.csv"), col_type = "ddcccnnnnncc") %>% 
  select(x,y, current_count) %>% 
  rasterFromXYZ()

pigs3 <- read_csv(file.path(prep,  "animal_farm/farm/pigs_industrial_meat_location_df.csv"), col_type = "ddcccnnnnncc") %>% 
  select(x,y, current_count) %>% 
  rasterFromXYZ()


pigs_ours <-  stack(pigs1, pigs2, pigs3) %>% 
  sum() %>% 
  cellStats(stat= 'sum')

 ## 978,734,721

```


```{r}
pig_back <- read_csv(file.path(prep, "animal_farm/farm/pigs_backyard_meat_location_df.csv"), col_types = "ddcccnnnnncc") %>% 
  select(x,y, current_count) %>% 
  rasterFromXYZ(crs = food_crs)

pig_inte <- read_csv(file.path(prep, "animal_farm/farm/pigs_intermediate_meat_location_df.csv"), col_types = "ddcccnnnnncc") %>% 
  select(x,y, current_count) %>% 
  rasterFromXYZ(crs = food_crs)

pig_indus <- read_csv(file.path(prep, "animal_farm/farm/pigs_industrial_meat_location_df.csv"), col_types = "ddcccnnnnncc") %>% 
  select(x,y, current_industrial_headcount) %>% 
  rasterFromXYZ(crs = food_crs)


plot(log(pig_back+1),
     main= "Log Pigs Backyard Meat")

plot(log(pig_inte+1),
     main= "Log Pigs Intermediate Meat")

plot(log(pig_indus+1),
     main= "Log Pigs Industrial Meat")


```

## Data Check
```{r}

fao_map_backyard_total <-  raster(file.path(raw, "FAO_livestock_maps/d2019/pigs/backyard_pigs/8_PgExt_2010_Da.tif")) %>% 
  cellStats(stat= 'sum')
#320,154,129

our_total <- cellStats(pig_back, stat= 'sum')
# 333,271,667

fao_map_intermediate_total <- raster(file.path(raw, "FAO_livestock_maps/d2019/pigs/semi_intensive_pigs/9_PgInt_2010_Da.tif")) %>% 
  cellStats(stat= 'sum')
# 106,285,361

our_total <- cellStats(pig_inte, stat= 'sum')
# 110,640,145
    
fao_map_industrial_total <- raster(file.path(raw, "FAO_livestock_maps/d2019/pigs/intensive_pigs/10_PgInd_2010_Da.tif")) %>% 
  cellStats(stat= 'sum')
#513,514,332

our_total <- cellStats(pig_indus, stat= 'sum')
# 534,554,334

```

