---
title: "STEP 4: Calculate Gleam Production Rates for milk/meat in dairy/non-dairy systems"
author: "Juliette"
date: "9/15/2020"
output: html_document
---

# Objectives
This markdown This markdown creates a data file with production/head of animal using the GLEAM data. We are not differentiating between grassland and mixed systems.

# Input
**Reference** GLEAM
**Description** Herd and emissions data
**Resolution** Country level
**Time Range** 2010
**Format** CSV file (after we scraped and wrangled)

## Output: ruminants_production_rates.csv

**

## Setup
Here we are exploring the production rate/head for each country for each of our categories. We want to determine if the production rate differs between system (grassland, mixed, etc).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(stringr)
source(here("_workflow/common.R"))

'%notin%' <- Negate('%in%')

ruminants_raw <- read_csv(here("animal_farm/farm/data/ruminants_GLEAMi_v2.csv"))

un <- read_csv(here("_spatial/_output/UNSD_Methodology.csv")) %>% 
  dplyr::select(iso3c, Global_Name, Region_Name, Sub_region_Name, Intermediate_Region_Name, Region_Name)

```

## Methods

Ruminants (cows, buffalo, sheep, goats)
**Note** we are lumping Feedlot cows in with mixed so we need to add them together

```{r}
ruminants <- ruminants_raw %>% 
  select(-Unit) %>% 
  ## combining the feedlot head counts and production in with the mixed
  mutate(Production_system = ifelse(Production_system == "Feedlot operations", "Mixed systems", Production_system),
         Herd_type = ifelse(Herd_type == "Feedlot", "Non-dairy", Herd_type)) %>% 
  group_by(Country, iso3c, Species, Production_system, Herd_type, Variable) %>% 
  dplyr::summarise(Value = sum(Value)) %>%
  ungroup() %>% 
  filter(Herd_type != "Whole herd",
         Production_system == "All systems") %>% 
  filter(Variable %in% c("HERD: total number of animals", "PROD: Milk - fresh weight", "PROD: Meat - carcass weight")) %>%
  pivot_wider(names_from = Variable, values_from = Value) %>% 
  select(animal = Species, herd_type = Herd_type, country = Country, iso3c, system = Production_system, count = 'HERD: total number of animals', kg_milk_year = 'PROD: Milk - fresh weight', kg_cw_year = 'PROD: Meat - carcass weight') %>% 
  mutate(tonnes_milk_head = (kg_milk_year/1000)/count,
         tonnes_meat_head = (kg_cw_year/1000)/count,
         tonnes_milk_head = ifelse(is.na(tonnes_milk_head), 0, tonnes_milk_head),
         tonnes_meat_head = ifelse(is.na(tonnes_meat_head), 0, tonnes_meat_head)) %>% 
  mutate(animal = tolower(animal),
         animal = ifelse(animal == "cattle", "cows", 
                         ifelse(animal == "buffalo", "buffaloes", animal)),
         herd_type = tolower(herd_type)) %>% 
  select(country, iso3c, animal, system_1 = herd_type, system_2 = system, tonnes_milk_head, tonnes_meat_head)%>% 
  pivot_longer(cols = 6:7, names_to = "product", values_to = "tonnes_per_head") %>% 
  mutate(product = ifelse(product == "tonnes_meat_head", "meat", "milk"))
  
```

Ruminants

We need to add in and gapfill missing countries using the UN Regions

```{r}

rum_miss_list <- setdiff(food_rgns$iso3c, ruminants$iso3c)

rum_miss <- as_tibble(rum_miss_list) %>% 
  rename(iso3c = value) %>% 
  slice(rep(1:n(), each =4)) %>% 
  mutate(animal = rep(c("cows", "buffaloes", "goats", "sheep"), times = 11)) %>% 
  slice(rep(1:n(), each =2)) %>% 
  mutate(system_1 = rep(c("dairy", "non-dairy"), times = 44)) %>% 
  slice(rep(1:n(), each =2)) %>% 
  mutate(product = rep(c("milk", "meat"), times = 88),
         system_2 = "All systems") %>% 
  left_join(food_rgns, by = "iso3c") %>% 
  select(iso3c, country = Country, animal, system_1, system_2, product) %>% 
  mutate(tonnes_per_head = NA)


ruminants_gf_missing <- rbind(ruminants, rum_miss) %>%
  group_by(system_1, system_2, product) %>%
  mutate(gapfilled = ifelse(is.na(tonnes_per_head), "rate gf with global mean", NA), 
  tonnes_per_head = ifelse(is.na(tonnes_per_head), mean(tonnes_per_head, na.rm = TRUE), tonnes_per_head)) %>%
  ungroup() %>%
  select(iso3c, country , animal, system_1, system_2, product, tonnes_per_head, gapfilled)

## gapfill the countries that have 0 animals according to GLEAM
ruminants_gf <- ruminants_gf_missing %>% 
  group_by(iso3c, animal) %>% 
  mutate(gf_int = sum(tonnes_per_head)) %>% 
  ungroup() %>% 
  group_by(animal, system_1, product) %>% 
  mutate(gapfilled = ifelse(gf_int == 0, "rate gf with global mean", gapfilled),
         tonnes_per_head =  ifelse(gf_int == 0, mean(tonnes_per_head, na.rm = TRUE), tonnes_per_head)) %>% 
  ungroup() %>% 
  group_by(iso3c, animal, product) %>% 
  mutate(gf_int_prod = sum(tonnes_per_head)) %>% 
  ungroup() %>% 
  group_by(animal, system_1, product) %>% 
  mutate(gapfilled = ifelse(gf_int == 0, "rate gf with global mean", gapfilled),
         tonnes_per_head =  ifelse(gf_int == 0, mean(tonnes_per_head, na.rm = TRUE), tonnes_per_head)) %>% 
  select(-gf_int, -gf_int_prod)
  

```

Save
```{r}
write_csv(ruminants_gf, here("animal_farm/farm/data/ruminants_production_rates.csv"))
```

check
```{r}
gf <- ruminants_gf %>% 
  group_by(iso3c, animal, product) %>% 
  mutate(check = sum(tonnes_per_head))
```



