---
title: "crustacean_nutrient_pollution"
author: "Gage Clawson (NCEAS, UCSB, OHI)"
date: "12/22/2020"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    includes: 
toc: true
---

This markdown uses models and species specific parameters from the paper *Aquaculture Production is a Large, Spatially Concentrated Source of Nutrients in Chinese Freshwater and Coastal Seas* (2020 Wang et al). 

## Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(sf)
library(raster)
library(readr)
library(data.table)

source(here("_spatial/template_raster.R"))

path_prep <- "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/crustaceans"
path_final <-"/home/shares/food-systems/Food_footprint/all_food_systems/datalayers"

production <- read_csv(file.path(path_prep, "updated/crustaceans_farm_updated.csv"))
food_raster <- raster(nrows=2160, ncols=4320, xmn=-180, xmx=180, ymn=-90, ymx=90)
```

crustacean N/P pollution values

N: 57 kg/tonne of growth
P: 9.2 kg/tonne of growth

Steps: 
1. Multiply tonnes production by 0.95 (this corrects for the growth that occurs outside the seapen on land), this is the value Lex uses
2. Then multiply growth by the N value and the P value
3. We report in total excess nutrients (N + P): Sum the N and P kg.
4. We report in tonnes, so divide by 1000. We will also so individual N and P rasters for reporting purposes.

```{r}

crustacean_pollution <- production %>%
  dplyr::mutate(prod_growth = tonnes_production*0.95) %>%
  dplyr::mutate(prod_growth_N_poll = prod_growth*57,
                prod_growth_P_poll = prod_growth*9.2) %>% ## in kg pollution now
  # dplyr::mutate(PO4eq_kg = prod_growth_N_poll*0.42 + prod_growth_P_poll*3.07) %>% Don't do this anymore (01/14/2021)
  dplyr::mutate(total_N_P_kg = prod_growth_N_poll + prod_growth_P_poll) %>%
  dplyr::mutate(total_N_P_t = total_N_P_kg/1000,
                N_poll_t = prod_growth_N_poll/1000, 
                P_poll_t = prod_growth_P_poll/1000) %>%
  dplyr::select(x, y, total_N_P_t, N_poll_t, P_poll_t)

sum(crustacean_pollution$total_N_P_t) # 25317.85
sum(production$tonnes_production) # 402573.6

## save to dataprep folder with the correct naming conventions: 
write.csv(crustacean_pollution, file.path(path_prep, "crustacean_nutrient.csv"))


## Save total N P raster
total_df <- crustacean_pollution %>% 
  dplyr::select(x, y, total_N_P_t)

final_rast <- rasterFromXYZ(total_df, crs = food_crs)
crs(final_rast)
plot(final_rast)
final_rast
cellStats(final_rast, "sum") # 25317.85

final_area_fix <- final_rast/raster::area(final_rast)

final_rast_proj <- projectRaster(final_area_fix, food_raster, method = "ngb")

final_rast <- final_rast_proj*raster::area(final_rast_proj)
cellStats(final_rast, "sum") # 25317.85
final_rast

## Save to datalayers 
writeRaster(final_rast, file.path(path_final, "marine_crustaceans_aquaculture_meat_nutrient.tif"), overwrite = T)



## Save total N raster
total_df <- crustacean_pollution %>% 
  dplyr::select(x, y, N_poll_t)

final_rast <- rasterFromXYZ(total_df, crs = food_crs)
crs(final_rast)
plot(final_rast)
final_rast
cellStats(final_rast, "sum") # 25317.85

final_area_fix <- final_rast/raster::area(final_rast)

final_rast_proj <- projectRaster(final_area_fix, food_raster, method = "ngb")

final_rast <- final_rast_proj*raster::area(final_rast_proj)
cellStats(final_rast, "sum") # 25317.85
final_rast

## Save to datalayers 
writeRaster(final_rast, file.path("/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers/marine_crustaceans_aquaculture_meat_nutrientN.tif"), overwrite = T)


## Save total P raster
total_df <- crustacean_pollution %>% 
  dplyr::select(x, y, P_poll_t)

final_rast <- rasterFromXYZ(total_df, crs = food_crs)
crs(final_rast)
plot(final_rast)
final_rast
cellStats(final_rast, "sum") # 3518.493

final_area_fix <- final_rast/raster::area(final_rast)

final_rast_proj <- projectRaster(final_area_fix, food_raster, method = "ngb")

final_rast <- final_rast_proj*raster::area(final_rast_proj)
cellStats(final_rast, "sum") # 3518.493
final_rast

## Save to datalayers 
writeRaster(final_rast, file.path("/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers/marine_crustaceans_aquaculture_meat_nutrientP.tif"), overwrite = T)
```

Datacheck:

Values from this paper: https://pubs.acs.org/doi/pdf/10.1021/acs.est.9b03340


   - 57 kg N / tonne
   - 9.2 kg P / tonne 

The estimates from their paper for "crustaceans" lumps shrimp in with that, so we just used the same values for shrimp and crustaceans. There are no issues with the code. 


```{r}
crustaceans_nutrient_N <- raster(file.path("/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers/marine_crustaceans_aquaculture_meat_nutrientN.tif"))
cellStats(crustaceans_nutrient_N, "sum") # 21753.09 N tonnes

crustaceans_nutrient_P <- raster(file.path("/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers/marine_crustaceans_aquaculture_meat_nutrientP.tif"))
cellStats(crustaceans_nutrient_P, "sum") # 3511.025 P tonnes

sum(production$tonnes_production) # 401719.1

## Total production is 401719.1 tonnes for 2017

## Used 57 N kg and 9.2 P kg for crustaceans

401719.1*0.95*57/1000 # 21753.09 N tonnes
401719.1*0.95*9.2/1000 # 3511.025 P tonnes

## both match our final layers 

final_rast <- raster(file.path(path_final, "marine_crustaceans_aquaculture_meat_nutrient.tif"))
cellStats(final_rast, "sum") # 25264.11
```
