---
title: "Calculate disturbance layers for pigs (industrial, intermediate, backyard)"
author: "Juliette"
date: "10/12/2020"
output: html_document
---

## Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)

source(here("_workflow/common.R"))
source(here("_workflow/common_spatial.R"))
source(here("_spatial/template_raster.R"))

conversion_values <- read_csv(here("animal_farm/disturbance/data/livestock_conversion_values.csv")) 
extra_barn_space <- as.numeric(conversion_values$value[conversion_values$animal == "extra barn space"])
extra_farm_space <- as.numeric(conversion_values$value[conversion_values$animal == "extra farm space"])

backyard_raster_space <- read_csv(here("animal_farm/disturbance/data/livestock_area_head_table.csv")) %>% 
  filter(animal == "pigs" & system == "backyard") %>% 
  right_join(food_rgns_xy, by = "iso3c") %>% 
  select(x,y, meters_sq_animal) %>% 
  rasterFromXYZ(crs = food_crs) 
#plot(backyard_space)

```

For pigs in the EU their LSU matches the USA for the handful of countries that I checked (france, hungary, slovenia, slovakia, uk, czetch republic, germany,norway) no match - cyprus. So we will use the across country conversions.

```{r}
country_conversions <- read_csv(here("animal_farm/disturbance/data/across_country_conversions.csv")) %>% 
  filter(animal == "pigs") %>% 
  right_join(food_rgns_xy, by = "iso3c") %>% 
  select(x,y, adjust_multiplier) %>% 
  rasterFromXYZ(crs = food_crs)

plot(country_conversions)
```




Industrial pigs
```{r}
industrial_df <- read_csv(file.path(prep, "animal_farm/farm/pigs_industrial_meat_location_df.csv"), col_types = "ddcccnnnnncc") 

industrial_raster <- industrial_df %>% 
  select(x,y, current_count) %>% 
  rasterFromXYZ(crs = food_crs)

## industrial market pigs
prop_market <- as.numeric(conversion_values$value[conversion_values$animal == "market pigs" & conversion_values$units == "proportion"]) ## prop market pigs
space_industrial_market <- as.numeric(conversion_values$value[conversion_values$animal == "market pigs" & conversion_values$system == "industrial"]) ## m2/head

industrial_market <- industrial_raster*country_conversions*prop_market*space_industrial_market*0.000001*extra_barn_space*extra_farm_space* country_conversions

## industrial breeder pigs
prop_breeder <- as.numeric(conversion_values$value[conversion_values$animal == "breeder pigs" & conversion_values$units == "proportion"]) ## prop market pigs
space_industrial_breeder <- as.numeric(conversion_values$value[conversion_values$animal == "breeder pigs" & conversion_values$system == "industrial and intermediate"]) ## m2/head

industrial_breeder <- industrial_raster*country_conversions*prop_breeder*space_industrial_breeder*0.000001*extra_barn_space*extra_farm_space

## add up the market and breeder industrial areas and save the total area raster

industrial_area <- industrial_market+industrial_breeder
plot(industrial_area)
plot(log(1+industrial_area))

writeRaster(industrial_area, file.path(prep, "animal_farm/disturbance/total_area_rasters/pigs_industrial_meat_disturbance_area.tif"), format = "GTiff", overwrite = TRUE)

##create disturbance layer

names(industrial_area) <- "km2_eq"

writeRaster(industrial_area, file.path(layers, "land_pigs_industrial_meat_disturbance.tif"), format = "GTiff", overwrite = TRUE)


```


Intermediate pigs
```{r}
intermediate_df <- read_csv(file.path(prep, "animal_farm/farm/pigs_intermediate_meat_location_df.csv"), col_types = "ddcccnnnnncc") 

intermediate_raster <- intermediate_df %>% 
  select(x,y, current_count) %>% 
  rasterFromXYZ(crs = food_crs)

## intermediate market pigs
prop_market <- as.numeric(conversion_values$value[conversion_values$animal == "market pigs" & conversion_values$units == "proportion"]) ## prop market pigs
space_intermediate_market <- as.numeric(conversion_values$value[conversion_values$animal == "market pigs" & conversion_values$system == "intermediate"]) ## m2/head

intermediate_market <- intermediate_raster*country_conversions*prop_market*space_intermediate_market*0.000001*extra_barn_space*extra_farm_space

## intermediate breeder pigs
prop_breeder <- as.numeric(conversion_values$value[conversion_values$animal == "breeder pigs" & conversion_values$units == "proportion"]) ## prop market pigs
space_intermediate_breeder <- as.numeric(conversion_values$value[conversion_values$animal == "breeder pigs" & conversion_values$system == "industrial and intermediate"]) ## m2/head

intermediate_breeder <- intermediate_raster*country_conversions*prop_breeder*space_intermediate_breeder*0.000001*extra_barn_space*extra_farm_space

## add up the market and breeder intermediate areas and save the total area raster

intermediate_area <- intermediate_market+intermediate_breeder
plot(intermediate_area)
plot(log(1+intermediate_area))

writeRaster(intermediate_area, file.path(prep, "animal_farm/disturbance/total_area_rasters/pigs_intermediate_meat_disturbance_area.tif"), format = "GTiff", overwrite = TRUE)

##create disturbance layer

names(intermediate_area) <- "km2_eq"

writeRaster(intermediate_area, file.path(layers, "land_pigs_intermediate_meat_disturbance.tif"), format = "GTiff", overwrite = TRUE)

```


backyard pigs
```{r}
backyard_df <- read_csv(file.path(prep, "animal_farm/farm/pigs_backyard_meat_location_df.csv"), col_types = "ddcccnnnnncc") 

backyard_raster <- backyard_df %>% 
  select(x,y, current_count) %>% 
  rasterFromXYZ(crs = food_crs)

backyard_area <- backyard_raster*backyard_raster_space*0.000001

plot(backyard_area)
plot(log(1+backyard_area))

writeRaster(backyard_area, file.path(prep, "animal_farm/disturbance/total_area_rasters/pigs_backyard_meat_disturbance_area.tif"), format = "GTiff", overwrite = TRUE)

##create disturbance layer

names(backyard_area) <- "km2_eq"

writeRaster(backyard_area, file.path(layers, "land_pigs_backyard_meat_disturbance.tif"), format = "GTiff", overwrite = TRUE)

```

## Do a data check
Do zonal stats for the number of chickens in each system, the space used, and calculate a space/chicken (to compare between the three. layers and broilers should be similar, backyard should be less)

```{r}

## industrial meat
ind_counts <- zonal(industrial_raster, food_rgns_tif, fun="sum", progress="text", na.rm=TRUE) %>% 
  data.frame() %>% 
  rename(ID_0 = zone,
         indu_counts = sum) %>%
  left_join(food_rgns, by="ID_0") %>% 
  select(-ID_0)

ind_area <- zonal(industrial_area, food_rgns_tif, fun="sum", progress="text", na.rm=TRUE) %>% 
  data.frame() %>% 
  rename(ID_0 = zone,
         indu_area = sum) %>%
  left_join(food_rgns, by="ID_0") %>% 
  select(-ID_0)

industrial_check <- left_join(ind_counts, ind_area, by = c("iso3c", "Country")) %>% 
  mutate(km2_per_industrial = indu_area/indu_counts)

## intermediate meat
inter_counts <- zonal(intermediate_raster, food_rgns_tif, fun="sum", progress="text", na.rm=TRUE) %>% 
  data.frame() %>% 
  rename(ID_0 = zone,
         inter_counts = sum) %>%
  left_join(food_rgns, by="ID_0") %>% 
  select(-ID_0)

inter_area <- zonal(intermediate_area, food_rgns_tif, fun="sum", progress="text", na.rm=TRUE) %>% 
  data.frame() %>% 
  rename(ID_0 = zone,
         inter_area = sum) %>%
  left_join(food_rgns, by="ID_0") %>% 
  select(-ID_0)

intermediate_check <- left_join(inter_counts, inter_area, by = c("iso3c", "Country")) %>% 
  mutate(km2_per_intermediate = inter_area/inter_counts)

## backyard
back_counts <- zonal(backyard_raster, food_rgns_tif, fun="sum", progress="text", na.rm=TRUE) %>% 
  data.frame() %>% 
  rename(ID_0 = zone,
         back_counts = sum) %>%
  left_join(food_rgns, by="ID_0") %>% 
  select(-ID_0)

back_area <- zonal(backyard_area, food_rgns_tif, fun="sum", progress="text", na.rm=TRUE) %>% 
  data.frame() %>% 
  rename(ID_0 = zone,
         back_area = sum) %>%
  left_join(food_rgns, by="ID_0") %>% 
  select(-ID_0)

backyard_check <- left_join(back_counts, back_area, by = c("iso3c", "Country")) %>% 
  mutate(km2_per_backyard = back_area/back_counts)

```

```{r}
library(plotly)

all_check <- left_join(industrial_check, intermediate_check) %>% 
  left_join(backyard_check)


check_graph <- ggplot(all_check) +
  geom_point(aes(x = km2_per_industrial, y = Country), 
             color = "brown") + 
  
  geom_point(aes(x = km2_per_intermediate, y = Country), 
             color = "blue") +
  
  geom_point(aes(x = km2_per_backyard, y = Country),
             color = "darkgreen") +
  labs(x = "km2_per_pig",
       title = "Area per pig",
       subtitle = "industrial (brown), intermediate (blue), and backyard (green)")
          

ggsave("plots/area_per_pig.jpg",width = 15, height = 30, dpi=300) 

```
