---
title: "Calculate disturbance layers for chicken (industrial and backyard)"
author: "Juliette"
date: "10/12/2020"
output: html_document
---

**Information on the extra building and farm space multipliers:**

We use a kg/m2 value to calculate the amount of space a chicken needs in its lifetime. However this does not account for the rest of building space or the rest of the farm space. From our research we concluded that adding an additional 10% more area for building walls, etc is appropriate. To do so we multiply the calculated area but 1.1. We also concluded that the building accounts for 40% of the space the entire chicken farm needs, meaning that 60% is missing. This is really 1/0.4 = 2.5, so to account for this we can multiply by 2.5. The math to prove this is below:
 
Proof: 
If you have a pen that is 100m2 there would be 10% more for the chicken building: 100m2 x 1.1 = 110m2
This would only be 40% of the farm
0.4 x ? = 110
? = 275 = total farm area
Can check: 110/275 = 40%


## Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)

source(here("_workflow/common.R"))
source(here("_workflow/common_spatial.R"))
source(here("_spatial/template_raster.R"))

industrial_meat_conversion_values <- read_csv(here("animal_farm/disturbance/data/chicken_pigs_live_weights.csv")) %>% filter(production_system == "broilers")

backyard_value <- read_csv(here("animal_farm/disturbance/data/space_rec.csv")) %>% 
  filter(animal == "chickens")

conversion_values <- read_csv(here("animal_farm/disturbance/data/livestock_conversion_values.csv")) %>% 
  filter(animal == "chicken")

```

## Industrial Meat Chickens
Let's do the industrial meat chickens first
```{r}

## load in the location df map and convert to raster format
industrial_meat <- read_csv(file.path(prep, "animal_farm/farm/chickens_industrial_meat_location_df.csv"), col_types = "ddnnnccnnccncnnc") 

ind_meat_raster <- industrial_meat %>% 
  select(x,y, current_count) %>% 
  rasterFromXYZ(crs = food_crs)

## create raster of head count to live weight conversion values
ind_meat_convert_raster <- industrial_meat_conversion_values %>% 
  right_join(food_rgns_xy) %>% 
  select(x, y, kg_live_wt) %>% 
  rasterFromXYZ(crs = food_crs)

## save the space values for math ease
broiler_space_m2 <- as.numeric(conversion_values$value[conversion_values$product == "meat"]) ## kg live weight/m2
building_space <- 1.1
rest_farm_space <- 2.5

## Calculate total area map, make sure it's in km2 not m2

area_chicken_industrial_meat <- (ind_meat_raster*ind_meat_convert_raster/broiler_space_m2)/(1000000)*building_space*rest_farm_space
plot(area_chicken_industrial_meat)
plot(log(1+area_chicken_industrial_meat))

writeRaster(area_chicken_industrial_meat, file.path(prep, "animal_farm/disturbance/total_area_rasters/chicken_industrial_meat_disturbance_area.tif"), format = "GTiff", overwrite = TRUE)

## save as a disturbance layer, scale to the area of each cell

names(area_chicken_industrial_meat) <- "km2_eq"

writeRaster(area_chicken_industrial_meat, file.path(layers, "land_chickens_industrial_meat_disturbance.tif"), format = "GTiff", overwrite = TRUE)
```


## Industrial egg chickens 
```{r}

## load in the location df map and convert to raster format
industrial_egg <- read_csv(file.path(prep, "animal_farm/farm/chickens_industrial_eggs_location_df.csv"), col_types = "ddnnnccnnccncnnc")

indus_egg_raster <- industrial_egg %>% 
  select(x,y, current_count) %>% 
  rasterFromXYZ(crs = food_crs)

## save the space values for math ease
layer_space_m2 <- as.numeric(conversion_values$value[conversion_values$product == "eggs"]) ## heads/m2
building_space <- 1.1
rest_farm_space <- 2.5

## Calculate total area map, make sure it's in km2 not m2

area_chicken_industrial_eggs <- (indus_egg_raster*(layer_space_m2/1000000))*building_space*rest_farm_space

plot(area_chicken_industrial_eggs)
plot(log(1+area_chicken_industrial_eggs))

## create a total area and a disturbance raster
writeRaster(area_chicken_industrial_eggs, file.path(prep, "animal_farm/disturbance/total_area_rasters/chicken_industrial_eggs_disturbance_area.tif"), format = "GTiff", overwrite = TRUE)

## save as a disturbance layer, scale to the area of each cell

names(area_chicken_industrial_eggs) <- "km2_eq"

writeRaster(area_chicken_industrial_eggs, file.path(layers, "land_chickens_industrial_eggs_disturbance.tif"), format = "GTiff", overwrite = TRUE)

```


backyard chickens next
```{r}
backyard <- read_csv(file.path(prep, "animal_farm/farm/chickens_backyard_eggs&meat_location_df.csv"), col_types = "ddnnnccnnccncnnc")

backyard_raster <- backyard %>% 
  select(x,y, current_count) %>% 
  rasterFromXYZ(crs = food_crs)

backyard_space_m2 <- as.numeric(backyard_value$meters_sq[backyard_value$system == "backyard"]) ## heads/m2

## Calculate total area map, make sure it's in km2 not m2

area_chicken_backyard <- backyard_raster*backyard_space_m2/1000000
# area_chicken_backyard[is.na(area_chicken_backyard)] <- 0
# area_chicken_backyard[!is.finite(area_chicken_backyard)] <- 0

plot(area_chicken_backyard)
plot(log(1+area_chicken_backyard))

## create a total area and a disturbance raster
writeRaster(area_chicken_backyard, file.path(prep, "animal_farm/disturbance/total_area_rasters/chicken_backyard_eggs&meat_disturbance_area.tif"), format = "GTiff", overwrite = TRUE)

## save as a disturbance layer, scale to the area of each cell

names(area_chicken_backyard) <- "km2_eq"

writeRaster(area_chicken_backyard, file.path(layers, "land_chickens_backyard_eggs&meat_disturbance.tif"), format = "GTiff", overwrite = TRUE)

```

## Do a data check
Do zonal stats for the number of chickens in each system, the space used, and calculate a space/chicken (to compare between the three. layers and broilers should be similar, backyard should be less)

```{r}

## industrial meat
ind_meat_counts <- zonal(ind_meat_raster, food_rgns_tif, fun="sum", progress="text", na.rm=TRUE) %>% 
  data.frame() %>% 
  rename(ID_0 = zone,
         meat_counts = sum) %>%
  left_join(food_rgns, by="ID_0") %>% 
  select(-ID_0)

ind_meat_area <- zonal(area_chicken_industrial_meat, food_rgns_tif, fun="sum", progress="text", na.rm=TRUE) %>% 
  data.frame() %>% 
  rename(ID_0 = zone,
         meat_area = sum) %>%
  left_join(food_rgns, by="ID_0") %>% 
  select(-ID_0)

meat_check <- left_join(ind_meat_counts, ind_meat_area, by = c("iso3c", "Country")) %>% 
  mutate(km2_per_chicken_meat = meat_area/meat_counts)

## industrial eggs
ind_eggs_counts <- zonal(indus_egg_raster, food_rgns_tif, fun="sum", progress="text", na.rm=TRUE) %>% 
  data.frame() %>% 
  rename(ID_0 = zone,
         egg_count = sum) %>%
  left_join(food_rgns, by="ID_0") %>% 
  select(-ID_0)

ind_eggs_area <- zonal(area_chicken_industrial_eggs, food_rgns_tif, fun="sum", progress="text", na.rm=TRUE) %>%
  data.frame() %>% 
  rename(ID_0 = zone,
         egg_area = sum) %>%
  left_join(food_rgns, by="ID_0") %>% 
  select(-ID_0)

eggs_check <- left_join(ind_eggs_counts, ind_eggs_area, by = c("iso3c", "Country")) %>% 
  data.frame() %>% 
  mutate(km2_per_chicken_eggs = egg_area/egg_count)

## backyard
backyard_counts <- zonal(backyard_raster, food_rgns_tif, fun="sum", progress="text", na.rm=TRUE) %>% 
  data.frame() %>% 
  rename(ID_0 = zone,
         backyard_count = sum) %>%
  left_join(food_rgns, by="ID_0") %>% 
  select(-ID_0)

backyard_area <- zonal(area_chicken_backyard, food_rgns_tif, fun="sum", progress="text", na.rm=TRUE) %>% 
  data.frame() %>% 
  rename(ID_0 = zone,
         backyard_area = sum) %>%
  left_join(food_rgns, by="ID_0") %>% 
  select(-ID_0)

backyard_check <- left_join(backyard_counts, backyard_area, by = c("iso3c", "Country")) %>% 
  mutate(km2_per_chicken_backyard = backyard_area/backyard_count)

```

```{r}
library(plotly)

all_check <- left_join(meat_check, eggs_check) %>% 
  left_join(backyard_check)


check_graph <- ggplot(all_check) +
  geom_point(aes(x = km2_per_chicken_meat, y = Country), 
             color = "brown") + 
  
  geom_point(aes(x = km2_per_chicken_eggs, y = Country), 
             color = "blue") +
  
  geom_point(aes(x = km2_per_chicken_backyard, y = Country),
             color = "darkgreen") +
  labs(x = "km2_per_chicken",
       title = "Area per chicken",
       subtitle = "industrial meat(brown), industrial eggs (blue), and backyard (green)")
          

ggsave("plots/area_per_chicken.jpg",width = 15, height = 30, dpi=300) 

```

Check the broiler total area from this project to the one from chicken salmon, should be similar except we used 2017 counts here instead of 2016
```{r}
old_area_csv <- read_csv("/home/verstaen/github/food_chicken_salmon/chicken_disturbance/data/summary_stats.csv")
global_old_area <-  old_area_csv %>% 
  summarise(total_area = sum(total_area),
            total_count = sum(count))
# 406,684.9 km2
# 15,059,060,451 heads

new_area <- cellStats(area_chicken_industrial_meat, 'sum')
new_count <- cellStats(ind_meat_raster, 'sum')
## 2,022,313 km2
## 15,803,802,499 heads


```



