---
title: "Calculate disturbance layers for grassland ruminants"
author: "Juliette"
date: "10/12/2020"
output: html_document
---

## Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(countrycode)

source(here("_workflow/common.R"))
source(here("_workflow/common_spatial.R"))
source(here("_spatial/template_raster.R"))

un_rgns <- read_csv(here("_spatial/_output/UNSD_Methodology.csv")) %>%
  filter(!is.na(iso3c)) %>%
  filter(iso3c %in% food_rgns$iso3c)

```


```{r}

ruminant_feed_intake <- read_csv(here("animal_farm/farm/data/ruminants_GLEAMi_v2.csv")) %>% 
  filter(Variable %in% c("HERD: total number of animals", "INTAKE: Total intake - Roughages"),
         Herd_type != "Whole herd",
         Production_system == "Grassland systems") %>% 
  select(-Unit) %>% 
  pivot_wider(names_from = Variable, values_from = Value) %>% 
  rename(headcount = 'HERD: total number of animals',
         kg_dm = 'INTAKE: Total intake - Roughages' ) %>% 
  rename(country = Country,
         animal = Species,
         system = Production_system,
         product = Herd_type) %>% 
  mutate(animal = case_when(animal == "Cattle" ~ "cows",
                            animal == "Buffalo" ~ "buffaloes",
                            animal == "Goats" ~ "goats",
                            animal == "Sheep" ~ "sheep"),
         product = case_when(product == "Dairy" ~ "milk",
                             product == "Non-dairy" ~ "meat"),
         system = "grassland") %>% 
 mutate(kg_dm_per_head = kg_dm/headcount,
         kg_dm_per_head = ifelse(headcount == 0 | kg_dm == 0, NA, kg_dm_per_head)) %>% 
  select(-country)

```

```{r}

missing_countries <- setdiff(food_rgns$iso3c, ruminant_feed_intake$iso3c) %>% 
  as_tibble() %>% 
  rename(iso3c = value) %>% 
  slice(rep(1:n(), each =4)) %>% 
  mutate(animal = rep(c("cows", "buffaloes", "goats", "sheep"), times = 11)) %>% 
  slice(rep(1:n(), each = 2)) %>% 
  mutate(product = rep(c("meat", "milk"), times = 44)) %>% 
  mutate(system = "grassland",
         headcount = NA,
         kg_dm = NA,
         kg_dm_per_head = NA)

ruminant_feed_intake_gf <- ruminant_feed_intake %>%
  rbind(missing_countries) %>% 
  left_join(un_rgns, by = "iso3c") %>% 
  group_by(animal, product, Intermediate_Region_Name) %>% 
  mutate(kg_dm_per_head = ifelse(is.na(kg_dm_per_head), mean(kg_dm_per_head, na.rm = TRUE), kg_dm_per_head)) %>% 
  ungroup() %>% 
  group_by(animal, product, Sub_region_Name) %>% 
  mutate(kg_dm_per_head = ifelse(is.na(kg_dm_per_head), mean(kg_dm_per_head, na.rm = TRUE), kg_dm_per_head)) %>% 
  ungroup() %>% 
  group_by(animal, product, Region_Name) %>% 
  mutate(kg_dm_per_head = ifelse(is.na(kg_dm_per_head), mean(kg_dm_per_head, na.rm = TRUE), kg_dm_per_head)) %>% 
  ungroup() %>% 
  group_by(animal, product, Global_Name) %>% 
  mutate(kg_dm_per_head = ifelse(is.na(kg_dm_per_head), mean(kg_dm_per_head, na.rm = TRUE), kg_dm_per_head)) %>% 
  ungroup() %>% 
  select(iso3c, animal, system, product, headcount, kg_dm, kg_dm_per_head)

write_csv(ruminant_feed_intake_gf, here("animal_farm/disturbance/data/grazing_ruminant_feed_intake.csv"))

```


## Calculate grazing intensity

1. Create a map for each ruminant of total 
2. Add them up - need units in C tonnes/km2/year. it will be in kg_dm_cell/year
3. Calculate the C dry matter using the 50% 
4. Calculate above ground NPP with 60% as above ground biomass
5. Calculate grazing intensity

Calculate total dry mass eaten
```{r}

kg_dm_per_head <- read_csv(here("animal_farm/disturbance/data/grazing_ruminant_feed_intake.csv"))

file_list <- list.files(file.path(prep, "animal_farm/farm/location_tifs"),  pattern = "grassland", full = TRUE)
rum_list <- c("sheep", "goats", "cows", "buffaloes")
product_list <- c("milk", "meat")

 for(ruminant_name in rum_list){
    for(product_name in product_list){

  # ruminant_name <- rum_list[1]
  # product_name <- product_list[1]
      
      file <- str_subset(file_list, ruminant_name) 
      file <- str_subset(file, product_name)
      
      headcount <- raster(file)
      
      perhead_tif <- kg_dm_per_head %>% 
        filter(animal == ruminant_name,
               product == product_name) %>% 
        select(iso3c, kg_dm_per_head) %>% 
        right_join(food_rgns_xy) %>% 
        select(x,y, kg_dm_per_head) %>% 
        rasterFromXYZ(crs = food_crs)
      
      total_eaten <- perhead_tif*headcount
      
      writeRaster(total_eaten, filename = paste(prep, "animal_farm/disturbance/dm_grazed/", ruminant_name, "_grassland", "_", product_name, "_grazed", sep = ""), format = "GTiff", overwrite = TRUE)
  
      }}

```


Add up all the grazing mass 

```{r}

files <- list.files(file.path(prep, "animal_farm/disturbance/dm_grazed/"), full = TRUE) 
files <- grep(files, pattern= 'buffaloes_grassland_meat_grazed', inv=T, value=T)

grazing_stack <- stack(lapply(files, raster))

total_raster <- sum(grazing_stack)

# C tonnes/km2/year. it will be in kg_dm_cell/year
C_dm <- (total_raster*0.5/1000)/area(total_raster)
# make NAs 0
C_dm <- calc(C_dm, fun = function(x){ifelse(is.na(x), 0, x)})

npp <- raster(file.path(prep, "animal_farm/disturbance/npp_2013_2015_rasters/mean_npp_2013_2015.tif")) 
npp_above <- npp*0.6 

grazing_intensity <- C_dm/npp_above
grazing_intensity <- calc(grazing_intensity, fun = function(x){ifelse(x > 0.8, 0.8, x)})
grazing_intensity[grazing_intensity < 0.1] <- 0

plot(grazing_intensity)

writeRaster(grazing_intensity, file.path(prep, ("animal_farm/disturbance/grazing_intensity_all.tif"), sep = ""), format = "GTiff", overwrite = TRUE ) 
```


Next we want to create rasters for the conversion value of all our grazing animals in each country


```{r}

grazing_animals <-  c("buffaloes", "cows", "sheep", "goats")

lsu_values <- read_csv(here("animal_farm/disturbance/data/across_country_conversions.csv")) %>% 
  select(iso3c, animal, lsu_value) %>% 
  filter(animal %in% grazing_animals) %>% 
  pivot_wider(names_from = animal, values_from = lsu_value) %>%
  right_join(food_rgns_xy) %>% 
  select(-iso3c, -cellindex, -ID_0, - Country)

for(animal in grazing_animals){
 
 # animal <- grazing_animals[1]
  
  lsu_raster <- lsu_values %>% 
    select(x,y, paste0(animal)) %>% 
    rasterFromXYZ(crs = food_crs)
  
  names(lsu_raster) <- paste0(animal, "lsu_conversios")
  
  writeRaster(lsu_raster, file.path(paste0(prep, "animal_farm/disturbance/lsu_conversion_rasters/", animal, "_lsu_conversions.tif"), sep = ""), format = "GTiff", overwrite = TRUE ) 
  
}
    
  
```

grab all the grassland animal rasters 
```{r}

location_list <- list.files(file.path(prep, "animal_farm/farm/location_tifs"), pattern = "grassland", full = TRUE)
location_list <- grep(location_list, pattern= 'buffaloes_grassland_meat', inv=T, value=T)

conversion_list <-list.files(file.path(prep, "animal_farm/disturbance/lsu_conversion_rasters"), full = TRUE)

stack1 <- stack(lapply(location_list, raster))
stack2 <- stack(lapply(conversion_list, raster))

stack_all <- stack(stack1, stack2) %>% 
  raster_df()

df_all <- stack_all %>% 
  rowwise() %>% 
  mutate(cows_milk_lsu = cows_grassland_milk*cows_lsu_conversions,
         cows_meat_lsu = cows_grassland_meat*cows_lsu_conversions,
         sheep_meat_lsu = sheep_grassland_meat*sheep_lsu_conversions,
         sheep_milk_lsu = sheep_grassland_milk*sheep_lsu_conversions,
         goats_milk_lsu = goats_grassland_milk*goats_lsu_conversions,
         goats_meat_lsu = goats_grassland_meat*goats_lsu_conversions,
         buffaloes_milk_lsu = buffaloes_grassland_milk*buffaloes_lsu_conversions) %>% 
  mutate(sum_lsu = sum(cows_milk_lsu, cows_meat_lsu, sheep_meat_lsu, sheep_milk_lsu, goats_milk_lsu, goats_meat_lsu, buffaloes_milk_lsu, na.rm = TRUE)) %>%
  mutate(prop_cows_milk = cows_milk_lsu/sum_lsu,
         prop_cows_meat = cows_meat_lsu/sum_lsu,
         prop_sheep_meat = sheep_meat_lsu/sum_lsu,
         prop_sheep_milk = sheep_milk_lsu/sum_lsu,
         prop_goats_milk = goats_milk_lsu/sum_lsu,
         prop_goats_meat = goats_meat_lsu/sum_lsu,
         prop_buffaloes_milk = buffaloes_milk_lsu/sum_lsu)


## do some checks


```

Save a proportion of the total lsu in each cell an animal is contributing to. Use this to allocate the total grazing intensity



```{r}

animals <- c("buffaloes", "cows", "goats", "sheep")
products <- c("milk", "meat")

for(animal in animals){
  for(product in products){
    
    # animal <- animals[1]
    # product <- products[1]
    
    if(animal == "buffaloes" && product == "meat") {
      
      print("meat buffaloes is not one of our categories")
      
    }else{
     raster <- df_all %>% 
      select(x,y, paste0("prop_", animal, "_", product)) %>% 
     rasterFromXYZ(crs = food_crs)
     
      writeRaster(raster, file.path(paste0(prep, "animal_farm/disturbance/grazing_intensity_allocation_rasters/", animal, "_", product, "_grazing_allocation.tif", sep = "")), format = "GTiff", overwrite = TRUE ) 
  
    } 
  }
}

```


Allocate grazing intensity to create disturbance layer

```{r}
animals <- c("buffaloes", "cows", "goats", "sheep")
products <- c("milk", "meat")

grazing_intensity <- raster(file.path(prep, "animal_farm/disturbance/grazing_intensity_all.tif"))

for(animal in animals){
  for(product in products){
    
    # animal <- animals[1]
    # product <- products[1]
    
    if(animal == "buffaloes" && product == "meat") {
      
      print("meat buffaloes is not one of our categories")
      
    }else{
      
      raster <- raster(file.path(prep, paste0("animal_farm/disturbance/grazing_intensity_allocation_rasters/", animal, "_", product, "_grazing_allocation.tif")))
      
      disturbance <- grazing_intensity*raster
      disturbance <- disturbance*area(disturbance)
      names(disturbance) <- "km2_eq"
      disturbance[is.na(disturbance)] <- 0
     
      writeRaster(disturbance, file.path(paste0(layers, "land_", animal, "_grassland_", product, "_disturbance.tif", sep = "")), format = "GTiff", overwrite = TRUE ) 
  
    
    } 
  }
}


```

Check 
```{r}
test <- raster(file.path(layers, "land_cows_grassland_meat_disturbance.tif"))
plot(test)


allocation <-   raster(file.path(prep, "animal_farm/disturbance/grazing_intensity_allocation_rasters/cows_meat_grazing_allocation.tif"))

```

