---
title: "salmon_nutrient_pollution"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    includes: 
toc: true
---

This markdown uses models and species specific parameters from the paper *Hindcasts and Future Projections of Global Inland and Coastal Nitrogen and Phosphorus Loads Due to Finfish Aquaculture* (2013 Bouwman et al)

## Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(sf)
library(raster)
library(readr)
library(data.table)

source(here("_spatial/template_raster.R"))

path_prep <- "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/salmon"
path_final <-"/home/shares/food-systems/Food_footprint/all_food_systems/datalayers"

production <- read_csv(file.path(path_prep, "updated/salmon_farm_updated.csv"))
food_raster <- raster(nrows=2160, ncols=4320, xmn=-180, xmx=180, ymn=-90, ymx=90)
```

Salmon N/P pollution values

N: 49 kg/tonne of growth
P: 8 kg/tonne of growth

Steps: 
1. Multiply tonnes production by 0.95 (this corrects for the growth that occurs outside the seapen on land), this is the value Lex uses
2. Then multiply growth by the N value and the P value
3. We report in total excess nutrients: Sum the N and P.
4. We report in tonnes, so divide by 1000. Save a total raster, a N raster, and a P raster. 

```{r}

salmon_pollution <- production %>%
  dplyr::mutate(prod_growth = tonnes_production*0.95) %>%
  dplyr::mutate(prod_growth_N_poll = prod_growth*49,
                prod_growth_P_poll = prod_growth*8) %>% ## in kg pollution now
  # dplyr::mutate(PO4eq_kg = prod_growth_N_poll*0.42 + prod_growth_P_poll*3.07) %>% cut (01/14/2021)
  dplyr::mutate(total_N_P_tonnes = (prod_growth_N_poll + prod_growth_P_poll)/1000,
                N_poll_t = prod_growth_N_poll/1000,
                P_poll_t = prod_growth_P_poll/1000) %>%
  dplyr::select(x, y, total_N_P_tonnes, N_poll_t, P_poll_t)

sum(salmon_pollution$total_N_P_tonnes) # 149245.8

## save to dataprep folder with the correct naming conventions: 
write.csv(salmon_pollution, file.path(path_prep, "salmon_nutrient.csv"))




## Save total N + P layer

N_P_df <- salmon_pollution %>%
  dplyr::select(x,y,total_N_P_tonnes)

final_rast <- rasterFromXYZ(N_P_df, crs = food_crs)
crs(final_rast)
plot(final_rast)
final_rast
cellStats(final_rast, "sum") # 149245.8

final_area_fix <- final_rast/raster::area(final_rast)

final_rast_proj <- projectRaster(final_area_fix, food_raster, method = "ngb")

final_rast <- final_rast_proj*raster::area(final_rast_proj)
cellStats(final_rast, "sum") # 149245.8
final_rast

## Save to datalayers 
writeRaster(final_rast, file.path(path_final, "marine_salmon_aquaculture_meat_nutrient.tif"), overwrite = T)



## Save total N layer

N_df <- salmon_pollution %>%
  dplyr::select(x,y,N_poll_t)

final_rast <- rasterFromXYZ(N_df, crs = food_crs)
crs(final_rast)
plot(final_rast)
final_rast
cellStats(final_rast, "sum") # 128187.2

final_area_fix <- final_rast/raster::area(final_rast)

final_rast_proj <- projectRaster(final_area_fix, food_raster, method = "ngb")

final_rast <- final_rast_proj*raster::area(final_rast_proj)
cellStats(final_rast, "sum") # 128299
final_rast

## Save to datalayers 

writeRaster(final_rast, file.path("/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers/marine_salmon_aquaculture_meat_nutrientN.tif"), overwrite = T)


## Save total P layer

P_df <- salmon_pollution %>%
  dplyr::select(x,y,P_poll_t)

final_rast <- rasterFromXYZ(P_df, crs = food_crs)
crs(final_rast)
plot(final_rast)
final_rast
cellStats(final_rast, "sum") # 20946.78

final_area_fix <- final_rast/raster::area(final_rast)

final_rast_proj <- projectRaster(final_area_fix, food_raster, method = "ngb")

final_rast <- final_rast_proj*raster::area(final_rast_proj)
cellStats(final_rast, "sum") # 20946.78
final_rast

## Save to datalayers 
writeRaster(final_rast, file.path("/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers/marine_salmon_aquaculture_meat_nutrientP.tif"), overwrite = T)
```

Datacheck: 

Values from this paper, table 7: https://www.tandfonline.com/doi/full/10.1080/10641262.2013.790340

The paper doesn't include actually nutrient output estimates, but it doesn't really matter, since we just use the numbers provided for the nutrient output factors. There are no mistakes in the code, the numbers all make sense to me. 

```{r}
salmon_nutrient_N <- raster(file.path("/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers/marine_salmon_aquaculture_meat_nutrientN.tif"))
cellStats(salmon_nutrient_N, "sum") # 128187.2 N tonnes

salmon_nutrient_P <- raster(file.path("/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers/marine_salmon_aquaculture_meat_nutrientP.tif"))
cellStats(salmon_nutrient_P, "sum") # 20928.52 P tonnes

sum(production$tonnes_production) # 2753752

## Total production is 2753752 tonnes for 2017

## Used 49 N kg and 8 P kg for salmon

2753752*0.95*49/1000 # 127830.3 N tonnes
2753752*0.95*8/1000 # 20870.25 P tonnes

## both match our final layers 

final_rast <- raster(file.path(path_final, "marine_salmon_aquaculture_meat_nutrient.tif"))
cellStats(final_rast, "sum") # 149115.7


```

