---
title: "STEP12_explore_plots_graphs"
author: "Juliette"
date: "5/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
source(here("_workflow/common.R"))

pressures <- read_csv(here::here("_efficiency/data/all_food_pressures.csv"))

```


"a scatter plot of each food, cumulate pressure on the x-axis, Kcal (or protein) content on the y-axis.  Could even potentiallly have each dot be a pie chart with four wedges, each wedge one of the four pressures.  Another way to think about efficiency . . ."

```{r}

## crops
scatter_protein_crops <- ggplot(filter(pressures, grouping == "produce")) +
  geom_point(aes(x = tonnes_protein, y = cumulative_pressure)) +
  facet_wrap(vars(product), scales = "free") +
  theme_minimal() +
  labs(title = "Crops Human Consumption Protein")
ggsave(here("_efficiency/output/efficiency_scatterplots/crop_protein.png"),width = 10, height = 5, units=c("in"))


scatter_kcals_crops <- ggplot(filter(pressures, grouping == "produce")) +
  geom_point(aes(x = million_kcal, y = cumulative_pressure)) +
  facet_wrap(vars(product), scales = "free") +
  theme_minimal() +
  labs(title = "Crops Human Consumption Kcals")
ggsave(here("_efficiency/output/efficiency_scatterplots/crop_calories.png"), width = 10, height = 5,units=c("in"))

## livestock
scatter_protein_livestock <- ggplot(filter(pressures, grouping == "livestock")) +
  geom_point(aes(x = tonnes_protein, y = cumulative_pressure)) +
  facet_wrap(vars(product), scales = "free") +
  theme_minimal() +
  labs(title = "Livestock Protein")
ggsave(here("_efficiency/output/efficiency_scatterplots/livestock_protein.png"), width = 10, height = 5,units=c("in"))

scatter_kcals_livestock <- ggplot(filter(pressures, grouping == "livestock")) +
  geom_point(aes(x = million_kcal, y = cumulative_pressure)) +
  facet_wrap(vars(product), scales = "free") +
  theme_minimal() +
  labs(title = "Livestock Kcals")
ggsave(here("_efficiency/output/efficiency_scatterplots/livestock_calories.png"),width = 10, height = 5, units=c("in"))

## fisheries
scatter_protein_livestock <- ggplot(filter(pressures, str_detect(grouping, "fisheries"))) +
  geom_point(aes(x = tonnes_protein, y = cumulative_pressure)) +
  facet_wrap(vars(product), scales = "free") +
  theme_minimal() +
  labs(title = "Fisheries Protein")
ggsave(here("_efficiency/output/efficiency_scatterplots/fisheries_protein.png"), width = 10, height = 5,units=c("in"))

scatter_kcals_livestock <- ggplot(filter(pressures, str_detect(grouping, "fisheries"))) +
  geom_point(aes(x = million_kcal, y = cumulative_pressure)) +
  facet_wrap(vars(product), scales = "free") +
  theme_minimal() +
  labs(title = "Fisheries Kcals")
ggsave(here("_efficiency/output/efficiency_scatterplots/fisheries_calories.png"), width = 10, height = 5,units=c("in"))

## mariculture
scatter_protein_mariculture <- ggplot(filter(pressures, grouping == "mariculture")) +
  geom_point(aes(x = tonnes_protein, y = cumulative_pressure)) +
  facet_wrap(vars(product), scales = "free") +
  theme_minimal() +
  labs(title = "Mariculture Protein")
ggsave(here("_efficiency/output/efficiency_scatterplots/mariculture_protein.png"), units=c("in"),width = 10, height = 5)

scatter_kcals_mariculture <- ggplot(filter(pressures, grouping == "mariculture")) +
  geom_point(aes(x = million_kcal, y = cumulative_pressure)) +
  facet_wrap(vars(product), scales = "free") +
  theme_minimal() +
  labs(title = "Mariculture Kcals")
ggsave(here("_efficiency/output/efficiency_scatterplots/mariculture_calories.png"), width = 10, height = 5, units=c("in"))
 
 
```


### digging to see if we can unearth cool things about the pressures, products,etc
Data
```{r}
product_tonnes <- read_csv(here("_efficiency/data/product_tonnes.csv"))
pressures_summary <- read_csv(here("_efficiency/data/pressures_summary.csv"))

crop_cal_pro <- read_csv(here::here("_efficiency", "data", "crops_protein_calories.csv"))
livestock_cal_pro <- read_csv(here::here("_efficiency", "data", "livestock_protein_calories.csv"))
mari_cal_pro <- read_csv(here::here("_efficiency", "data", "mariculture_protein_calories.csv"))

```


## Side analysis for pressures 

Crops first
```{r}
pressures_country <- pressures_summary %>% 
  filter(product == "produce") %>% 
  left_join(product_tonnes) %>% 
  group_by(category1_production, pressure) %>% 
  dplyr::summarise(pressure_sum = sum(pressure_rescaled),
                   tonnes_sum = sum(tonnes, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(eff = pressure_sum/tonnes_sum)

ggplot(pressures_country, aes(x = eff, y = pressure, color= pressure)) +
  geom_point() +
  facet_wrap(vars(category1_production)) #,scales = "free"

## kcals and protein
crop_cal_pro_global  <- pressures_summary %>% 
  filter(product == "produce") %>% 
  select(-product) %>% 
  rename(product = organism) %>% 
  left_join(crop_cal_pro) %>% 
  group_by(category1_production, pressure) %>% 
  dplyr::summarise(pressure_sum = sum(pressure_rescaled),
                   tonnes_sum = sum(tonnes, na.rm = TRUE),
                   kcals_sum = sum(million_kcal, na.rm = TRUE),
                   protein_sum = sum(tonnes_protein, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(eff_t = pressure_sum/tonnes_sum,
         eff_c = pressure_sum/kcals_sum,
         eff_p = pressure_sum/protein_sum)

ggplot(crop_cal_pro_global, aes(x = eff_t, y = pressure, color= pressure)) +
  geom_point() +
  facet_wrap(vars(category1_production))

ggplot(crop_cal_pro_global, aes(x = eff_c, y = pressure, color= pressure)) +
  geom_point() +
  facet_wrap(vars(category1_production))

ggplot(crop_cal_pro_global, aes(x = eff_p, y = pressure, color= pressure)) +
  geom_point() +
  facet_wrap(vars(category1_production))

```

Now livestock
```{r}

## livestock 
pressures_livestock  <- pressures_summary %>% 
  filter(category2_production %in% c("cows_meat", "large_ruminant_milk", "large_ruminant_meat", "small_ruminant_milk", "small_ruminant_meat", "chickens_eggs&meat", "chickens_eggs", "chickens_meat", "pigs_meat")) %>% 
  left_join(product_tonnes) %>% 
  group_by(category1_production, pressure) %>% 
  dplyr::summarise(pressure_sum = sum(pressure_rescaled),
                   tonnes_sum = sum(tonnes, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(eff = pressure_sum/tonnes_sum)

ggplot(pressures_livestock, aes(x = eff, y = pressure, color= pressure)) +
  geom_point() +
  facet_wrap(vars(category1_production))

## kcals and protein
livestock_cal_pro_global  <- pressures_summary %>% 
  filter(category2_production %in% c("cows_meat", "large_ruminant_milk", "large_ruminant_meat", "small_ruminant_milk", "small_ruminant_meat", "chickens_eggs&meat", "chickens_eggs", "chickens_meat", "pigs_meat")) %>% 
  unite(product, c("organism", "product")) %>% 
  left_join(livestock_cal_pro) %>% 
  group_by(category1_production, pressure) %>% 
  dplyr::summarise(pressure_sum = sum(pressure_rescaled),
                   tonnes_sum = sum(tonnes, na.rm = TRUE),
                   kcals_sum = sum(million_kcal, na.rm = TRUE),
                   protein_sum = sum(tonnes_protein, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(eff_t = pressure_sum/tonnes_sum,
         eff_c = pressure_sum/kcals_sum,
         eff_p = pressure_sum/protein_sum)

ggplot(livestock_cal_pro_global, aes(x = eff_t, y = pressure, color= pressure)) +
  geom_point() +
  facet_wrap(vars(category1_production))

ggplot(livestock_cal_pro_global, aes(x = eff_c, y = pressure, color= pressure)) +
  geom_point() +
  facet_wrap(vars(category1_production))

ggplot(livestock_cal_pro_global, aes(x = eff_p, y = pressure, color= pressure)) +
  geom_point() +
  facet_wrap(vars(category1_production))
```


Mariculture

```{r}

## mariculture 
pressures_mari  <- pressures_summary %>% 
  filter(category2_production == "mariculture") %>% 
  left_join(product_tonnes) %>% 
  group_by(category1_production, pressure) %>% 
  dplyr::summarise(pressure_sum = sum(pressure_rescaled),
                   tonnes_sum = sum(tonnes, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(eff = pressure_sum/tonnes_sum)

ggplot(pressures_mari, aes(x = eff, y = pressure, color= pressure)) +
  geom_point() +
  facet_wrap(vars(category1_production))

## kcals and protein
mari_cal_pro_global  <- pressures_summary %>% 
  filter(category2_production == "mariculture") %>% 
  select(-product) %>% 
  left_join(mari_cal_pro) %>% 
  group_by(category1_production, pressure) %>% 
  dplyr::summarise(pressure_sum = sum(pressure_rescaled),
                   tonnes_sum = sum(tonnes_product, na.rm = TRUE),
                   kcals_sum = sum(million_kcals, na.rm = TRUE),
                   protein_sum = sum(tonnes_protein, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(eff_t = pressure_sum/tonnes_sum,
         eff_c = pressure_sum/kcals_sum,
         eff_p = pressure_sum/protein_sum)

ggplot(mari_cal_pro_global, aes(x = eff_t, y = pressure, color= pressure)) +
  geom_point() +
  facet_wrap(vars(category1_production))

ggplot(mari_cal_pro_global, aes(x = eff_c, y = pressure, color= pressure)) +
  geom_point() +
  facet_wrap(vars(category1_production))

ggplot(mari_cal_pro_global, aes(x = eff_p, y = pressure, color= pressure)) +
  geom_point() +
  facet_wrap(vars(category1_production))
```


## addressing direct ms question
In Niger pulses are 2 fold less efficient than millet
```{r}
pressure_nig <- pressures_summary %>% 
  filter(iso3c == "NER",
         organism %in% c("xmil", "xpul")) %>% 
  group_by(organism) %>% 
  mutate(organism_sum_press = sum(pressure_rescaled))%>% 
  left_join(product_tonnes) %>% 
  ungroup() %>% 
  mutate(eff = pressure_rescaled/tonnes)

ggplot(pressure_nig, aes(x = eff, y = pressure, color= pressure)) +
  geom_point() +
  facet_wrap(vars(category1_production))
  
```

