---
title: "Create Pressures Summary DF"
author: "Gage"
date: "05/10/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---
This script gets:
 - total production for each group: crops include all production including human food, animal feed, other.
 Feed crops are counted twice (crops) so these files should only be used to calculate efficiency.
 - rescaled pressures for each group.
    Rescaled crop pressures are calculated in _analysis/figures/paper/data_prep/crop_pressures_rescaled.Rmd because
    the "other" portion is not included in the total pressures and has to be calculated a bit differently.


```{r setup, include=FALSE}

library(ggplot2)
library(tidyverse)
library(here)
library(countrycode)
library(cowplot)

source(here("_workflow/common.R"))
source(here("_workflow/common_spatial.R"))

```

Get production data for each group.

### Livestock

```{r}
## Livestock animal products
livestock_cats <- read_csv(here("_analysis/paper_stats/data/FAO_livestock_categories.csv")) %>%
  filter(!is.na(category1_production))
table(livestock_cats$category1_production)
table(livestock_cats$category2_production)

## read in tonnes of production
livestock <- read_csv(here("_analysis/paper_stats/data/FAOSTAT_data_12-18-2020_livestock_primary.csv")) %>%
  filter(Area != "China") %>%
  filter(!(is.na(Value))) %>%
  filter(Item %in% livestock_cats$Item) %>%
  left_join(livestock_cats, by="Item") %>%
  group_by(Area, category1_production, category2_production) %>%
  summarize(tonnes = sum(Value)) %>%
  ungroup()


livestock_iso3c <- livestock %>%
    mutate(Area = iconv(Area, "UTF-8", "UTF-8",sub=''),
         Area = ifelse(Area == "Eswatini", "Swaziland", Area),
         Area = ifelse(Area == "Netherlands Antilles (former)", "Bonaire, Sint Eustatius and Saba", Area),
         Area = ifelse(Area == "French Guyana", "French Guiana", Area),
         Area = ifelse(Area == "China, mainland", "China", Area))

## add standardized country names
livestock_iso3c$iso3c<- countrycode(as.character(livestock_iso3c$Area), origin="country.name", destination = "iso3c")

## make sure all countries have an iso3c code
check <- filter(livestock_iso3c, is.na(iso3c))
unique(check$Area) # these are all higher level area, safe to delete NAs

livestock_iso3c <- livestock_iso3c %>%
  select(iso3c, category1_production, category2_production, tonnes)

```

### Marine fisheries

```{r}

## human component of fofm replaces the total (feed is counted below)
human_fofm <- read_csv("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/catch_human_x_fofm.csv") %>% 
  mutate(category1_production = "fofm_meat",
         category2_production = "marine_fisheries") %>% 
  select(iso3c, category1_production, category2_production, tonnes = catch)

marine_fish <- read_csv("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/catch_extracted_final.csv") %>%
  mutate(category1_production = recode(species_class_final, "Benthic" = "benthic_meat",
             "Demersal" = "demersal_meat",
             "forage_fish" = "fofm_meat",
             "Large pelagic" = "large-pelagic_meat",
             "Medium pelagic" = "medium-pelagic_meat",
             "Reef-associated" = "reef_meat",
             "Small pelagic" = "small-pelagic_meat")) %>%
#  group_by(country_corrected, iso3c) %>%
#  summarize(tonnes = sum(catch_tonnes, na.rm=TRUE)) %>%
  mutate(category2_production = "marine_fisheries") %>%
  select(iso3c, category1_production, category2_production, tonnes=catch) %>% 
  filter(category1_production != "fofm_meat") %>% 
  rbind(human_fofm)

```

### Crops (includes crops for human food, feed, and other)
```{r}
fao_prod <- read_csv(here("feed/data/MAPSPAMcrop_production.csv"))

crop_tonnes <- fao_prod %>% 
  mutate(tonnes = tonnes_producing_crop) %>%
  mutate(category2_production = "human_crop") %>%
  mutate(category1_production = paste(SPAM_super, "produce", sep = "_")) %>%
  select(iso3c = iso3c_producing, category1_production, category2_production, tonnes) %>% 
  filter(!category1_production %in% c("othr_produce", "teas_produce", "toba_produce", "xcof_produce"))


```


### Mariculture: 
```{r}
library(countrycode)

mariculture <- read_csv("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/production_extracted_final.csv") %>%
 # filter(!(aq_group %in% "bivalves")) %>%
  mutate(aq_group = ifelse(aq_group=="marine_fish_general", "marine-fish-general", aq_group),
         aq_group = ifelse(aq_group=="salmonids", "salmon", aq_group),
         aq_group = ifelse(aq_group=="shrimps_prawns", "shrimp", aq_group)) %>%
  select(Country, species = aq_group, tonnes=total_tonnes) %>% 
  mutate(category1_production = paste(species, "_meat", sep = ""),
         category2_production = "mariculture",
         iso3c = countrycode(as.character(Country), origin="country.name", destination = "iso3c"),
         iso3c = ifelse(Country == "High Seas", "HSX",
                        ifelse(Country == "Kosovo", "XKO",
                               ifelse(Country == "Madeira Island", "XMI",
                                      ifelse(Country == "Micronesia", "FSM",
                                             ifelse(Country == "Saint-Martin", "MAF", iso3c))))))  %>% 
  select(iso3c, category1_production, category2_production, tonnes)

```


### Feed: Crops
```{r}

livestock_cats <- read_csv(here("_analysis/paper_stats/data/raster_livestock_cats.csv"))
table(livestock_cats$category2_production, livestock_cats$category1_production)

mapspam_feed <- read_csv(here("feed/data/tonnes_feedproduced_per_country_system.csv")) %>% 
  left_join(livestock_cats, by="animal_system") %>%
  filter(!(is.na(category2_production))) %>%
    mutate(feed_source = "crops") %>%
  group_by(iso3c_producing, feed_source, category1_production, category2_production) %>%
  summarize(tonnes = sum(consumed_tonnes, na.rm=TRUE)) %>%
  dplyr::select(iso3c=iso3c_producing, category1_production, category2_production, feed_source, tonnes) %>%
  ungroup() 


```

### Feed: fodder

```{r}

# remove old files in folder that are created here
do.call(file.remove, list(list.files("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/figures/fodder_tonnes", full.names = TRUE)))

fodder_prod <- sum(stack(raster(file.path(prep, "crop/farm/scaled_maps_2017/crop_fodd_I_scaled.tif")), raster(file.path(prep, "crop/farm/scaled_maps_2017/crop_fodd_H_scaled.tif"))), na.rm=TRUE)

fodder_props <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/datalayers", pattern="_x_land_fodd", full=TRUE)

for(fodder_prop in fodder_props){ #fodder_prop = fodder_props[1]
  fodder_raster <- raster(fodder_prop)
  saveName <- basename(fodder_prop)
  saveName <- gsub("_x_land_fodd_crop_feed", "", saveName)
  saveName <- gsub("_land_", "", saveName)
  fodder_tonnes <- fodder_raster * fodder_prod
  writeRaster(fodder_tonnes, sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/figures/fodder_tonnes/%s", saveName), overwrite=TRUE)
}


lf <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/figures/fodder_tonnes", full=TRUE) 
# lf <- lf[!str_detect(lf, ".tif.aux.xml")] # for if it write .tif.aux.xml files 

fodd_tonnes_stack <- stack(lf)

rgns_eez_land <- raster(file.path(prep, "spatial/land_eez_rgns.tif"))
# plot(rgns_eez_land)

country_sum <- zonal(fodd_tonnes_stack, rgns_eez_land, fun="sum", progress="text", na.rm=TRUE) %>% 
    data.frame() %>%
    rename(ID_0 = zone) %>%
    left_join(food_rgns, by="ID_0") %>%
    select(-ID_0)

country_fodder <- country_sum %>%
   pivot_longer(cols=c(-Country, -iso3c), "animal_system") %>%
   mutate(animal_system = gsub("^land_", "", animal_system)) %>%
     mutate(animal_system = gsub("eggs.meat", "eggs&meat", animal_system)) %>%
   left_join(livestock_cats, by="animal_system") %>%
  mutate(feed_source = "fodder") %>%
  select(iso3c, category1_production, category2_production, feed_source, tonnes=value)

```

Feed: FOFM

```{r}
fofm <- read_csv(here("feed/data/FMFO_bySource_2021Apr.csv")) %>%
  mutate(feed_source = "feedfofm") %>%
  rename(animal_system = system) %>%
  left_join(livestock_cats, by="animal_system") %>%
  filter(!is.na(category2_production)) %>%
  select(iso3c, category1_production, category2_production, feed_source, tonnes) %>% 
  group_by(iso3c, category1_production, category2_production, feed_source) %>% 
  dplyr::summarise(tonnes = sum(tonnes))

```

# Calculate total tonnes of feed per country for each animal group:
```{r}

total_feed <- rbind(country_fodder, fofm, mapspam_feed) %>%
  group_by(iso3c, category1_production, category2_production) %>%
  summarize(tonnes = sum(tonnes, na.rm=TRUE)) %>%
  mutate(category1_production = paste("feed_", category1_production, sep="")) %>%
  mutate(category2_production = paste("feed_", category2_production, sep="")) %>%
  filter(category1_production != "feed_NA")

```


Freshwater fish: 
```{r}
fresh_fish <- read_csv(here("fisheries/freshwater/data/fw_catch_per_iso3c.csv")) %>% 
  select(-ID_0, - Country) %>% 
  rename(tonnes = fw_catch_tonnes) %>% 
  mutate(category1_production  = "fish_meat",
         category2_production = "freshwater_fisheries")

setdiff(food_rgns$iso3c, fresh_fish$iso3c)
setdiff(fresh_fish$iso3c, food_rgns$iso3c)
```

## Join products
```{r}

product_tonnes <- rbind(livestock_iso3c, marine_fish, crop_tonnes, fresh_fish, mariculture, total_feed)

## deal with feed backyard chickens
product_tonnes_feed_chickens <- product_tonnes %>%
  filter(category1_production == "feed_chickens_backyard_eggs&meat") %>%
  mutate(tonnes = tonnes/2) 

product_tonnes_feed_chickens_eggs <- product_tonnes_feed_chickens %>%
  mutate(category1_production = "chickens_eggs")
product_tonnes_feed_chickens_meat <- product_tonnes_feed_chickens %>%
  mutate(category1_production = "chickens_meat")

product_tonnes <- product_tonnes %>%
  filter(category1_production != "feed_chickens_backyard_eggs&meat") %>% 
  rbind(product_tonnes_feed_chickens_eggs, product_tonnes_feed_chickens_meat) %>% 
  mutate(category1_production = ifelse(category1_production == "bivalves_meat", "bivalve_meat", category1_production)) 

write_csv(product_tonnes, here("_efficiency/data/product_tonnes.csv"))

```

### Get product tonnes with crop portion for feed/other removed
```{r}

## deal with feed
total_feed_v2 <- rbind(country_fodder, fofm, mapspam_feed) %>%
  select(iso3c, category1_production, category2_production, feed_source, tonnes) %>%
  mutate(category1_production = paste("feed_", feed_source, "_", category1_production, sep="")) %>%
  mutate(category2_production = paste("feed_", feed_source, "_", category2_production, sep="")) %>%
  select(-feed_source) %>%
  filter(category1_production != "feed_fodder_NA")

## only crops going to humans
fao_prod <- read_csv(here("feed/data/MAPSPAMcrop_production.csv"))
human_crop_prop <- read_csv("feed/data/human_crop_prop_consume.csv")

crop_tonnes_v2 <- left_join(human_crop_prop, fao_prod, by=c("iso3c_producing", "SPAM_super")) %>%
  rowwise() %>%
  mutate(tonnes = human_feed_prop*tonnes_producing_crop) %>%
  #group_by(iso3c_producing) %>%
  #summarize(tonnes = sum(tonnes, na.rm=TRUE)) %>%
  mutate(category2_production = "human_crop") %>%
  mutate(category1_production = paste(SPAM_super, "produce", sep = "_")) %>%
  select(iso3c = iso3c_producing, category1_production, category2_production, tonnes)


product_tonnes_v2 <- rbind(livestock_iso3c, marine_fish, crop_tonnes_v2, fresh_fish, mariculture, total_feed_v2)

## deal with feed backyard chickens
product_tonnes_feed_chickens <- product_tonnes_v2 %>%
  filter(category1_production == "feed_chickens_backyard_eggs&meat") %>%
  mutate(tonnes = tonnes/2) 

product_tonnes_feed_chickens_eggs <- product_tonnes_feed_chickens %>%
  mutate(category1_production = "chickens_eggs")
product_tonnes_feed_chickens_meat <- product_tonnes_feed_chickens %>%
  mutate(category1_production = "chickens_meat")

product_tonnes_v2 <- product_tonnes_v2 %>%
  filter(category1_production != "feed_chickens_backyard_eggs&meat") %>% 
  rbind(product_tonnes_feed_chickens_eggs, product_tonnes_feed_chickens_meat) %>% 
  mutate(category1_production = ifelse(category1_production == "bivalves_meat", "bivalve_meat", category1_production)) 

write_csv(product_tonnes_v2, here("_efficiency/data/product_tonnes_human_crop_consumption.csv"))

check <- read_csv(here("_efficiency/data/product_tonnes_human_crop_consumption.csv"))
table(check$category1_production)

stats <- check %>%
  filter(!(str_detect(category1_production, "feed_"))) %>%
  group_by(category2_production) %>%
  summarize(tonnes = sum(tonnes, na.rm=TRUE)) %>%
  mutate(global_tonnes = sum(tonnes),
         prop_global_tonnes = tonnes/global_tonnes)

stats <- check %>%
  filter(!(str_detect(category1_production, "feed_crop|feed_fodder"))) %>%
  group_by(category2_production) %>%
  summarize(tonnes = sum(tonnes, na.rm=TRUE)) %>%
  mutate(global_tonnes = sum(tonnes),
         prop_global_tonnes = tonnes/global_tonnes)
0.00369+0.0124
tmp <- stats %>%
filter(str_detect(category2_production, "feed_|marine_fisheries|mariculture")) 
sum(tmp$prop_global_tonnes)
```

### Get costs
pressures for all production.  In this case, we replace the pressures with the ones calculated in analysis/figures/paper/dataprep/
crop_pressure_rescaled.Rmd which calculates pressures for all crops (feed, food, other)
```{r}

pressures <- read_csv("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/raw_df_summary/rgn_raw_summary.csv")

pressures <- pressures %>%
  mutate(feed = case_when(category %in% c("feedfofm", "feedfodd", "feedcrop") ~ "feed_",
                          T ~ "")) %>% 
  group_by(iso3c, organism, product, feed, pressure) %>%
  summarize(pressure_value = sum(sum, na.rm=TRUE)) %>%
  mutate(category1_production = paste(organism, product, sep="_")) %>% 
  unite(category1_production, c(feed, category1_production), sep = "")

## divide backyard chicken pressures equally between meat/eggs. 
pressures_chickens <- pressures %>%
  filter(category1_production == "chickens_eggs&meat") %>%
  mutate(pressure_value = pressure_value/2) 

pressures_chickens_eggs <- pressures_chickens %>%
  mutate(product = "eggs",
         category1_production = "chickens_eggs")

pressures_chickens_meat <- pressures_chickens %>%
  mutate(product = "meat",
         category1_production = "chickens_meat")

## for backyard chicken feed
pressures_chickens <- pressures %>%
  filter(category1_production == "feed_chickens_eggs&meat") %>%
  mutate(pressure_value = pressure_value/2) 

pressures_feed_chickens_eggs <- pressures_chickens %>%
  mutate(product = "eggs",
         category1_production = "feed_chickens_eggs")

pressures_feed_chickens_meat <- pressures_chickens %>%
  mutate(product = "meat",
         category1_production = "feed_chickens_meat")

## update the backyards:
pressures_chicken_update <- pressures %>%
  filter(category1_production != "chickens_eggs&meat",
         category1_production != "feed_chickens_eggs&meat") %>%
  rbind(pressures_chickens_meat) %>%
  rbind(pressures_chickens_eggs) %>%
  rbind(pressures_feed_chickens_meat) %>%
  rbind(pressures_feed_chickens_eggs)


setdiff(pressures_chicken_update$category1_production, product_tonnes$category1_production) # 
setdiff(product_tonnes$category1_production, pressures_chicken_update$category1_production) # ofib_produce (not a food item)

```

## combine and rescale pressures
```{r}

category_list <- product_tonnes %>%
  select(category1_production, category2_production) %>%
  unique()

rescaling_values <- read_csv(here("_analysis/rescale_values.csv")) 

pressures_summary_int <- pressures_chicken_update %>%
  left_join(rescaling_values, by="pressure") %>%
  rowwise() %>%
  mutate(pressure_rescaled = pressure_value/global_total) %>%
  left_join(category_list, by="category1_production") 

## the crop specific one'
crops_scaled_raw <- read_csv(here::here("_analysis/figures/paper/data_prep/crops_rescaled.csv"))
crops_scaled_pressures <- crops_scaled_raw %>%
  unite(category1_production, c("organism", "product"), sep = "_", remove = FALSE) %>% 
  mutate(pressure_value = NA,
         global_total = NA,
         category2_production = "human_crop") %>% 
  rename(pressure_rescaled = prop_global) %>% 
    select(-system)

pressures_summary <- pressures_summary_int %>% 
  filter(category2_production != "human_crop") %>% 
  rbind(crops_scaled_pressures) %>% 
  ungroup()

write_csv(pressures_summary , here("_efficiency/data/pressures_summary.csv"))
```


