---
title: "Calculate efficiency rates for figures"
author: "Gage"
date: "05/10/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(tidyverse)
library(here)
library(countrycode)
library(cowplot)
library(vroom)

source(here("_workflow/common.R"))
source(here("_workflow/common_spatial.R"))

pressures_summary <- read_csv(here::here("_efficiency/data/pressures_summary.csv"))
product_tonnes <- read_csv(here::here("_efficiency/data/product_tonnes.csv"))


'%!in%' <- function(x,y)!('%in%'(x,y))

livestock_cal_pro <- read_csv(here::here("_efficiency", "data", "livestock_protein_calories.csv"))
crop_cal_pro <- read_csv(here::here("_efficiency", "data", "crops_protein_calories.csv"))
marine_cal_pro <- read_csv(here::here("_efficiency", "data", "marine_fisheries_protein_calories.csv"))
fresh_cal_pro <- read_csv(here::here("_efficiency", "data", "freshwater_fisheries_protein_calories.csv")) %>% 
  mutate(organism = "fish")
mari_cal_pro <- read_csv(here::here("_efficiency", "data", "mariculture_protein_calories.csv"))

```

# Summary 

In this script we will calculate all pressure efficiencies for tonnes, protein, and kcals for each food group. 


# Methods


## Tonnes efficiencies


### Crops

```{r}

pressures_summary_crops <- pressures_summary %>%
  filter(product == "produce") %>%
  group_by(iso3c, category1_production) %>%
  summarize(cumulative_pressure = sum(pressure_rescaled, na.rm=TRUE)) %>%
  ungroup() %>%
  separate(col=category1_production, c("organism", "product"), sep="_") %>%
  filter(product == "produce")

tmp <- filter(pressures_summary_crops, iso3c=="USA") %>% data.frame()

product_crops <- product_tonnes %>%
  group_by(iso3c, category1_production) %>%
  summarize(tonnes = sum(tonnes, na.rm=TRUE)) %>%
  ungroup() %>%
#  filter(grepl("feed", category1_production))
 # filter(., !grepl("feed",category1_production)) %>%
  separate(col=category1_production, c("organism", "product"), sep="_") %>%
  filter(product == "produce")  
tmp <- filter(product_crops, iso3c=="USA") %>% data.frame()

#setdiff(product_crops$organism, pressures_summary_crops$organism)

cp_rate_crops_int <- left_join(pressures_summary_crops, product_crops, by=c("iso3c", "organism", "product")) %>%
  filter(!is.na(tonnes)) %>%
    filter(!is.na(cumulative_pressure)) %>%
  filter(cumulative_pressure>0) %>%
    filter(tonnes>0) %>%
  rowwise() %>%
  mutate(pressure_per_tonne = cumulative_pressure/tonnes * 1000000) %>% 
  ungroup() %>% 
  ## remove other fibre crops
  filter(organism != "ofib")

long_name <- vroom::vroom(here::here("crop/farm/data/crop_codes_updated.csv")) %>%
  select(organism = SPAM_super,
         long = SPAM_full_name) %>%
  mutate(long = ifelse(organism == "tnut", "tree nuts",
                       ifelse(organism == "spis", "spices",
                              ifelse(organism == "xfru", "fruits",
                                     ifelse(organism == "xmil", "millet",
                                            ifelse(organism == "xoil", "oil",
                                                   ifelse(organism == "xpul", "pulses", long))))))) %>%
  unique()

cp_rate_crops_tonnes <- cp_rate_crops_int %>%
  left_join(long_name) %>%
  mutate(long = ifelse(organism == "xfru", "other fruits",
                          ifelse(organism == "xmil", "millet",
                                 ifelse(organism == "xoil", "other oil crops",
                                        ifelse(organism == "xpul", "pulses", long))))) %>%
  dplyr::select(-product) %>%
  rename(product = organism) %>%
  rename(plot_names = long)
```

### Fisheries 


```{r}
fisheries_list <- c("fish", "demersal", "reef", "large-pelagic", "small-pelagic", "medium-pelagic", "benthic", "fofm")

pressures_summary_fisheries <- pressures_summary %>%
  group_by(iso3c, category1_production) %>%
  summarize(cumulative_pressure = sum(pressure_rescaled, na.rm=TRUE)) %>%
  ungroup() %>%
  separate(col=category1_production, c("organism", "product"), sep="_") %>%
  filter(organism %in% fisheries_list)

tmp <- filter(pressures_summary_fisheries, iso3c=="USA") %>% data.frame()

product_fisheries <- product_tonnes %>%
  group_by(iso3c, category1_production) %>%
  summarize(tonnes = sum(tonnes, na.rm=TRUE)) %>%
  ungroup() %>%
  filter(., !grepl("feed",category1_production)) %>%
  separate(col=category1_production, c("organism", "product"), sep="_") %>%
  filter(organism %in% fisheries_list)

tmp <- filter(product_fisheries, iso3c=="USA") %>% data.frame()

cp_rate_fisheries_tonnes <- left_join(pressures_summary_fisheries, product_fisheries, by=c("iso3c", "organism", "product")) %>%
  filter(!is.na(tonnes)) %>%
    filter(!is.na(cumulative_pressure)) %>%
  filter(cumulative_pressure>0) %>%
    filter(tonnes>0) %>%
  rowwise() %>%
  mutate(pressure_per_tonne = cumulative_pressure/tonnes * 1000000) %>%
  dplyr::select(-product) %>%
  rename(product = organism) %>%
    mutate(plot_names = ifelse(product == "fofm", "forage fish",
                           ifelse(product == "fish", "freshwater fish", product)), 
           product = ifelse(product == "fofm", "forage fish",
                           ifelse(product == "fish", "freshwater fish", product)))
```



### Mariculture 

```{r}

mariculture_list <- c("salmon", "shrimp", "tuna", "crustaceans", "marine-fish-general", "bivalve")

pressures_summary_mariculture <- pressures_summary %>%
  group_by(iso3c, category1_production) %>%
  summarize(cumulative_pressure = sum(pressure_rescaled, na.rm=TRUE)) %>%
  ungroup() %>%
  separate(col=category1_production, c("organism", "product"), sep="_") %>%
  filter(organism %in% mariculture_list)

tmp <- filter(pressures_summary_mariculture, iso3c=="USA") %>% data.frame()

product_mariculture <- product_tonnes %>%
  group_by(iso3c, category1_production) %>%
  summarize(tonnes = sum(tonnes, na.rm=TRUE)) %>%
  ungroup() %>%
  filter(., !grepl("feed",category1_production)) %>%
  separate(col=category1_production, c("organism", "product"), sep="_") %>%
  filter(organism %in% mariculture_list) 

tmp <- filter(product_mariculture, iso3c=="USA") %>% data.frame()

cp_rate_mariculture_tonnes <- left_join(pressures_summary_mariculture, product_mariculture, by=c("iso3c", "organism", "product")) %>%
  filter(!is.na(tonnes)) %>%
    filter(!is.na(cumulative_pressure)) %>%
  filter(cumulative_pressure>0) %>%
    filter(tonnes>0) %>%
  mutate(iso3c = ifelse(iso3c == "HKG", "CHN", iso3c)) %>% 
  group_by(iso3c, organism, product) %>% 
  dplyr::summarise(cumulative_pressure = sum(cumulative_pressure),
                   tonnes = sum(tonnes)) %>% 
  rowwise() %>%
  mutate(pressure_per_tonne = cumulative_pressure/tonnes * 1000000) %>%
  dplyr::select(-product) %>%
  rename(product = organism) %>%
  mutate(plot_names = product)


```


### Livestock 

groups:  cows_meat, cows_milk, buffalo_milk, goats_milk, goats_meat, sheep_milk, sheep_meat, chickens_eggs, chickens_meat, pigs_meat

```{r}
categories <- read_csv(here("_analysis/figures/paper/data/grouping_naming_structure.csv"))

livestock_list <- c("cows_meat", "cows_milk", "buffaloes_milk", "goats_milk", "goats_meat", "sheep_milk", "sheep_meat", "chickens_eggs", "chickens_meat", "pigs_meat")

pressures_summary_livestock <- pressures_summary %>%
  group_by(iso3c, category1_production) %>%
  summarize(cumulative_pressure = sum(pressure_rescaled, na.rm=TRUE)) %>%
  ungroup() %>%
  filter(category1_production %in% livestock_list) 

tmp <- filter(pressures_summary_livestock, iso3c=="USA") %>% data.frame()

product_livestock <- product_tonnes %>%
  group_by(iso3c, category1_production) %>%
  summarize(tonnes = sum(tonnes, na.rm=TRUE)) %>%
  ungroup() %>%
  filter(., !grepl("feed",category1_production))%>%
  filter(category1_production %in% livestock_list)

tmp <- filter(product_livestock, iso3c=="USA") %>% data.frame()

cp_rate_livestock_tonnes <- left_join(pressures_summary_livestock, product_livestock, by=c("iso3c", "category1_production")) %>%
  filter(!is.na(tonnes)) %>%
    filter(!is.na(cumulative_pressure)) %>%
  filter(cumulative_pressure>0) %>%
    filter(tonnes>0) %>%
  rowwise() %>%
  mutate(pressure_per_tonne = cumulative_pressure/tonnes * 1000000) %>% 
  rename(product = category1_production) %>%
  mutate(product = ifelse(product == "cows_meat", "cattle_meat", product)) %>%
  mutate(plot_names = as.factor(str_replace(product, "_", " ")))
```

### Feed 
Need to adjust pressure/tonne to reflect tonnes of final animal product. 

```{r}
categories <- read_csv(here("_analysis/figures/paper/data/grouping_naming_structure.csv"))

pressures_summary_feed <- pressures_summary %>% 
  filter(str_detect(category1_production, "feed_")) %>% 
  group_by(iso3c, category1_production) %>%
  summarize(cumulative_pressure = sum(pressure_rescaled, na.rm=TRUE)) %>%
  ungroup() 

product_feed <- product_tonnes %>%
  group_by(iso3c, category1_production) %>%
  summarize(tonnes = sum(tonnes, na.rm=TRUE)) %>%
  ungroup() %>%
  filter(str_detect(category1_production, "feed_")) 

cp_rate_feed_tonnes <- left_join(pressures_summary_feed, product_feed, by=c("iso3c", "category1_production")) %>%
  filter(!is.na(tonnes)) %>%
    filter(!is.na(cumulative_pressure)) %>%
  filter(cumulative_pressure>0) %>%
    filter(tonnes>0) %>%
  rowwise() %>%
  mutate(pressure_per_tonne = cumulative_pressure/tonnes * 1000000) %>%
  mutate(product = category1_production) %>%
  mutate(category1_production = str_replace(category1_production, "_", " "),
         category1_production = str_replace(category1_production, "_", " ")) %>%
  mutate(category1_production = str_remove(category1_production, "feed ")) %>%
  separate(category1_production, into = c("animal", "prod"), sep = " ", remove = TRUE) %>%
  mutate(prod = ifelse(animal %in% c("buffaloes", "chickens", "cows", "pigs", "goats", "sheep"), prod, NA)) %>%
  unite(category1_production,  c("animal", "prod"), sep = " ", na.rm = TRUE) %>%
  mutate(category1_production = paste(category1_production, "feed", sep = " ")) %>%
  rename(plot_names = category1_production) %>%
  mutate(plot_names = str_replace(plot_names, "cows meat", "cattle meat"), 
         product = str_replace(product, "cows_meat", "cattle_meat"))
  
```

Save the tonnes CP rates 

```{r}

all_tonnes <- rbind(cp_rate_crops_tonnes, cp_rate_fisheries_tonnes, cp_rate_mariculture_tonnes, cp_rate_livestock_tonnes, cp_rate_feed_tonnes)

write_csv(all_tonnes, here::here("_efficiency/data/all_pressure_per_tonnes.csv"))

```


## Protein and Kcal efficiencies 

### Livestock 

```{r}

live_prod_list <- pull(livestock_cal_pro, product) %>%  unique()

## remove feed for now until we clean that up
pressures_raw <- read_csv(here("_efficiency/data/pressures_summary.csv")) 

pressures <- pressures_raw %>% 
  unite(product, c(organism, product)) %>% 
  filter(product %in% live_prod_list,
         !str_detect(category1_production, "feed")) %>% 
  group_by(iso3c, product) %>% 
  summarize(cumulative_pressure = sum(pressure_rescaled, na.rm=TRUE)) %>%
  ungroup()

## finish up prepping the data
cp_rate_livestock_pro_cal <- left_join(pressures, livestock_cal_pro) %>%
  filter(!is.na(tonnes)) %>%
  filter(!is.na(cumulative_pressure)) %>%
  filter(cumulative_pressure>0) %>%
  filter(tonnes >= 10) %>% 
  rowwise() %>%
  mutate(pressure_per_tonne_protein = cumulative_pressure/tonnes_protein * 1000000,
         pressure_per_mill_kcal = cumulative_pressure/million_kcal * 1000000) %>% 
  mutate(product = ifelse(product == "cows_meat", "cattle_meat", product)) %>%
  select(-tonnes, -product_tonnes_to_mill_kcal, -product_tonnes_to_tonnes_protein) %>% 
    mutate(plot_names = as.factor(str_replace(product, "_", " ")))

```


### Crops

```{r}

long_name <- vroom::vroom(here::here("crop/farm/data/crop_codes_updated.csv")) %>% 
  select(organism = SPAM_super,
         long = SPAM_full_name) %>% 
  mutate(long = ifelse(organism == "tnut", "tree nuts",
                       ifelse(organism == "spis", "spices",
                              ifelse(organism == "xfru", "other fruits",
                                     ifelse(organism == "xmil", "millet",
                                            ifelse(organism == "xoil", "other oil crops",
                                                   ifelse(organism == "xpul", "pulses", long))))))) %>% 
  unique() %>% 
  rename(product = organism)

spam_names_df <- vroom::vroom(here::here("crop/farm/data/crop_codes_updated.csv")) 
spam_names <- unique(spam_names_df$SPAM_super)

pressures_crop <- pressures_raw %>% 
  filter(organism %in% spam_names) %>% 
  group_by(iso3c, organism) %>% 
  summarize(cumulative_pressure = sum(pressure_rescaled, na.rm=TRUE)) %>%
  ungroup() %>% 
  rename(product = organism)

cp_rate_crops_pro_cal <- left_join(crop_cal_pro, pressures_crop, by = c("iso3c", "product")) %>%
  filter(!is.na(tonnes)) %>%
  filter(!is.na(cumulative_pressure)) %>%
  filter(cumulative_pressure>0) %>%
  filter(tonnes >= 10) %>% 
  rowwise() %>%
  mutate(
         pressure_per_tonne_protein = cumulative_pressure/tonnes_protein * 1000000,
         pressure_per_mill_kcal = cumulative_pressure/million_kcal * 1000000) %>%
  dplyr::select(-tonnes) %>%
    left_join(long_name) %>%
  rename(plot_names = long)
  
```


### Fisheries
Some wrangling
```{r}
fresh_cal_pro_2 <- fresh_cal_pro %>% 
  mutate(organism = "freshwater fish",
         species_class_final = "Freshwater fish") %>%
  rename(tonnes_product = freshwater_tonnes) 

fisheries_cal_pro <- marine_cal_pro %>% 
  mutate(organism = case_when(species_class_final == "Demersal" ~ "demersal",
                              species_class_final == "Large pelagic" ~ "large-pelagic",
                              species_class_final == "Medium pelagic" ~ "medium-pelagic",
                              species_class_final == "Small pelagic" ~ "small-pelagic",
                              species_class_final == "Reef-associated" ~ "reef",
                              species_class_final == "Benthic" ~ "benthic",
                              species_class_final == "forage_fish" ~ "fofm")) %>% 
  rbind(fresh_cal_pro_2)

fish_cat <- pull(fisheries_cal_pro,organism) %>% 
  unique()

pressures_fisheries <- pressures_raw %>% 
  mutate(organism = ifelse(organism == "fish", "freshwater fish", organism)) %>% 
  filter(organism %in% fish_cat) %>% 
  group_by(iso3c, organism) %>% 
  summarize(cumulative_pressure = sum(pressure_rescaled, na.rm=TRUE)) %>%
  ungroup()

cp_rate_fish_pro_cal <- left_join(fisheries_cal_pro, pressures_fisheries, by = c("iso3c", "organism")) %>%
  filter(!is.na(million_kcal)) %>%
  filter(!is.na(cumulative_pressure)) %>%
  filter(cumulative_pressure>0) %>%
    mutate(organism = ifelse(organism == "fofm", "forage fish", organism)) %>% 
  #filter(tonnes_prod >= 10) %>% 
  rowwise() %>%
  mutate(
         pressure_per_tonne_protein = cumulative_pressure/tonnes_protein * 1000000,
         pressure_per_mill_kcal = cumulative_pressure/million_kcal * 1000000) %>%
   dplyr::select(-tonnes_product, -species_class_final) %>%
  mutate(plot_names = organism) %>%
  rename(product = organism)
  
```



### Mariculture
Some wrangling
```{r}

pressures_mari <- pressures_raw %>% 
  filter(category2_production == "mariculture") %>% 
  group_by(iso3c, organism) %>% 
  summarize(cumulative_pressure = sum(pressure_rescaled, na.rm=TRUE)) %>%
  ungroup()

cp_rate_mari_pro_cal <- left_join(mari_cal_pro, pressures_mari, by = c("iso3c", "organism")) %>%
  filter(!is.na(million_kcals)) %>%
  filter(!is.na(cumulative_pressure)) %>%
  filter(cumulative_pressure>0) %>% 
  rowwise() %>%
  mutate(pressure_per_tonne_protein = cumulative_pressure/tonnes_protein * 1000000,
         pressure_per_mill_kcal = cumulative_pressure/million_kcals * 1000000) %>%
  select(-tonnes_product) %>%
  rename(product = organism) %>%
  mutate(plot_names = product) %>%
  rename(million_kcal = million_kcals)
  
```


Save the calories and protein CP rates 


```{r}
all_cal_prot <- rbind(cp_rate_livestock_pro_cal, cp_rate_mari_pro_cal, cp_rate_fish_pro_cal, cp_rate_crops_pro_cal) %>%
   ungroup() %>% 
  filter(!is.infinite(pressure_per_tonne_protein))

write_csv(all_cal_prot, here::here("_efficiency/data/all_pressure_per_cal_prot.csv"))

```



Combine them into one df we can use for plotting figures and supplementary data

```{r}

all_cal_prot_cp <- read_csv(here::here("_efficiency/data/all_pressure_per_cal_prot.csv"))

all_tonnes_cp <- read_csv(here::here("_efficiency/data/all_pressure_per_tonnes.csv"))

name_fig <- unique(all_cal_prot$product)
name_si <- unique(all_tonnes_cp$product)

setdiff(name_fig, name_si)
setdiff(name_si, name_fig)

combine <- left_join(all_tonnes_cp, all_cal_prot_cp) %>% 
  left_join(food_rgns) %>% 
  select(country = Country, iso3c, product, plot_names, tonnes, cumulative_pressure, pressure_per_tonne, million_kcal, tonnes_protein, pressure_per_tonne_protein, pressure_per_mill_kcal) %>% 
  mutate(country = ifelse(iso3c == "HSX", "High Seas", country)) 


chk <- combine %>% 
  filter(is.na(million_kcal),
         !str_detect(product, "feed")) %>% 
  mutate(remove = "yes")

combine_rm_na <- left_join(combine, chk) %>% 
  filter(is.na(remove)) %>% 
  select(-remove)

## This is the data we will use for the SI_data and for all of the efficiency figures
write_csv(combine_rm_na, here::here("_analysis/SI_data/output/efficiency_df.csv"))

```

