---
title: "Create data frame for other analysis"
author: "Juliette"
date: "5/12/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

Current weirdness with this data:
- there are some countries we have CP for livestock products but no kcals or protein. this is probably because FAO didn't report those products. I need to go back and make sure i gapfilled for those


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
source(here("_workflow/common.R"))
livestock_cal_pro <- read_csv(here::here("_efficiency", "data", "livestock_protein_calories.csv")) %>% 
  select(-product_tonnes_to_mill_kcal, -product_tonnes_to_tonnes_protein)
crop_cal_pro <- read_csv(here::here("_efficiency", "data", "crops_protein_calories.csv"))
marine_cal_pro <- read_csv(here::here("_efficiency", "data", "marine_fisheries_protein_calories.csv")) %>% 
  rename(product = species_class_final)  %>% 
  mutate(product = ifelse(product == "forage_fish", "Forage fish", product))
fresh_cal_pro <- read_csv(here::here("_efficiency", "data", "freshwater_fisheries_protein_calories.csv"))  %>% 
  mutate(product = "freshwater fish") %>% 
  rename(tonnes = freshwater_tonnes)
mari_cal_pro <- read_csv(here::here("_efficiency", "data", "mariculture_protein_calories.csv")) %>% 
  rename(product = organism,
         tonnes = tonnes_product,
         million_kcal = million_kcals)
```

Need to add back in total tonnes product for fisheries
```{r}
edible_tonnes <- read_csv(here("_efficiency/data/fisheries_edible_tonnes.csv")) %>% 
  group_by(iso3c, species_class_final) %>% 
  dplyr::summarise(tonnes = sum(edible_tonnes, na.rm = TRUE) )%>% 
  rename(product = species_class_final) %>%
  mutate(product = ifelse(product == "forage_fish", "Forage fish", product))
marine_cal_pro_2 <- left_join(marine_cal_pro, edible_tonnes, by = c("iso3c", "product")) %>% 
  mutate(product = tolower(product)) %>%
  dplyr::select(-tonnes) %>%
  rename(tonnes = tonnes_product)
```


Read in and wrangle the the cumulative pressures

```{r}
pressures_raw <- read_csv(here("_efficiency/data/pressures_summary.csv")) 
pressures <- pressures_raw %>% 
  filter(!str_detect(category1_production, "feed")) %>% 
  mutate(product_2 = ifelse(str_detect(organism, c("chickens|cows|buffaloes|pigs|sheep|goats")), category1_production, organism)) %>% 
  group_by(iso3c, product_2) %>% 
  dplyr::summarise(cumulative_pressure = sum(pressure_rescaled, na.rm = TRUE)) %>% 
  rename(product = product_2) %>% 
  ## fix fisheries naming
  mutate(product = ifelse(product =="large-pelagic", "large pelagic",
                          ifelse(product == "medium-pelagic", "medium pelagic",
                                 ifelse(product == "small-pelagic", "small pelagic",
                                        ifelse(product == "reef", "reef-associated",
                                               ifelse(product == "fofm", "forage fish", product)))))) %>% 
  ## fix freshwater fisheries name
  mutate(product = ifelse(product == "fish", "freshwater fish", product))
```

Combine all foods and presssures together
```{r}
all_food <- rbind(livestock_cal_pro, crop_cal_pro, marine_cal_pro_2, fresh_cal_pro, mari_cal_pro) 
all_food_pressures <- left_join(pressures, all_food)
## check the weird ones where there is a teeeennny amount of CP but no tonnage

ch <- all_food_pressures %>% 
  filter(is.na(tonnes))

all_food_pressures_final <- all_food_pressures %>% 
  filter(!is.na(tonnes)) %>% 
  ## add in grouping names
  mutate(grouping = case_when(product %in% c("benthic", "demersal", "forage fish", "large pelagic", "medium pelagic", "reef-associated", "small pelagic") ~ "marine fisheries",
                              product == "freshwater fish" ~ "freshwater fisheries",
                              product %in% c("bivalve", "crustaceans", "salmon", "shrimp", "marine-fish-general", "tuna") ~ "mariculture",
                              str_detect(product, c("pigs|chickens|cows|buffaloes|sheep|goats")) ~ "livestock",
                              T ~ "produce")) %>% 
  ungroup() %>% 
  mutate(global_pressure = sum(cumulative_pressure),
         global_tonnes = sum(tonnes),
         global_kcal = sum(million_kcal),
         global_protein = sum(tonnes_protein),
         
         percent_pressure = cumulative_pressure/global_pressure,
         percent_tonnes = tonnes/global_tonnes,
         percent_kcal = million_kcal/global_kcal,
         percent_protein = tonnes_protein/global_protein) 

write_csv(all_food_pressures_final, here::here("_efficiency/data/all_food_pressures.csv"))

all_food_pressures_summary <- all_food_pressures_final %>% 
  group_by(grouping) %>% 
  dplyr::summarise(percent_pressure = sum(percent_pressure),
                   percent_tonnes = sum(percent_tonnes),
                   percent_kcal = sum(percent_kcal),
                   percent_protein = sum(percent_protein)) %>% 
  ungroup()

all_food_pressures_summary_foods <- all_food_pressures_final %>% 
  group_by(product) %>% 
  dplyr::summarise(percent_pressure = sum(percent_pressure),
                   percent_tonnes = sum(percent_tonnes),
                   percent_kcal = sum(percent_kcal),
                   percent_protein = sum(percent_protein)) %>% 
  ungroup()
```

Combine all foods and presssures together: this time only include the human consumed portion.
```{r}

crop_cal_pro_humans <- read_csv(here::here("_efficiency/data/crops_protein_calories_humans_only.csv")) 

all_food <- rbind(livestock_cal_pro, crop_cal_pro_humans, marine_cal_pro_2, fresh_cal_pro, mari_cal_pro) 

all_food_pressures_final_humans <- all_food %>% 
  filter(!is.na(tonnes)) %>% 
  ## add in grouping names
  mutate(grouping = case_when(product %in% c("benthic", "demersal", "forage fish", "large pelagic", "medium pelagic", "reef-associated", "small pelagic") ~ "marine fisheries",
                              product == "freshwater fish" ~ "freshwater fisheries",
                              product %in% c("bivalve", "crustaceans", "salmon", "shrimp", "marine-fish-general", "tuna") ~ "mariculture",
                              str_detect(product, c("pigs|chickens|cows|buffaloes|sheep|goats")) ~ "livestock",
                              T ~ "produce")) %>% 
  ungroup() %>% 
  mutate(global_tonnes = sum(tonnes),
         global_kcal = sum(million_kcal),
         global_protein = sum(tonnes_protein),
         
      
         percent_tonnes = tonnes/global_tonnes,
         percent_kcal = million_kcal/global_kcal,
         percent_protein = tonnes_protein/global_protein) 

write_csv(all_food_pressures_final_humans, here::here("_efficiency/data/all_food_pressures_human_crop_consume.csv"))


```


Table for the SI
```{r}
si_table <-all_food_pressures_final %>% 
  group_by(iso3c) %>% 
  dplyr::summarise(global_prop_product = sum(percent_tonnes),
                   global_prop_kcal = sum(percent_kcal),
                   global_prop_protein = sum(percent_protein)) %>% 
  ungroup()
write_csv(si_table, here::here("_efficiency/data/SI_results_global_props_total.csv"))
```
