---
title: "Wrangle and calculate FAO protein and calories CF for crops"
author: "Juliette"
date: "3/8/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(countrycode)
source(here("_workflow/common.R"))

un <- read_csv(here("_spatial/_output/UNSD_Methodology.csv")) %>% 
  dplyr::select(iso3c, Global_Name, Region_Name, Sub_region_Name, Intermediate_Region_Name) %>% 
  filter(iso3c %in% food_rgns$iso3c)

crops_raw <- read_csv(file.path(raw, "FAO_data/v2021/supply_utlilization_accounts/crops/FAOSTAT_data_2-5-2021.csv"))
```

# Crops

## Wrangle, calculate conversion factors, and calculate tonnes product to tonnes protein and million kcals
```{r}

nutri_contents_crops <- crops_raw %>% 
  select(Area, Element, Item, Year, Unit, Value) %>% 
  filter(Year == 2017,
         Element %in% c("Food supply quantity (tonnes)", "Calories/Year", "Proteins/Year")) %>% 
  
  ## country code wrangling
   filter(Area != "China") %>% 
   mutate(Area = ifelse(Area == "Eswatini", "Swaziland", Area),
         Area = ifelse(Area == "Netherlands Antilles (former)", "Bonaire, Sint Eustatius and Saba", Area),
         Area = ifelse(Area == "French Guyana", "French Guiana", Area),
         Area = ifelse(Area == "China, mainland", "China", Area),
         iso3c = countrycode(Area, origin="country.name", destination = "iso3c")) %>% 
  
  ## final wrangling
  select(-Area, -Year, -Unit) %>% ##units for production are tonnes, protein is tonnes, and calories is million kcals
  pivot_wider(names_from = "Element", values_from = "Value") %>% 
  select(iso3c, item = Item, production = 'Food supply quantity (tonnes)' , million_kcal_year = 'Calories/Year', tonnes_protein_year = 'Proteins/Year') %>% 
  filter(!is.na(million_kcal_year),
         !is.na(production)) %>% 
  
  ## calculate the mult factor
  mutate(million_kcal_per_prod_tonne = million_kcal_year/production,
         tonne_protein_per_prod_tonne = tonnes_protein_year/production)

## see how well the fao to mapsam csv works
temp <- nutri_contents_crops %>% 
  select(item) %>% 
  unique()

spam_fao_names <- read_csv(here("feed/data/MapSPAM_to_FAO.csv"))

##doesnt seem very well aligned
setdiff(spam_fao_names$product_description_FAOSTAT, temp$item) 
save <- setdiff(temp$item, spam_fao_names$product_description_FAOSTAT) %>% as_tibble()
#write_csv(save, here("_efficiency/data/save_fao_missing_crops.csv"))
## eh not great... so I'm just going to go through and create a file that will work best for this

#write_csv(temp, here("_efficiency/data/temp.csv"))

# names <- pressures_summary %>% 
#   ungroup() %>% 
#   filter(product == "produce") %>% 
#   select(organism) %>% 
#   unique()

```

Assign SPAM names to the fao product descriptions
```{r}
fix_missing_names <- read_csv(here("_efficiency/data/missing_fao_names.csv")) %>% 
  rename(item = fao_name)

name_match <- spam_fao_names %>% 
  select(spam_name = SPAM_super, item = product_description_FAOSTAT)

nutri_contents_crops_named <- left_join(nutri_contents_crops, name_match) 

fix_missing <- nutri_contents_crops_named %>% 
  filter(is.na(spam_name)) %>% 
  left_join(fix_missing_names) %>% 
  select(-spam_name) %>% 
  rename(spam_name = SPAM_super)

nutri_contents_crops_named_2 <- nutri_contents_crops_named %>% 
  filter(!is.na(spam_name)) %>% 
  rbind(fix_missing) %>% 
  group_by(iso3c, spam_name) %>% 
  dplyr::summarise(tonnes = sum(production, na.rm = TRUE),
                   million_kcal = sum(million_kcal_year, na.rm = TRUE),
                   tonnes_protein = sum(tonnes_protein_year, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(protein_cf = tonnes_protein/tonnes,
         protein_cf = ifelse(is.na(protein_cf) | is.infinite(protein_cf), 0, protein_cf),
         calorie_cf = million_kcal/tonnes,
         calorie_cf = ifelse(is.na(calorie_cf) | is.infinite(calorie_cf), 0, calorie_cf))
```


Lets see what it looks like

```{r}
ggplot(nutri_contents_crops_named_2, aes(x = tonnes, y = million_kcal)) +
  geom_point()+
  theme_minimal()+
  geom_smooth(method='lm', formula= y~x) +
  facet_wrap(vars(spam_name), scales = "free")

ggplot(nutri_contents_crops_named_2, aes(x = tonnes, y = tonnes_protein)) +
  geom_point()+
  theme_minimal()+
  geom_smooth(method='lm', formula= y~x) +
  facet_wrap(vars(spam_name), scales = "free")
```


```{r}
crop_protein_cfs <- nutri_contents_crops_named_2 %>% 
  group_by(spam_name) %>% 
  filter(protein_cf >0) %>% 
  dplyr::summarise(protein_cf = mean(protein_cf, na.rm = TRUE))

crop_calorie_cfs <- nutri_contents_crops_named_2 %>% 
  group_by(spam_name) %>% 
  filter(calorie_cf >0) %>% 
  dplyr::summarise(calorie_cf = mean(calorie_cf, na.rm = TRUE))

crop_cfs <- left_join(crop_calorie_cfs, crop_protein_cfs) %>% 
  filter(!is.na(spam_name)) %>%  ## is is mate
  add_row(spam_name = "oilp", calorie_cf = 4.289197, protein_cf = 0.1540623) %>%  ## gapfilled with xoil values
  rename(product = spam_name)
```


## Do the conversion for the crops 

This one  contains all the crop production (used for general efficiency calculations).
```{r}
crop_product_tonnes <- read_csv(here("_efficiency/data/product_tonnes.csv"))
filter(crop_product_tonnes, iso3c=="USA" & category1_production=="maiz_produce")

crop_product_tonnes <- read_csv(here("_efficiency/data/product_tonnes.csv")) %>% 
  rename(product = category1_production) %>% 
  separate(product, c("product", "source"), sep = "_") %>% 
  filter(product %in% crop_cfs$product) %>% 
  dplyr::select(-category2_production) %>% 
  left_join(crop_cfs) %>% 
  mutate(million_kcal = tonnes*calorie_cf,
         tonnes_protein = tonnes*protein_cf) %>% 
  select(iso3c, product, tonnes, million_kcal, tonnes_protein) 

write_csv(crop_product_tonnes, here::here("_efficiency", "data", "crops_protein_calories.csv"))

```

This one contains the crop production only consumed by humans.
This is used in figure 4 for the points.
```{r}


crop_product_tonnes <- read_csv(here("_efficiency/data/product_tonnes_human_crop_consumption.csv"))
filter(crop_product_tonnes, iso3c=="USA" & category1_production=="maiz_produce")

crop_product_tonnes_humans <- crop_product_tonnes %>% 
  rename(product = category1_production) %>% 
  separate(product, c("product", "source"), sep = "_") %>% 
  filter(product %in% crop_cfs$product) %>% 
  dplyr::select(-category2_production) %>% 
  left_join(crop_cfs) %>% 
  mutate(million_kcal = tonnes*calorie_cf,
         tonnes_protein = tonnes*protein_cf) %>% 
  select(iso3c, product, tonnes, million_kcal, tonnes_protein) 

write_csv(crop_product_tonnes_humans, here::here("_efficiency", "data", "crops_protein_calories_humans_only.csv"))




```
