---
title: "aligning_EEZ"
author: "Juliette"
date: "4/14/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

This markdown aligns EEZs to our region list.


```{r}
library(tidyverse)
library(dplyr)
library(raster)
library(rgdal)
library(sf)
library(fasterize)
library(readr)
library(here)
library(mapview)
library(countrycode)

source(here::here("_workflow/common.R"))

eez <- st_read(file.path(prep, "spatial/eez_regions_2017_update.shp"))
food_rgns <- read_csv(here("_spatial/_output/food_rgns.csv"))
```


## Prepare eez raster
The first step is to merge the eez spatial data with the eez_food_rgns.csv.  Then rasterize using the ID_0 value.  This value is mostly the same as the ID_0 values in gadm36, however, there are a few mismatches:

* a couple eezs combine multiple regions (eventually we will want to separate these out)
Addioional ocean areas
* ID_0 10 is Antarctica oceans
* ID_0 355 is disputed eezs
* >=2600 FAO areas
* 400, 244 (US outlying islands), 82 (French Southern Territories) are all uninhabited islands
* 183251 Puerto Rico and US Virgin Islands
* 16893 Northern Mariana and Guam
* 92141 Guadaloupe and Martinique
* 49 or China includes Macao and Hong Kong

```{r}
## this was a list of of EEZ regions that we modified to match our food systems regions
eez_translate <- read_csv(here("_spatial/_output/eez_food_rgns_translate.csv"), col_types = "dccdc") %>%
  dplyr::select(rgn_id, iso3c, ID_0)
eez_rgns <- eez %>% 
  mutate(rgn_id = as.numeric(as.character(rgn_id))) %>%
  dplyr::select(rgn_typ, rgn_id) %>%
  filter(rgn_id %in% eez_translate$rgn_id) %>%
  left_join(eez_translate, by="rgn_id") %>%
  filter(rgn_typ %in% c("eez", "eez-disputed", "fao"))
summary(eez_rgns)
## check that everything matches as well as can be expected
# additional EEZs
setdiff(eez_rgns$iso3c, food_rgns$iso3c) # these should not have matches additional EEZS: [1] "uninhabited_island" # "PRI/VIR"            "UMI"                "MNP/GUM"            "GLP/MTQ"           
#  [6] "disputed"           "CHN/MAC/HKG"        "ATA"                "FAO"                "ATF" 
no_eez <- setdiff(food_rgns$iso3c, eez_rgns$iso3c)
landlocked <- read_csv(here("_spatial/_output/UNSD_Methodology.csv")) %>%
  filter(Land_Locked == "x") 
setdiff(no_eez, landlocked$iso3c)
# known problems"ALA" "CHN" "GLP" "GUM" "HKG" "IMN" "MAC" "MTQ" "MNP" "PSE" "PRI" "BLM" "SJM" "VIR"
```

Rasterize

```{r}
EEZ_raster <- fasterize(eez_rgns, food_raster, field="ID_0")
writeRaster(EEZ_raster, "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/spatial/eez_food_rgns.tif", overwrite=TRUE)
# save food eez regions file
eez_translate <- read_csv(here("_spatial/_output/eez_food_rgns_translate.csv"), col_types = "dccdc") %>%
  dplyr::select(ID_0, iso3c) %>%
  unique()
write_csv(eez_translate, here("_spatial/_output/eez_food_rgns.csv"))
```

