---
title: "Crop Disturbance"
output: html_document
editor_options: 
  chunk_output_type: console
---
## Paul-Eric Rayner

## NOTE: if updtes, rerun: crops/farm/step7...


# Script Description
# Calculate the area of each raster cell that is occupied by cropland.

# Preamble
```{r setup, include = FALSE}
# getting packages we want
library(tidyverse)
library(here)
library(raster)
library(sf)
library(janitor)
library("furrr")
library(vroom)

# Raster templates
source(here("_spatial/template_raster.R"))
source(here("_workflow/common.R")) 
source(here("_workflow/common_spatial.R"))

# Import food_rgns xy df
food_rgns_xy <- 
  food_rgns_xy %>% 
  select(x, y, iso3c)  
``` 

# Disturbance calculator
# Unit conversion: crop measured in hectare, cell is measured in km2. 1ha = 0.01 km2

```{r}
scaling_coef <- vroom::vroom(here("crop/farm/data/scaling_coef.csv"))

disturbance_calculator <- function(crop) { #crop="maiz"
  
crop_disturbance <-
    raster(file.path(prep,
                     "crop/farm/MapSPAM_correct_extent/spam2010V2r0_global_A_",
                     toupper(crop),
                     "_A.tif", 
                     fsep = "")) / 100
  
  # Rename layer
  names(crop_disturbance) <- paste("land_", crop, "_crop_produce_disturbance_km2eq", sep = "")
  
  # Save tif
  writeRaster(crop_disturbance, 
              file.path(prep, "crop/farm/stressors/land_", crop, "_crop_produce_disturbance", fsep = ""), 
              format    = "GTiff", 
              overwrite = TRUE)
}
plan(multisession, workers = 12)

options <- furrr_options(seed = 54232)

future_map(unique(scaling_coef$SPAM_short_name),
                  disturbance_calculator,
           .options = options)

delete_aux <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/crop/farm/stressors", pattern="aux", full=TRUE)
file.remove(delete_aux)
```

# Rest categories.

```{r}
rest_category_proportion <- 
  vroom(here("crop/farm/data/rest_category_proportion.csv")) %>% 
  select(!tonnes)
rest_disturbance_map <- 
  raster(file.path(prep, 
                   "crop/farm/stressors/land_rest_crop_produce_disturbance.tif", 
                   fsep = ""))
rest_disturbance <- function(crop) {
  crop_map <- 
    rest_category_proportion %>% 
    filter(SPAM_super %in% crop) %>% 
    left_join(food_rgns_xy, .) %>% 
    select(x, y, prop) %>% 
    rasterFromXYZ(., crs = food_crs)
    
  overlay(rest_disturbance_map,
          crop_map,
          fun = function(r1, r2) {
              
              return(r1 * r2)
              
              }) %>%
    
    writeRaster(.,
                file.path(prep, 
                          "crop/farm/stressors/land_",
                          crop,
                          "_crop_produce_disturbance.tif", 
                          fsep = ""),
                format = "GTiff", 
                overwrite = TRUE
                )
}
future_map(unique(rest_category_proportion$SPAM_super),
           rest_disturbance,
           .options = options)
r1 <- raster(file.path(prep,
                "crop/farm/stressors/land_tnut_crop_produce_disturbance.tif")) %>% 
  cellStats(., stat = "sum")
r2 <- raster(file.path(prep,
                "crop/farm/stressors/land_othr_crop_produce_disturbance.tif")) %>% 
  cellStats(., stat = "sum")
raster("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/crop/farm/stressors/land_othr_crop_produce_disturbance.tif")
r3 <- raster(file.path(prep,
                "crop/farm/stressors/land_spis_crop_produce_disturbance.tif")) %>% 
  cellStats(., stat = "sum")
r1 + r2 + r3
cellStats(rest_disturbance_map, stat = "sum")
```

Cotton fiber vs. seed correction.
```{r}

cotton_correction = 0.65

cotton_list <- 
  list.files(file.path(prep, "crop/farm/stressors"),
             pattern = "land_.*_crop_produce_disturbance.tif", 
             full.names = TRUE) %>% 
  .[grepl("cott_",.)]


file.copy(from = cotton_list,
to = gsub("_cott_", "_cott-seedlint_", cotton_list), overwrite=TRUE)


cotton_seedlint_list <- 
  list.files(file.path(prep, "crop/farm/stressors"),
             pattern = "disturbance", 
             full.names = TRUE) %>% 
  .[grepl("seedlint",.)]

for(crop in cotton_seedlint_list){ # crop = cotton_seedlint_list[2]
  data <- raster(crop) # raster::cellStats(data, "sum", na.rm=TRUE)
  data_correct <- data*cotton_correction # raster::cellStats(data_correct, "sum", na.rm=TRUE)
  writeRaster(data_correct, gsub("cott-seedlint", "cott", crop), overwrite=TRUE)
}

delete_aux <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/crop/farm/stressors", pattern="aux", full=TRUE)
file.remove(delete_aux)

cellStats(raster("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/crop/farm/stressors/land_cott-seedlint_crop_produce_disturbance.tif"),
          "sum", na.rm=TRUE)
cellStats(raster("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/crop/farm/stressors/land_cott_crop_produce_disturbance.tif"),
          "sum", na.rm=TRUE)


```



# Check number of files produced
```{r}
length(list.files(file.path(prep, "crop/disturbance/", fsep = ""))) # 42
```

# Map fodder crop disturbance
```{r}
# read in and convert from 1,000ha to km2
fodd_raster <- 
  overlay(raster(file.path(raw, "GAEZAct_2015/GAEZAct2015_HarvArea_Foddercrops_Irrigated.tif")) / 10,
          raster(file.path(raw, "GAEZAct_2015/GAEZAct2015_HarvArea_Foddercrops_Rainfed.tif"))   / 10,
          
          fun = function(r1, r2) {
             
               return(r1 + r2)
             
               }) 

names(fodd_raster) <- "land_fodd_crop_disturbance_km2eq"
  
writeRaster(fodd_raster, 
            file.path(prep, 
                      "crop/farm/stressors/land_fodd_crop_produce_disturbance", 
                      fsep = ""),
            format    = "GTiff",
            overwrite = TRUE)

# check output
file.path(prep,
          "crop/farm/stressors/land_fodd_crop_produce_disturbance.tif", 
          fsep = "") %>% 
  raster() %>% 
  plot()


delete_aux <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/crop/farm/stressors", pattern="aux", full=TRUE)
file.remove(delete_aux)

```


# checking
```{r}

tmp <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/crop/farm/MapSPAM_correct_extent", pattern = "_A.tif", full=TRUE)
tmp <- grep("_A_", tmp, value=TRUE)

raster_stack <- stack(tmp) %>%
  calc(., sum, na.rm = TRUE)

plot(raster_stack)
cellStats(raster_stack, "sum")

tmp <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/crop/farm/stressors", pattern = "_disturbance.tif", full=TRUE)

raster_stack <- stack(tmp) %>%
  calc(., sum, na.rm = TRUE)

cellStats(raster_stack, "sum") # https://www.nature.com/articles/s43016-021-00429-z 1,244 Mha

```