---
title: "Shrimp disturbance"
author: "Gage Clawson (NCEAS, OHI, UCSB)"
date: "01/05/2021"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r read in libaries and paths, eval = FALSE, include = FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
library(tidyverse)
library(raster)
library(data.table)
library(raster)
library(rgdal)
library(here)

path_final <- "/home/shares/food-systems/Food_footprint/all_food_systems/datalayers"

path_prep <- "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/shrimp/"

food_crs <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"

food_raster <- raster(nrows=2160, ncols=4320, xmn=-180, xmx=180, ymn=-90, ymx=90)

```

Assumptions: 

Source: http://www.fao.org/3/ac210e/AC210E02.htm#ch2

**Assuming Traditional or extensive culture**

stocking density = 3000-5000 fry per hectare: assume 4000 fry per hectare 

area: 1 - 10 hectares per pond recommended, actual sizes range from 3 - 20 hectares : assume 10 hectare ponds 

marketable size: 30 pieces/kg : Assuming that all "fry" grow into marketable size pieces 

Conversions: 

$ stocking \ density = \frac{400000 \ fry}{km2} \ *  \frac{1 kg }{30 \ pieces\ (or fry)} = \frac{13333.33 kg}{km^2} = \frac{0.0133 kg}{m^2} $ 

$ area = 0.10\  km^2 = 100000 \ m^2 $ 


Steps: 

1. calculate kg/pond:
   $ area \ m^2 * stocking\  density \frac{kg}{m^2} = kg/pond $
2. Calculate number of ponds:
   $ \frac{FAO \ production \ raster \ (kg)}{kg / pond} = ponds$

3. calculate surface area disturbance, assume ponds lie next to each other, with no buffer. Pond area will be 0.1 km2:

  $ disturbance \ (km^2) = cages*area $
  

## Method 

Establish the necessary parameters from above
```{r parameters}

stocking_density = (400000/30)/1000000
pond_area_km2 = 0.1
pond_area_m2 = 100000

```

###Calculate the total number of fish in the water (occupancy)
Read in production map

```{r}
production_raster <- rasterFromXYZ(read_csv(paste0(path_prep, "updated/shrimp_farm_updated.csv")), crs = food_crs) 

#plot(production_raster)
cellStats(production_raster, stat = "sum") # 5097823

```

```{r calculate shrimp occupancy}
# Assume all kg are live weight 
# as live weight of shrimp, in kg
shrimp_live_kg  <- production_raster*1000
shrimp_live_tonnes  <- production_raster

cellStats(shrimp_live_kg, "sum") # 5097822760
```


### Calculate the total area (from an aerial view) used for shrimp aquaculture

First we determine how many ponds are required given pond dimensions, stocking densities, and kg living fish. 

We used FAO data to estimate ~1333 kg shrimp (harvest weight) per pond.

This is based on ~100000 m2 pond.  We assume a stocking density of 0.013 kg/m2

```{r}

kg_per_pond <- pond_area_m2*stocking_density
t_per_pond <- kg_per_pond/1000

# check
(5097823/t_per_pond)/4 # 955841.8 if harvest happens 4 times a year


```

Given the kg of shrimp per cell we determine the number of ponds needed as:

We round this value up to the nearest whole number.

```{r}

ponds <- ceiling(shrimp_live_kg/kg_per_pond)
cellStats(ponds, "sum") # 3824733

```

2. We the determine how much surface area the ponds disturb, and area for infrastructure.

We will assume a 1.5 increase in area for infrastructure (buildings, drain/supply canals, etc.). 

Justification: 
Pond dimensions of 200 by 500 m2 seems reasonable based on figures. http://www.fao.org/3/ac210e/AC210E05.htm#ch5 (chapter 5 pond figures)

Then increase by 20 on all edges, so 240 * 540 = ~130000

So that is 30%, but then there is additional buildings, and I am sure there is disturbance beyond the canals…etc

```{r}
pond_disturbance_m2 <- pond_area_m2*1.5

total_disturbance_km2 <- (ponds * pond_disturbance_m2)/1000000
total_disturbance_prop <- total_disturbance_km2/raster::area(total_disturbance_km2)
plot(total_disturbance_prop)
plot(total_disturbance_km2)
cellStats(total_disturbance_km2, "sum") # 573709.9

shrimp_aquaculture_area_df <- as.data.frame(total_disturbance_prop, xy=TRUE) %>% 
  dplyr::rename(shrimp_disturbance_prop = layer )

write_csv(shrimp_aquaculture_area_df, file.path(path_prep, "shrimp_disturbance_prop.csv"))

shrimp_aquaculture_area_df_km2 <- as.data.frame(total_disturbance_km2, xy = TRUE) %>%
  dplyr::rename(shrimp_disturbance_km2 = layer)

write_csv(shrimp_aquaculture_area_df_km2, file.path(path_prep, "shrimp_disturbance_km2.csv"))


## save to datalayers folder with the correct naming conventions: 
final_layer_dist <- read_csv(file.path(path_prep, "shrimp_disturbance_km2.csv"))

final_rast <- rasterFromXYZ(final_layer_dist, crs = food_crs)
crs(final_rast)
plot(final_rast)
plot(log(final_rast + 1))
cellStats(final_rast, "sum") # 573709.9
final_rast
cellStats(production_raster, "sum") # 573709.9

final_area_fix <- final_rast/raster::area(final_rast)

final_rast_proj <- projectRaster(final_area_fix, food_raster, method = "ngb")

final_rast <- final_rast_proj*raster::area(final_rast_proj)
cellStats(final_rast, "sum") # 573709.9
final_rast

writeRaster(final_rast, file.path(path_final, "marine_shrimp_aquaculture_meat_disturbance.tif"), overwrite = TRUE)
```


```{r}
philippines <- raster::crop(total_disturbance_km2, extent(115, 130, 5, 20))
plot(philippines)

indo <- raster::crop(total_disturbance_km2, extent(95, 140, -15, 5))
plot(indo)


sri <- raster::crop(total_disturbance_km2, extent(79, 82, 6, 10))
plot(sri)
```

  

