---
title: "crop_water_separate.Rmd"
output: html_document
editor_options: 
  chunk_output_type: console
---

## NOTE: If this file is updated, run crop/farm/step7 again

# Script objective
Calculate and map green and blue WF of crops separately.
### NOTE: this is now serving as the final data layers! (vs. blue + green water use in previous iterations)


# Data description
WF data sourced from Mekonnen and Hoekstra 2011 Appendix 2 Table 2 [Accessed 02/09/2020]
Shapefiles for FIPS codes:http://www.fao.org/geonetwork/srv/en/resources.get?id=29036&fname=rwdb_ad2_py.z
ip&access=private [Accessed: 02/20/2020]


# Set-up
```{r setup, include = FALSE}

# library and package
library(here); library(janitor); library(fasterize); library(furrr); library(sf)
source(here("_spatial/template_raster.R"))
source(here("_workflow/common.R"))
source(here("_workflow/common_spatial.R"))

# Import food_rgns xy df
food_rgns_xy <- food_rgns_xy %>% dplyr::select(x, y, iso3c)

``` 

# Script-specific requisites
```{r}
## Water footprint data
wf_df <- vroom::vroom(here("crop/water/data/wf_df.csv"))

# A df through which to cycle functions
distinct_vars <- 
  expand.grid(
    unique(wf_df$split_rest_names),
     c("blue", "green"))

global_average <- read_csv(here("crop/water/data/global_average.csv"))

## Import shp files for FIPS codes
fips_shp <- 
  read_sf("/home/shares/food-systems/Food_footprint/_raw_data/GADM36/rwdb_ad2_py/RWDB_Ad2-Py.shp", 
          stringsAsFactors = FALSE) %>% 
  clean_names() %>% 
  select("geometry", "fips_code")
```

# Create water footprint maps
```{r}
wf_raster <- function(crop, colour) {

  # Filter by crop and wf_type
  wf_df_filtered <- 
    wf_df %>% 
    filter(split_rest_names %in% crop & 
           wf_type %in% colour) %>% 
    dplyr::select(fips_code, 
                  m3_per_ton) %>% 
    left_join(fips_shp, 
              ., 
              by = "fips_code") %>% 
    fasterize(., 
              food_raster, 
              field = "m3_per_ton")

  # Assign global average value to empty raster cells
  wf_df_filtered[is.na(wf_df_filtered)] <- 
    global_average %>% 
    filter(split_rest_names %in% crop &
           wf_type %in% colour) %>% 
    .$global_average

  # Overlay to calculate m3 per raster
  water_map <- 
    overlay(
      wf_df_filtered, 
      raster(file.path(prep,
                       "crop/farm/scaled_maps_2017/crop_",
                       crop,
                       "_A_scaled.tif", 
                       fsep = "")),
      fun = function(r1, r2) {
     
       return(r1 * r2)
     
       })

 # Change layer names
 names(water_map) <- "m3"
 
 # Save rasters
 if(colour=="blue"){
 writeRaster(water_map, 
             file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/crop/farm/stressors/land_",
                       crop,
                       "_crop_produce_water.tif", 
                       fsep = ""), 
             format    = "GTiff", 
             overwrite = TRUE)} else{
writeRaster(water_map, 
             file.path("/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers/land_",
                       crop,
                       "_crop_produce_",
                       colour,
                       "_water.tif", 
                       fsep = ""), 
             format    = "GTiff", 
             overwrite = TRUE)               
             }
}

map2(distinct_vars$Var1, 
     distinct_vars$Var2, 
     wf_raster)

delete_aux <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers", pattern="aux", full=TRUE)
file.remove(delete_aux)
delete_aux <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/crop/farm/stressors", pattern="aux", full=TRUE)
file.remove(delete_aux)


```

# Map fodder water consumption
```{r}
fodder_green_water <- 
  raster(file.path(prep,
                   "crop/farm/scaled_maps_2017/crop_fodd_rainfed_scaled.tif",
                   fsep = "")) * 
  207

fodder_blue_water <- 
  raster(file.path(prep, 
                   "crop/farm/scaled_maps_2017/crop_fodd_I_scaled.tif",
                   fsep = "")) * 
  27

# Change layer names
names(fodder_green_water) <- "m3"
names(fodder_blue_water)  <- "m3"
plot(fodder_green_water)
plot(fodder_blue_water)

writeRaster(fodder_green_water,
            "/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers/land_fodd_crop_produce_green_water.tif",
            format    = "GTiff",
            overwrite = TRUE)

writeRaster(fodder_blue_water,
            "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/crop/farm/stressors/land_fodd_crop_produce_water.tif", 
            format    = "GTiff",
            overwrite = TRUE)

delete_aux <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers", pattern="aux", full=TRUE)
file.remove(delete_aux)
delete_aux <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/crop/farm/stressors", pattern="aux", full=TRUE)
file.remove(delete_aux)

```


Exploring a cotton correction.
```{r}

cotton <- read_csv(here("crop/farm/data/FAOSTAT_cotton_seed_vs_lint_5-11-2022.csv")) %>%
  select(Area, Item, Value) %>%
  filter(!is.na(Value)) %>%
  filter(Value > 0) %>%
  pivot_wider(names_from="Item", values_from="Value") %>%
  rowwise() %>%
  mutate(percent_retained = (`Cotton lint` + Cottonseed)/(`Seed cotton`),
         percent_seed = Cottonseed/(`Cotton lint` + Cottonseed))

summary(cotton)
(0.65 * 0.12) / (0.35 * 0.80)

```

Cotton correction: by weight
```{r}

cotton_correction = 0.65


## correct the main data layer
cotton_list <- 
  list.files("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/crop/farm/stressors",
             pattern = "water", 
             full.names = TRUE) %>% 
  .[grepl("cott",.)]


file.copy(from = cotton_list,
to = gsub("_cott_", "_cott-seedlint_", cotton_list))


cotton_seedlint_list <- 
  list.files("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/crop/farm/stressors",
             pattern = "water", 
             full.names = TRUE) %>% 
  .[grepl("seedlint",.)]

for(crop in cotton_seedlint_list){ # crop = cotton_seedlint_list[2]
  data <- raster(crop) # raster::cellStats(data, "sum", na.rm=TRUE)
  data_correct <- data*cotton_correction # raster::cellStats(data_correct, "sum", na.rm=TRUE)
  writeRaster(data_correct, gsub("cott-seedlint", "cott", crop), overwrite=TRUE)
}

delete_aux <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/crop/farm/stressors", pattern="aux", full=TRUE)
file.remove(delete_aux)


cellStats(raster("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/crop/farm/stressors/land_cott_crop_produce_water.tif"),
          "sum", na.rm=TRUE)
cellStats(raster("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/crop/farm/stressors/land_cott-seedlint_crop_produce_water.tif"),
          "sum", na.rm=TRUE)



## correct green data
cotton_list <- 
  list.files("/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers",
             pattern = "water", 
             full.names = TRUE) %>% 
  .[grepl("cott_crop",.)]


file.copy(from = cotton_list,
to = gsub("_cott_", "_cott-seedlint_", cotton_list))


cotton_seedlint_list <- 
  list.files("/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers",
             pattern = "water", 
             full.names = TRUE) %>% 
  .[grepl("seedlint",.)]

for(crop in cotton_seedlint_list){ # crop = cotton_seedlint_list[2]
  data <- raster(crop) # raster::cellStats(data, "sum", na.rm=TRUE)
  data_correct <- data*cotton_correction # raster::cellStats(data_correct, "sum", na.rm=TRUE)
  writeRaster(data_correct, gsub("cott-seedlint", "cott", crop), overwrite=TRUE)
}

delete_aux <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers", pattern="aux", full=TRUE)
file.remove(delete_aux)


cellStats(raster("/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers/land_cott_crop_produce_green_water.tif"),
          "sum", na.rm=TRUE)
cellStats(raster("/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers/land_cott-seedlint_crop_produce_green_water.tif"),
          "sum", na.rm=TRUE)


```


# to here, label as all_water:
/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers


# move "blue_water" files from here: 
/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers

# to here, and change name: "land_?crop?_produce_water.tif
/home/shares/food-systems/Food_footprint/all_food_systems/datalayers

```
