---
title: "wf_df_mapping"
output: html_document
editor_options: 
  chunk_output_type: console
---
## NOTE: originally we conceived this stressor to be the sum of blue and green water use.
Now we provide blue and green water use separately and the pressure is based on blue water consumption.


# Script objective
Rasterize WF data.
Calculate and map WF of crops.

# Data description
WF data sourced from Mekonnen and Hoekstra 2011 Appendix 2 Table 2 [Accessed 02/09/2020]
Shapefiles for FIPS codes:http://www.fao.org/geonetwork/srv/en/resources.get?id=29036&fname=rwdb_ad2_py.z
ip&access=private [Accessed: 02/20/2020]

# Set-up
```{r setup, include = FALSE}

# library and package
library(here); library(janitor); library(fasterize); library(furrr)
source(here("_spatial/template_raster.R"))
source(here("_workflow/common.R"))
source(here("_workflow/common_spatial.R"))

# Import food_rgns xy df
food_rgns_xy <- food_rgns_xy %>% dplyr::select(x, y, iso3c)
``` 

# Load dependencies
```{r}
## Water footprint data
wf_df <- 
  read_csv(here("crop/water/data/wf_df.csv")) %>%
  group_by(fips_code, country, split_rest_names) %>%
  summarize(m3_per_ton = sum(m3_per_ton, na.rm=TRUE)) %>% 
  ungroup()

global_average <- 
  read_csv(here("crop/water/data/global_average.csv")) %>%
  group_by(split_rest_names) %>%
  summarize(global_average = sum(global_average, na.rm = TRUE)) %>% 
  ungroup()

## Import shp files for FIPS codes
fips_shp <- 
  read_sf("/home/shares/food-systems/Food_footprint/_raw_data/GADM36/rwdb_ad2_py/RWDB_Ad2-Py.shp", 
          stringsAsFactors = FALSE) %>% 
  clean_names() %>% 
  select("geometry", "fips_code")

# A df through which to cycle functions
distinct_vars <- unique(wf_df$split_rest_names)
```

# Create water footprint maps
```{r}
wf_raster <- function(crop) {

  # Filter by crop and wf_type
  wf_df_filtered <- 
    wf_df %>% 
    filter(split_rest_names %in% crop) %>% 
    dplyr::select(fips_code, 
                  m3_per_ton, 
                  split_rest_names) %>% 
    left_join(fips_shp, 
              ., 
              by = "fips_code") %>% 
    fasterize(., 
              food_raster, 
              field = "m3_per_ton")

  # Assign global average value to empty raster cells
  wf_df_filtered[is.na(wf_df_filtered)] <- 
    global_average %>% 
    filter(split_rest_names %in% crop) %>% 
    .$global_average

  # Overlay to calculate m3 per raster
  water_map <- 
    overlay(
      wf_df_filtered, 
      raster(file.path(prep,
                       "crop/farm/scaled_maps_2017/crop_",
                       crop,
                       "_A_scaled.tif", 
                       fsep = "")),
      fun = function(r1, r2) {
     
       return(r1 * r2)
     
       })

   # Change layer names
 names(water_map) <- "m3"
 
 # Save rasters
   writeRaster(water_map, 
               file.path(prep, 
                         "crop/farm/stressors/land_",
                         crop,
                         "_crop_produce_water.tif", 
                         fsep = ""), 
               format    = "GTiff", 
               overwrite = TRUE)
}
map(distinct_vars, wf_raster)
```

# Map fodder water consumption
```{r}
fodder_water <- 
  overlay(raster(file.path(prep,
                           "crop/farm/scaled_maps_2017/crop_fodd_rainfed_scaled.tif",
                           fsep = "")),
          raster(file.path(prep,
                           "crop/farm/scaled_maps_2017/crop_fodd_I_scaled.tif",
                           fsep = "")),
          fun = function(r1, r2) {
            
            return((r1 + r2) * (207 + 27))
            
            })

# Change layer names
names(fodder_water) <- "m3"

plot(fodder_water)

writeRaster(fodder_water,
            file.path(prep, "crop/farm/stressors/land_fodd_crop_produce_water.tif", fsep = ""), 
            format    = "GTiff",
            overwrite = TRUE)
```

```{r}
map(.x = list.files(file.path(prep, "crop/farm/stressors/"),
                    pattern = "water",
                    full.names = TRUE),
    ~ cellStats(raster(.x),
                stat = "sum"))
```

