---
title: "crop_pressures_rescaled"
author: "Juliette"
date: "6/10/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

Determine rescaled pressures for all crops (including: human food, animal feed, other).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(doParallel)
library(raster)

source(here("_workflow/common.R"))
source(here("_workflow/common_spatial.R"))

#This is normally what we would use, but need to use a slightly different approach to account for pressures
# lost to "other"
#rescale <- read_csv(here::here("_analysis/rescale_values.csv"))

```


Use land/eez raster to extract the summed pressures on land in eez areas.

```{r}
rgns_ezz_land <- raster(file.path(prep, "spatial/land_eez_rgns.tif"))
rgns_ezz_land <-calc(rgns_ezz_land, fun = function(x){ifelse(is.na(x), 9999, x)})

par(mar=c(1,1,1,1))
plot(rgns_ezz_land)

food_rgns <- food_rgns %>% 
  add_row(iso3c = "HSX", ID_0 = 9999, Country = "High Seas")
```

```{r}

crop_files_list <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/datalayers", full.names = TRUE, pattern = "crop_produce_")

crop_summary_df <- data.frame(country = NA,
                              iso3c = NA,
                              category = NA,
                              organism=NA,
                              system = NA, 
                              product = NA,
                              pressure = NA,
                              sum = NA)

for(file in crop_files_list){
  
#  file = crop_files_list[1]
  beginCluster(n=8) ## lets see if we can parallelize this 
  country_sum <- zonal(raster(file), rgns_ezz_land, fun="sum", progress="text", na.rm=TRUE) %>% 
    data.frame() %>%
    rename(ID_0 = zone) %>%
    left_join(food_rgns, by="ID_0")
  
  endCluster()
  
    cat <- str_split(str_split(file, pattern = "/")[[1]][8], pattern = "_")[[1]][1]
    org <- str_split(str_split(file, pattern = "/")[[1]][8], pattern = "_")[[1]][2]
    sys <- str_split(str_split(file, pattern = "/")[[1]][8], pattern = "_")[[1]][3]
    pro <- str_split(str_split(file, pattern = "/")[[1]][8], pattern = "_")[[1]][4]
    pre <- str_split(str_split(file, pattern = "/")[[1]][8], pattern = "_")[[1]][5]
    pre <- str_split(pre, pattern = ".tif")[[1]][1]
  
country_sum_clean <- country_sum %>% 
  mutate(category = cat,
          organism = org,
          system = sys,
          product = pro,
          pressure = pre) %>% 
  select(-ID_0) %>% 
  rename(country = Country) %>% 
  relocate(sum, .after = last_col()) 

  crop_summary_df <- rbind(crop_summary_df, country_sum_clean) %>% 
  filter(!is.na(product))
}

## make high seas regions 0
crop_summary_df <- crop_summary_df %>%
  mutate(sum= ifelse(iso3c=="HSX", 0, sum))

```

# correct pressure values for non-feed/food portion of crops
```{r}
# correction <- read_csv(here("feed/data/proportion_feed_and_other_per_country_system.csv"))
# 
# crop_summary_df_correct <- crop_summary_df %>%
#   mutate(iso3c_producing = iso3c,
#          SPAM_super = organism) %>%
#   left_join(correction, by=c("SPAM_super", "iso3c_producing")) %>%
#   filter(!is.na(total_non_human_prop)) %>%   #high seas
#   mutate(food_feed_prop = 1 - other_prop) %>%
#   mutate(sum_correct = sum*food_feed_prop) %>%
#   select(iso3c, organism, system, product, pressure, sum=sum_correct)
```

# rescale 
In this case, this includes all the crop (pressures are calculated without the other), so we must use a slightly different
rescaling value (pressures from entire crop including "other" portion)
```{r}

rescale <- read_csv(here("_analysis/checking_data/data/raster_check.csv")) %>%
  select(file, sum_cells) %>%
  filter(!(str_detect(file, "_x_"))) %>%
  separate(file, c("origin", "organism", "category", "product", "pressure"), sep="_") %>%
  group_by(pressure) %>%
  summarize(global_total = sum(sum_cells)) %>%
  mutate(pressure = gsub(".tif", "", pressure))
  
crop_pressures_rescaled <- crop_summary_df %>%
  left_join(rescale, by="pressure") %>% 
  mutate(prop_global = sum/global_total) %>% 
  group_by(iso3c, organism, system, product, pressure) %>% 
  dplyr::summarise(prop_global = sum(prop_global, na.rm = TRUE))

```

```{r}
write_csv(crop_pressures_rescaled, here::here("_analysis/figures/paper/data_prep/crops_rescaled.csv" ))

```

