---
title: "Marine finfish disturbance"
author: "Gage Clawson (UCSB, NCEAS, OHI)"
date: "12/29/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r read in libaries and paths}
library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
library(tidyverse)
library(raster)
library(data.table)
library(raster)
library(rgdal)
library(here)

path_final <- "/home/shares/food-systems/Food_footprint/all_food_systems/datalayers"

path_prep <- "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/marine_fish_general/"

food_crs <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"

food_raster <- raster(nrows=2160, ncols=4320, xmn=-180, xmx=180, ymn=-90, ymx=90)

```

Because there are so many species of finfish included in our analysis, we will treat them all the same (i.e. same stocking density, cage volume, etc.). Although some fish likely do rear for longer than 1 year, we will assume that most are 1 year or less, and that the loss is minimal. Meaning that the production reported from the FAO is the true live weight tonnes. 


Assumptions: 

stocking density = 20 kg/m3 (Froelich et al. 2017)

depth = 10 m (Froelich et al. 2017)

volume = 9000 m3

##Methods

Read in the appropriate values 
```{r parameters from conversion table}


stocking_density <- 20
cage_volume <- 9000
```

###Calculate the total number of fish in the water (occupancy)
Read in production map

```{r}
production_raster <- rasterFromXYZ(read_csv(paste0(path_prep, "updated/marine_fish_general_farm_updated.csv")), crs = food_crs) 

#plot(production_raster)
cellStats(production_raster, stat = "sum") # 3909255

```

```{r calculate finfish}
# Assume all kg are live weight 
# as live weight of finfish, in kg
finfish_live_kg  <- production_raster*1000
finfish_live_tonnes  <- production_raster

writeRaster(finfish_live_tonnes, file.path(path_prep, "finfish_occupancy/finfish_live_tonnes.tif"), format = "GTiff", overwrite = TRUE)
```


### Calculate the total area (from an aerial view) used for general finfish aquaculture

First we determine how many cages are required given cage dimensions, stocking densities, and kg living fish. 

We used Gentry 2017 data to estimate 180,000 kg finfish (harvest weight) per cage.

This is based on 9000 m3 cage that we assume to be circular with a depth of 10 m.  We assume a stocking density of 20kg/m3.

```{r}

kg_per_cage <- cage_volume*stocking_density

```

Given the kg of finfish per cell we determine the number of cages needed as:

kg_finfish/180,000

```{r}

cages <- ceiling(finfish_live_kg/kg_per_cage)
cellStats(cages,"sum") # 22701

```

We the determine how much surface area the cages disturb including the inner cage area, a 5 meter buffer around the cage, and area for infrastructure.

```{r}
cage_depth_m <- 10
buffer <- 5
infrastructure <- 1.5 #50% of area to infrastructure

cage_area <- cage_volume/cage_depth_m
cage_radius <- sqrt(cage_area/pi)
cage_radius_buffer <- cage_radius + buffer
cage_area_with_buffer <- pi*cage_radius_buffer^2
cage_disturbance_m2 <- cage_area_with_buffer*infrastructure

total_disturbance_km2 <- cages * cage_disturbance_m2/1000000
total_disturbance_prop <- total_disturbance_km2/raster::area(total_disturbance_km2)
plot(total_disturbance_prop)
plot(total_disturbance_km2)
cellStats(total_disturbance_km2, "sum") # 51.42716

finfish_aquaculture_area_df <- as.data.frame(total_disturbance_prop, xy=TRUE) %>% 
  dplyr::rename(finfish_disturbance_prop = layer )

write_csv(finfish_aquaculture_area_df, file.path(path_prep, "marine_fish_general_disturbance_prop.csv"))

finfish_aquaculture_area_df_km2 <- as.data.frame(total_disturbance_km2, xy=TRUE) %>% 
  dplyr::rename(finfish_disturbance_km2 = layer)

write_csv(finfish_aquaculture_area_df_km2, file.path(path_prep, "marine_fish_general_disturbance_km2.csv"))

## save to datalayers folder with the correct naming conventions: 
final_layer_dist <- read_csv(file.path(path_prep, "marine_fish_general_disturbance_km2.csv"))

final_rast <- rasterFromXYZ(final_layer_dist, crs = food_crs)
crs(final_rast)
plot(final_rast)
cellStats(final_rast, "sum") # 51.42716
final_rast

final_area_fix <- final_rast/raster::area(final_rast)

final_rast_proj <- projectRaster(final_area_fix, food_raster, method = "ngb")

final_rast <- final_rast_proj*raster::area(final_rast_proj)
cellStats(final_rast, "sum") # 51.42716
final_rast

writeRaster(final_rast, file.path(path_final, "marine_marine-fish-general_aquaculture_meat_disturbance.tif"), overwrite = TRUE)
```


Look at the pieces more closely
```{r}
norway <- raster::crop(total_disturbance_km2, extent(-25, 50, 50, 75))
plot(norway)
writeRaster(norway, file.path(path_prep, "finfish_occupancy/norway.tif"), overwrite=TRUE)

```


