---
title: "Create production rasters and extract per food raster"
author: "Gage Clawson (UCSB, NCEAS, OHI)"
date: "May 13, 2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

Here we need to take the raw watson production data and rasterize it (save those files, for each species class, and for each species). Then we will extract the production values from the rasters using our food systems and zonal() with file.path(prep, "spatial/land_eez_rgns.tif")), so that we avoid any weird mismatches between our pressure rasters and production reporting. (The output of this will really only be used for the efficiency calculations).

```{r, eval = FALSE}
#load relevant packages, etc.

library(here)
library(raster)
library(tidyverse)
library(doParallel)
library(foreach)
library(parallel)

# raster template
r <- raster::raster(ncol=720, nrow=360, vals=c(1:259200))
food_raster <- raster(nrows=2160, ncols=4320, xmn=-180, xmx=180, ymn=-90, ymx=90)

prep <- file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep")


source(here("_workflow/common_spatial.R"))

food_rgns <- food_rgns %>% 
  add_row(iso3c = "HSX", ID_0 = 9999, Country = "High Seas")


cat_msg <- function(x, ...) {
  if(is.null(knitr:::.knitEnv$input.dir)) {
    ### not in knitr environment, so use cat()
    cat(x, ..., '\n')
  } else {
    ### in knitr env, so use message()
    message(x, ...)
  }
  return(invisible(NULL))
}

```

Save catch rasters for each species class (IUU + Reported)
```{r, eval = FALSE}

## filter this for 2017... our assessment year 
catch_full_final <- read_csv("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/disturbance/catch_spp_class_2017.csv") 

catch_summary <- catch_full_final %>%
  mutate(catch = ReportedIND + IUUIND + ReportedNIND + IUUNIND)  %>%
  group_by(Cell, species_class_fin) %>%
  summarize(catch = sum(catch, na.rm=TRUE)) %>%
  ungroup() 

check <- catch_summary %>%
  group_by(species_class_fin) %>%
  summarise(check = sum(catch, na.rm = TRUE))

## get species, description, taxon ids
catch_cats <- catch_summary %>%
  dplyr::select(species_class_fin) %>%
  distinct() %>%
  mutate(species_class_fix = case_when(
    species_class_fin == "Demersal" ~ "demersal",
    species_class_fin == "Small pelagic" ~ "small-pelagic", 
    species_class_fin == "Medium pelagic" ~ "medium-pelagic", 
    species_class_fin == "Large pelagic" ~ "large-pelagic", 
    species_class_fin == "Reef-associated" ~ "reef", 
    species_class_fin == "Benthic" ~ "benthic", 
    species_class_fin == "forage_fish" ~ "fofm"
  ))

reload <- FALSE ## change this to true if you want to rerun all species

## rasterize and extract catch using zonal
for(i in 1:nrow(catch_cats)){
  
  # i = 1

  ## get the ids for this i  
  spp_class_id <- catch_cats$species_class_fin[i]
  
  spp_class_id_fix <- catch_cats$species_class_fix[i]
  
  ## this will be the final file name
  spp_final_file <- file.path(sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/production_rasters/tonnes_%s.tif", spp_class_id))
  
  spp_predatalayers_file <- file.path(sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers/marine_%s_fisheries_meat_tonnes_production.tif", spp_class_id_fix))
    
if(!file.exists(spp_final_file) | reload) { ## if the file already exists, skip over it (unless otherwise specified)
  
## filter for correct observations  
catch_summary_i <- catch_summary %>%
  filter(species_class_fin == spp_class_id) %>%
  group_by(Cell) %>%
  summarise(catch_final = sum(catch, na.rm = TRUE)) %>%
  ungroup()

print(sum(catch_summary_i$catch_final)) # 45088617

## rasterize
tonnes_raster <- raster::subs(r, catch_summary_i, by = "Cell", which = "catch_final", subsWithNA=TRUE)
# plot(log(tonnes_raster+1))

# tonnes_raster

print(cellStats(tonnes_raster, "sum")) # 45088617


### DIVIDE BY AREA HERE for reprojection
tonnes_raster_fix <- tonnes_raster/raster::area(tonnes_raster)
# plot(tonnes_raster_fix)

## reproject
tonnes_raster_proj <- projectRaster(tonnes_raster_fix, food_raster, method="ngb")
# tonnes_raster_proj

### MULTIPLY BY AREA HERE for reprojection
tonnes_raster_proj_final <- tonnes_raster_proj*raster::area(tonnes_raster_proj)
# plot(log(tonnes_raster_proj_final + 1))
print(cellStats(tonnes_raster_proj_final, "sum", na.rm = TRUE)) # 45088512 ### close enough

# save raster
writeRaster(tonnes_raster_proj_final, spp_final_file,
    overwrite=TRUE)

writeRaster(tonnes_raster_proj_final, spp_predatalayers_file,
    overwrite=TRUE)

cat_msg('Finished ', i, ' out of ', nrow(catch_cats), 'species')

} else {
      
      cat_msg('species file ', spp_final_file, ' already exists; skipping these spp')
      
    }
}
```

Save and extract catch rasters for each individual species of fish
```{r, eval = FALSE}

catch_summary <- catch_full_final %>%
  mutate(catch = ReportedIND + IUUIND + ReportedNIND + IUUNIND)  %>%
  group_by(Cell, species_class_fin, Descript, TaxonName) %>%
  summarize(catch = sum(catch, na.rm=TRUE)) %>%
  ungroup() %>%
  mutate(TaxonName = ifelse(TaxonName == "Homaridae/Palinuridae", "Homaridae_Palinuridae", TaxonName))

check <- catch_summary %>%
  group_by(species_class_fin) %>%
  summarise(check = sum(catch, na.rm = TRUE))


test <- catch_summary %>%
  filter(str_detect(TaxonName, "/"))

## get species, description, taxon ids
catch_cats <- catch_summary %>%
  dplyr::select(species_class_fin, Descript, TaxonName) %>%
  distinct()

reload <- FALSE ## change this to true if you want to rerun all species

## rasterize and extract catch using zonal
for(i in 1:nrow(catch_cats)){
  
  # i = 364

  ## get the ids for this i  
  spp_class_id <- catch_cats$species_class_fin[i]
  descript_id <- catch_cats$Descript[i]
  taxon_id <- catch_cats$TaxonName[i]
  
  ## this will be the final file name
  spp_final_file <- file.path(sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/production_csvs/%s/tonnes_%s_%s_%s.csv", spp_class_id, spp_class_id, descript_id, taxon_id))
    
if(!file.exists(spp_final_file) | reload) { ## if the file already exists, skip over it (unless otherwise specified)
  
## filter for correct observations  
catch_summary_i <- catch_summary %>%
  filter(species_class_fin == spp_class_id,
         Descript == descript_id,
         TaxonName == taxon_id) %>%
  group_by(Cell) %>%
  summarise(catch_final = sum(catch, na.rm = TRUE)) %>%
  ungroup()

## rasterize
tonnes_raster <- raster::subs(r, catch_summary_i, by = "Cell", which = "catch_final", subsWithNA=TRUE)
# plot(log(tonnes_raster+1))

# tonnes_raster

print(cellStats(tonnes_raster, "sum")) # 8187537


### DIVIDE BY AREA HERE for reprojection
tonnes_raster_fix <- tonnes_raster/raster::area(tonnes_raster)
# plot(tonnes_raster_fix)
# cellStats(tonnes_raster_fix, "sum", na.rm = TRUE) # 2832.451

## reproject
tonnes_raster_proj <- projectRaster(tonnes_raster_fix, food_raster, method="ngb")
# tonnes_raster_proj
# print(cellStats(tonnes_raster_proj, "sum")) # 101968.2

### MULTIPLY BY AREA HERE for reprojection
tonnes_raster_proj_final <- tonnes_raster_proj*raster::area(tonnes_raster_proj)
# plot(log(tonnes_raster_proj_final + 1))
print(cellStats(tonnes_raster_proj_final, "sum", na.rm = TRUE)) # 8187512 ### close enough

# save raster
writeRaster(tonnes_raster_proj_final, sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/production_rasters/%s/tonnes_%s_%s_%s.tif", spp_class_id, spp_class_id, descript_id, taxon_id),
    overwrite=TRUE)

# read in spatial file
land_eez_rgns <- raster(file.path(prep, "spatial/land_eez_rgns.tif"))

## tag high seas areas
land_eez_rgns_hsx <- calc(land_eez_rgns, fun = function(x){ifelse(is.na(x), 9999, x)})

# extract with zonal
tonnes_df <- as.data.frame(raster::zonal(tonnes_raster_proj_final, land_eez_rgns_hsx, fun = "sum"))

## reattach species id tags
tonnes_df_species <- tonnes_df %>%
  mutate(species_class_fin = spp_class_id,
         Descript = descript_id, 
         TaxonName = taxon_id) %>%
  left_join(food_rgns, by = c("zone" = "ID_0")) %>%
  rename("catch" = "sum", rgn_id = "zone")

print(sum(tonnes_df_species$catch)) # 8187512 a little less 

# save dataset
write.csv(tonnes_df_species, sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/production_csvs/%s/tonnes_%s_%s_%s.csv", spp_class_id, spp_class_id, descript_id, taxon_id), row.names = FALSE)

cat_msg('Finished ', i, ' out of ', nrow(catch_cats), 'species', "species_name", taxon_id)

} else {
      
      cat_msg('species file ', spp_final_file, ' already exists; skipping these spp')
      
    }
}

## compile all the files together
catch_files <- list.files(file.path(prep, 'fisheries/marine/production_csvs'), 
                                    pattern = 'tonnes_', 
                                    full.names = TRUE,
                          recursive = TRUE)

## read in the csvs and bind them all together
all_catch <- lapply(catch_files, FUN = function(x) {
    read.csv(x)}) %>%
    bind_rows() %>%
  mutate(TaxonName = ifelse(TaxonName == "Homaridae_Palinuridae", "Homaridae/Palinuridae", TaxonName)) ## re-fix the fix earlier 

sum(all_catch$catch) # 111225909

45088617 +1021491 + 39054068 + 7874482 + 3621795 + 8187512 + 6378153 # 111226118 nearly perfect 

## save final csv
write.csv(all_catch, file.path(prep, "fisheries/marine/catch_extracted_final.csv"), row.names = FALSE)
```

## FOFM human consumption
Calculate the amount of forage fish catch that goes to human consumption, not feed. 
```{r}
## read in total FOFM tonnage raster
fofm_tonnes <- raster("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/production_rasters/tonnes_forage_fish.tif")
cellStats(fofm_tonnes, "sum") # 39053973
  
## read in raster for human consumption of fofm
human <- raster("/home/shares/food-systems/Food_footprint/all_food_systems/datalayers/land_human_human_human_x_marine_fofm_fisheries_meat.tif")

## Multiply human consumption raster here
fofm_human_tonnes <- fofm_tonnes*human
cellStats(fofm_human_tonnes, "sum") # 27381757

# save the raster
writeRaster(fofm_human_tonnes, "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/production_rasters/tonnes_fofm_human.tif", overwrite=TRUE)


land_eez_rgns <- raster(file.path(prep, "spatial/land_eez_rgns.tif"))

land_eez_rgns_hsx <- calc(land_eez_rgns, fun = function(x){ifelse(is.na(x), 9999, x)})

## extract per each region
tonnes_df <- as.data.frame(raster::zonal(fofm_human_tonnes, land_eez_rgns_hsx, fun = "sum"))


tonnes_df_species <- tonnes_df %>%
  mutate(species_class_fin = "forage_fish") %>%
  left_join(food_rgns, by = c("zone" = "ID_0")) %>%
  rename("catch" = "sum", rgn_id = "zone")

print(sum(tonnes_df_species$catch)) # 27381757 perfect

## save the csv
write.csv(tonnes_df_species, "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/catch_human_x_fofm.csv", row.names = FALSE)

```




