---
title: "Untitled"
output: html_document
---

Citation: Bouwman, A. F., A. H. W. Beusen, and G. Billen (2009), Human alteration of the global nitrogen and phosphorus soil balances for the period 1970 – 2050, Global Biogeochem. Cycles, 23, GB0A04, doi:10.1029/2009GB003576.

```{r}
library(here)
library(vroom)
library(janitor)
source(here("_spatial/template_raster.R"))
source(here("_workflow/common.R"))
source(here("_workflow/common_spatial.R"))

# Import food_rgns xy df
food_rgns_xy <- 
  food_rgns_xy %>% 
  dplyr::select(x, y, iso3c)
```

We need to convert table 2 “Overall efficiency of N and P Use of the Total Agricultural System" to a dataframe with all our iso3c regions and gapfilling any missing regions.

# Browse regions
```{r}
regions <- 
  vroom::vroom(here("_spatial/_output/UNSD_Methodology.csv"), 
               .name_repair = make_clean_names)
unique(regions$intermediate_region_name)
```

# Wrangle dataframe with iso3c according to each 
```{r}
# Add UN `region` column to dataframe
bouwman_2009_regions <- 
  vroom::vroom(here("_spatial/_output/UNSD_Methodology.csv"), 
               .name_repair = make_clean_names) %>%
  mutate(IMAGE_region = 
           case_when(sub_region_name %in% "Northern America" ~ "north_america",
                     region_name     %in% "Africa" ~ "africa",
                     region_name     %in% "Oceania" ~ "oceania",
                     region          %in% c("Russian Federation", "Belarus", 
                                            "Ukraine", "Republic of Moldova") ~ 
                       "north_asia",
                     region_name     %in% "Europe" ~ "europe",
                     region_name     %in% "Asia" ~ "south_asia",
                     sub_region_name %in% "Latin America and the Caribbean" ~ 
                       "central_and_south_america")) %>% 
  select(sub_region_name, region_name, region, IMAGE_region, iso3c) %>% 
  left_join(food_rgns, ., by = "iso3c") %>% 
  select(iso3c, IMAGE_region)
```

# Import nutrient dataset table from Bouwman 2009
```{r}
bouwman_2009_tbl2 <- 
  vroom(here("crop/nutrient/data/bouwman_2009_tbl2_N_raw.csv"),
             .name_repair = make_clean_names) %>% 
  mutate(nutrient = "N") %>% 
  bind_rows(., vroom(here("crop/nutrient/data/bouwman_2009_tbl2_P_raw.csv"),
                 .name_repair = make_clean_names)) %>% 
  mutate(nutrient = if_else(is.na(nutrient), 
                            "P",
                            nutrient)) %>% 
  pivot_longer(data = .,
               cols = !c(year_scenario, nutrient),
               names_to = "IMAGE_region", 
               values_to = "efficiency") %>% 
  mutate(year_scenario = case_when(
    year_scenario %in% c("2030-GO", 
                         "2030-AM",
                         "2030-OS",
                         "2030-TG") ~ "2030",
    TRUE ~ year_scenario)) %>% 
  filter(year_scenario %in% c("2000", "2030")) %>%
  group_by(year_scenario, IMAGE_region, nutrient) %>%
  summarise(efficiency = mean(efficiency)) %>% 
  ungroup() %>% 
  mutate(efficiency = if_else(year_scenario == 2000,
                              efficiency * (12.5 / 31),
                              efficiency * (18.5 / 31))) %>% 
  group_by(IMAGE_region, nutrient) %>% 
  summarise(efficiency = sum(efficiency)) %>% 
  ungroup()
```

# Which data is most representative of 2017?
we only have 2000 data and projected data for 2030. We could potentially find some
middle ground value?

```{r}
# Rasterize the change in values between 1970, 2000 and 2030
bouwman_2009_tbl2_tidy <- 
  left_join(bouwman_2009_tbl2, 
            bouwman_2009_regions, 
            by = "IMAGE_region") %>% 
  select(!IMAGE_region)

write_csv(bouwman_2009_tbl2_tidy,
         here("crop/nutrient/data/bouwman_2009_tbl2.csv"))
```

# I visually compared the numbers to see if they were in the right ballpark.
```{r}
view(bouwman_2009_tbl2)
view(vroom::vroom(here("crop/nutrient/data/bouwman_2009_tbl2_N_raw.csv"),
                  .name_repair = make_clean_names))
```

-------------------------------------------------------------------------------

# Wrangle dataframe with iso3c according to each 
```{r}
# Add UN `region` column to dataframe
bouwman_2005_regions <- 
  vroom::vroom(here("_spatial/_output/UNSD_Methodology.csv"), 
               .name_repair = make_clean_names) %>%
  mutate(IMAGE_region = 
           case_when(sub_region_name %in% "Northern America" ~ "north_america",
                     region %in% c("Albania", "Bosnia and Herzegovina",
                                   "Montenegro", "Serbia", "Macedonia", 
                                   "Armenia", "Azerbaijan", "Belarus", 
                                   "Georgia", "Kazakhstan", "Kyrgyzstan", 
                                   "Republic of Moldova", "Russian Federation",
                                   "Tajikistan", "Turkmenistan", "Ukraine",
                                   "Uzbekistan") ~ "transition_countries",
                     sub_region_name %in% "Northern Africa" ~ "north_africa",
                     sub_region_name %in% "Sub-Saharan Africa" ~ "subsaharan_africa",
                     intermediate_region_name %in% "Western Europe" ~ "western_europe",
                     sub_region_name %in% "South-eastern Asia" ~ "southeast_asia",
                     sub_region_name %in% "Southern Asia" ~ "south_asia",
                     sub_region_name %in% "Eastern Asia" ~ "east_asia",
                     sub_region_name %in% "Latin America and the Caribbean" ~ "latin_america",
                     intermediate_region_name %in% "Western Asia"  ~ "middle_east",
                     developed_developing %in% "Developed" ~ "industrialized",
                     developed_developing %in% "Developing" ~ "developing",
                     iso3c %in% "IRN" ~ "middle_east",
                     TRUE ~ "world")) %>% 
  select(sub_region_name, region_name, region, IMAGE_region, iso3c) %>% 
  left_join(food_rgns, ., by = "iso3c") %>% 
  select(iso3c, IMAGE_region, region)
```

```{r}
bouwman_2005_tbl5 <- 
  tibble(IMAGE_region = rep(c("north_america",
                              "western_europe",
                              "transition_countries",
                              "latin_america",
                              "middle_east",
                              "north_africa",
                              "subsaharan_africa",
                              "south_asia",
                              "east_asia",
                              "southeast_asia",
                              "world",
                              "developing",
                              "industrialized"),
                              2),
         year = c(rep("1995", 13), 
                  rep("2030", 13)),
         NUE  = c(48, 54, 67, 49, 58, 58, 108, 41, 48, 78, 52, 51, 49, 
                  63, 68, 83, 66, 63, 63, 131, 58, 42, 90, 61, 58, 62),
         system_n_recovery = c(43, 49, 46, 43, 47, 47, 51, 31, 43, 55, 43, 42, 45,
                               51, 58, 55, 49, 50, 50, 58, 38, 39, 61, 47, 45, 52))

bouwman_2005_tbl5_final <- 
  bouwman_2005_tbl5 %>% 
  mutate(NUE = if_else(year == 1995,
                       NUE * (12.5 / 36),
                       NUE * (23.5 / 36)),
         system_n_recovery = if_else(year == 1995,
                                     system_n_recovery * (12.5 / 36),
                                     system_n_recovery * (23.5 / 36))) %>% 
  group_by(IMAGE_region) %>% 
  summarise(NUE = sum(NUE),
            system_n_recovery = sum(system_n_recovery)) %>% 
  ungroup() %>% 
  left_join(bouwman_2005_regions, .) %>% 
  na.omit()

write_csv(bouwman_2005_tbl5_final,
          here("crop/nutrient/data/bouwman_2005_tbl5.csv"))
```

-------------------------------------------------------------------------------

Boumwan et al 2002 - 
"Estimation of global NH3 volatilization loss from synthetic fertilizers and 
animal manure applied to arable lands and grasslands

Table 4 from here (all three categories, columns "N use, kt" and "NH3-N loss, kt"):

```{r}
bouwman_2002_regions <- 
  vroom::vroom(here("_spatial/_output/UNSD_Methodology.csv"), 
               .name_repair = make_clean_names) %>%
  mutate(IMAGE_region = 
           case_when(region %in% c("Albania", "Bosnia and Herzegovina",
                                   "Montenegro", "Serbia", "Macedonia", 
                                   "Armenia", "Azerbaijan", "Belarus", 
                                   "Georgia", "Kazakhstan", "Kyrgyzstan", 
                                   "Republic of Moldova", "Russian Federation",
                                   "Tajikistan", "Turkmenistan", "Ukraine",
                                   "Uzbekistan") ~ "former_ussr",
                     iso3c %in% c("GRL", "XMI") ~ "oecd_europe",
                     sub_region_name %in% "Western Europe" ~ "oecd_europe",
                     sub_region_name %in% "Northern Europe" ~ "oecd_europe",
                     sub_region_name %in% "Southern Europe" ~ "oecd_europe",
                     sub_region_name %in% "Eastern Europe" ~ "eastern_europe",
                     iso3c %in% "SDN" ~ "eastern_africa",
                     iso3c %in% c("TZA", "AGO", "ZMB", "ZWE", "MWI", "MOZ") ~ "southern_africa",
                     intermediate_region_name %in% "Northern Africa" ~ "north_africa",
                     intermediate_region_name %in% "Western Africa" ~ "western_africa",
                     intermediate_region_name %in% "Eastern Africa" ~ "eastern_africa",
                     intermediate_region_name %in% "Southern Africa" ~ "southern_africa",
                     intermediate_region_name %in% "Middle Africa" ~ "western_africa",
                     iso3c %in% c("USA", "SPM") ~ "usa",
                     region %in% "Canada" ~ "canada",
                     iso3c %in% "BMU" ~ "central_america",
                     intermediate_region_name %in% "Caribbean" ~ "central_america",
                     intermediate_region_name %in% "Central America" ~ "central_america",
                     intermediate_region_name %in% "South America" ~ "south_america",
                     region %in% "Japan" ~ "japan",
                     iso3c %in% "IRN" ~ "middle_east",
                     intermediate_region_name %in% "Western Asia"  ~ "middle_east",
                     sub_region_name %in% "South-eastern Asia" ~ "southeast_asia",
                     sub_region_name %in% "Southern Asia" ~ "south_asia",
                     sub_region_name %in% "Eastern Asia" ~ "east_asia",
                     region_name %in% "Oceania" ~ "oceania")) %>% 
  select(sub_region_name, region_name, region, IMAGE_region, iso3c) %>% 
  left_join(food_rgns, ., by = "iso3c") %>% 
  select(iso3c, IMAGE_region, region)
```

# Merge iso3c with regions in the table
```{r}
bouwman_2002_tbl4 <- 
  vroom(here("crop/nutrient/data/bouwman_2002_tbl4_raw.csv"),
        .name_repair = make_clean_names)[c(1,9,10)] %>% 
  rename(IMAGE_region = region)
bouwman_2002_tbl4$IMAGE_region <- make_clean_names(bouwman_2002_tbl4$IMAGE_region)

bouwman_2002_tbl4_tidy <- 
  left_join(bouwman_2002_tbl4,
          bouwman_2002_regions,
          by = "IMAGE_region") %>% 
  na.omit()

write_csv(bouwman_2002_tbl4_tidy,
            here("crop/nutrient/data/bouwman_2002_tbl4.csv"))
```

# grasslands
```{r}
bouwman_2002_tbl4_grasslands <- 
  vroom(here("crop/nutrient/data/bouwman_2002_tbl4_raw.csv"),
        .name_repair = make_clean_names)[c(1:3)] %>% 
  mutate(prop_nh3_total_grasslands = nh3_n_grassland_loss / n_use_fertilized_grasslands_kt) %>% 
  mutate(prop_nh3_total_grasslands = if_else(is.finite(prop_nh3_total_grasslands),
                                             prop_nh3_total_grasslands,
                                             0)) %>% 
  rename(IMAGE_region = region) %>% 
  select(IMAGE_region, prop_nh3_total_grasslands)

bouwman_2002_tbl4_grasslands$IMAGE_region <- 
  make_clean_names(bouwman_2002_tbl4_grasslands$IMAGE_region)

bouwman_2002_tbl4_grasslands_tidy <- 
  left_join(bouwman_2002_tbl4_grasslands,
            bouwman_2002_regions,
            by = "IMAGE_region") %>% 
  na.omit()

write_csv(bouwman_2002_tbl4_grasslands_tidy,
            here("crop/nutrient/data/bouwman_2002_tbl4_grasslands.csv"))
```


# Merge iso3c with regions in the table
```{r}
bouwman_2002_tbl5 <- 
  vroom(here("crop/nutrient/data/bouwman_2002_tbl5_raw.csv"),
        .name_repair = make_clean_names,
        skip = 1)[c(1,9,10,11,14)] %>% 
  rename(IMAGE_region = region)
bouwman_2002_tbl5$IMAGE_region <- make_clean_names(bouwman_2002_tbl5$IMAGE_region)

bouwman_2002_tbl5_tidy <- 
  left_join(bouwman_2002_tbl5,
            bouwman_2002_regions,
            by = "IMAGE_region")[2:6] %>% 
  na.omit()

write_csv(bouwman_2002_tbl5_tidy,
            here("crop/nutrient/data/bouwman_2002_tbl5.csv"))
```