---
title: "fao_nutrient_wrangling"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Script Description
# Clean FAO data describing nutrient application in each country.
# Data information
# FAOSTAT category selection: 2017 | Fertilizer by Nutrient | Agricultural use | All countries
# Extracted: 11/4/2020

# Preamble
```{r setup, include = FALSE}
# getting packages we want
library(here)
library(janitor)
library(vroom)
library(cowplot)

# Raster templates
source(here("_spatial/template_raster.R"))
source(here("_workflow/common.R"))
source(here("_workflow/common_spatial.R"))

# Import food_rgns xy df
food_rgns_xy <- 
  food_rgns_xy %>% 
  select(x, y, iso3c)
``` 

## Integrate iso3c codes
```{r}
# Import country-level fertilizer data
fao_nutrient <- 
  vroom::vroom(here("crop/nutrient/data/FAOSTAT_2017_nutrient_agri_use.csv"), 
               .name_repair = make_clean_names) %>%  
  rename(nutrient = item, country_application = value) %>% 
  mutate(nutrient = recode(nutrient,
                           "Nutrient nitrogen N (total)"     = "N",
                           "Nutrient phosphate P2O5 (total)" = "P2O5",
                           "Nutrient potash K2O (total)"     = "K2O"))


```

# Match FAOSTAT regions with food_rgns list
```{r}
# Remove China and re-allocate Serbia data to Kosovo and Serbia  
fao_nutrient <- 
  fao_nutrient %>%   # SRB without XKO
  mutate(area_code = if_else(area_code %in% 272, 275, area_code)) %>%  # XKO
  bind_rows(fao_nutrient) %>%
  mutate(area_code = if_else(area_code %in% 272, 286, area_code)) %>%
  unique() %>% 
  filter(!area_code %in% 351) # Code 351 values are for all of China, including MAC, HKG, TWN.

# Global tonnes of each fertilizer
sum_before_wrangling <- 
  fao_nutrient %>% 
  group_by(nutrient) %>% 
  summarize(global_tonnes = sum(country_application))
sum_before_wrangling
```
  
# Add iso3c column and split serbia data into serbia and kosovo
```{r}  
# Convert fao rgns to iso3c
fao_rgn_codes <- vroom::vroom(here("_spatial/_output/FAO_rgn_codes.csv"))

fao_nutrient <- 
  left_join(fao_nutrient, fao_rgn_codes, by = "area_code") %>% 
  dplyr::select(nutrient, country_application, iso3c) %>% 
  unique()

# Allocate fertilizer data to Serbia and Kosovo
SRB_XKO <- 
  vroom::vroom(here("crop/farm/data/prod_crop_rgns_2017.csv")) %>% 
  filter(iso3c %in% c("XKO", "SRB") & prod_system %in% "A") %>% 
  group_by(iso3c) %>% 
  summarize(production = sum(production)) %>% 
  ungroup() %>% 
  mutate(relative_proportion = production / sum(production))

fao_nutrient$country_application[fao_nutrient$iso3c %in% "SRB"] <- 
  fao_nutrient$country_application[fao_nutrient$iso3c %in% "SRB"] * 
  SRB_XKO$relative_proportion[SRB_XKO$iso3c %in% "SRB"]

fao_nutrient$country_application[fao_nutrient$iso3c %in% "XKO"] <- 
  fao_nutrient$country_application[fao_nutrient$iso3c %in% "XKO"] * 
  SRB_XKO$relative_proportion[SRB_XKO$iso3c %in% "XKO"]
```

## Fill data for missing_rgns
```{r}
# Extract iso3c codes not included in fao_nutrient dataset
missing_rgns <- setdiff(unique(food_rgns$iso3c), unique(fao_nutrient$iso3c))
duplicated(fao_rgn_codes$area_code)
length(missing_rgns) # 72 missing rgns

# I gave these autonomous regions the same area_codes as POR and GBR respectively.
fao_nutrient$country_application[fao_nutrient$iso3c %in% c("XMI", "GGY")] <- 0

# Calculate global tonnes of each fertilizer
sum_before_allocating <- 
  fao_nutrient %>% 
  group_by(nutrient) %>% 
  summarize(global_tonnes_before = sum(country_application))
sum_before_allocating

fao_nutrient <- 
  expand_grid(iso3c = missing_rgns,
              nutrient = c("N", "P2O5", "K2O"),
              country_application = NA,
              reported_by_fao = "No") %>% 
  bind_rows(fao_nutrient) %>% 
  mutate(reported_by_fao = if_else(is.na(reported_by_fao),
                                   "Yes",
                                   reported_by_fao))
fao_nutrient %>% group_by(nutrient) %>%
  summarize(total = sum(country_application, na.rm=TRUE))

```


# Remove the proportion of fertilizer used for grasslands from the FAO data.
This will get counted with the grazing animal.

```{r}
fubc_grass <- vroom(here("crop/nutrient/data/fubc_2017_grassland.csv"))
  
fao_nutrient_adjusted <- 
  left_join(fao_nutrient, fubc_grass) %>% 
  mutate(country_application = country_application - (country_application * grassland)) %>% 
  select(- grassland)

fao_nutrient_grassland <- 
  left_join(fao_nutrient, fubc_grass) %>% 
  mutate(country_application = country_application * grassland) %>% 
  select(- grassland)

fao_nutrient_adjusted %>%
  group_by(nutrient) %>%
  summarize(total = sum(country_application, na.rm=TRUE))

```

# Save

```{r}
write_csv(fao_nutrient_adjusted, here("crop/nutrient/data/fao_nutrient_application.csv"))

write_csv(fao_nutrient_grassland, here("crop/nutrient/data/fao_nutrient_grassland_1.csv"))
```


