---
title: "grassland_nutrient_mapping"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Script objective
Map nutrient leaching for grassland fertilizer applications.

# Preamble
```{r setup, include = FALSE}
# getting packages we want
library(here);library(janitor);library("future");library(furrr);library(vroom)

# Prevent writing xml files
rgdal::setCPLConfigOption("GDAL_PAM_ENABLED", "FALSE")

# Raster templates
source(here("_spatial/template_raster.R"))
source(here("_workflow/common.R"))
source(here("_workflow/common_spatial.R"))

# Import food_rgns xy df
food_rgns_xy <- 
  food_rgns_xy %>% 
  dplyr::select(x, y, iso3c)
``` 

# Withdrawal of N

Make a raster that describes proportion withdrawal by plants (Table V, NUE values, Bouwman et al. 2005). NOTE: When NUE > 1, use proportion 1. Multiply this raster by the tonnes N applied. Save this raster in dataprep. Check: The sum of the entire raster should be somewhere around 50% of total global fertilizer application.

```{r}
N_withdrawal_raster <- 
  vroom(here("crop/nutrient/data/bouwman_2005_tbl5.csv"))[,c(1,4)] %>% 
  mutate(NUE = if_else(NUE >= 100, 1, NUE / 100)) %>% 
  left_join(food_rgns_xy, ., by = "iso3c") %>% 
  select(x, y, NUE) %>% 
  rasterFromXYZ(crs = food_crs)
plot(N_withdrawal_raster)
```

# Volatilization of N as NH3
Make a raster that describes proportion of N lost to volatilization as NH3. This is based on Table 4 of Bouwman et al. 2002. Ultimately, I decided the best bet is to add the grassland, flooded rice, and crops to get the average values based on how the fertilizer is regionally applied (so this will be slightly different from what you created already). I have done the calculations in the attached spreadsheet (this column will create the raster: prop_NH3-N total). Multiply this raster by the tonnes N applied. Save this raster in the predata folder (people will want this one). Check: The sum of the entire raster should be somewhere around 14% of total global fertilizer application.

```{r}
N_volatilized_raster <- 
  vroom(here("crop/nutrient/data/bouwman_2002_tbl4_grasslands.csv")) %>%
  select(prop_nh3_total_grasslands, iso3c) %>% 
  left_join(food_rgns_xy, ., by = "iso3c") %>% 
  select(x, y, prop_nh3_total_grasslands) %>% 
  rasterFromXYZ(crs = food_crs)
plot(N_volatilized_raster)
```

# Map excess fertilizer applications  (similar to FAO estimates for manure)

```{r}
nutrient_application <- function(this_grazing_animal, this_nutrient) {

  # import grazing animal rasters
  grazing_animal_raster <- 
    raster(file.path(prep,
                     "animal_farm/nutrient/grassland_fertilizer_application/",
                     str_extract(this_grazing_animal, ".+?(?=_)"), 
                     "_grassland_", 
                     str_extract(this_grazing_animal, "(?<=grassland_).*$"),
                     "_nutrient_application_",
                     this_nutrient,
                     ".tif",
                     fsep = ""))


  if (this_nutrient == "N") {
  
  # Withdrawal of N
  
  N_withdrawal_raster_outcome <-
    overlay(N_withdrawal_raster,
            grazing_animal_raster,
            fun = function(r1, r2){
              
              r1[is.na(r1)] <- 0
              r2[is.na(r2)] <- 0
              
              r1 * r2
              
            })
  
    writeRaster(N_withdrawal_raster_outcome,
                file.path(prep,
                          "animal_farm/nutrient/grassland_nutrient_pollution/",
                          str_extract(this_grazing_animal, ".+?(?=_)"), 
                          "_grassland_", 
                          str_extract(this_grazing_animal, "(?<=grassland_).*$"),
                          "_withdrawal_N.tif", 
                          fsep = ""), 
                format = "GTiff", 
                overwrite = TRUE)
  
  # Volatilization of N as NH3
  
    N_volatilized_raster_outcome <- 
      overlay(N_volatilized_raster,
              grazing_animal_raster,
              fun = function(r1, r2){
                
                r1[is.na(r1)] <- 0
                r2[is.na(r2)] <- 0
                
                r1 * r2
              }) 
      
      writeRaster(N_volatilized_raster_outcome,
                  file.path(prep,
                            "animal_farm/nutrient/grassland_nutrient_pollution/",
                            str_extract(this_grazing_animal, ".+?(?=_)"), 
                            "_grassland_", 
                            str_extract(this_grazing_animal, "(?<=grassland_).*$"),
                            "_volatilized_N.tif", 
                            fsep = ""), 
                  format = "GTiff", 
                  overwrite = TRUE) 
    
    # Denitrification, N2 and N20 emissions
    ## We are ignoring NO emissions, which are about 0.7% of N application.
    ## Multiply the tonnes applied raster by 0.01 and then by 10.2 to get the 
    ## amount of N lost to denitrification. 
    
    N_denitrification_raster <- 
      grazing_animal_raster *
      0.01 *
      10.2
        
    writeRaster(N_denitrification_raster,
                file.path(prep,
                          "animal_farm/nutrient/grassland_nutrient_pollution/",
                          str_extract(this_grazing_animal, ".+?(?=_)"), 
                          "_grassland_", 
                          str_extract(this_grazing_animal, "(?<=grassland_).*$"),
                          "_denitrification_N.tif", 
                          fsep = ""), 
                format = "GTiff", 
                overwrite = TRUE)
      
   # Leaching/runoff
   ## Sum the tonnes N rasters for: withdrawal, volatilization, and denitrification.
   ## Leaching/runoff: Tonnes applied minus N - (withdrawal + volatilization + denitrification)
   ## NOTE: when leaching/runoff < 0, make 0
   ## Save this raster in predata folder. Report on global average of leaching/runoff
      
   N_leaching <- 
      overlay(grazing_animal_raster,
              N_withdrawal_raster_outcome,
              N_volatilized_raster_outcome,
              N_denitrification_raster,
              fun = function(applied,
                             withdrawn,
                             volatilized,
                             denitrified) { 
                
                leached <- applied - (withdrawn + volatilized + denitrified)
                leached[leached < 0] <- 0
                return(leached)
                
                })

    writeRaster(N_leaching,
                file.path(prep, 
                          "animal_farm/nutrient/grassland_nutrient_pollution/",
                          str_extract(this_grazing_animal, ".+?(?=_)"), 
                          "_grassland_", 
                          str_extract(this_grazing_animal, "(?<=grassland_).*$"),
                          "_leached_N", 
                          fsep = ""), 
                format = "GTiff", 
                overwrite = TRUE)
    
  } else if (this_nutrient == "P2O5") {
    
    leaching_raster_P <-
      grazing_animal_raster /
      2.29 * # Convert to P
      0.065 # Leaching proportion
      
  leaching_raster_P[is.na(leaching_raster_P)] <- 0

  writeRaster(leaching_raster_P,
              file.path(prep, 
                        "animal_farm/nutrient/grassland_nutrient_pollution/",
                          str_extract(this_grazing_animal, ".+?(?=_)"), 
                          "_grassland_", 
                          str_extract(this_grazing_animal, "(?<=grassland_).*$"),
                        "_leached_P", 
                        fsep = ""),
              format = "GTiff",
              overwrite = TRUE)
  }
}

plan(multisession, workers = 12)
options <- furrr_options(seed = 54232)

grazer_list <- 
  list.files(file.path(prep, "animal_farm/nutrient/grassland_fertilizer_application"),
           pattern = ".(N|P2O5)\\.tif$") %>% 
  str_extract(., pattern = ".+?((meat|milk))") %>% 
  unique()

distinct_vars <- 
  expand_grid(grazer_list = grazer_list,
            nutrient = c("N", "P2O5"))

future_map2(distinct_vars$grazer_list,
            distinct_vars$nutrient,
            nutrient_application,
            .options = options)
```

-------------------------------------------------------------------------------

# Data check
```{r}
total_N <- 
  list.files(file.path(prep,
                     "animal_farm/nutrient/grassland_fertilizer_application/",
                     fsep = ""),
             pattern = "_N",
             full.names = TRUE) %>% 
  stack() %>% 
  stackApply(., 1, fun = "sum") %>% 
  cellStats(., "sum")
total_N
withdrawn_N <- 
  list.files(file.path(prep,
                     "animal_farm/nutrient/grassland_nutrient_pollution",
                     fsep = ""),
             pattern = "withdrawal",
             full.names = TRUE) %>% 
  stack() %>% 
  stackApply(., 1, fun = "sum") %>% 
  cellStats(., "sum")


withdrawn_N / total_N # should be approx 50% of total N

# NH3 volatilization
volatilized_N <- 
  list.files(file.path(prep,
                     "animal_farm/nutrient/grassland_nutrient_pollution",
                     fsep = ""),
             pattern = "volatilized",
             full.names = TRUE) %>%  
  stack() %>% 
  stackApply(., 1, fun = "sum") %>% 
  cellStats(., "sum")

volatilized_N / total_N # should be approx 14% of total N

# Denitrification, N2 and N20 emissions
denitrification_N <- 
  list.files(file.path(prep,
                     "animal_farm/nutrient/grassland_nutrient_pollution",
                     fsep = ""),
             pattern = "denitrification",
             full.names = TRUE) %>% 
  stack() %>% 
  stackApply(., 1, fun = "sum") %>% 
  cellStats(., "sum")

denitrification_N / total_N # should be approx 10% of total N

leached_N <- 
  list.files(file.path(prep,
                       "animal_farm/nutrient/grassland_nutrient_pollution",
                       fsep = ""),
             pattern = "leached_N",
             full.names = TRUE) %>% 
  stack() %>% 
  stackApply(., 1, fun = "sum") %>% 
  cellStats(., stat = "sum")


total_N - (withdrawn_N + volatilized_N + denitrification_N)
leached_N / total_N

# Check P
P_leached <- 
  list.files(file.path(prep, "animal_farm/nutrient/grassland_nutrient_pollution"),
             pattern = "_leached_P",
             full.names = TRUE) %>% 
  stack()  %>% 
  stackApply(., 1, fun = "sum") %>% 
  cellStats(., stat = "sum")

(leached_N + P_leached + volatilized_N) / total_N
leached_N 
           
total_P <- 
  list.files(file.path(prep,
                     "animal_farm/nutrient/grassland_fertilizer_application",
                     fsep = ""),
             pattern = "P2O5",
             full.names = TRUE) %>% 
  stack() %>% 
  stackApply(., 1, fun = "sum") %>% 
  cellStats(., "sum")

(total_P / 2.29) * 0.065
P_leached / total_P
```

# Nutrient pollution per animal
```{r}
for(val in grazer_list) {

  print(val)
  
leached_N <- 
  list.files(file.path(prep, "animal_farm/nutrient/grassland_nutrient_pollution"),
           pattern = paste(val, ".*_leached_N", sep = ""),
           full.names = TRUE) %>% 
  raster() %>% 
  cellStats(., stat = "sum")
leached_N  

P_leached <- 
  list.files(file.path(prep, "animal_farm/nutrient/grassland_nutrient_pollution"),
           pattern = paste(val, ".*_leached_P", sep = ""),
           full.names = TRUE) %>% 
  raster() %>% 
  cellStats(., stat = "sum")
P_leached

volatilized_N <- 
  list.files(file.path(prep, "animal_farm/nutrient/grassland_nutrient_pollution"),
           pattern = paste(val, ".*volatilized", sep = ""),
           full.names = TRUE) %>% 
  raster() %>% 
  cellStats(., stat = "sum")
volatilized_N

total_applied <- 
  list.files(file.path(prep, "animal_farm/nutrient/grassland_fertilizer_application",
                       fsep = ""),
             pattern = paste(val, ".*_(N|P2O5)\\.tif", sep = ""),
           full.names = TRUE) %>% 
  stack() %>% 
  stackApply(., 1, fun = "sum") %>% 
  cellStats(., "sum")
total_applied

print(leached_N + P_leached + volatilized_N)
}
```

# Nutrient pollution per animal
```{r}
for(val in grazer_list) {

print(val)
  
withdrawn_N <- 
  list.files(file.path(prep, "animal_farm/nutrient/grassland_nutrient_pollution"),
           pattern = paste(val, ".*withdrawal", sep = ""),
           full.names = TRUE) %>% 
  raster() %>% 
  cellStats(., stat = "sum")
withdrawn_N  

denitrification_N <- 
  list.files(file.path(prep, "animal_farm/nutrient/grassland_nutrient_pollution"),
           pattern = paste(val, ".*denitrification", sep = ""),
           full.names = TRUE) %>% 
  raster() %>% 
  cellStats(., stat = "sum")
denitrification_N

volatilized_N <- 
  list.files(file.path(prep, "animal_farm/nutrient/grassland_nutrient_pollution"),
           pattern = paste(val, ".*volatilized", sep = ""),
           full.names = TRUE) %>% 
  raster() %>% 
  cellStats(., stat = "sum")
volatilized_N

total_N <- 
  list.files(file.path(prep, "animal_farm/nutrient/grassland_fertilizer_application",
                       fsep = ""),
             pattern = paste(val, ".*_N", sep = ""),
           full.names = TRUE) %>% 
  stack() %>% 
  stackApply(., 1, fun = "sum") %>% 
  cellStats(., "sum")
total_N

print(1- ((withdrawn_N + volatilized_N + denitrification_N) / total_N))

}
```

