---
title: "Checking tif files"
output: html_document
editor_options: 
  chunk_output_type: console
---
* final edits to files

* Check that all files have been created.
* Create a csv file describing all combinations of systems and pressures
* Create a csv with tif parameters to check that all is good

2-3 hour runtime
```{r}

library(rgdal)
library(sp)
library(raster)
library(tidyverse)
library(here)

```



## last minute change to name of fodder crop related files
changing: land_fodd_crop_produce to 
          land_fodd_crop_feed

```{r}

fodder_files <- list.files('/home/shares/food-systems/Food_footprint/all_food_systems/datalayers', full=TRUE, pattern = "land_fodd_crop_produce")


file.rename(from= fodder_files, to = 
                gsub("land_fodd_crop_produce", "land_fodd_crop_feed", fodder_files))



```

## check against master list to make sure everything is there
Add to this list by hand: _analysis/food_system_categories.csv"
```{r}

complete_list <- read_csv(here("_analysis/food_system_categories.csv")) %>% data.frame()

# all farm pressures
farm_pressures <- paste(complete_list$origin,
                        complete_list$organism,
                        complete_list$system,
                        complete_list$product, sep="_")

farm_pressures <- expand.grid(farm_pressures, c("disturbance", "ghg", "water", "nutrient"))
pressures <- paste(farm_pressures$Var1, farm_pressures$Var2, sep="_")

## feed proportion files
fed <- filter(complete_list, fed=="fed") %>%
  mutate(animal_system = paste(origin, organism, system, product, sep="_")) %>%
  select(animal_system) %>%
  data.frame()

crops <- filter(complete_list, subcategory=="crops") %>%
  mutate(feed_system = paste(origin, organism, system, product, sep="_")) %>%
  select(feed_system)

animal_feed_combos <- merge(data.frame(fed), data.frame(crops)) %>%
  mutate(pressures = paste(animal_system, feed_system, sep="_x_")) %>%
  select(pressures) %>%
  pull(.)

# human crop consumption
human_crop_consumption <- paste("land_human_human_human", crops$feed_system, sep="_x_") 
human_crop_consumption <- grep("fodd", human_crop_consumption, value=TRUE, invert=TRUE) # eliminate the fodder category

#add in fofm (only these animals)
fed_fofm_system <- grep(pattern="pig|salmon|chicken|tuna|crustaceans|shrimp|marine-fish-general", fed$animal_system, value=TRUE)
fed_fofm_system <- grep("eggs&meat", fed_fofm_system, value=TRUE, invert=TRUE)
fed_fofm <- paste(fed_fofm_system, "marine_fofm_fisheries_meat", sep="_x_")

# human and extra fofm consumption
human_fofm_consumption <- paste("land_human_human_human", "marine_fofm_fisheries_meat", sep="_x_") 



# combine everything:
all_tiffs <- c(pressures, fed_fofm, animal_feed_combos, human_crop_consumption, human_fofm_consumption)
all_tiffs <- paste(all_tiffs, "tif", sep=".")

```


Check to see if anything is missing...

```{r}
missing <- setdiff(all_tiffs, 
        list.files('/home/shares/food-systems/Food_footprint/all_food_systems/datalayers', pattern = ".tif"))
## we can ignore these things:
missing <- grep("marine_salmon_aquaculture_meat_x_land_fodd_crop_feed.tif", missing, value=TRUE, invert=TRUE)
missing <- grep("marine_shrimp_aquaculture_meat_x_land_fodd_crop_feed.tif", missing, value=TRUE, invert=TRUE)
missing <- grep("marine_crustaceans_aquaculture_meat_x_land_fodd_crop_feed.tif", missing, value=TRUE, invert=TRUE)
missing <- grep("marine_marine-fish-general_aquaculture_meat_x_land_fodd_crop_feed.tif", missing, value=TRUE, invert=TRUE)
missing <- grep("marine_tuna_aquaculture_meat_x_land_fodd_crop_feed.tif", missing, value=TRUE, invert=TRUE)
missing  # nothing should be left!

setdiff(list.files('/home/shares/food-systems/Food_footprint/all_food_systems/datalayers', pattern = ".tif"), all_tiffs)  ## ignoring buffaloes/meat  and ofib categories of crops.

```


adjust all_tiffs a bit to remove files that we didn't produce:
```{r}

final_tiffs <- all_tiffs 

final_tiffs <- grep("marine_salmon_aquaculture_meat_x_land_fodd_crop_feed.tif", final_tiffs, value=TRUE, invert=TRUE)
final_tiffs <- grep("marine_shrimp_aquaculture_meat_x_land_fodd_crop_feed.tif", final_tiffs, value=TRUE, invert=TRUE)
final_tiffs <- grep("marine_marine-fish-general_aquaculture_meat_x_land_fodd_crop_feed.tif", final_tiffs, value=TRUE, invert=TRUE)
final_tiffs <- grep("marine_crustaceans_aquaculture_meat_x_land_fodd_crop_feed.tif", final_tiffs, value=TRUE, invert=TRUE)
final_tiffs <- grep("marine_tuna_aquaculture_meat_x_land_fodd_crop_feed.tif", final_tiffs, value=TRUE, invert=TRUE)

setdiff(final_tiffs, 
        list.files('/home/shares/food-systems/Food_footprint/all_food_systems/datalayers', pattern = ".tif"))

final_tiffs
# should be 982 files!
write_csv(data.frame(raw_tiffs_list = final_tiffs), here("_analysis/final_list_raw_tiffs.csv"))
```


# Check rasters to make sure extents and such are correct
```{r}

tif_check_list <- list.files('/home/shares/food-systems/Food_footprint/all_food_systems/datalayers', full=TRUE, pattern = ".tif")

# empty data.frame to add to
raster_check <- data.frame()


for(file in tif_check_list){
  # file <- "/home/shares/food-systems/Food_footprint/all_food_systems/datalayers/marine_demersal_fisheries_meat_disturbance.tif" 
name <- basename(file)
raster_file <- raster(file)
tmp_values <- raster::values(raster_file)
extent_data <- data.frame(file = name, 
                          cellN = ncell(raster_file),
                          xmin = extent(raster_file)@xmin, 
                          xmax = extent(raster_file)@xmax, 
                          ymin = extent(raster_file)@ymin, 
                          ymax = extent(raster_file)@ymax,
                          variable = names(raster_file),
                          min = minValue(raster_file),
                          max = maxValue(raster_file),
                          sum_cells = cellStats(raster_file, sum, na.rm=TRUE),
                          quant_50 = quantile(raster_file, c(0.5)),
                          mean = cellStats(raster_file, mean, na.rm=TRUE),
                          total_zero = sum(tmp_values==0, na.rm=TRUE),
                          total_NA = sum(is.na(tmp_values)),
                          total_greater_zero = sum(tmp_values>0, na.rm=TRUE))
raster_check <- rbind(raster_check, extent_data)
}

write_csv(raster_check, here("_analysis/checking_data/data/raster_check.csv"))

# some checks
## make sure these are equal:
length(tif_check_list)
dim(raster_check)

filter(raster_check, cellN!=9331200)
summary(raster_check)
filter(raster_check, is.na(mean))

new <- read_csv(here("_analysis/checking_data/data/raster_check.csv")) %>%
  select(file, sum_cells)
old <- read_csv(here("_analysis/checking_data/data/raster_check_May28_2022.csv")) %>%
  select(file, old_sum_cells = sum_cells) %>%
  left_join(new, by="file") %>%
  mutate(difference = old_sum_cells - sum_cells) %>%
  filter(difference != 0) %>% 
  mutate(prop_diff = difference/old_sum_cells*100) %>%
  data.frame()
plot(old$old_sum_cells, old$sum_cells)
abline(0,1, col="red")

new_crop_nutrients <- old %>%
  filter(str_detect(file, "crop")) %>%
  filter(str_detect(file, "nutrient")) %>%
  filter(!str_detect(file, "_x_")) %>%
  data.frame()
sum(new_crop_nutrients$old_sum_cells)
sum(new_crop_nutrients$sum_cells)  
plot(new_crop_nutrients$old_sum_cells, new_crop_nutrients$sum_cells)
abline(0,1, col="red")

new_crop_disturbance <- old %>%
  filter(str_detect(file, "crop")) %>%
  filter(str_detect(file, "disturbance")) %>%
  filter(!str_detect(file, "_x_")) %>%
  data.frame()
sum(new_crop_disturbance$old_sum_cells)
sum(new_crop_disturbance$sum_cells)
plot(new_crop_disturbance$old_sum_cells, new_crop_disturbance$sum_cells)
abline(0,1, col="red")


new_crop_ghg <- old %>%
  filter(str_detect(file, "crop")) %>%
  filter(str_detect(file, "ghg")) %>%
  filter(!str_detect(file, "_x_")) %>%
  data.frame()
sum(new_crop_ghg$old_sum_cells)
sum(new_crop_ghg$sum_cells)
plot(new_crop_ghg$old_sum_cells, new_crop_ghg$sum_cells)
abline(0,1, col="red")
```

