---
title: "Finfish mariculture diets"
author: "Gage Clawson"
date: "01/12/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

Determining the percent composition of finfish diet.

1. Filter out the production data that gage gave me by xxxx
2. We have diet compositions of xxxxx species, look to see how different they are
3. Look to see how much of total production do these species account for 
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(readxl)

diet <- read_excel(here("feed/diet_composition_wrangling/Final_DietCrop_Numbers.xlsx"), sheet = "% type of crop input")
finfish_prod <- read_csv(here('aquaculture/marine/marine_fish_general/finfish_farms/data/fao_mariculture_finfish_taxa_production.csv'))

```


```{r}

all_finfish <- unique(finfish_prod$species)
## there are 39 species

over_k_finfish_global <- finfish_prod %>% 
  group_by(species) %>% 
  dplyr::summarise(sum_global = sum(fao_tonnes_production, na.rm = TRUE)) %>% 
  filter(sum_global > 1000) %>% 
  select(species) %>% 
  unique() %>% 
  pull()

## there are 34 species where global production is 1000 tonnes or more. This is the same as per country, so I think that we can figure out diet composition for all finfish from these species

```

Explore the diet compositon for the species we have from Halley
```{r}

## definitely filter for "Flounders, halibuts, soles", "Miscellaneous coastal fishes", "Miscellaneous demersal fishes", "Miscellaneous diadromous fishes", "Miscellaneous pelagic fishes"


## what we have covered, that almost match exactly in common name (10 out of 39 FAO species): 
# FAO Common name - ISCAAP group (halley common name)
# Righteye flounders nei - Flounders, halibuts, soles (European flounder) X 
# European seabass - Miscellaneous coastal fishes (European seabass) X 
# Gilthead seabream - Miscellaneous coastal fishes (Gilthead seabream) X 
# Groupers nei - Miscellaneous coastal fishes (Grouper) X 
# Red drum - Miscellaneous coastal fishes (Red drum) X 
# Mullets nei - Miscellaneous coastal fishes (Mullet) X 
# Korean rockfish - Miscellaneous demersal fishes (Wolfishes & rockfish) X 
# Milkfish - Miscellaneous diadromous fishes (Milk fish) X 
# Barramundi(=Giant seaperch) - Miscellaneous diadromous fishes (Barramundi) X 
# Japanese amberjack - Miscellaneous pelagic fishes (Japanese amberjack) X 


## what we could add to that potentially (adding this would cover 21/39 FAO species)
# common_name : Category
# Atlantic halibut - Flounders, halibuts, soles (same as European flounder) X 
# Bastard halibut - Flounders, halibuts, soles (same as European flounder) X 
# Amberjacks nei - Miscellaneous pelagic fishes (same as japanese amberjack) X 
# Blackhead seabream - Miscellaneous coastal fishes (same as gilthead seabream) X 
# Jacks crevalles nei - Miscellaneous pelagic fishes (same as japanese amberjack) X 
# Japanese seabass - Miscellaneous coastal fishes (same as european seabass) X 
# Japanese seabream - Miscellaneous coastal fishes (same as gilthead seabream) X 
# Porgies, seabreams nei - Miscellaneous coastal fishes (same as gilthead seabream) X 
# Senegalese sole - Flounders, halibuts, soles (same as European flounder) X 
# Silver seabream - Miscellaneous coastal fishes (same as gilthead seabream) X
# Marine fishes nei - same as japanese amberjack (this is just the average of marine species listed) X 

# then the rest will be given the average (which is the japanese amberjack number)



diet_finfish_1 <- diet %>% 
  filter(ISSCAAP_Group %in% c("Flounders, halibuts, soles", "Miscellaneous coastal fishes", "Miscellaneous demersal fishes", "Miscellaneous diadromous fishes")) %>% 
  select(common_name, crop, prop_crop_inputs)

diet_finfish_2 <- diet %>%
  filter(ISSCAAP_Group == "Miscellaneous pelagic fishes") %>%
  group_by(common_name) %>% 
  mutate(total_input_prop = sum(prop_crop_inputs)) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(prop_diet = prop_crop_inputs/total_input_prop) %>%
  dplyr::select(common_name, crop, prop_crop_inputs = prop_diet)

diet_finfish <- rbind(diet_finfish_1, diet_finfish_2)

##compare the diet compositions

ggplot(diet_finfish) +
  geom_col(aes(x = prop_crop_inputs, y = common_name , fill = crop)) +
  theme_minimal() + ggtitle("Diet compositon crop breakdown")
## ok so there are for sure some differences, so we should probably account for that with a weighted proportion of raised species in each country. 

```


Now we need to figure out how to assign diet compositions to the species we don't have info one. 

I will use the gapfilling outlined above. I.e. we dont have diet comp data for Silver seabream, but I will assign it the gilthead seabream values... anything that falls out of what i have outlined will be given the marine fish average 


Which countries and which species are countries raising?

```{r}

ggplot(finfish_prod)  +
  geom_col(aes(x = fao_tonnes_production, y = country , fill = species)) +
  theme_minimal() +
  ggtitle("Tonnes fish aquaculture production")



```


## Create a diet compositon for all finfish using weighted proportions
```{r}

species_prop <- finfish_prod %>% 
  mutate(diet_species = case_when(species == "Milkfish"  ~ "Milk fish",
                                  species == "Righteye flounders nei" ~ "European flounder",
                                  species == "Groupers nei" ~ "Grouper",
                                  str_detect(species, "Mullets nei|mullet") ~ "Mullet", 
                                  species == "Korean rockfish" ~ "Wolfishes & rockfish",
                                  species == "Barramundi(=Giant seaperch)" ~ "Barramundi",
                                  str_detect(species, "seabream") ~ "Gilthead seabream",
                                  str_detect(species, "Jack|jack|Marine fishes nei") ~ "Japanese amberjack",
                                  str_detect(species, "seabass|bass") ~ "European seabass",
                                  str_detect(species, "halibut|sole") ~ "European flounder",
                                  species == "Red drum" ~ "Red drum",
                                  species == "Gilthead seabream" ~ "Gilthead seabream",
                                  species == "European seabass" ~ "European seabass", 
                                  species == "Japanese amberjack" ~ "Japanese amberjack",
                                  TRUE ~ "Japanese amberjack" # This is just the average of marine fish
                                  )) %>% 
  group_by(diet_species, iso3c) %>% 
  dplyr::summarise(species_rgn_tonnes = sum(fao_tonnes_production)) %>% 
  ungroup() %>% 
  group_by(iso3c) %>% 
  mutate(rgn_tonnes = sum(species_rgn_tonnes)) %>% 
  ungroup() %>% 
  rowwise() %>% 
  mutate(prop = species_rgn_tonnes/rgn_tonnes) %>% 
  ungroup() %>% 
  select(iso3c, common_name = diet_species, prop)
  
check <- species_prop %>%
  group_by(iso3c) %>%
  summarise(sum(prop))

## fofm is 1 - 0.885 for everything

prop_fofm = 1 - 0.885

# prop_diadromous = 1 - 0.975

weighted_diet <- diet_finfish %>% 
  left_join(species_prop, by ="common_name") %>% 
  rowwise() %>% 
  mutate(weight_prop_int = prop_crop_inputs*prop) %>% 
  group_by(iso3c, crop) %>%
  dplyr::summarise(weighted_crop_prop = sum(weight_prop_int) * (1 - prop_fofm)) %>%
  ungroup()


check <- weighted_diet %>% 
  group_by(iso3c) %>% 
  dplyr::summarise(perc_check = sum(weighted_crop_prop))
  ## all per_check values should be 0.885 after incorporating fofm!

```

Adding in fofm component to diet.  This will be later adjusted to match Halley final tonnes. 
```{r}
fofm_addition <- dplyr::select(weighted_diet, iso3c) %>% 
  unique() %>%
  mutate(crop = "fofm") %>%
  mutate(weighted_crop_prop = prop_fofm)

weighted_diet <- rbind(weighted_diet, fofm_addition)
unique(weighted_diet$crop)

check <- weighted_diet %>% 
  group_by(iso3c) %>% 
  dplyr::summarise(perc_check = sum(weighted_crop_prop))
  ## all per_check values should be 1, after incorporating fofm!


## need to make sure these are added to the feed_category_table.csv, product_to_fao_spam.csv
weighted_diet %>% select(crop) %>% unique()

finfish_diet_comp <- weighted_diet %>% 
  rename(raw_name = crop, prop_diet = weighted_crop_prop) %>% 
  mutate(production_system ="aquaculture",
         species = "marine-fish-general")

write_csv(finfish_diet_comp, here("feed/diet_composition_wrangling/marine-fish-general_aquaculture_diet_composition.csv"))

```
