---
title: "MapSPAM_scaling_2010_to_2017"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Script Description
Rescale MapSPAM data to year 2017.

# Data information
We do this using FAOSTAT crop production data trends between year 2010 and 2017.

# Preamble
```{r setup, include = FALSE}
# Raster templates
library(here)
source(here("_workflow/common.R"))
source(here("_workflow/common_spatial.R"))

# getting packages we want
library(janitor)

# Import food_rgns xy df
food_rgns_xy <- 
  food_rgns_xy %>% 
  dplyr::select(x, y, iso3c)

predatalayers_path <- "/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers/"

```

# Import and wrangle FAO data for 2010 and 2017 
```{r}
# SPAM names
crop_codes_updated <- 
  vroom(here("crop/farm/data/crop_codes_updated.csv"))  %>% 
  dplyr::select(SPAM_short_name, item_code)

scaling_coef_tidy <- 
  vroom::vroom(here("crop/farm/data/FAOSTAT_crop_production_2010_2017.csv")) %>% 
  clean_names() %>% 
  dplyr::select(area_code, item_code, year, value) %>% 
  rename(tonnes = value) %>% 
  na.omit()
```

# Special Sudanese case

Sudan was split into two countries in 2010, so now we need to scale the difference from 2010 to 2017. To do this, I am extracting production data for SPAM's 2010 map as opposed to splitting FAO data.

```{r}
SDN_SSD_2010 <- 
  vroom::vroom(here("crop/farm/data/prod_crop_rgns_2010.csv")) %>% 
  filter(iso3c %in% c("SSD", "SDN")) %>% 
  group_by(iso3c, SPAM_short_name) %>% 
  summarize(`2010` = sum(production)) %>% 
  ungroup() 

SDN_SSD_2017 <- 
  scaling_coef_tidy %>% 
  filter(area_code %in% c(276, 277)) %>% # Filter Sudan
  left_join(crop_codes_updated) %>%   
  left_join(read_csv(here("_spatial/_output/FAO_rgn_codes.csv")), 
            by = "area_code") %>% # Add iso3c
  group_by(SPAM_short_name, iso3c) %>% 
  summarise(`2017` = sum(tonnes, na.rm = TRUE)) %>% 
  ungroup() %>% 
  na.omit()

sudan_coef <- 
  full_join(SDN_SSD_2010, SDN_SSD_2017) %>% 
  mutate_all(~replace(., is.na(.), 0)) %>% 
  mutate(coef = `2017` / `2010`) %>% 
  select(!contains("tonnes"))
```

# Wrangle FAO dataset 
```{r}
scaling_coef_tidy <- 
  scaling_coef_tidy %>% 
  mutate(area_code = if_else(area_code == 206, 277, area_code)) %>% # Re-allocate former Sudan's 2010 values.
  bind_rows(scaling_coef_tidy) %>%
  mutate(area_code = if_else(area_code == 206, 276, area_code)) %>% 
  unique() %>% 
  filter(!area_code %in% 351) %>% # Remove China
  left_join(read_csv(here("_spatial/_output/FAO_rgn_codes.csv")), 
            by = "area_code") %>% # Add iso3c
  left_join(crop_codes_updated) %>% # Add SPAM_short_name
  dplyr::select(year, tonnes, iso3c, SPAM_short_name) %>% 
  na.omit() # Remove NA created by milled rice.

# Distinct area_codes
n_distinct(scaling_coef_tidy$iso3c) # 214

missing_rgns <- setdiff(food_rgns$iso3c, scaling_coef_tidy$iso3c)
missing_rgns
food_rgns %>% filter(iso3c %in% missing_rgns) # Nothing unusual
```

# Calculate scaling from 2010 to 2017 for all crops and countries.
```{r}
scaling_coef <- 
  scaling_coef_tidy %>% 
  group_by(SPAM_short_name, year, iso3c) %>%
  summarize(tonnes = mean(tonnes)) %>% # Calculate mean tonnes for each crop, year and country.
  ungroup() %>% 
  spread(., year, tonnes) %>% 
  filter(!iso3c %in% c("SDN", "SSD")) %>% # Replace Sudan values
  mutate_all(~replace(., is.na(.), 0)) %>% 
  group_by(SPAM_short_name, iso3c, `2017`, `2010`) %>%
  summarize(coef = `2017` / `2010`, .groups = "keep") %>% 
  ungroup() %>% 
  bind_rows(sudan_coef)

# Make sure that countries with more than 0 production in SPAM 2010 but
# no data in 2017 have 0 value for 2017.
scaling_coef$coef[scaling_coef$`2010` > 0 & scaling_coef$`2017` == 0] <- 0

# Some countries will have data in 2017, according to the FAO, but none in 2010. We need to make sure that we manually insert these values in the scaled maps.

inf_values <- 
  scaling_coef %>% 
  filter(`2010` == 0 & `2017` > 0) %>% 
  select(SPAM_short_name, iso3c, tonnes = `2017`)
inf_values

scaling_coef$coef[scaling_coef$`2010` == 0 & scaling_coef$`2017` > 0] <- 0

# All infinite numbers assigned a coef of 1.
scaling_coef$coef[which(!is.finite(scaling_coef$coef))] <- 0
```

# Save
```{r}
vroom_write(inf_values, here("crop/farm/data/inf_values.csv"))
write_csv(scaling_coef, here("crop/farm/data/scaling_coef.csv"))
```

-------------------------------------------------------------------------------

# Map fodder crop disturbance
```{r}
bovine_coefs <- 
  vroom::vroom(here("crop/farm/data/FAOSTAT_bovine_production.csv"), 
               .name_repair = janitor::make_clean_names) %>% 
  select("area_code", "value", "year") %>% 
  filter(!area_code %in% 351) %>% # Remove China
  group_by(year, area_code) %>% 
  summarise(heads = sum(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(central_year = case_when(year %in% c(2014, 2015, 2016) ~ 2015,
                                  year %in% c(2016, 2017, 2018) ~ 2017)) %>% 
  group_by(central_year, area_code) %>% 
  summarise(heads = mean(heads, na.rm = TRUE)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = central_year,
              values_from = heads) %>% 
  group_by(area_code) %>%
  summarize(coef = `2017` / `2015`) %>% 
  ungroup() %>%
  rename(ID_0 = area_code) %>% 
  left_join(food_rgns) %>% 
  na.omit() %>% 
  left_join(food_rgns_xy, ., by = "iso3c") %>% 
  select(x, y, coef) %>% 
  mutate(coef = ifelse(is.na(coef), 1, coef)) %>% 
  rasterFromXYZ()
plot(bovine_coefs)

fodder_irrigated <- raster(file.path(raw, "GAEZAct_2015/GAEZAct2015_Production_Foddercrops_Irrigated.tif"))
fodder_rainfed   <- raster(file.path(raw, "GAEZAct_2015/GAEZAct2015_Production_Foddercrops_Rainfed.tif"))

overlay(fodder_irrigated,
        bovine_coefs,
        fun = function(r1, r2) {
             
          return(r1 * r2 * 1000) # Convert to tonnes
             
             }) %>% 
  writeRaster(.,
              file.path(
                prep,
                "crop/farm/scaled_maps_2017/crop_fodd_I_scaled.tif",
                fsep = ""),
              format = "GTiff", 
              overwrite = TRUE) %>% 
  writeRaster(.,
              file.path(
                predatalayers_path,
                "fodd_I_production_2017.tif",
                fsep = ""),
              format = "GTiff", 
              overwrite = TRUE)

overlay(fodder_rainfed,
        bovine_coefs,
        fun = function(r1, r2) {
             
          return(r1 * r2 * 1000) # Fodder map is in 1,000 tonnes, so convert to tonnes
             
             }) %>% 
  writeRaster(.,
              file.path(
                prep,
                "crop/farm/scaled_maps_2017/crop_fodd_rainfed_scaled.tif",
                fsep = ""),
              format = "GTiff", 
              overwrite = TRUE) %>% 
  writeRaster(.,
              file.path(
                predatalayers_path,
                "fodd_rainfed_production_2017.tif",
                fsep = ""),
              format = "GTiff", 
              overwrite = TRUE)

```
