---
title: "Backcalculating emissions intensities from FUI"
author: "Gage Clawson (UCSB, NCEAS, OHI)"
date: "September 17, 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---
The purpose of this script is to prep the emissions intensity estimate from the [Seafood Carbon Emissions Tool](http://seafoodco2.dal.ca/(overlay:menu/5bcb48abaaea53205a2de526). We will back calculate to exlcude the up- and downstream emissions, only accounting for direct fuel use emissions in our estimate. 

We use the data from:

 - [Seafood Carbon Emissions Tool](http://seafoodco2.dal.ca/(overlay:menu/5bcb48abaaea53205a2de526)

 - Parker, R.W.R., Blanchard, J.L., Gardner, C., Green, B.S., Hartmann, K., Tyedmers, P.H., and Watson, R.A. (2018). Fuel use and greenhouse gas emissions of world fisheries. Nature Climate Change 8, 333.


Explain numbers in paper and provide an example of how to back calculate one emissions intensity estimate:

We use the data for the species Common name "Blue grenadier" sector (Table 1).


The value provided in the data is 3.4 to 4.6 CO2eq/kg.  But this includes direct and indirect (shipbuilding, fuel extraction, etc) costs, and is a range, when we need a mean.  Due to the difficulty of estimating the location of indirect costs, we excluded them. This required a bit of backcalculating.

Direct fuel costs are: 
2.8 kg CO2eq/ lfuel (from IPCC 2013 characterization factors)

Calculated as:
Fuel density = 0.9kg/l (in Parker methods section)
Carbon content of fuel = 860g/kg (in Parker methods section)
C02/C ratio = 50/14 
So:
0.9 * 0.860 * 50/14 = 2.8 kg CO2eq/lfuel


We estimated 25% of fuel costs are for refrigeration.  This paper uses 25% for all additional costs, but suggests that refrigeration is the main one.  Similarly, another paper estimates 30% for refrigeration: https://onlinelibrary.wiley.com/doi/epdf/10.1111/j.1530-9290.2011.00344.x
> Refrigeration during transportation was modeled by the addition of 30% extra fuel use for the transport on the basis of the literature data (Garnett 2007)


2.8/0.75 = 3.7 kg CO2eq/l fuelß
check: 3.7*0.25=0.925  and 2.8+0.925=3.7

Blue grenadier FUI: 666 l fuel/tonne

3.7 kgCO2eq/l fuel * 0.666 l fuel/kg catch = 2.4642 kgCO2eq/kgcatch 

# Setup
```{r, eval = FALSE}
#load relevant packages, etc.

library(here)
library(raster)
library(tidyverse)
library(janitor)

raw_folder <- file.path("/home/shares/food-systems/Food_footprint/_raw_data/")
```

# Methods


**Back calculate all emissions intensities from original Parker values**
```{r, eval = FALSE}
## Read in parker emissions intensity estimates
parker_ei <- read_csv(here("fisheries/marine/ghg/data/parker_estimates.csv"))

## read in emissions intensity data 
emissions_intensity_df <- read_csv(file.path(raw_folder, "seafood_carbon_emmisions_tool/d2020/Specie_List_-_9_22_2020,_12_56_43_PM_Wild.csv")) %>%
  dplyr::select(1:17) %>%
  clean_names() %>%
  dplyr::select(common_name, scientific_name, gear, ghg_low, ghg_high, fui_l_t)


## Convert fuel use intensity to L/kgcatch and add in average cost of fuel as a column
food_ei <- emissions_intensity_df %>%
  mutate(FUI_l_kg = fui_l_t/1000) %>% ## 1000 kg in 1 MT
  mutate(fuel_cost = 3.7) ## fuel cost is 3.7 kgCO2eq/l

## calculate direct emissions intensity 
food_ei <- food_ei %>%
  mutate(ei_direct = fuel_cost*FUI_l_kg)

## save intermediate file for later use
food_ei <- food_ei %>%
  dplyr::select(common_name, scientific_name, gear, ghg_low,  ghg_high, FUI_l_kg, fuel_cost, fui_l_t, ei_direct_kgco2_kgcatch = ei_direct)

DT::datatable(food_ei)

write.csv(food_ei, "fisheries/marine/ghg/int/direct_ei.csv", row.names = FALSE)

```



