---
title: "Data Download"
author: "Gage Clawson (UCSB, NCEAS, OHI)"
date: "September 17, 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
library(xlsx)
```

# New Global Fisheries Catch Data

Catch (1950 - 2017) - reported, iuu, and discard catch data for each cell location and unique identifier

* DATA CODE DEFINITIONS (gear/taxa/country codes and cell lat/lon references)

* For this, we will only download the final data year... 2017. So it will be the Catch2015_2019 file.

* The download link is in the README on aurora..

```{r, eval = FALSE}

## Download catch data from web and save into mazu

web_year <- "Catch2015_2019"

data <- read.csv(sprintf("URLgoeshere/%s.csv", web_year)) ## the download URL is located in the README on aurora... place it right before the %s

write.csv(data, file.path( sprintf("/home/shares/food-systems/Food_footprint/_raw_data/watson_data/v5.0/%s.csv", web_year)), row.names = FALSE) 
```

# How to download the data code definitions. 

For this, it is saved as an .xlsx file (Codes.xlsx). This file has 4 sheets in it which contain meta data that explain columns in the Index files, and one sheet with is "Spatial Cells Reference - contains geospatial information associated wtih the Industrial Catch data". 
To download this data, go to the IMAS website: http://data.imas.utas.edu.au/portal/search?uuid=ff1274e1-c0ab-411b-a8a2-5a12eb27f2c0 and dowload it manually.

This .xlsx file contains these sheets:

* Spatial Cells Reference (Cells) - contains geospatial information associated wtih the Industrial Catch data
* Gear Reference (Gear) - contains information regarding how different fishing gear is classified in the index datasets.
* Taxa Reference (Taxa) - contains information regarding how different taxa is classified in the index datasets.
* Country Reference (Country) - contains informations regarding how different countries are labeled in the index datasets. 

I split these up into separate .csv files in the folder on aurora. 



