---
title: "Creating global emissions from marine fisheries rasters"
author: "Gage Clawson (UCSB, NCEAS, OHI)"
date: "September 25, 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

The purpose of this script is to create rasters for global emissions of marine fisheries. One will be created for all FOFM emissions, and one will be created for all catch minus the FOFM emissions.

We use the data and information from:

1. Watson, R. A. and Tidd, A. 2018. Mapping nearly a century and a half of global marine fishing: 1869–2015. Marine Policy, 93, pp. 171-177. [(Paper URL)](https://doi.org/10.1016/j.marpol.2018.04.023)

2. [Seafood Carbon Emissions Tool](http://seafoodco2.dal.ca/(overlay:menu/5bcb48abaaea53205a2de526)

## Setup
```{r, eval = FALSE}
#load relevant packages, etc.

library(here)
library(raster)
library(tidyverse)
library(countrycode)

# function to project raster
## Use correct CRS

watson_raster_template <- raster::raster(ncol=720, nrow=360, vals=c(1:259200))
source(here("fisheries/marine/ghg/raster_crs.R"))


raw_folder <- file.path("/home/shares/food-systems/Food_footprint/_raw_data")
raw_folder_watson <- file.path("/home/shares/food-systems/Food_footprint/_raw_data/watson_data/v5.0/")
ghg_folder <- file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/ghg")
final <- file.path("/home/shares/food-systems/Food_footprint/all_food_systems/datalayers")
```

Calculate emissions
```{r, eval = FALSE}
## read in emissions data
catch_emissions <- read_csv(file.path(ghg_folder, "watson_v5_emissions/all_catch_emissions_2017.csv")) 

catch_emissions_df <- catch_emissions %>%
  mutate(catch_kg = catch*1000, 
         FOFM_catch_kg = FOFM_catch*1000)



## Multiply catch by the emissions intensities to get an emissions estimates

### get an estimate in kg co2eq and t co2eq 
catch_emissions_all <- catch_emissions_df %>%
  mutate(emissions_fofm_kgco2 = FOFM_catch_kg*ei_kgco2_kgcatch) %>%
  mutate(emissions_fofm_tco2 = emissions_fofm_kgco2/1000) %>%
  mutate(emissions_kgco2 =  catch_kg*ei_kgco2_kgcatch) %>%
  mutate(emissions_tco2 = emissions_kgco2/1000)  # %>%
  # mutate(emissions_no_fofm_kgco2 = emissions_kgco2 - emissions_fofm_kgco2,
  #        emissions_no_fofm_tco2 = emissions_tco2 - emissions_fofm_tco2)

sum(catch_emissions_all$emissions_tco2) #382823674


```

Save rasters for supplementary info by intermediate species classes
```{r}
#### Now lets create rasters for each intermediate category and visualize and save
food_crs <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
food_raster <- raster(nrows=2160, ncols=4320, xmn=-180, xmx=180, ymn=-90, ymx=90)

## Forage fish
allforagefish_raster <- catch_emissions_all %>%
  dplyr::group_by(Cell) %>%
  dplyr::summarize(forage_emissions_tonnes = sum(emissions_fofm_tco2, na.rm=TRUE), 
                   forage_emissions_kg = sum(emissions_fofm_kgco2, na.rm = TRUE))

sum(allforagefish_raster$forage_emissions_tonnes) #67529789 - 2017

forage_fish_raster_t <- raster::subs(watson_raster_template, allforagefish_raster, by = "Cell", which = "forage_emissions_tonnes", subsWithNA=TRUE)

plot(log(forage_fish_raster_t + 1))
cellStats(forage_fish_raster_t, "sum", na.rm=TRUE) #67529789 - 2017

forage_raster_fix_area <- forage_fish_raster_t/raster::area(forage_fish_raster_t)

## Fix CRS 
forage_fish_raster_t <- projectRaster(forage_raster_fix_area, food_raster, method = "ngb")

crs(forage_fish_raster_t)
plot(log(forage_fish_raster_t+1))
cellStats(forage_fish_raster_t, "sum", na.rm = TRUE) # 994640.6

## Now multiply by area again
forage_fish_raster_t_final <- forage_fish_raster_t*raster::area(forage_fish_raster_t)
crs(forage_fish_raster_t_final)
plot(log(forage_fish_raster_t_final+1))
cellStats(forage_fish_raster_t_final, "sum", na.rm = TRUE) # 67529616



## save to intermediate dataprep folder on aurora
writeRaster(forage_fish_raster_t_final, file.path(ghg_folder, "watson_v5_emissions/emissions_rasters/tonnes/marine_fofm_fisheries_meat_ghg.tif"), overwrite=TRUE)



## Large pelagics 
large_pelagic_raster <- catch_emissions_all %>%
  dplyr::filter(species_class_int == "Large pelagic", forage_fish ==0) %>%
  dplyr::group_by(Cell) %>%
  dplyr::summarize(emissions_tonnes = sum(emissions_tco2, na.rm=TRUE), 
                   emissions_kg = sum(emissions_kgco2, na.rm = TRUE))

large_pelagic_raster_t <- raster::subs(watson_raster_template, large_pelagic_raster, by = "Cell", which = "emissions_tonnes", subsWithNA=TRUE)

plot(log(large_pelagic_raster_t + 1))
cellStats(large_pelagic_raster_t, "sum", na.rm=TRUE) #31362801 - 2017


large_pelagic_area_fix <- large_pelagic_raster_t/raster::area(large_pelagic_raster_t)
cellStats(large_pelagic_area_fix, "sum")

## Fix CRS 
large_pelagic_raster_t <- projectRaster(large_pelagic_area_fix, food_raster, method = "ngb")

crs(large_pelagic_raster_t)
plot(large_pelagic_raster_t)
large_pelagic_raster_t
cellStats(large_pelagic_raster_t, "sum")

large_pelagic_raster_t_final <- large_pelagic_raster_t*raster::area(large_pelagic_raster_t)
cellStats(large_pelagic_raster_t_final, "sum") # 31362707 - close enough

## save the intermediate data prep on aurora
writeRaster(large_pelagic_raster_t_final, file.path(ghg_folder, "watson_v5_emissions/emissions_rasters/tonnes/marine_large-pelagic_fisheries_meat_ghg.tif"), overwrite=TRUE)


## Medium pelagics 
med_pelagic_raster <- catch_emissions_all %>%
  dplyr::filter(species_class_int == "Medium pelagic", forage_fish == 0) %>%
  dplyr::group_by(Cell) %>%
  dplyr::summarize(emissions_tonnes = sum(emissions_tco2, na.rm=TRUE), 
                   emissions_kg = sum(emissions_kgco2, na.rm = TRUE))

med_pelagic_raster_t <- raster::subs(watson_raster_template, med_pelagic_raster, by = "Cell", which = "emissions_tonnes", subsWithNA=TRUE)

plot(log(med_pelagic_raster_t + 1))
cellStats(med_pelagic_raster_t, "sum", na.rm=TRUE) # 26422906 - 2017

med_pelagic_area_fix <- med_pelagic_raster_t/raster::area(med_pelagic_raster_t)
cellStats(med_pelagic_area_fix, "sum") # 9944.209


## Fix CRS 
med_pelagic_raster_t <- projectRaster(med_pelagic_area_fix, food_raster, method = "ngb")

crs(med_pelagic_raster_t)
plot(med_pelagic_raster_t)
med_pelagic_raster_t
cellStats(med_pelagic_raster_t, "sum") # 357991.5

med_pelagic_raster_t_final <- med_pelagic_raster_t*raster::area(med_pelagic_raster_t)
cellStats(med_pelagic_raster_t_final, "sum") # 26422832

## save the intermediate data prep on aurora
writeRaster(med_pelagic_raster_t_final, file.path(ghg_folder, "watson_v5_emissions/emissions_rasters/tonnes/marine_medium-pelagic_fisheries_meat_ghg.tif"), overwrite=TRUE)



## Small pelagics 
small_pelagic_raster <- catch_emissions_all %>%
  dplyr::filter(species_class_int == "Small pelagic", forage_fish == 0) %>%
  dplyr::group_by(Cell) %>%
  dplyr::summarize(emissions_tonnes = sum(emissions_tco2, na.rm=TRUE), 
                   emissions_kg = sum(emissions_kgco2, na.rm = TRUE))

small_pelagic_raster_t <- raster::subs(watson_raster_template, small_pelagic_raster, by = "Cell", which = "emissions_tonnes", subsWithNA=TRUE)

plot(log(small_pelagic_raster_t + 1))
cellStats(small_pelagic_raster_t, "sum", na.rm=TRUE) #4524598 - 2017

small_pelagic_area_fix <- small_pelagic_raster_t/raster::area(small_pelagic_raster_t)

## Fix CRS 
small_pelagic_raster_t <- projectRaster(small_pelagic_area_fix, food_raster, method = "ngb")

crs(small_pelagic_raster_t)
plot(log(small_pelagic_raster_t+1))
small_pelagic_raster_t
cellStats(small_pelagic_raster_t, "sum")

small_pelagic_raster_t_final <- small_pelagic_raster_t*raster::area(small_pelagic_raster_t)
cellStats(small_pelagic_raster_t_final, "sum") # 4524590

## save the intermediate data prep on aurora
writeRaster(small_pelagic_raster_t_final, file.path(ghg_folder, "watson_v5_emissions/emissions_rasters/tonnes/marine_small-pelagic_fisheries_meat_ghg.tif"), overwrite=TRUE)



## Large demersal
large_dem_raster <- catch_emissions_all %>%
  dplyr::filter(species_class_int == "Large demersal", forage_fish == 0) %>%
  dplyr::group_by(Cell) %>%
  dplyr::summarize(emissions_tonnes = sum(emissions_tco2, na.rm=TRUE), 
                   emissions_kg = sum(emissions_kgco2, na.rm = TRUE))

large_dem_raster_t <- raster::subs(watson_raster_template, large_dem_raster, by = "Cell", which = "emissions_tonnes", subsWithNA=TRUE)

plot(log(large_dem_raster_t + 1))
cellStats(large_dem_raster_t, "sum", na.rm=TRUE) #24044356 - 2017

plot(log(large_dem_raster_t+1))


large_dem_area_fix <- large_dem_raster_t/raster::area(large_dem_raster_t)

## Fix CRS 
large_dem_raster_t <- projectRaster(large_dem_area_fix, food_raster, method = "ngb")

crs(large_dem_raster_t)
plot(log(large_dem_raster_t + 1))
large_dem_raster_t
cellStats(large_dem_raster_t, "sum")

large_dem_raster_t_final <- raster::area(large_dem_raster_t)*large_dem_raster_t
cellStats(large_dem_raster_t_final, "sum") # 24044309

## save the intermediate data prep on aurora
writeRaster(large_dem_raster_t_final, file.path(ghg_folder, "watson_v5_emissions/emissions_rasters/tonnes/marine_large-demersal_fisheries_meat_ghg.tif"), overwrite=TRUE)



## Medium demersal
med_dem_raster <- catch_emissions_all %>%
  dplyr::filter(species_class_int == "Medium demersal", forage_fish == 0) %>%
  dplyr::group_by(Cell) %>%
  dplyr::summarize(emissions_tonnes = sum(emissions_tco2, na.rm=TRUE), 
                   emissions_kg = sum(emissions_kgco2, na.rm = TRUE))

med_dem_raster_t <- raster::subs(watson_raster_template, med_dem_raster, by = "Cell", which = "emissions_tonnes", subsWithNA=TRUE)

plot(log(med_dem_raster_t + 1))
cellStats(med_dem_raster_t, "sum", na.rm=TRUE) #26855176 - 2017

med_dem_area_fix <- med_dem_raster_t/raster::area(med_dem_raster_t)
cellStats(med_dem_area_fix, "sum") # 11347.59

## Fix CRS 
med_dem_raster_t <- projectRaster(med_dem_area_fix, food_raster, method = "ngb")

crs(med_dem_raster_t)
plot(log(med_dem_raster_t + 1))
med_dem_raster_t
cellStats(med_dem_raster_t, "sum") # 408513.1

med_dem_raster_t_final <- med_dem_raster_t*raster::area(med_dem_raster_t)
cellStats(med_dem_raster_t_final, "sum") # 26855110

## save the intermediate data prep on aurora
writeRaster(med_dem_raster_t_final, file.path(ghg_folder, "watson_v5_emissions/emissions_rasters/tonnes/marine_medium-demersal_fisheries_meat_ghg.tif"), overwrite=TRUE)


## Small demersal
small_dem_raster <- catch_emissions_all %>%
  dplyr::filter(species_class_int == "Small demersal", forage_fish == 0) %>%
  dplyr::group_by(Cell) %>%
  dplyr::summarize(emissions_tonnes = sum(emissions_tco2, na.rm=TRUE), 
                   emissions_kg = sum(emissions_kgco2, na.rm = TRUE))

small_dem_raster_t <- raster::subs(watson_raster_template, small_dem_raster, by = "Cell", which = "emissions_tonnes", subsWithNA=TRUE)

plot(log(small_dem_raster_t + 1))
cellStats(small_dem_raster_t, "sum", na.rm=TRUE) #136464334 - 2017

small_dem_area_fix <- small_dem_raster_t/raster::area(small_dem_raster_t)
cellStats(small_dem_area_fix, "sum") # 53246.32

## Fix CRS 
small_dem_raster_t <- projectRaster(small_dem_area_fix, food_raster, method = "ngb")

crs(small_dem_raster_t)
plot(log(small_dem_raster_t + 1))
small_dem_raster_t
cellStats(small_dem_raster_t, "sum")

small_dem_raster_t_final <- small_dem_raster_t*raster::area(small_dem_raster_t)
cellStats(small_dem_raster_t_final, "sum") # 136463965

## save the intermediate data prep on aurora
writeRaster(small_dem_raster_t_final, file.path(ghg_folder, "watson_v5_emissions/emissions_rasters/tonnes/marine_small-demersal_fisheries_meat_ghg.tif"), overwrite=TRUE)



## Large reef-associated
large_reef_associated_raster <- catch_emissions_all %>%
  dplyr::filter(species_class_int == "Large reef-associated", forage_fish == 0) %>%
  dplyr::group_by(Cell) %>%
  dplyr::summarize(emissions_tonnes = sum(emissions_tco2, na.rm=TRUE), 
                   emissions_kg = sum(emissions_kgco2, na.rm = TRUE))

large_rf_raster_t <- raster::subs(watson_raster_template, large_reef_associated_raster, by = "Cell", which = "emissions_tonnes", subsWithNA=TRUE)

plot(log(large_rf_raster_t + 1))
cellStats(large_rf_raster_t, "sum", na.rm=TRUE) # 7220947 - 2017

large_rf_area_fix <- large_rf_raster_t/raster::area(large_rf_raster_t)
cellStats(large_rf_area_fix, "sum") # 2537.094


## Fix CRS 
large_rf_raster_t <- projectRaster(large_rf_area_fix, food_raster, method = "ngb")

crs(large_rf_raster_t)
plot(log(large_rf_raster_t + 1))
large_rf_raster_t
cellStats(large_rf_raster_t, "sum")

large_rf_raster_t_final <- large_rf_raster_t*raster::area(large_rf_raster_t)
cellStats(large_rf_raster_t_final, "sum") # 7220925

## save the intermediate data prep on aurora
writeRaster(large_rf_raster_t_final, file.path(ghg_folder, "watson_v5_emissions/emissions_rasters/tonnes/marine_large-reef_fisheries_meat_ghg.tif"), overwrite=TRUE)


## Medium reef-associated
med_rf_raster <- catch_emissions_all %>%
  dplyr::filter(species_class_int == "Medium reef-associated", forage_fish == 0) %>%
  dplyr::group_by(Cell) %>%
  dplyr::summarize(emissions_tonnes = sum(emissions_tco2, na.rm=TRUE), 
                   emissions_kg = sum(emissions_kgco2, na.rm = TRUE))

med_rf_raster_t <- raster::subs(watson_raster_template, med_rf_raster, by = "Cell", which = "emissions_tonnes", subsWithNA=TRUE)

plot(log(med_rf_raster_t + 1))
cellStats(med_rf_raster_t, "sum", na.rm=TRUE) #16888647 - 2017

med_rf_area_fix <- med_rf_raster_t/raster::area(med_rf_raster_t)
cellStats(med_rf_area_fix, "sum")


## Fix CRS 
med_rf_raster_t <- projectRaster(med_rf_area_fix, food_raster, method = "ngb")

crs(med_rf_raster_t)
plot(log(med_rf_raster_t + 1))
med_rf_raster_t
cellStats(med_rf_raster_t, "sum")


med_rf_raster_t_final <- med_rf_raster_t*raster::area(med_rf_raster_t)
cellStats(med_rf_raster_t_final, "sum") # 16888594


## save the intermediate data prep on aurora
writeRaster(med_rf_raster_t_final, file.path(ghg_folder, "watson_v5_emissions/emissions_rasters/tonnes/marine_medium-reef_fisheries_meat_ghg.tif"), overwrite=TRUE)


## Small reef-associated
small_rf_raster <- catch_emissions_all %>%
  dplyr::filter(species_class_int == "Small reef-associated", forage_fish == 0) %>%
  dplyr::group_by(Cell) %>%
  dplyr::summarize(emissions_tonnes = sum(emissions_tco2, na.rm=TRUE), 
                   emissions_kg = sum(emissions_kgco2, na.rm = TRUE))

small_rf_raster_t <- raster::subs(watson_raster_template, small_rf_raster, by = "Cell", which = "emissions_tonnes", subsWithNA=TRUE)

plot(log(small_rf_raster_t + 1))
cellStats(small_rf_raster_t, "sum", na.rm=TRUE) #4085745 - 2017

small_rf_area_fix <- small_rf_raster_t/raster::area(small_rf_raster_t)
cellStats(small_rf_area_fix, "sum")


## Fix CRS 
small_rf_raster_t <- projectRaster(small_rf_area_fix, food_raster, method = "ngb")

crs(small_rf_raster_t)
plot(log(small_rf_raster_t + 1))
small_rf_raster_t
cellStats(small_rf_raster_t, "sum")

small_rf_raster_t_final <- small_rf_raster_t*raster::area(small_rf_raster_t)
cellStats(small_rf_raster_t_final, "sum") # 4085734


## save the intermediate data prep on aurora
writeRaster(small_rf_raster_t_final, file.path(ghg_folder, "watson_v5_emissions/emissions_rasters/tonnes/marine_small-reef_fisheries_meat_ghg.tif"), overwrite=TRUE)



## Benthic
benthic_raster <- catch_emissions_all %>%
  dplyr::filter(species_class_int == "Benthic", forage_fish == 0) %>%
  dplyr::group_by(Cell) %>%
  dplyr::summarize(emissions_tonnes = sum(emissions_tco2, na.rm=TRUE), 
                   emissions_kg = sum(emissions_kgco2, na.rm = TRUE))

benthic_raster_t <- raster::subs(watson_raster_template, benthic_raster, by = "Cell", which = "emissions_tonnes", subsWithNA=TRUE)

plot(log(benthic_raster_t + 1))
cellStats(benthic_raster_t, "sum", na.rm=TRUE) # 37424377 - 2017

benthic_area_fix <- benthic_raster_t/raster::area(benthic_raster_t)
cellStats(benthic_area_fix, "sum")



## Fix CRS 
benthic_raster_t  <- projectRaster(benthic_area_fix , food_raster, method = "ngb")

crs(benthic_raster_t)
plot(log(benthic_raster_t  + 1))
benthic_raster_t 
cellStats(benthic_raster_t, "sum")

benthic_raster_t_final <- benthic_raster_t*raster::area(benthic_raster_t)
cellStats(benthic_raster_t_final, "sum") # 37424276

## save the intermediate data prep on aurora
writeRaster(benthic_raster_t_final, file.path(ghg_folder, "watson_v5_emissions/emissions_rasters/tonnes/marine_benthic_fisheries_meat_ghg.tif"), overwrite=TRUE)
   
             
```

Save final rasters by final species classes

```{r}
## FOFM
## save to final datalayer folder on aurora
cellStats(forage_fish_raster_t_final, "sum") # 67529616
writeRaster(forage_fish_raster_t_final, file.path(final, "marine_fofm_fisheries_meat_ghg.tif"), overwrite=TRUE)


## Large pelagics
## save the final data layer to aurora
cellStats(large_pelagic_raster_t_final, "sum") # 31362707
writeRaster(large_pelagic_raster_t_final, file.path(final, "marine_large-pelagic_fisheries_meat_ghg.tif"), overwrite=TRUE)

## Medium pelagics
## save the final data layer to aurora
cellStats(med_pelagic_raster_t_final, "sum") # 26422832
writeRaster(med_pelagic_raster_t_final, file.path(final, "marine_medium-pelagic_fisheries_meat_ghg.tif"), overwrite=TRUE)

## Small pelagics 
## save the final data layer to aurora
cellStats(small_pelagic_raster_t_final, "sum") # 4524590
writeRaster(small_pelagic_raster_t_final, file.path(final, "marine_small-pelagic_fisheries_meat_ghg.tif"), overwrite=TRUE)


## Demersal
dem_raster <- catch_emissions_all %>%
  dplyr::filter(species_class_fin == "Demersal", forage_fish == 0) %>%
  dplyr::group_by(Cell) %>%
  dplyr::summarize(emissions_tonnes = sum(emissions_tco2, na.rm=TRUE), 
                   emissions_kg = sum(emissions_kgco2, na.rm = TRUE))

dem_raster_t <- raster::subs(watson_raster_template, dem_raster, by = "Cell", which = "emissions_tonnes", subsWithNA=TRUE)

plot(log(dem_raster_t + 1))
cellStats(dem_raster_t, "sum", na.rm=TRUE) #187363867 - 2017
dem_raster_t

dem_area_fix <- dem_raster_t/raster::area(dem_raster_t)
cellStats(dem_area_fix, "sum")


## Fix CRS
dem_raster_t  <- projectRaster(dem_area_fix , food_raster, method = "ngb")

crs(dem_raster_t)
plot(log(dem_raster_t  + 1))
dem_raster_t
cellStats(dem_raster_t, "sum")

dem_raster_t_final <- dem_raster_t*raster::area(dem_raster_t)
cellStats(dem_raster_t_final, "sum") # 187363385

## save the final data layer to aurora
writeRaster(dem_raster_t_final, file.path(final, "marine_demersal_fisheries_meat_ghg.tif"), overwrite=TRUE)


## reef-associated
rf_raster <- catch_emissions_all %>%
  dplyr::filter(species_class_fin == "Reef-associated", forage_fish == 0) %>%
  dplyr::group_by(Cell) %>%
  dplyr::summarize(emissions_tonnes = sum(emissions_tco2, na.rm=TRUE), 
                   emissions_kg = sum(emissions_kgco2, na.rm = TRUE))

rf_raster_t <- raster::subs(watson_raster_template, rf_raster, by = "Cell", which = "emissions_tonnes", subsWithNA=TRUE)

plot(log(rf_raster_t + 1))
cellStats(rf_raster_t, "sum", na.rm=TRUE) #28195338 - 2017

rf_area_fix <- rf_raster_t/raster::area(rf_raster_t)
cellStats(rf_area_fix, "sum")


## Fix CRS
rf_raster_t  <- projectRaster(rf_area_fix, food_raster, method = "ngb")

crs(rf_raster_t)
plot(log(rf_raster_t  + 1))
rf_raster_t
cellStats(rf_raster_t, "sum")

rf_raster_t_final <- rf_raster_t*raster::area(rf_raster_t)
cellStats(rf_raster_t_final, "sum") # 28195253


## save the final data layer to aurora
writeRaster(rf_raster_t_final, file.path(final, "marine_reef_fisheries_meat_ghg.tif"), overwrite=TRUE)


## Benthic
## save the final data layer to aurora
cellStats(benthic_raster_t_final, "sum") # 37424276 
writeRaster(benthic_raster_t_final, file.path(final, "marine_benthic_fisheries_meat_ghg.tif"), overwrite=TRUE)
```

Datacheck
Lets do some comparing to Parker 2018 estimates 
```{r, eval = FALSE}

#load relevant packages, etc.

library(here)
library(raster)
library(tidyverse)
library(countrycode)

# function to project raster
## Use correct CRS

watson_raster_template <- raster::raster(ncol=720, nrow=360, vals=c(1:259200))
source(here("fisheries/marine/ghg/raster_crs.R"))


raw_folder <- file.path("/home/shares/food-systems/Food_footprint/_raw_data")
raw_folder_watson <- file.path("/home/shares/food-systems/Food_footprint/_raw_data/watson_data/v5.0/")
ghg_folder <- file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/fisheries/marine/ghg")
final <- file.path("/home/shares/food-systems/Food_footprint/all_food_systems/datalayers")


## read in emissions data
catch_emissions <- read_csv(file.path(ghg_folder, "watson_v5_emissions/all_catch_emissions_2017.csv")) 

sum(catch_emissions$catch)
# [1] 111226186 tonnes = 111226186000 kg

## values from final ghg rasters 
# benthic  # reef     # demersal  #small pel # med pel # large pel # fofm
37424276 + 28195253 + 187363385 + 4524590 + 26422832 + 31362707 + 67529616 # we estimate 382822659 tonnes co2eq total

## quick check using our mean EI 
111226186000*3.50801/10000 # 390182573 -- makes sense... the estimate seen above is using a more specific per species/rgn EI allocation, not just the mean. 


#### Using the EI from parker et al. 
# lets guess that with non industrial, parker wouldve had ~100 million landings. How much co2eq t can we guess from ~ 20 million tonnes landings
mean(catch_emissions$ei_kgco2_kgcatch) # mean emissions intensities kgco2/kgcatch is 3.50801
20*1000 #20000 million kg catch non industrial in 2011
20000*2.2 ## estimate 44000 million kgco2eq emissions in 2011 from non industrial 
44000/1000 ## 44 million non industrial tco2eq in 2011
81000*2.2/1000 # 178.2 million non industrial tco2eq in 2011

178.2+44 # 222.2 million tonnes co2eq in 2011 using parkers EI of 2.2 kg CO2eq per kg

## this is a large difference from our estimate

## this is because the EI that they used was much lower than ours. They had a mean EI of 2.2 kg CO2-eq per kg... we had a mean of 3.50801 kg CO2-eq per kg... what if we used 2.2 kg CO2eq per kg on our catch estimate: 

111226186000*2.2/1000 # 244697609 ## makes sense 


# Use our mean EI on the tonnes of production from Parker paper
# lets guess that with non industrial, parker wouldve had ~100 million landings. How much co2eq t can we guess from ~ 20 million tonnes landings
mean(catch_emissions$ei_kgco2_kgcatch) # mean emissions intensities kgco2/kgcatch is 3.50801
20*1000 #20000 million kg catch non industrial in 2011
(81000+20000)*3.50801 ## estimate 354309 kgco2eq emissions in 2011 
354309/1000 ## 354.309 million tco2eq in 2011 from parkers tonnes estimates using OUR EI

## this matches pretty well to our estimate
```

