---
title: "Modeling salmon production (tonnes) to total number"
author: "Juliette Verstaen"
date: "1/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries and paths}
library(dplyr)
library(tidyr)
library(data.table)
library(countrycode)
library(readxl)
library(stringi)
library(stringr)
library(ggplot2)
library(readr)
library(here)

raw_path <- "/home/shares/food-systems/Food_footprint/_raw_data/"
```

## About 

Calculating conversion factor converting production of salmon reported in tonnage to the number of individual salmon.

## Data

1. For this we are going to use the data found [here](https://www.ssb.no/en/statbank/table/09259/), which documents for each county: 
- The stock number as of Jan 1st
- Supplies - how many additional fish were added
- Delivered for sale - how many fish were removed for consumption
- Loss - number of fish died
- The stock number as of Dec 31st
- Percent loss

2. [Aquaculture sales](https://www.ssb.no/en/statbank/table/07326)


## Methods

### Norwegian Fisheries Directorate data wrangling
First we're going to read in the sub regional production data we have, and keep all the years. We are also going to gapfill this information. 

# Norway
Data from:
https://www.ssb.no/en/statbank/table/09259/
09259: Aquaculture. Stock and loss of live fish for food (rough classification) (C) 2010 - 2018 (replaced ":" with NA)

https://www.ssb.no/en/statbank/table/07326
07326: Aquaculture. Sales of slaughtered fish for food, by region, fish species, contents and year
(replaced "-" with 0; replaced, ":" with NA)

Downloaded: Jan 22 2020


Gapfilling missing data.  loss-percent variable is cut because it cannot be gapfilled.
```{r}

#gapfill missing data

norway_mar <- read.csv("/home/shares/food-systems/Food_footprint/_raw_data/Regional_aquaculture_production/Norway/norway_mariculture.csv") %>%
  gather(key = "year", value = "value", -region, -species, -variable, -units) %>%
  mutate(year = gsub("X", "", year)) %>%
  filter(variable != "loss_percent")


# first round of gapfilling if every year is NA or 0, then 0
norway_mar_gf <- norway_mar %>%
  group_by(region, species, variable) %>%
  mutate(total_sp = sum(value, na.rm=TRUE)) %>%   # if all NA or 0 for a species across years, then 0
  mutate(gapfilled = ifelse(total_sp==0 & is.na(value), "gapfilled", NA)) %>%
    mutate(value = ifelse(total_sp == 0 & is.na(value), 0, value)) %>%
  dplyr::select(-total_sp)

# second round of gapfilling, gapfill based on unaccounted fish and avg proportion of aquaculture in previous years

avg_totals <- filter(norway_mar, species == "Total") %>%
  group_by(region, variable) %>%
  summarize(total_mean = mean(value))

totals <- filter(norway_mar, species=="Total") %>%
  dplyr::select(-species, total_value=value)


norway_mar_gf <- norway_mar_gf %>%
  left_join(avg_totals) %>%               
  left_join(totals) %>%
  group_by(region, species, variable) %>%
  mutate(avg_value = mean(value, na.rm=TRUE)) %>%
  ungroup() %>%
  mutate(proportion = avg_value/total_mean) %>%
  group_by(region, variable, year) %>%
  mutate(total_obs = sum(value[.$species != "Total"], na.rm=TRUE)) %>%
  mutate(unaccounted = total_value - total_obs) %>%
  mutate(unaccounted = ifelse(unaccounted < 0, 0, unaccounted))  %>%     # some negatives due to rounding and possible data entry error
  mutate(avg_value_missing = proportion * ifelse(is.na(value), 1, 0)) %>%                            # figure out proportion of only missing
   mutate(prop_value_missing = avg_value_missing/sum(avg_value_missing, na.rm=TRUE)) %>%
  mutate(gf_value = prop_value_missing * unaccounted) %>%                           # gapfill missing data with unaccounted production multiplied by the average proportion of production for each species in previous years 
mutate(gapfilled = ifelse(is.na(value) & !is.na(gf_value), "gapfilled", gapfilled)) %>%
  mutate(value = ifelse(is.na(value), gf_value, value))

## watch these be gapfilled
summary(norway_mar_gf) # make sure no NA values
data.frame(filter(norway_mar_gf, is.na(value)))
filter(norway_mar_gf, year==2017)
filter(norway_mar_gf, region=="Rogaland", variable == "tonnes_fish")

summary(norway_mar_gf)
dim(norway_mar_gf)
summary(norway_mar)
dim(norway_mar)

## clean and save data
norway_mar_gf <- norway_mar_gf %>%
dplyr::select(region, species, variable, units, year, value, gapfilled) %>%
  filter(species != "Total")
write.csv(norway_mar_gf, here("aquaculture/marine/salmon/salmon_disturbance/data/norway_mariculture_gf.csv"), row.names=FALSE ) 
  
```

```{r wrangling production data}
nor_prod <- read_csv(here("aquaculture/marine/salmon/salmon_disturbance/data/norway_mariculture_gf.csv")) %>% 
  filter(species == "Salmon") %>% 
  dplyr:: select(year, region,variable, value) %>% 
  spread(variable, value) %>% 
  rename(salmon_tonnes = tonnes_fish) %>% 
  mutate(Stock_Jan1_count = Stock_Jan1_count*1000,
         loss_count = loss_count*1000,
         stock_dec_31_count = stock_dec_31_count*1000,
         delivered_for_sale_count = delivered_for_sale_count*1000,
         supplies_count = supplies_count*1000)

```

Checking the total annual sum of salmon production in Norway
```{r annual totals}
## combine them all together for total salmon in all of Norway.
nor_pro_county<- nor_prod %>% 
  group_by(year) %>% 
  summarize(total_salmon_tonnes = sum(salmon_tonnes))

nor_pro_county
```

Because graphs are fun!!
```{r graph production by region}
# graph!
plot <- ggplot(nor_prod)+
  geom_line(aes(x=year, y = salmon_tonnes, color=region))+
  ggtitle("Salmon Production in Norwegian Counties") +
  ylab("Tonnage of Salmon") +
  xlab("Year") +
  theme_classic() 
plot

```

###SSB Data wrangling

**Pulling out Nord and Sor** 
Up until 2017 Trøndelag was Nord-Trøndelag and Sør-Trøndelag. Since the production data we have combines both of these counties during that time, we are going to do the same here to avoid unnecessary errors in our modeling. Here I pull out that data to work on it so I can better see what's happening. 

```{r nord_tron and sor_tron wrangling}

trond <- nor_prod %>% 
  filter(region == "Nord-Trøndelag" | region == "Sør-Trøndelag" | region == "Trøndelag") %>% 
  group_by(year) %>%
  
  mutate(t_stock = sum(Stock_Jan1_count),
         t_prod = sum(salmon_tonnes),
         t_loss_count = sum(loss_count),
         t_stock_dec_31_count = sum(stock_dec_31_count),
         t_sale = sum(delivered_for_sale_count),
         t_supplies_count = sum(supplies_count)) %>% 
  
  dplyr::select(year, t_stock, t_prod, t_loss_count, t_stock_dec_31_count, t_sale, t_supplies_count) %>% 
  
  rename(Stock_Jan1_count = t_stock,
         salmon_tonnes = t_prod,
         delivered_for_sale_count = t_sale,
         stock_dec_31_count = t_stock_dec_31_count,
         loss_count = t_loss_count,
         supplies_count = t_supplies_count) %>% 
  
  mutate(region = "Trøndelag") %>% 
    unique() 

```

**Finishing up SSB data wrangling**
```{r finalizing ssb data wrangling}
nor_prod <- nor_prod %>% 
  ### get rid of the Trondelag, Sor, and Nord that we will add back in consolidated
  filter(region != "Nord-Trøndelag" & region != "Sør-Trøndelag" & region != "Trøndelag") %>%
  bind_rows(trond)

plot <- ggplot(nor_prod)+
  geom_line(aes(x=year, y = salmon_tonnes, color=region))+
  ggtitle("Salmon Production in Norwegian Counties") +
  ylab("Tonnage of Salmon") +
  xlab("Year") +
  theme_classic() 
plot

```

###Combine both data sets


```{r combine data sets}
all_salmon <- nor_prod %>% 
  mutate(salmon_kg = salmon_tonnes*1000,
         avg_harvest_wgt_kg = salmon_kg/delivered_for_sale_count,
         mortality_rate = loss_count/(Stock_Jan1_count)) 
#%>%   filter(num_loss != is.na(num_loss))

write.csv(all_salmon, file = "data/all_salmon_norway.csv")
```

Check to see if the math for adding the number of salmon at the end of the year matches the reported values
```{r}
sal_check <- all_salmon %>% 
  mutate(test = Stock_Jan1_count  + supplies_count - loss_count - delivered_for_sale_count,
         diff = test - stock_dec_31_count,
         perdiff = diff/stock_dec_31_count)
```

Some of these values are off by either 1000 or 2000 fish (weird that they are such even numbers...) but this is mostly a drop in the bucket in the ground scheme of things, so we won't worry about it too much


```{r explore data}
### does larger production mean increased mortality rate?
plot(all_salmon$salmon_kg, all_salmon$mortality_rate) #no

### which counties have higher mortality rates
ggplot(all_salmon)+
  geom_line(aes(x= year, y = mortality_rate, color= region))

salmon_model <- lm(delivered_for_sale_count ~salmon_tonnes, data = all_salmon)
salmon_model
summary(salmon_model)
plot(salmon_model)

ggplot(all_salmon) +
  geom_point(aes(x= salmon_tonnes, y = delivered_for_sale_count))+
  xlab("Tonnage of Salmon") +
  ylab("Number Salmon Sold") +
  theme_classic() +
  stat_smooth(aes(x= salmon_tonnes, y = delivered_for_sale_count), method = "lm")

plot(all_salmon$loss_count, all_salmon$delivered_for_sale_count)
```

Simple model conversion production to numbers
tanding_stock_beginning_year ~ number_produced/sold - 1
```{r conversion production to numbers}

salm_conv_model <- lm(Stock_Jan1_count ~ delivered_for_sale_count - 1, data = sal_check)
salm_conv_model
summary(salm_conv_model)

ggplot(all_salmon)+
  geom_point(aes(x= delivered_for_sale_count, y = Stock_Jan1_count)) + 
  geom_abline(intercept=0, slope = 1.477)
```



Making histogram graphs for harvest weights 
- mean and standard deviations

```{r}

all_salmon_2 <- all_salmon %>% 
  mutate(sd = sd(avg_harvest_wgt_kg),
         mean = mean(avg_harvest_wgt_kg))


hist_harv_weight <- ggplot(all_salmon) +
  geom_histogram(aes(x= avg_harvest_wgt_kg), fill = "gray65") +
  ggtitle("Norwegian Salmon Harvest Weight") +
  ylab("Count") +
  xlab("Harvest Weight (kg)") +
  theme_classic() +
  geom_vline(aes(xintercept = mean(avg_harvest_wgt_kg)) ,col='dodgerblue2',size=1)+
  geom_vline(aes(xintercept = mean(avg_harvest_wgt_kg) + sd(avg_harvest_wgt_kg)), col='seagreen3', size = 0.5)+
  geom_vline(aes(xintercept = mean(avg_harvest_wgt_kg) - sd(avg_harvest_wgt_kg)), col='seagreen3', size = 0.5)+
  geom_vline(aes(xintercept = mean(avg_harvest_wgt_kg) - 2*sd(avg_harvest_wgt_kg)), col='seagreen3', size = 0.5)+
  geom_vline(aes(xintercept = mean(avg_harvest_wgt_kg) + 2*sd(avg_harvest_wgt_kg)), col='seagreen3', size = 0.5) +
  theme(legend.position = "none") 
  
hist_harv_weight

mean(all_salmon$avg_harvest_wgt_kg)
#4.907344
sd(all_salmon$avg_harvest_wgt_kg)
#0.3927127
min(all_salmon$avg_harvest_wgt_kg)
#3.30712
max(all_salmon$avg_harvest_wgt_kg)
#6.53237

```




















