---
title: "occupancy_distribution"
author: "Juliette Verstaen"
date: "November 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r read in libaries and paths}
library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
library(tidyverse)
library(raster)
library(data.table)
library(raster)
library(rgdal)
library(here)

path_final <- "/home/shares/food-systems/Food_footprint/all_food_systems/datalayers"

path_prep <- "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/salmon/"

path_updated <- "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/salmon/updated"

food_crs <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"

food_raster <- raster(nrows=2160, ncols=4320, xmn=-180, xmx=180, ymn=-90, ymx=90)

```

In this markdown we are calculating:

1. How many live fish are in the water at a given time.
2. The total volume (and area) used by salmon farming

##Methods

##Conversion Table
Read in and save the necessary values from the salmon conversion table
```{r parameters from conversion table}
salmon_conversion <- read_csv(here("aquaculture/marine/salmon/salmon_disturbance/data/salmon_conversion_factors.csv"))

stocking_density <- as.numeric(salmon_conversion[salmon_conversion$name == "stocking_density", "value"])
live_harvested_conversion <- as.numeric(salmon_conversion[salmon_conversion$name == "live_harvested", "value"])
harvest_weight <- as.numeric(salmon_conversion[salmon_conversion$name == "mean_harvest_weight", "value"])
cage_volume <- as.numeric(salmon_conversion[salmon_conversion$name == "cage_volume", "value"])
```

###Calculate the total number of fish in the water (occupancy)
Read in production map

```{r}
production_raster <- rasterFromXYZ(read_csv(paste0(path_updated, "/salmon_farm_updated.csv")), crs = food_crs) 

#plot(production_raster)
cellStats(production_raster, stat = "sum") # 2756155

## check to make sure that number is correct - it is close enough. The difference is due to rounding
salmon_fao <- read_csv("aquaculture/marine/salmon/data/fao_mariculture_salmon.csv") %>%
  filter(year == 2017)
sum(salmon_fao$fao_tonnes_production) # 2753752 

```

```{r calculate salmon occupancy}
#as total number of salmon
salmon_occupancy <- (production_raster*1000)/harvest_weight * live_harvested_conversion
writeRaster(salmon_occupancy, file.path(path_prep, "salmon_occupancy/salmon_occupancy.tif"), format = "GTiff", overwrite = TRUE)

# as live weight of salmon, in kg
salmon_live_kg  <- salmon_occupancy*harvest_weight
salmon_live_tonnes  <- salmon_occupancy*harvest_weight/1000

writeRaster(salmon_live_tonnes, file.path(path_prep, "salmon_occupancy/salmon_live_tonnes.tif"), format = "GTiff", overwrite = TRUE)
```


### Calculate the total area (from an aerial view) used for salmon aquaculture

First we determine how many cages are required given cage dimensions, stocking densities, and kg living fish. 

We used Gentry 2017 data to estimate 180,000 kg salmon (harvest weight) per cage.

This is based on 9000 m3 cage that we assume to be circular with a depth of 10 m.  We assume a stocking density of 20kg/m3.

```{r}

kg_per_cage <- cage_volume*stocking_density

```

Given the kg of salmon per cell we determine the number of cages needed as:

kg_salmon/180,000

We round this value up to the nearest whole number and multiply by 1.3 to include fallowing cages.

```{r}

cages <- ceiling(salmon_live_kg/kg_per_cage)

cages_pls_fallow <- ceiling(cages * 1.3)

```

2. We the determine how much surface area the cages disturb including the inner cage area, a 5 meter buffer around the cage, and area for infrastructure.

```{r}
cage_depth_m <- 10
buffer <- 5
infrastructure <- 1.5 #50% of area to infrastructure

cage_area <- cage_volume/cage_depth_m
cage_radius <- sqrt(cage_area/pi)
cage_radius_buffer <- cage_radius + buffer
cage_area_with_buffer <- pi*cage_radius_buffer^2
cage_disturbance_m2 <- cage_area_with_buffer*infrastructure

total_disturbance_km2 <- cages_pls_fallow * cage_disturbance_m2/1000000
total_disturbance_prop <- total_disturbance_km2/raster::area(total_disturbance_km2)
plot(total_disturbance_prop)

cellStats(total_disturbance_km2, "sum") # 71.99486

salmon_aquaculture_area_df <- as.data.frame(total_disturbance_prop, xy=TRUE) %>% 
  dplyr::rename(salmon_disturbance_prop = layer )

write_csv(salmon_aquaculture_area_df, file.path(path_prep, "salmon_disturbance_prop.csv"))

salmon_aquaculture_area_df_km2 <- as.data.frame(total_disturbance_km2, xy=TRUE) %>% 
  dplyr::rename(salmon_disturbance_km2 = layer)

write_csv(salmon_aquaculture_area_df_km2, file.path(path_prep, "salmon_disturbance_km2.csv"))

## save to datalayers folder with the correct naming conventions: 
final_layer_dist <- read_csv(file.path(path_prep, "salmon_disturbance_km2.csv"))

final_rast <- rasterFromXYZ(final_layer_dist, crs = food_crs)
crs(final_rast)
plot(final_rast)
cellStats(final_rast, "sum") # 71.99486
final_rast

final_area_fix <- final_rast/raster::area(final_rast)

final_rast_proj <- projectRaster(final_area_fix, food_raster, method = "ngb")

final_rast <- final_rast_proj*raster::area(final_rast_proj)
cellStats(final_rast, "sum") # 71.99486
final_rast

writeRaster(final_rast, file.path(path_final, "marine_salmon_aquaculture_meat_disturbance.tif"), overwrite = TRUE)
```


Look at the pieces more closely
```{r}
norway <- raster::crop(total_disturbance_km2, extent(-25, 50, 50, 75))
plot(norway)
writeRaster(norway, file.path(path_prep, "salmon_occupancy/norway.tif"), overwrite=TRUE)

chile <- raster::crop(total_disturbance_km2, extent(-80, -60, -60, -35))
plot(chile)
writeRaster(chile, file.path(path_prep, "salmon_occupancy/chile.tiff"), overwrite=TRUE)
 
```

Datacheck: 
There isn't a great way to check this, but for now I will compare our estimates to the chicken salmon paper results. Our final disturbance should be a little bit bigger than theirs, since we include more species. 

```{r}
## makes a lot of sense

new_km2 <- read_csv(file.path(path_prep, "salmon_disturbance_km2.csv"))
sum(new_km2$salmon_disturbance_km2) # 71.99486

old_km2 <- raster("/home/shares/food-systems/Food_footprint/all_food_systems/datalayers/marine_salmon_aquaculture_meat_disturbance.tif")

cellStats(old_km2, "sum") # 71.80003

```
