---
title: "Overall paper stats version 2"
output: html_document
---

This script contains code and data to fill out the various statistics requested for the main manuscript. Above each chunk is either an excerpt from the paper, or a specific request for a statistic.

```{r setup, include=FALSE, eval = FALSE, eval = FALSE}
knitr::opts_chunk$set(eval = TRUE)

library(here)
library(tidyverse)

food_raster <- raster::raster(nrows=2160, ncols=4320, xmn=-180, xmx=180, ymn=-90, ymx=90)

library(scales)
library(rnaturalearth)
library(rnaturalearthdata)
library(raster)
library(sf)
library(mapview)
library(doParallel)
source(here("_workflow/common.R"))
source(here("_workflow/common_spatial.R"))

### save the gall peters projection
gall_peters <- "+proj=cea +lon_0=0 +x_0=0 +y_0=0 +lat_ts=45 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

### create a water raster
water <- readOGR(dsn = here("_analysis/raw/rgn_all_gcs_med_res.shp"))
water <- water[!water$rgn_nam == "Antarctica",]
proj4string(water) <- CRS("+proj=longlat +datum=WGS84")
water <- spTransform(water, CRS("+proj=cea +lon_0=0 +x_0=0 +y_0=0 +lat_ts=45 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"))
#plot(water)

### create a land raster
land <- ne_countries(scale = "medium", returnclass = "sp") 
land <- land[!land$sov_a3 == "ATA",]
land <- st_as_sf(land)
land <- st_transform(land, crs=gall_peters)
land <- as(land, "Spatial")

rgn_raw <- read_csv(file.path("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/raw_df_summary/rgn_raw_summary.csv"))

```


## Check the origin of cumulative pressures

**Aquatic systems produce only 1.1% of food but 9.9% of global footprint.**

**Contributions from land (89.9% of global cumulative pressure) vastly outweigh those from oceans (9.9%) or freshwater ecosystems (0.2%), yet these ocean pressures are substantial given that relatively little (1.1%, by tonnes) food and feed for fed animals comes from the sea.**

```{r, eval = FALSE}
rescale <- read_csv("../rescale_values.csv")

### do overall, including feed and fodder
percent_pressure_rescale <- rgn_raw %>%
  left_join(rescale, by = "pressure") %>%
  mutate(global_pressure_prop = sum/global_total) %>%
  group_by(origin) %>%
  summarise(cum_pressure_prop_rescaled = sum(global_pressure_prop),
            cum_pressure = sum(sum)) %>%
  ungroup()

percent_pressure_origin <- percent_pressure_rescale %>%
  mutate(total_cum_prop = sum(cum_pressure_prop_rescaled)) %>%
  mutate(cum_pressure_prop_final = cum_pressure_prop_rescaled/total_cum_prop) %>%
  group_by(origin) %>%
  summarise(prop_final = sum(cum_pressure_prop_final)) %>%
  ungroup() %>%
  mutate(percent_final = round(prop_final*100,2))


write.csv(percent_pressure_origin, "output/percent_cum_pressure_origin.csv")


```


## Check the top 1% of pixels for cumulative pressures

**The top 1% of pixels with respect to cumulative pressures (5,114,880km2 total) fall nearly entirely on land (only 94,608km2, or 1.8% of this top 1%, fall in the ocean, and none in the high seas) and produce 39.4% of food’s global cumulative pressure and 30.9% of assessed tonnage of food. They occur primarily in India, China, the U.S., Brazil, and Indonesia (Fig. 2a). Nearly all pressures (92.5%) are exerted in just 10% of pixels.**


```{r, eval = FALSE}
# Find out what the top 1% of raster cells scores are and get the area of those.

cum_stress <- raster(file.path("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/stressor_summary/equal_area/rescaled/cumulative_stress/all_systems_cumulative_stress.tif"))
plot(cum_stress)
cellStats(cum_stress, "sum", na.rm = TRUE) # 4026176

cum_stress_df <- as.data.frame(cum_stress, xy= TRUE)

top_1_per <- quantile(cum_stress, 0.99)  

top_1_perc <- cum_stress_df[cum_stress_df$all_systems_cumulative_stress > top_1_per,] ### 142080 cells in the top 1%

cum_stress2 <- cum_stress
cum_stress2[cum_stress2 < top_1_per] <- NA ### assign anything not in the top 1% an NA 
plot(cum_stress2) ### looks similar to the dark figures on figure 1

cellStats(cum_stress2, "sum", na.rm = TRUE) # 1584422

1584422/4026176 # 0.3935302 - prop of footprint in the top 1%

cum_stress3 <- cum_stress
cum_stress3[cum_stress3 < top_1_per] <- NA ### assign anything not in the top 1% an NA 
cum_stress3[!is.na(cum_stress3)] <- 1
plot(cum_stress3) ### matches exactly the figure

### area in km2 should be:
# there are 142080 cells in the top 1%
142080*36 # 5114880 km2


### now extract the values from this which are in the ocean
eez_rgns <- raster("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/spatial/gall_peters/eez_rgns_gall_peters.tif")
plot(eez_rgns)

ocean_stress <- data.frame(zonal(cum_stress2, eez_rgns, fun = "sum")) ### ok so some of the cumulative stressors are in the ocean... 

eez_rgns[eez_rgns > 0 ] <- 1
plot(eez_rgns)

land_footprint <- eez_rgns*cum_stress3

plot(land_footprint)
cellStats(land_footprint, "sum") 
# NEW: 2628 ### so the ocean area is 2628*36 = 94608 km2 
94608/5114880 # 0.01849662 = 1.8%

### top 10% calculations for paper 
top_10_per <- quantile(cum_stress, 0.90)
chi_10 <- raster("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/stressor_summary/equal_area/rescaled/cumulative_stress/all_systems_cumulative_stress.tif")
chi_10[chi_10 < top_10_per ] <- 0
cellStats(chi_10, stat = 'sum')


3725460/4026176
# 0.9253098
# 92.5
```

##Determine the production associated with the top 1% of pressures

Reproject the production rasters - added parallel processing with 6 cores (you could do even more to speed up), and now takes ~20 mins for reprojection

```{r}
rgdal::setCPLConfigOption("GDAL_PAM_ENABLED", "FALSE")

prod_list <- list.files(path = "/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers", pattern="_production", full.names = TRUE)
prod_list

# Control for differences in raster cells area.  
for(file in prod_list){ # 
  
  # file = prod_list[4] 
  
  prod_raster <- raster(file)
  
 crs(prod_raster) <- "+proj=longlat +datum=WGS84 +no_defs"
  
  name <- basename(file)
  name <- gsub(".tif", "", name)
  
  raster_area <- raster::area(prod_raster)
  #plot(raster_area)

  
   prod_area_adjust <- prod_raster/raster_area
   
   writeRaster(prod_area_adjust,
               sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/prod_reproject/%s_per_area.tif", name),
               overwrite=TRUE)

}


# Next: project rasters to equal area gall peters so we can more easily scale data and such.

gall_peters <- "+proj=cea +lon_0=0 +x_0=0 +y_0=0 +lat_ts=45 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
#raster template 
template_eq_area <- raster(res=0.083333333333333333)
extent(template_eq_area) <- c(-180, 180, -90, 90)
template_eq_area <- projectRaster(template_eq_area, crs=gall_peters, res=3000) 



rasters_to_new_crs <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/prod_reproject", pattern=".tif", full=TRUE)

# rasters_to_new_crs_vector <- c(grep(list.files("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/prod_reproject", full=TRUE), pattern = "gall_peter", invert = TRUE))

# rasters_to_new_crs_test <- rasters_to_new_crs[c(1:35, 106:121)]
# 
# rasters_to_new_crs_final <- rasters_to_new_crs_test[36:51]


registerDoParallel(cores = 8)
#for(rast in rasters_to_new_crs_test){ # rast <- rasters_to_new_crs[4]
foreach(rast = rasters_to_new_crs) %dopar% { 
 
rast_name <- basename(rast)
 getraster <- raster(rast)
 #beginCluster(n=8)
 highres_raster <- disaggregate(getraster, fact=2)
 projectRaster(highres_raster, template_eq_area, method="ngb", over=TRUE, progress="text",
                       filename=sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/prod_reproject/hi_res_prod_gall_peter_%s", rast_name), overwrite=TRUE)

 gp_raster <- raster(sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/prod_reproject/hi_res_prod_gall_peter_%s", rast_name))
#plot(gp_raster)
 
 aggregate(gp_raster, fact=2, fun=mean, filename=sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/prod_reproject/gall_peter_%s", rast_name), overwrite=TRUE)
 #endCluster()
}



# Delete the hi_res version which is just a temp file.

remove <- list.files(file.path("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/prod_reproject"), full=TRUE, pattern="hi_res")
file.remove(remove)



# Check that all went well. Compare raw rasters to the values here. These should basically match, but there will be some error due to estimates relating to altering the projection.

rasters <- list.files(path = "/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers", pattern="_production", full.names = TRUE)
# rasters <- grep("_consumption", rasters, value=TRUE, invert=TRUE)
check_data <- data.frame()
for(rast in rasters){ # rast=rasters[18]
  file_name <- basename(rast)
  file_name <- gsub(".tif", "", file_name)
  org_raster <- raster(rast)
  org_data <- cellStats(org_raster, "sum", na.rm=TRUE)
  
  
new_raster_path <- list.files(path = "/home/shares/food-systems/Food_footprint/all_food_systems/analysis/prod_reproject", pattern=sprintf("gall_peter_%s", file_name), full.names = TRUE)
  
  new_raster <- raster(new_raster_path)
  new_data_pre <- new_raster*raster::area(new_raster)/1000000
  new_data <- cellStats(new_data_pre, "sum", na.rm=TRUE)
  new_dataframe <- data.frame(name=file_name, org_data=org_data, new_data=new_data)
  check_data <- rbind(check_data, new_dataframe)
}


check_data <- check_data %>%
  mutate(diff = org_data - new_data)

plot(log(check_data$org_data +1), log(check_data$new_data +1))
abline(0,1, col="red")
check_data_compare <- check_data %>%
  mutate(percent = org_data/new_data)
hist(check_data_compare$percent)



## convert to total per cell
per_km_rasts <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/prod_reproject", full=TRUE, pattern = "gall_peter")

registerDoParallel(cores = 10)
#for(rast in per_km_rasts){ # rast = per_km_rasts[1]
foreach(rast = per_km_rasts) %dopar% {
 save_name <- basename(rast)
 save_name <- gsub("_per_area", "_per_cell", save_name)
    new_raster <- raster(rast)
  per_cell_rast <- new_raster*6000*6000/1000000
writeRaster(per_cell_rast, sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/prod_reproject/%s", save_name), overwrite=TRUE)
#check <- raster(sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/prod_reproject/%s", save_name))
#cellStats(check, "sum", na.rm=TRUE)
}


## Now stack those and multiply by the 1% of pressures raster and sum 

per_cell_files <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/prod_reproject", full=TRUE, pattern = "per_cell")

per_cell_rasts <- stack(per_cell_files)


total <- calc(per_cell_rasts, sum, na.rm = TRUE)

cellStats(total, "sum", na.rm = TRUE) # 13098756583

final_1_perc_area <- per_cell_rasts*cum_stress3

sum_rast_1_perc <- calc(final_1_perc_area, sum, na.rm = TRUE) # this raster is correct...

plot(sum_rast_1_perc)
plot(cum_stress3)

cellStats(sum_rast_1_perc, "sum") # 4055154578

4055154578/13098756583 # 0.3095832 - the top 1% of cumulative pressure pixels account for 31.9% of production assessed
```


## Check the proportion of cumulative pressures by country

**The cumulative pressure imposed by food production is greatest in India, China, Brazil, the U.S., and Pakistan (Fig. 4). These 5 countries alone contribute nearly half (43.8%) of global cumulative pressure.**


**Fig. 4. Proportional contribution to the cumulative food footprint in the highest ranking countries and the high seas for A) each pressure summed across all food types, or B) each food type summed across four pressure classes. These areas have the highest proportion of cumulative environmental pressure, and collectively account for 70.9% of the global footprint of food production.**
 
```{r}

summary_df_raw <- read_csv(file.path("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/raw_df_summary/rgn_raw_summary.csv"))

rescaling_values <- read_csv(here("_analysis/rescale_values.csv"))

un_geopolitical <- read_csv(here("_spatial/_output/UNSD_Methodology.csv")) %>%
  dplyr::select(iso3c, georegion=Region_Name)

summary_df <- summary_df_raw %>% 
  mutate(origin = ifelse(category == "feedfofm", "marine", origin)) %>% 
  mutate(land_ocean = case_when(origin %in% c("land", "freshwater") ~ "land",
                                origin == "marine" ~ "ocean")) %>% 
  mutate(sum = ifelse(iso3c == "HSX" & category == "feedfodd", 0, sum)) %>% 
  group_by(country, iso3c, land_ocean, pressure) %>% 
  dplyr::summarise(pressure_sum = sum(sum)) %>% 
  left_join(rescaling_values, by = "pressure") %>% 
  ungroup() %>% 
  rowwise() %>% 
  mutate(prop_of_global = pressure_sum/global_total)  %>% 
  unite(source, c("land_ocean", "pressure"), sep = "-", remove = FALSE) %>% 
  left_join(un_geopolitical, by = "iso3c") %>% 
  mutate(georegion = ifelse(country == "High Seas", "High Seas", georegion))


test <- summary_df %>%
  group_by(country) %>%
  summarise(sum = sum(prop_of_global)/4) %>%
  arrange(-sum) %>%
  head(5) 

sum(test$sum) # 0.4379168


## Figure 4 stats; there are 23 regions in it

test <- summary_df %>%
  group_by(country) %>%
  summarise(sum = sum(prop_of_global)/4) %>%
  arrange(-sum) %>%
  head(23) 

sum(test$sum) # 0.7023417

test2 <- summary_df %>%
  group_by(country) %>%
  summarise(sum = sum(prop_of_global)/4) %>%
  arrange(-sum) %>%
  filter(country == "High Seas")

test3 <- rbind(test, test2)

sum(test3$sum)# 0.7090657

```

## Check the origin of cumulative pressure by country and origin (land, freshwater, ocean)

**Country-level cumulative pressure derives almost entirely from land-based food production, with the exception of island nations and some countries with extensive coastlines, such as Norway (88% from oceans), Japan (40%), Chile (38%), the U.K. (37%), Indonesia (33%), and Vietnam (26%) (Fig. 4; Data S3).**
  
```{r, eval = FALSE}
unique(rgn_raw$origin)
# [1] "land"       "marine"     "freshwater"


rescale <- read_csv("../rescale_values.csv")


percent_pressure_ocean <- rgn_raw %>%
  left_join(rescale, by = "pressure") %>%
  mutate(global_pressure_prop = sum/global_total) %>%
  group_by(iso3c, country, origin) %>%
  summarise(cum_pressure_prop_rescaled = sum(global_pressure_prop),
            cum_pressure = sum(sum)) %>%
  ungroup()


### now save one that also splits it by fw
percent_pressure_origin_countries <- percent_pressure_ocean %>%
  select(-cum_pressure) %>%
  group_by(iso3c, country) %>%
  mutate(total_cum_prop = sum(cum_pressure_prop_rescaled)) %>%
  ungroup() %>%
  mutate(cum_pressure_prop_final = cum_pressure_prop_rescaled/total_cum_prop) %>%
  mutate(cum_pressure_percent_final = round(cum_pressure_prop_final*100,2)) %>%
  arrange(iso3c) %>%
  dplyr::select(iso3c, country, origin, cum_pressure_prop_final, cum_pressure_percent_final)


write.csv(percent_pressure_origin_countries, "output/percent_cum_pressure_country_origin_fw.csv", row.names = FALSE)

```

## Check the contributions of fisheries and aquaculture to total pressures

**Marine fisheries and aquaculture contribute >25% of total pressures in 90 countries, primarily in island nations (61 countries).**

```{r, eval = FALSE}
rescale <- read_csv("../rescale_values.csv")

percent_pressure_fisheries <- rgn_raw %>%
  left_join(rescale, by = "pressure") %>%
  mutate(global_pressure_prop = sum/global_total) %>%
  group_by(iso3c, system, origin) %>%
  summarise(cum_pressure_prop_rescaled = sum(global_pressure_prop)) %>%
  ungroup()

percent_pressure_fisheries_2 <- percent_pressure_fisheries %>%
  group_by(iso3c) %>%
  mutate(total_cum_prop = sum(cum_pressure_prop_rescaled)) %>%
  ungroup() %>%
  mutate(cum_pressure_prop_final = cum_pressure_prop_rescaled/total_cum_prop) %>%
  mutate(percent_final = round(cum_pressure_prop_final*100,1))
  # group_by(organism) %>%
  # summarise(prop_final = sum(cum_pressure_prop_final)) %>%
  # ungroup() %>%
  # mutate(percent_final = round(prop_final*100,1))

write.csv(percent_pressure_fisheries_2, "output/percent_cum_pressure_country_fisheries.csv", row.names = FALSE)


marine_fisheries_pressures <- percent_pressure_fisheries_2 %>%
  left_join(food_rgns, by = "iso3c") %>%
  filter(system %in% c("fisheries", "aquaculture"), origin == "marine") %>%
  group_by(iso3c, Country) %>%
  summarise(percent_final = sum(percent_final)) %>%
  filter(percent_final >= 25)
count(marine_fisheries_pressures) ### 94 countries (95 with high seas)
94/245 ### about 0.3836735 of countries have marine fisheries accounting for >25% of their pressures

islands <- read_csv("data/island_nations.csv")
islands_vec <- islands$country

marine_fis_islands <- marine_fisheries_pressures %>% 
  filter(Country %in% islands_vec) ### at least 31... there are a lot of name mismatches though

setdiff(marine_fisheries_pressures$Country, islands_vec)

marine_fis_islands2 <- marine_fisheries_pressures %>%
  filter(!(Country %in% islands_vec)) %>%
  filter(str_detect(Country, "Island|island")) ### 11


marine_fis_islands3 <- marine_fisheries_pressures %>%
  filter(!(Country %in% islands_vec)) %>%
    filter(!str_detect(Country, "Island|island")) # now lets hand check marine_fis_islands3 


# 24

31 + 11 + 24 # 66 are island nations
``` 

## Check statistics that correspond to figure 5 (the barchart)

**For example, the ruminant digestive system of cows makes their GHG emissions noteworthy (53% of their cumulative pressures), along with nutrient emissions from their wastes and feed production (31%). The footprint of rice, and wheat crops more strongly reflects water use and disturbed land area (Fig. 5). Assessing the cumulative pressures of different foods by country also reveals that crop production, consumed by both people and livestock, dominates overall pressure in nearly all countries, but there are some exceptions such as Brazil, which has relatively high cumulative pressures from meat production (Fig. 4b). **

 - code from fig 4
 

OLD: 

**Importantly, food types often rank differently in their global cumulative pressure compared to ratings derived from per-unit assessments of individual classes of pressure. For example, we find that the cumulative footprint of catching demersal fishes is 2x that of raising sheep for meat, which is counter to common generalisations.**
 

new: 
**In this figure, pigs have a total cumulative pressure of ~0.15X (i.e. 15% of the total pressure of all foods globally), while rice is ~9%. The reason pigs are higher is because its pressure from excess nutrients encompasses so much of the total pressure of all foods globally (9.4% of the total pressure of all foods globally, versus 1.2%). Rice has higher GHG emissions (3.1% for rice, 1.7% for pigs) and water use (2.8% for water, 1.9% for pigs). Disturbance is about the same (1.5% for rice, 1.7% for pigs). These kinds of comparisons are a key reason we did our study and are only possible with our approach, i.e. this is the first time such comparisons have been possible.**

 
```{r}

'%!in%' <- function(x,y)!('%in%'(x,y))

summary_df_raw <- read_csv(file.path("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/raw_df_summary/rgn_raw_summary.csv")) 

rescaling_values <- read_csv(here("_analysis/rescale_values.csv"))

food_groupings <- read_csv(here("_analysis/figures/paper/data/grouping_naming_structure.csv")) 

mapspam_names <- read_csv(here("crop/farm/data/MapSPAM_names.csv"))%>% 
  select(sub_tier = SPAM_short_name, SPAM_full_name) %>% unique()


summary_df_int <-summary_df_raw %>% 
  mutate(feed_item = case_when(category == "feedcrop"|category =="feedfodd"|category =="feedfofm" ~ "feed",
                               T ~ " ")) %>% 
  unite(product, c("product", "feed_item"), sep = " ") %>% 
  mutate(product = str_trim(product, side = c("right"))) %>% 
  select(organism, product, pressure, sum) %>% 
  ## fresh water fisheries is listed as fish and there is still salmon instead of salmonids so we will change that here 
  mutate(organism = ifelse(organism == "fish", "river fish", 
                           ifelse(organism == "salmon", "salmonids", organism))) %>% 
  group_by(organism, product, pressure) %>% 
  dplyr::summarise(sum = sum(sum)) %>% 
  unite(tier_5, c("organism", "product"), sep = " ") %>% 
  ungroup() 

## need to split the back yard chickens eggs&meat
fix_chickens_1 <- summary_df_int %>% 
  filter(tier_5 == "chickens eggs&meat") 
fix_chickens_1 <- rbind(fix_chickens_1, fix_chickens_1) %>% 
  arrange(pressure) %>% 
  mutate(tier_5 = rep(c("chickens meat", "chickens eggs"), times = 4, each = 1))

fix_chickens_2 <- summary_df_int %>% 
  filter(tier_5 == "chickens eggs&meat feed") 
fix_chickens_2 <- rbind(fix_chickens_2, fix_chickens_2) %>% 
  arrange(pressure) %>% 
  mutate(tier_5 = rep(c("chickens meat", "chickens eggs"), times = 4, each = 1))

## add back
summary_df <- summary_df_int %>% 
  filter(tier_5 != "chickens eggs&meat",
         tier_5 != "chickens eggs&meat feed") %>% 
  rbind(fix_chickens_1, fix_chickens_2) %>% 
  group_by(tier_5, pressure) %>% 
  dplyr::summarise(sum = sum(sum)) %>% 
  left_join(rescaling_values, by = "pressure") %>% 
  ungroup() %>% 
  rowwise() %>% 
  mutate(prop_of_global = sum/global_total,
         prop_of_global = ifelse(tier_5 == "fofm meat", prop_of_global*0.1, prop_of_global)) %>% 
  left_join(food_groupings, by = "tier_5")

## quick check
sum(summary_df$prop_of_global) #  3.966979 looks good!



bar_df_int <- summary_df %>% 
  select(tier_2, tier_5, pressure, prop_of_global) %>% 
  mutate(sub_tier = str_remove(tier_5, " feed"),
         pressure_bar = ifelse(str_detect(tier_5, "feed"), paste(pressure, "_feed", sep = ""), pressure)) %>%
  group_by(tier_2, sub_tier) %>% 
  mutate(food_cat_sum = sum(prop_of_global)) %>% 
  ungroup()  %>% 
  
  mutate(facet_tier = ifelse(str_detect(tier_2, "fisheries"), "fisheries", tier_2),
         facet_tier = case_when(facet_tier == "livestock" ~ "Livestock",
                                facet_tier == "crop human consumption" ~ "Human Crop Consumption",
                                facet_tier == "fisheries" ~ "Marine and Freshwater Fisheries",
                                facet_tier == "mariculture" ~ "Mariculture")) %>% 
  mutate(prop_of_global = prop_of_global/4)

## make naming changes to make it look better
bar_df <- bar_df_int%>% 
  mutate(sub_tier = ifelse(facet_tier ==  "Human Crop Consumption", str_remove(tier_5, " produce"), sub_tier)) %>% 
  left_join(mapspam_names, by = "sub_tier") %>% 
  mutate(SPAM_full_name = ifelse(sub_tier == "xfru", "other fruits",
                                 ifelse(sub_tier == "xmil", "millet",
                                        ifelse(sub_tier == "xoil", "other oil crops",
                                               ifelse(sub_tier == "xpul", "pulses", SPAM_full_name))))) %>% 
  mutate(sub_tier = ifelse(tier_2 == "crop human consumption", SPAM_full_name, sub_tier)) %>% 
  select(-SPAM_full_name) %>% 
  mutate(sub_tier = ifelse(tier_2 == "marine fisheries", str_replace(sub_tier, "meat", "fish"), sub_tier),
         sub_tier = ifelse(tier_2 == "freshwater fisheries", "river fish*", sub_tier),
         sub_tier = ifelse(sub_tier == "fofm fish", "FOFM", sub_tier),
         sub_tier = ifelse(tier_2 == "mariculture", str_replace(sub_tier, "meat", ""), sub_tier)) %>% 
  mutate(tier_5 = str_replace(tier_5, "bivalve", "bivalves"),
         sub_tier = str_replace(sub_tier, "bivalve", "bivalves**"),
         sub_tier = ifelse(sub_tier == "FOFM", "forage fish", sub_tier),
         tier_5 = str_replace(tier_5, "fofm", "forage fish"))


bar_df$sub_tier <- factor(bar_df$sub_tier)

final_df <- bar_df %>%
  group_by(sub_tier) %>%
  mutate(total_prop = sum(prop_of_global)) %>%
  ungroup() %>%
  mutate(final_prop = prop_of_global/total_prop)

# the ruminant digestive system of cows makes their GHG emissions noteworthy (54% of their cumulative pressures), along with nutrient emissions from their wastes and feed production (31%).


# The footprint of rice, and wheat crops more strongly reflects water use and disturbed land area (Fig. 5).


## Everything below here in this chunk is obselete. 
cows_meat <- final_df %>%
  filter(sub_tier == "cows meat")

## cows ghg emissions are 0.53923501 or ~54% of their cumulative pressures
## cows nutrient (including feed) emissions are ~31% of their cumulative pressures

# In contrast, the footprint of chickens more strongly reflects water use from the feed (33% of their cumulative pressures) and land area disturbed from the feed (25%) for these animals that is driven in particular by their very large production levels


chickens_meat <- final_df %>%
  filter(sub_tier == "chickens meat")

## chickens meat water use from feed is 0.33070470 or ~33% of their cumulative pressure
## chickens meat disturbance from feed is 0.24794869 or ~25% of their cumulative pressure


pigs_meat <- final_df %>%
  filter(sub_tier == "pigs meat")

## pigs meat water use from feed is 0.1271276234 or ~13% of their cumulative pressure
## pigs meat disturbance from feed is 0.1157697775 or ~12% of their cumulative pressure


by_organisms <- final_df %>%
  group_by(sub_tier) %>%
  summarise(total_prop = sum(prop_of_global)) %>%
  ungroup()

test <- final_df %>%
  filter(is.na(sub_tier))

sum(test$prop_of_global) # 0.01619444

sum(by_organisms$total_prop) # 0.9921254

sum(test$prop_of_global) + sum(by_organisms$total_prop) # 1.00832 close enough

## look at pigs: 1.478654e-01  or pigs represent ~15% of total pressures


## look at rice: 8.602024e-02 or rice represents ~10% of total pressures

by_organisms_pressures <- final_df %>%
  group_by(sub_tier, pressure_bar) %>%
  summarise(total_prop = sum(prop_of_global)) %>%
  ungroup() %>%
  filter(sub_tier %in% c("pigs meat", "rice"))

#### water
## 2.8% of global pressure comes from rice water
# 0.01918562
## 1.9% of global pressure comes from pig water 

#### ghg
## 3.1% of global pressure comes from rice ghg
#0.01724006
## 1.7% of global pressure comes from pig ghg

#### disturbance 
## 1.5% rice
## 1.7% pigs
# 0.01720437

#### nutrients 
## 1.2% rice
#0.09423532
## 9.4% pigs 

```



## compare origin of pressure by organisms

 - compare magnitudes of pigs/chickens cumulative pressures on land to that of mariculture species 
    - pigs ocean cumulative pressure is similar to that of crustaceans, salmon, and marine fish general in ocean 
    - chickens ocean cumulative pressure is similar to that of tuna in water.
    

"The cumulative pressure for fed animals spreads far beyond the farm where they are raised. For example, because marine forage fish comprise an average of ~0.15% of chicken and ~0.02% of pig feed, these livestock have similar cumulative ocean footprints to that of some mariculture species (Fig. 5). Feed for mariculture species increasingly includes crops, and all fed species have >98% of their footprint on land (Data S4)." 

```{r}
rescale <- read_csv("../rescale_values.csv")

### rescale pressures 
percent_pressure_rescale <- rgn_raw %>%
  left_join(rescale, by = "pressure") %>%
  mutate(global_pressure_prop = sum/global_total) %>%
  group_by(organism, origin) %>%
  summarise(cum_pressure = sum(sum)) %>%
  ungroup()

### group by food group
percent_pressure_organisms_origin <- percent_pressure_rescale %>%
  group_by(organism) %>%
  mutate(total_organism_pressure = sum(cum_pressure)) %>%
  ungroup() %>%
  mutate(proportion_final = cum_pressure/total_organism_pressure) %>%
  filter(organism %in% c("chickens", "pigs", "crustaceans", "marine-fish-general", "bivalve", "salmon", "shrimp", "tuna"))
  


```

## Check the displacement of cumulative pressure that happens due to crop exports 

**More broadly, use of feed for livestock displaces on average X% of a country’s cumulative footprint to other countries (range: X-X%; Suppl. Table XX).**


```{r}
ex_im_stress <- read_csv(here("_analysis/export_import_stressors/data/cumulative_stress_imports_exports.csv"))

# ds indicates "domestic supply"

# To get the actual percent of a country’s displacement of their cumulative pressure, you need to multiply prop_exported x prop_ds_produced

# To get just the exported of the crops actually produced in the country, you need to multiply it by the proportion of domestic supply that is produced (prop_ds_produced)

ex_im_crops <- ex_im_stress %>%
  mutate(prop_exportc = prop_exported*prop_ds_produced)

mean(ex_im_stress$prop_exported, na.rm = TRUE) # 0.1511889

mean(ex_im_crops$prop_exportc, na.rm = TRUE) # 0.1165849

# so this means that on average 12% of a country's cumulative pressure is exported 

# range
range(ex_im_crops$prop_exportc, na.rm = TRUE) # 0.0000000 0.6478093

# 0% - 65%
```

 

## compare demersal and sheep production

**In some cases these differences are driven by the scale of production, for example demersal fishing is 4x greater production than sheep, so even though per-unit sheep are less efficient, demersal fishing exerts overall more cumulative pressure.**


```{r}
all_prod <- read_csv("output/all_production.csv")

by_product <- all_prod %>%
  group_by(category1_production) %>%
  summarise(total_tonnes = sum(tonnes, na.rm = TRUE))


# dem fish prod: 45088512
# sheep prod: 11212054

45088512/11212054 # 4x
```


## Find out what the maximum cumulative pressure score (across all foods) for a pixel was and what country it was in. 

**The resulting total cumulative pressure across all the global pixels equal 4 (by definition), and the maximum observed pixel value was 2.305x10-4, near xx in Israel.**

```{r, eval = FALSE}
# Find out what the maximum cumulative pressure score (across all foods) for a pixel was and what country it was in. 

cum_stress <- raster(file.path("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/stressor_summary/equal_area/rescaled/cumulative_stress/all_systems_cumulative_stress.tif"))
plot(cum_stress)

maxValue(cum_stress) 
# # NEW: 230.5179
# 2.305 * 10^-4

cellStats(cum_stress, "sum") # 4026176

land_eez_rgns_gp <- raster(file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/spatial/gall_peters/land_eez_rgns_gall_peters.tif"))
plot(land_eez_rgns_gp)

### fix food_rgns to include high seas
food_rgns <- rbind(food_rgns, data.frame(iso3c = "high_seas", ID_0 = 500, Country = "High Seas"))

test_df <- as.data.frame(zonal(cum_stress, land_eez_rgns_gp, "max")) %>%
   left_join(food_rgns, by = c("zone" = "ID_0"))

test2 <- as.data.frame(cum_stress, xy= TRUE)

#  2739569.68, 4760818.1 is where the max pressure is 

extent(cum_stress) <- c(xmin= -14207430, xmax= 14208570, ymin= 0.99*(-9000182), ymax= 0.99*(8999818)) ### cut the extent a little bit so that the reprojection will work.. 

test4 <- projectRaster(cum_stress, food_raster, method = "ngb")

maxValue(test4) 
# 230.152 

test5 <- as.data.frame(test4, xy = TRUE)
# this says max value is x = 34.708333 and y = 31.791667 but that isnt a point on a map... however if you flip the two, then it is a point in Israel?


test4[test4 < 200] <- NA

test55 <- rasterToPoints(test4, spatial = TRUE, centroids = TRUE)
mapview(test55) ### this matches above point. In Isreal, near Ashdod

 
```


## What % of total reported global food production is included in our assessment.
 - 85.7%

**Here we advance understanding of environmental consequences of global food production in three crucial ways: 1) expanding standardised assessment of food types to incorporate most marine, freshwater, and terrestrial foods, representing 85.7% of total reported global production (Supplementary Methods); **
  
```{r, eval = FALSE}

###### Mariculture production included: 
mar_prep <- "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine"

marine_fish <- read_csv(file.path(mar_prep, "marine_fish_general/updated/marine_fish_general_farm_updated.csv")) %>%
  mutate(type = "marine_fish_general")

bivalve <- read_csv(file.path(mar_prep, "bivalve/updated/bivalve_farm_updated.csv")) %>%
  mutate(type = "bivalve")

crustaceans <- read_csv(file.path(mar_prep, "crustaceans/updated/crustaceans_farm_updated.csv")) %>%
  mutate(type = "crustaceans")

salmon <- read_csv(file.path(mar_prep, "salmon/updated/salmon_farm_updated.csv")) %>%
  mutate(type = "salmon")

tuna <- read_csv(file.path(mar_prep, "tuna/updated/tuna_farm_updated.csv")) %>%
  mutate(type = "tuna")

shrimp <- read_csv(file.path(mar_prep, "shrimp/updated/shrimp_farm_updated.csv")) %>%
  mutate(type = "shrimp")


all_mar <- rbind(marine_fish, bivalve, crustaceans, salmon, tuna, shrimp) %>%
  group_by(type) %>%
  summarise(production = sum(tonnes_production))


sum(all_mar$production) # 29639681/30991774 animal mariculture production included
# 0/32560138 seaweed mariculture included

###### Marine fisheries production included (Reported + IUU, since discards are unlikely to be consumed by humans): 
# We include all from Watson
# 72172118/72172118 million tonnes of non-FOFM
# 39054068/39054068 million tonnes of FOFM

###### Freshwater aquaculture: 
# 0/48671218 freshwater aquaculture included

###### Freshwater wild-caught fish:
# 7134914/12700000 freshwater wildcaught fisheries included


fis_aq_categories <- data.frame(tonnes_included = c(29623515, 0, 72172118, 39054068, 0, 7134914),
                             tonnes_total = c(30991774, 32560138, 72172118, 39054068, 48671218, 12700000),
                             type = c("animal_mariculture", "seaweed_mariculture", "non-fofm_fis", "fofm-fis", "freshwater_aquaculture", "freshwater_fis")) %>%
  mutate(percent = tonnes_included/tonnes_total)

############# Crops
crop_translate<- read_csv(here("feed/data/MapSPAM_to_FAO.csv")) %>%
  select(SPAM_short_name, SPAM_super, FAO_item_code) %>%
  unique()

crop_super_cats <- read_csv(here("feed/data/MapSPAM_crop_info.csv"))
Mapspam_prod <- read_csv(here("feed/data/MAPSPAMcrop_production.csv"))
fao_crop_prod <- read_csv(here("_analysis/checking_data/data/FAOSTAT_data_12-21-2020_crops.csv")) %>%
  select(Area, FAO_item_code=`Item Code`, Item, Unit, Value)

### ID crop categories not counted in Mapspam but in FAO:
fao_extra <- left_join(fao_crop_prod, crop_translate, by="FAO_item_code")
filter(fao_extra, is.na(SPAM_super))
fao_tonnes <- left_join(fao_extra, crop_super_cats, by="SPAM_super") %>%
  dplyr::filter(!(is.na(food_group))) %>%
  dplyr::filter(food_group != "exclude") %>%
  group_by(food_group) %>%
  summarize(tonnes_fao = sum(Value, na.rm=TRUE))

spam_tonnes <- left_join(Mapspam_prod, crop_super_cats, by="SPAM_super") %>%
  group_by(food_group) %>%
  summarize(tonnes_mapspam = sum(tonnes_producing_crop, na.rm=TRUE))

compare <- left_join(spam_tonnes, fao_tonnes) %>%
  mutate(spam_percent = tonnes_mapspam/tonnes_fao)
plot(compare$tonnes_mapspam, compare$tonnes_fao)
abline(0,1, col="red")


compare_crops <- compare %>%
  rename("type" = "food_group", "tonnes_included" = "tonnes_mapspam", "tonnes_total" = "tonnes_fao", "percent" = "spam_percent") %>%
  dplyr::filter(!is.na(tonnes_total))


###### Livestock animals: 
livestock_cats <- read_csv(here("_analysis/paper_stats/data/FAO_livestock_categories.csv")) %>%
  filter(is.na(non_food_exclude))

livestock <- read_csv(here("_analysis/paper_stats/data/FAOSTAT_data_12-18-2020_livestock_primary.csv")) %>%
  filter(Area != "China") %>%
  filter(!(is.na(Value))) %>%
  filter(Item %in% livestock_cats$Item) %>%
  left_join(livestock_cats, by="Item") %>%
  group_by(included_for_us, category) %>%
  summarize(tonnes = sum(Value)) 

livestock_compare <- livestock %>%
  spread(included_for_us, tonnes) %>%
  mutate(yes = ifelse(is.na(yes), 0, yes)) %>%
  rowwise() %>%
  mutate(total = no+yes) %>%
  mutate(percent_counted = yes/(yes + no)) %>%
  dplyr::select("type" = "category", "tonnes_included" = "yes", "tonnes_total" = "total") %>%
  dplyr::mutate(percent = tonnes_included/tonnes_total)


###### Combine together 
all_food_groups <- rbind(livestock_compare, compare_crops, fis_aq_categories)

sum(all_food_groups$tonnes_included)/sum(all_food_groups$tonnes_total) # 0.8585466

```

## Related, what % of global production comes from the ocean (fisheries, aquaculture, FOFM)?

**yet these ocean pressures are non-trivial given that relatively little (1.1%, by tonnes) food and feed for fed animals comes from the sea**

**Although only 1.1% of food comes from aquatic systems**

```{r, eval = FALSE}
### Crops and meat comparisons
### What we include 

included <- as.data.frame(all_food_groups) %>%
  dplyr::filter(type %in% c("animal_mariculture", "seaweed_mariculture", "non-fofm_fis", "fofm-fis")) %>%
  dplyr::select(type, tonnes_included)

sum(included$tonnes_included)
sum(all_food_groups$tonnes_included)
### 140849701/13343822915 = 0.01055542


### Overall
overall <- included <- as.data.frame(all_food_groups) %>%
  dplyr::filter(type %in% c("animal_mariculture", "seaweed_mariculture", "non-fofm_fis", "fofm-fis")) %>%
  dplyr::select(type, tonnes_total)

sum(overall$tonnes_total)
sum(all_food_groups$tonnes_total)

### 174778098/15542340143 = 0.01124529


### Marine fish to meat (no eggs or milk) comparison
fish_meat <- as.data.frame(all_food_groups) %>%
  dplyr::filter(type %in% c("animal_mariculture", "non-fofm_fis", "fofm-fis", "meat"))

sum(fish_meat$tonnes_total) # 512155040

fish <- as.data.frame(all_food_groups) %>%
  dplyr::filter(type %in% c("animal_mariculture", "non-fofm_fis", "fofm-fis"))

sum(fish$tonnes_total) # 142217960

# 142217960/512155040 = 0.2776854

```



## Extract data on how much of freshwater aquaculture comes from Asia (should be something like 80%) and how much of the global total comes from just China.

**We excluded wild-harvest and mariculture of seaweed and freshwater aquaculture because no comprehensive data exist for farm locations; however, the vast majority of freshwater aquaculture occurs in Asia (77.6% of global production in tonnes, with China producing 59.8%), and so inclusion of these data would primarily increase pressures in Asia.**
 
```{r, eval = FALSE}
fao_mariculture_fw <- read_csv("../../aquaculture/data/fao_mariculture_clean.csv") %>%
  dplyr::filter(year == 2017, 
                environment == "Freshwater")
sum(fao_mariculture_fw$value, na.rm = TRUE) #48.7 mil

fw_asia <- fao_mariculture_fw %>%
  dplyr::filter(fao_major_area == "Asia - Inland waters") %>%
  dplyr::filter(!(country %in% c("Afghanistan", "Armenia", "Azerbaijan", "Cyprus", "Georgia", "Iran (Islamic Rep. of)", "Iraq", "Israel", "Jordan", "Kazakhstan", "Kyrgyzstan", "Lebanon", "Oman", "Pakistan", "Palestine", "Qatar", "Saudi Arabia", "Syrian Arab Republic", "Tajikistan", "Turkey", "Turkmenistan", "United Arab Emirates", "Uzbekistan", "Bhutan", "Bangladesh", "India", "Nepal", "Sri Lanka")))

fw_china <- fao_mariculture_fw %>%
  dplyr::filter(country == "China")

### Total tonnes fw 
sum(fao_mariculture_fw$value, na.rm = TRUE) # 48671218

### Total asia fw tonnes 
sum(fw_asia$value, na.rm = TRUE) # 37746079

### Total china fw tonnes 
sum(fw_china$value, na.rm = TRUE) # 29117696

# Percent fw aquaculture from Asia: 37746079/48671218 = 0.7755318

# Percent fw aquaculture from China: 29117696/48671218 = 0.5982529

unique(fw_asia$country)

```

