---
title: 'Figure 1: Cumulative footprint maps'
author: "Juliette"
date: "11/17/2020"
output: html_document
---

Figure 1. Cumulative footprint maps. a) map of full (‘comprehensive’) cumulative stressors from all foods; b-e) global maps of footprint of each stressor across all foods

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(sf)
library(scales)
library(rnaturalearth)
library(rnaturalearthdata)

source(here("_workflow/common.R"))
source(here("_workflow/common_spatial.R"))
source(here("_analysis/figures/paper/colors_stressors_fig_1_2.R"))

## save the gall peters projection
gall_peters <- "+proj=cea +lon_0=0 +x_0=0 +y_0=0 +lat_ts=45 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

## create a water raster
water <- readOGR(dsn = here("_analysis/raw/rgn_all_gcs_med_res.shp"))
water <- water[!water$rgn_nam == "Antarctica",]
proj4string(water) <- CRS("+proj=longlat +datum=WGS84")
water <- spTransform(water, CRS("+proj=cea +lon_0=0 +x_0=0 +y_0=0 +lat_ts=45 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"))
plot(water)

## create a land raster
land <- ne_countries(scale = "medium", returnclass = "sp") 
land <- land[!land$sov_a3 == "ATA",]
land <- st_as_sf(land)
land <- st_transform(land, crs=gall_peters)
land <- as(land, "Spatial")
plot(land)

library(RColorBrewer)
display.brewer.all()

stressor_df <- data.frame(stressor = c("ghg", "water", "nutrient", "disturbance"),
                          stressor_unit = c("CO2eq_tonnes", "m3", "PO4eq_tonnes", "proportion"))

```


## Cummulative stressors all food systems (livestock, aquaculture, fisheries, human)

Create the color pallette
```{r}

d_brown <- "#515256"
m_brown <- "#B27B54"
l_brown <- "#BC995F"
green <- "#A8B072"
yellow<- "#EFCE71"
l_yellow <- "#F7F6C1"

jv_pal <- c(l_yellow, yellow, green, l_brown, m_brown, d_brown)

continuous_pal <-  colorRampPalette(jv_pal, space="Lab", bias = 3.5)(10000)
image(volcano, asp=1, col=continuous_pal)

# jv_pal_new <- jv_pal[7000:length(continuous_pal)]
# image(volcano, asp=1, col=jv_pal_new)

light_gray <- "#F8F9FA"
final_palette <- c(light_gray, continuous_pal)
scales::show_col(final_palette)

```



```{r}

chi <- raster("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/stressor_summary/equal_area/rescaled/cumulative_stress/all_systems_cumulative_stress.tif")
#plot(log(1+chi_raster_raw))

## quantile
chi_999_quant <- quantile(chi, 0.999)
# chi_9999_quant <- quantile(chi, 0.9999)

#chi_cols = c("NA", paste0(final_palette, "BF"))
chi[chi > chi_999_quant ] <- chi_999_quant
raster_data = chi

title = "" 
cols = final_palette

png(here(sprintf("_analysis/figures/paper/output/fig_1a_all_cumulative_map_99_9.png", system)), res=500, width=7, height=3, units="in")

par(oma=c(1,1,1,0), new = TRUE) # bottom, left, top, and right
par(mfrow=c(1, 1), mar=c(1, 0, 1, 2))

plot(water, border= "#F8F9FA", col="#F8F9FA", lwd=0.5, main=sprintf("Total cumulative stress (99.9th quant)", system), cex=0.75)
plot(land, border="#BCBEB1", col="#F8F9FA", lwd=0.2, add=TRUE)
plot(raster_data, 
    col=cols, 
     axes=FALSE, 
     box=FALSE, 
     add=TRUE,  
     legend=TRUE, 
     legend.shrink = 0.7, 
     legend.width =.4, 
     axis.args = list(cex.axis = 0.6, cex.sub=0.6, cex=0.6, lwd.ticks=1, at = c(0, round(chi_999_quant/3, 0), round((chi_999_quant/3)*2, 0), round(chi_999_quant, 0)-1) ,
                       labels=c("0", round(chi_999_quant/3, 0), round((chi_999_quant/3)*2, 0), paste0(">", round(chi_999_quant, 0)))) ,
     legend.args=list(text = "", cex = 0.7, side = 3, line = 1))
dev.off()


```

### Make a histogram distribution of scores, and color code it like the legend

```{r}
library(scales)
library(ggforce)
library(hrbrthemes)

chi <- raster("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/stressor_summary/equal_area/rescaled/cumulative_stress/all_systems_cumulative_stress.tif")

chi_99_quant <- quantile(chi, 0.99)
raster_data_99 = chi
raster_data_99[raster_data_99 > chi_99_quant ] <- chi_99_quant

number_cells <- 14208000
set.seed(30)
raster_data_sample <- sample(raster_data_99, number_cells*0.5) 

set.seed(30)
raster_brown_data_sample <- sample(chi, number_cells*0.5) 

## first we need to create a df of the values
chi_df_sample <- as.data.frame(raster_data_sample)
max <- round(max(chi_df_sample$raster_data_sample))

chi_brown_df_sample <- as.data.frame(raster_brown_data_sample) %>% 
  filter(raster_brown_data_sample > max(chi_df_sample$raster_data_sample))
max_brown <- round(max(chi_brown_df_sample$raster_brown_data_sample))

## assign color breaks
pal_hist <-  colorRampPalette(jv_pal, space="Lab", bias = 3.5)(max/0.1) ## max of the 99th quant
light_gray <- "#F8F9FA"
pal_hist <- c(light_gray, pal_hist)
image(volcano, asp=1, col=pal_hist)
scales::show_col(pal_hist)
length(pal_hist)
# need to make all the values between 99th quant and max of the rest the dark brown
# time_rep <- round(max-6)
# add_brown<- c(rep("#515256", time_rep))
# pal_hist <- c(pal_hist, add_brown)

## for the colors that are included in the legend
col.brks <- seq(0, max, by = 0.1)
col.assign <- cut(col.brks, breaks=col.brks, include.lowest = TRUE)

colors_df <- data.frame(colors_plot = pal_hist, col.assign)
colors_df$col.assign <- as.character(colors_df$col.assign) 
 colors_df <- colors_df %>% 
   mutate(col.assign = ifelse(colors_plot == "#F8F9FA", '[0,0]',
                              ifelse(col.assign == '[0,1]', '(0,1]', col.assign)))

chi_df_sample <- chi_df_sample %>% 
  arrange(raster_data_sample)
chi_df_sample$col.assign <- cut(chi_df_sample$raster_data_sample, breaks=col.brks, include.lowest=TRUE) 
chi_df_sample$col.assign <- as.character(chi_df_sample$col.assign)
chi_df_sample <- chi_df_sample %>% 
   mutate(col.assign = ifelse(raster_data_sample == 0, '[0,0]', 
                              ifelse( col.assign == '[0,1]', '(0,1]', col.assign)))

brown <- chi_brown_df_sample %>% 
  mutate(col.assign = '(5.9,6]', 
         raster_data_sample = 5.902967) %>% 
  select(-raster_brown_data_sample)

chi_df_sample_hist <- rbind(chi_df_sample,brown) %>% 
  left_join(colors_df, by="col.assign") %>%
  mutate(colors_plot = as.character(colors_plot)) %>%
  arrange(raster_data_sample) %>% 
  group_by(col.assign) 

#class(chi_df_sample_hist$raster_data_sample)

cols <- chi_df_sample_hist %>% 
  mutate(colors_plot = as.character(colors_plot)) %>% 
  select(col.assign, colors_plot) %>% 
  unique() %>% 
  pull(colors_plot) 

image(volcano, asp=1, col=cols)
scales::show_col(cols)
```
 

facet zoom method
```{r}

## create a ggplot histogram
hist_facet <- ggplot(chi_df_sample_hist) +
  geom_histogram(aes(x = raster_data_sample, fill = colors_plot), bins = 30) +
  scale_fill_identity() +
  facet_zoom(xlim = c(0.6, 6), ylim = c(0,180000), zoom.size = 1, horizontal = TRUE)+
  #theme(zoom.y = element_blank(), validate = FALSE)
  labs(title = "",
       y = "Number Raster Cells",
       x = "Cumulative Impact")+
  theme_minimal() +
  theme(text = element_text(size=25),
        axis.title.x = element_text(hjust=0.2))
       #  panel.background = element_rect(fill = "transparent", colour = "transparent"), # bg of the panel
       #  plot.background = element_rect(fill = "transparent", color = NA), # bg of the plot
       #  #panel.grid.major = element_blank(), # get rid of major grid
       # # panel.grid.minor = element_blank(), # get rid of minor grid
       #  legend.background = element_rect(fill = "transparent"), # get rid of legend bg
       #  legend.box.background = element_rect(fill = "transparent"))
       # 

hist_facet

ggsave(here("_analysis/figures/paper/output/fig_1a_histogram.jpg"), width = 30, height = 10, dpi=300) # bg = "transparent"

```




## Make the map top 1% a red color
```{r}
chi <- raster("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/stressor_summary/equal_area/rescaled/cumulative_stress/all_systems_cumulative_stress.tif")
#plot(log(1+chi_raster_raw))

#hist(chi)

## quantile
chi_99_quant <- quantile(chi, 0.99)

palette_red <- c(final_palette, "#B90000")
scales::show_col(palette_red)

# chi_cols = c("NA", paste0(palette_red))

chi[chi > chi_99_quant ] <- chi_99_quant
#chi[chi > chi_999_quant] <- chi_999_quant
raster_data = chi

title = "" 
cols = palette_red

png(here(sprintf("_analysis/figures/paper/output/fig_1a_all_cumulative_map_99_red_top.png", system)), res=500, width=7, height=3, units="in")

par(oma=c(1,1,1,0), new = TRUE) # bottom, left, top, and right
par(mfrow=c(1, 1), mar=c(1, 0, 1, 2))

plot(water, border= "#F8F9FA", col="#F8F9FA", lwd=0.5, cex=0.75)
plot(raster_data, 
     col=cols, 
     axes=FALSE, 
     box=FALSE, 
     add=TRUE,  
     legend=TRUE, 
     legend.shrink = 0.7, 
     legend.width =.4, 
     margin = FALSE,
     axis.args = list(cex.axis = 0.6, cex.sub=0.6, cex=0.6, lwd.ticks=1, at = c(0,
                                                                                round(chi_99_quant/3, 0),
                                                                                round((chi_99_quant/3)*2, 0),
                                                                                round(chi_99_quant, 0)- 0.3), 
                       labels=c("0",
                                round(chi_99_quant/3, 0),
                                round((chi_99_quant/3)*2, 0),
                                paste0(">", round(chi_99_quant, 2)))), 

      legend.args=list(text = "Proportion of Global\nPressures(millionths)", cex = 0.7, side = 3, line = 1))
plot(land, border = "#BCBEB1", col= "#F8F9FA00", lwd=0.2, add=TRUE)

dev.off()

# Proportion of Global\nPressures(millionth
```




## Summary of each stressors of all food systems (livestock, aquaculture, fisheries, human)

```{r}

chi_plot_stressor <- function(stressor){
  
  ## stressor = "water"
  ## get the max value
  ## chi

  chi <- raster(list.files("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/stressor_summary/equal_area/stressor_per_km2/stressor", pattern = stressor, full=TRUE))
  
  # chi_cols = c("NA", paste0(final_palette, "BF"))
  
 #quant <- quantile(chi, 0.9999)
 #quant <- quantile(chi, 0.999)

#chi[chi > quant ] <- quant

  raster_data=chi
  
  title="" 
  
if(stressor == "disturbance"){
  cols <- dist_palette
}else if (stressor == "water"){
  cols <- wat_palette
  } else if (stressor == "nutrient"){
    cols <- nut_palette
      }else{
        cols <- ghg_palette}
      
# if(stressor == "ghg"){
#     unit <- "'co'[2]"
#   }else{
#     if(stressor == "water"){
#        unit <-  "'m'^3"
#         }else{
#           if(stressor == "nutrient"){
#             unit <- "'po'[4]"
#           }else{
#             unit <- "'km'[2]"
#           }}}

  if(stressor == "water"){
   # stressor <- "water"
    raster_data <- raster_data/1000
    
      png(here(sprintf("_analysis/figures/paper/output/fig_1b_%s_cumulative.png", stressor)), res=500, width=7, height=3, units="in")
  par(oma=c(1,0,1,0), new = TRUE, mar=c(0, 0, 1, 0)) # bottom, left, top, and right (oma=c(1,0,1,0), )
 # par(mfrow=c(1, 1), mar=c(1, 3, 1, 3))
  
plot(water, border = "#FFFFFF", col= "#FFFFFF", lwd=0.5,  cex= 0.75, main= "Water") 
plot(raster_data, 
     col=cols, 
     axes=FALSE, 
     box=FALSE, 
     add=TRUE,  
     legend=TRUE, 
     legend.shrink= 0.7, 
     legend.width=.4,
    # axis.args = list(cex.axis = 0.6, cex.sub=0.6, cex=0.6, lwd.ticks=1, at = c(0, round(max/3, 0), round(max/3*2, 0), round(max, 0)-1)),
                        # labels= c("0", round(max/3, 0), round(max/3*2, 0), round(max, 0))),
      legend.args=list(text = expression(paste("Thousand ", 'm'^3)), cex = 0.7, side = 3, line = 1))
plot(land, border = "#BCBEB1", col= "#F8F9FA00", lwd=0.2, add=TRUE)
dev.off()
    

  }else if(stressor == "ghg"){
    
    png(here(sprintf("_analysis/figures/paper/output/fig_1b_%s_cumulative.png", stressor)), res=500, width=7, height=3, units="in")
  par(oma=c(1,0,1,0), new = TRUE, mar=c(0, 0, 1, 0)) # bottom, left, top, and right
  #par(mfrow=c(1, 1), mar=c(1, 0, 1, 2))
  
plot(water, border = "#FFFFFF", col= "#FFFFFF", lwd=0.5, main="GHG", cex=0.75)
plot(raster_data, 
     col=cols, 
     axes=FALSE, 
     box=FALSE, 
     add=TRUE,  
     legend=TRUE, 
     legend.shrink= 0.7, 
     legend.width=.4,
     # axis.args = list(cex.axis = 0.6, cex.sub=0.6, cex=0.6, lwd.ticks=1, at = c(0, round(max(raster_data)/3, 0), round(max(raster_data)/3*2, 0), round(max(raster_data), 0)-1)),
      #                   labels=c("0", round(quant/3, 0), round((quant/3)*2, 0), paste0(">", round(quant, 0)))) ,
      legend.args = list(text = expression(paste('CO'[2], "-eq")), cex = 0.7, side = 3, line = 1))
plot(land, border = "#BCBEB1", col= "#F8F9FA00", lwd=0.2, add=TRUE)
dev.off()

  }else if(stressor == "disturbance"){
    
    png(here(sprintf("_analysis/figures/paper/output/fig_1b_%s_cumulative.png", stressor)), res=500, width=7, height=3, units="in")
  par(oma=c(1,0,1,0), new = TRUE, mar=c(0, 0, 1, 0)) # bottom, left, top, and right
  #par(mfrow=c(1, 1), mar=c(1, 0, 1, 2))
  
plot(water, border = "#FFFFFF", col= "#FFFFFF", lwd=0.5, main="Disturbance", cex=0.75)
plot(raster_data, 
     col=cols, 
     axes=FALSE, 
     box=FALSE, 
     add=TRUE,  
     legend=TRUE, 
     legend.shrink= 0.7, 
     legend.width=.4,
     # axis.args = list(cex.axis = 0.6, cex.sub=0.6, cex=0.6, lwd.ticks=1, at = c(0, round(max(raster_data)/3, 0), round(max(raster_data)/3*2, 0), round(max(raster_data), 0)-1)),
      #                   labels=c("0", round(quant/3, 0), round((quant/3)*2, 0), paste0(">", round(quant, 0)))) ,
      legend.args = list(text = expression(paste('km'^2, "-eq")), cex = 0.7, side = 3, line = 1)) 
plot(land, border = "#BCBEB1", col= "#F8F9FA00", lwd=0.2, add=TRUE)
dev.off()

  
}else{
  png(here(sprintf("_analysis/figures/paper/output/fig_1b_%s_cumulative.png", stressor)), res=500, width=7, height=3, units="in")
  par(oma=c(1,0,1,0), new = TRUE, mar=c(0, 0, 1, 0))  # bottom, left, top, and right
  #par(mfrow=c(1, 1), mar=c(1, 0, 1, 2))
  
plot(water, border = "#FFFFFF", col= "#FFFFFF", lwd=0.5, main="Nutrient", cex=0.75)
plot(raster_data, 
     col=cols, 
     axes=FALSE, 
     box=FALSE, 
     add=TRUE,  
     legend=TRUE, 
     legend.shrink= 0.7, 
     legend.width=.4,
     # axis.args = list(cex.axis = 0.6, cex.sub=0.6, cex=0.6, lwd.ticks=1, at = c(0, round(max(raster_data)/3, 0), round(max(raster_data)/3*2, 0), round(max(raster_data), 0)-1)),
      #                   labels=c("0", round(quant/3, 0), round((quant/3)*2, 0), paste0(">", round(quant, 0)))) ,
      legend.args = list(text = "Tonnes P and N", cex = 0.7, side = 3, line = 1))
plot(land, border = "#BCBEB1", col= "#F8F9FA00", lwd=0.2, add=TRUE)
plot(land, border = "#BCBEB1", col= "#F8F9FA00", lwd=0.2, add=TRUE)
dev.off()
}
}

chi_plot_stressor(stressor = "ghg")
chi_plot_stressor(stressor = "water")
chi_plot_stressor(stressor = "nutrient")
chi_plot_stressor(stressor = "disturbance")

```

Extra code that we aren't currently using

Some code for if we want to do them separably instead of with the facet for the CI histrogram legend
```{r}

## the subset plot
 chi_df_sample_hist_subset <-chi_df_sample_hist  %>%
   filter(raster_data_sample >= 4) %>% ## will need to check this math with someone: 
   arrange(raster_data_sample)
 
 subset_col <- chi_df_sample_hist_subset %>% 
  mutate(colors_plot = as.character(colors_plot)) %>% 
  select(col.assign, colors_plot) %>% 
  unique() %>% 
  pull(colors_plot) 
 scales::show_col(subset_col)

## separated
hist_1 <- ggplot(chi_df_sample_hist)  +
  geom_histogram(aes(x = raster_data_sample, fill = colors_plot), bins = 30) +
  scale_fill_manual("", values = rev(cols)) +
  # guides(fill=FALSE)
  theme_bw()

hist_2 <- ggplot(chi_df_sample_hist_subset)  +
  geom_histogram(aes(x = raster_data_sample, fill = colors_plot), bins = 43) +
  scale_fill_identity()
  
summary(chi_df_sample_hist_subset)  

  scale_fill_manual("", values = rev(cols)) +
  theme_bw()

hist_2

chi_df_sample_hist_subset$colors_plot
#AA7754
#imits = seq(0:38), breaks = seq(0:38)


check <- chi_df_sample_hist %>% 
  mutate(test = round(raster_data_sample, digits = 1)) %>% 
  select(test, col.assign, test) %>% 
  unique()

```


Calculate interesting stats
Summarize top 1% of pixels contribute to X % total impact
Do this for 1%, 5%, and 10%
Sym the total in the top X percent and divide by sum across all pixels

```{r}

chi <- raster("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/stressor_summary/equal_area/rescaled/cumulative_stress/all_systems_cumulative_stress.tif")
cellStats(chi, stat = 'sum')
##3855119

## top 1%
top_1_per <- quantile(chi, 0.99)
chi_1 <- raster("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/stressor_summary/equal_area/rescaled/cumulative_stress/all_systems_cumulative_stress.tif")

chi_1[chi_1 < top_1_per ] <- 0
cellStats(chi_1, stat = 'sum')
##1372980
1372980/3855119
## 35.6%

## top 5%
top_5_per <- quantile(chi, 0.95)
chi_5 <- raster("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/stressor_summary/equal_area/rescaled/cumulative_stress/all_systems_cumulative_stress.tif")
chi_5[chi_5 < top_5_per ] <- 0
cellStats(chi_5, stat = 'sum')
# 3083729
3083729/3855119
## 80%

## top 10%
top_10_per <- quantile(chi, 0.90)
chi_10 <- raster("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/stressor_summary/equal_area/rescaled/cumulative_stress/all_systems_cumulative_stress.tif")
chi_10[chi_10 < top_10_per ] <- 0
cellStats(chi_10, stat = 'sum')
# 3729393
3729393/3855119
## 96.7%

```

