---
title: "Summarizing relevant rescaled rasters"
output: html_document
editor_options: 
  chunk_output_type: console
---

We combine the rescaled data in various ways that will be of general interest.

```{r setup, include=FALSE}
library(rgdal)
library(sp)
library(raster)
library(tidyverse)
library(here)

categories <- read_csv(here("_analysis/food_system_categories.csv"))

```

## 1. Combine all systems for each stressor, true values (N=4 rasters)

I think it will be best to use actual values per km2. 
This means using:
- equal area raster
- pressure per km2
- non-rescaled data

```{r}

stressor_sum <- function(stressor = "disturbance"){

  all_rasters <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/_tif_equal_area_proj_per_km2", pattern=".tif", full=TRUE)
  
  stressor_rasts <- grep(stressor, all_rasters, value=TRUE)  
  
  # farm first:
food_files_farm <- grep("farm", stressor_rasts, value=TRUE)
stack_files_farm <- stack(food_files_farm)

raster::calc(stack_files_farm, fun=sum, na.rm=TRUE, 
filename = sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/stressor_summary/equal_area/stressor_per_km2/stressor/tmp_farm_%s.tif", stressor), overwrite=TRUE, progress="text")

# food second:
food_files_feed <- grep("feedcrop|feedfodd|feedfofm", stressor_rasts, value=TRUE)
stack_files_feed <- stack(food_files_feed)

raster::calc(stack_files_feed, fun=sum, na.rm=TRUE, 
filename = sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/stressor_summary/equal_area/stressor_per_km2/stressor/tmp_feed_%s.tif", stressor), overwrite=TRUE, progress="text")  
  
# wildcaught third:
food_files_fish <- grep("wildcaught", stressor_rasts, value=TRUE)
stack_files_fish <- stack(food_files_fish)

raster::calc(stack_files_fish, fun=sum, na.rm=TRUE, filename = sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/stressor_summary/equal_area/stressor_per_km2/stressor/tmp_fish_%s.tif", stressor), overwrite=TRUE, progress="text")  

## combine farm, feed, fish
farm <-   raster(sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/stressor_summary/equal_area/stressor_per_km2/stressor/tmp_farm_%s.tif", stressor))
feed <- raster(sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/stressor_summary/equal_area/stressor_per_km2/stressor/tmp_feed_%s.tif", stressor))
fish <- raster(sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/stressor_summary/equal_area/stressor_per_km2/stressor/tmp_fish_%s.tif", stressor))

stressor_rasts_stack <- stack(farm, feed, fish)

calc(stressor_rasts_stack, fun=sum, na.rm=TRUE, 
       filename = sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/stressor_summary/equal_area/stressor_per_km2/stressor/%s.tif", stressor), overwrite=TRUE)

delete_files <- as.list(list.files("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/stressor_summary/equal_area/stressor_per_km2/stressor", pattern = "tmp", full=TRUE))  
do.call(file.remove, delete_files)    
}


stressor_sum("ghg")
stressor_sum("disturbance")
stressor_sum("water")
stressor_sum("nutrient")



```

## 2. Combine all systems for each stressor, rescaled values (N=4 rasters)

I think it will be best to use actual values per km2. 
This means using:
- equal area raster
- rescaled data



```{r}

rescaled_stressor_sum <- function(stressor = "disturbance"){

  all_rasters <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/_tif_equal_area_proj_per_cell_rescaled", pattern=".tif", full=TRUE)
  
stressor_rasts <- grep(stressor, all_rasters, value=TRUE)  

    # farm first:
food_files_farm <- grep("farm", stressor_rasts, value=TRUE)
stack_files_farm <- stack(food_files_farm)

raster::calc(stack_files_farm, fun=sum, na.rm=TRUE, 
filename = sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/stressor_summary/equal_area/rescaled/stressors/tmp_farm_%s.tif", stressor), overwrite=TRUE, progress="text")

# food second:
food_files_feed <- grep("feedcrop|feedfodd|feedfofm", stressor_rasts, value=TRUE)
stack_files_feed <- stack(food_files_feed)

raster::calc(stack_files_feed, fun=sum, na.rm=TRUE, 
filename = sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/stressor_summary/equal_area/rescaled/stressors/tmp_feed_%s.tif", stressor), overwrite=TRUE, progress="text")  
  
# wildcaught third:
food_files_fish <- grep("wildcaught", stressor_rasts, value=TRUE)
stack_files_fish <- stack(food_files_fish)

raster::calc(stack_files_fish, fun=sum, na.rm=TRUE, filename = sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/stressor_summary/equal_area/rescaled/stressors/tmp_fish_%s.tif", stressor), overwrite=TRUE, progress="text")  

## combine farm, feed, fish
farm <-   raster(sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/stressor_summary/equal_area/rescaled/stressors/tmp_farm_%s.tif", stressor))
feed <- raster(sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/stressor_summary/equal_area/rescaled/stressors/tmp_feed_%s.tif", stressor))
fish <- raster(sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/stressor_summary/equal_area/rescaled/stressors/tmp_fish_%s.tif", stressor))

stressor_rasts_stack <- stack(farm, feed, fish)

  calc(stressor_rasts_stack, fun=sum, na.rm=TRUE, 
       filename = sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/stressor_summary/equal_area/rescaled/stressors/%s.tif", stressor), overwrite=TRUE)

  delete_files <- as.list(list.files("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/stressor_summary/equal_area/rescaled/stressors", pattern = "tmp", full=TRUE))  
do.call(file.remove, delete_files)      
}


rescaled_stressor_sum("ghg")
rescaled_stressor_sum("disturbance")
rescaled_stressor_sum("water")
rescaled_stressor_sum("nutrient")

## These should be close to 1e6 (can be a little different due to changes in CRS)
files <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/stressor_summary/equal_area/rescaled/stressors/", full=TRUE)
for(file in files){ # file=files[1]
  check_raster <- raster(file) 
  totalsum <- cellStats(check_raster, sum, na.rm=TRUE)
  cat(sprintf("checking sum of cells for %s: %s\n", basename(file), totalsum))
}


```


## 3. Combine all rescaled stressor to get cumulative stress (N=1 rasters)
Sum the rescaled rasters for each pressure to get cumulative pressure.

```{r}
all_stressors <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/stressor_summary/equal_area/rescaled/stressors", full=TRUE)

  stressor_rasts_stack <- stack(all_stressors)
  
  calc(stressor_rasts_stack, fun=sum, na.rm=TRUE, 
       filename = "/home/shares/food-systems/Food_footprint/all_food_systems/analysis/stressor_summary/equal_area/rescaled/cumulative_stress/all_systems_cumulative_stress.tif", overwrite=TRUE)
  
  
check <- raster("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/stressor_summary/equal_area/rescaled/cumulative_stress/all_systems_cumulative_stress.tif")
cellStats(check, sum, na.rm=TRUE)
cellStats(check, max, na.rm=TRUE)
```


## 4. Combine rescaled rasters for each animal/product

This means using:
- equal area raster
- rescaled data

Categories (includes feed and fodder and fofm):
livestock categories
combined human consumed crops
marine wildcaught fisheries
freshwater fisheries
all aquaculture



```{r}

cats <- read_csv(here("_analysis/food_system_categories.csv"))

# get list of categories to cycle through:
livestock <- filter(cats, subcategory=="livestock") %>%
  select(organism, product) %>%
  unique()

marine_fish <- filter(cats, origin == "marine" & subcategory == "fisheries") %>%
  select(organism, product = subcategory) %>%
  unique()
  
crops <- data.frame(organism = "farm", product="crop")

fresh_fish <- data.frame(organism = "wildcaught", product="freshwater")
mariculture <- data.frame(organism = "aquaculture", product="marine")

combined_cats <- rbind(livestock, crops, marine_fish, fresh_fish, mariculture)

## function
rescaled_animal_product <- function(search1 = "chickens", search2="eggs"){

all_rasters <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/_tif_equal_area_proj_per_cell_rescaled", pattern=search1, full=TRUE)

if(search1=="chickens" & search2 == "eggs" | search2=="meat"){  
subset_rasts <- grep(search2, all_rasters, value=TRUE)
subset_rasts <- grep("eggs&meat", subset_rasts, invert=TRUE, value=TRUE)
cat(sprintf("checking: %s, %s, N= %s \n", search1, search2, length(subset_rasts)))
}else{
subset_rasts <- grep(search2, all_rasters, value=TRUE)
cat(sprintf("checking: %s, %s, N= %s \n", search1, search2, length(subset_rasts)))
}

stack_files <- stack(subset_rasts)

raster::calc(stack_files, fun=sum, na.rm=TRUE, 
filename = sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/figures/animal_product/cpi/%s_%s.tif", search1, search2), overwrite=TRUE)
}



for(i in 1:dim(combined_cats)[1]){
  org <- pull(combined_cats[i,1])
  product <- pull(combined_cats[i,2])
## function
rescaled_animal_product(search1 = org, search2=product)
}


```
checking: sheep, meat, N= 24 
checking: sheep, milk, N= 24 
checking: goats, meat, N= 24 
checking: goats, milk, N= 24 
checking: cows, meat, N= 36 
checking: cows, milk, N= 24 
checking: buffaloes, milk, N= 24 
checking: pigs, meat, N= 48 
checking: chickens, eggs&meat, N= 12 
checking: chickens, eggs, N= 16 
checking: chickens, meat, N= 16 
checking: farm, crop, N= 108 
checking: fofm, fisheries, N= 4 
checking: reef, fisheries, N= 4 
checking: benthic, fisheries, N= 4 
checking: demersal, fisheries, N= 4 
checking: medium-pelagic, fisheries, N= 4 
checking: small-pelagic, fisheries, N= 4 
checking: large-pelagic, fisheries, N= 4 
checking: wildcaught, freshwater, N= 4 
checking: aquaculture, marine, N= 64 


## 4. Combine rescaled rasters for each animal/product

This means using:
- equal area raster
- rescaled data

Categories (includes feed and fodder and fofm):
livestock categories
combined human consumed crops
marine wildcaught fisheries
freshwater fisheries
all aquaculture


Combining data for the SI pressure plots for each animal/product.
Will need to do crops separately.

```{r}

cats <- read_csv(here("_analysis/food_system_categories.csv"))

# get list of categories to cycle through:
livestock <- filter(cats, subcategory=="livestock") %>%
  select(organism, product) %>%
  unique()

marine_fish <- filter(cats, origin == "marine" & subcategory == "fisheries") %>%
  select(organism, product = subcategory) %>%
  unique()
  
fresh_fish <- data.frame(organism = "wildcaught", product="freshwater")

mariculture <- filter(cats, origin == "marine" & subcategory == "aquaculture") %>%
  select(organism, product = subcategory) %>%
  unique()

combined_cats <- rbind(livestock, marine_fish, fresh_fish, mariculture) %>% data.frame()

## function
rescaled_animal_product <- function(search1 = "fofm", search2="fisheries"){

all_rasters <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/_tif_equal_area_proj_per_cell_rescaled", pattern=search1, full=TRUE)

if(search1=="chickens" & search2 == "eggs" | search2=="meat"){  
subset_rasts <- grep(search2, all_rasters, value=TRUE)
subset_rasts <- grep("eggs&meat", subset_rasts, invert=TRUE, value=TRUE)
cat(sprintf("checking: %s, %s, N= %s \n", search1, search2, length(subset_rasts)))
}else{
subset_rasts <- grep(search2, all_rasters, value=TRUE)
cat(sprintf("checking cpi: %s, %s, N= %s \n", search1, search2, length(subset_rasts)))
}

stack_files <- stack(subset_rasts)

raster::calc(stack_files, fun=sum, na.rm=TRUE, 
filename = sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/figures/SI_pressure_maps/cpi_%s_%s.tif", search1, search2), overwrite=TRUE)

pressure_list <- c("ghg", "nutrient", "disturbance", "water")

for(pressure in pressure_list){ # pressure = "ghg"

    subset_pressure_rasts <- grep(pressure, subset_rasts, value=TRUE)
  
  if(length(subset_pressure_rasts)==1){
 file.copy(from = subset_pressure_rasts, 
           to = sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/figures/SI_pressure_maps/%s_%s_%s.tif", pressure, search1, search2), 
           overwrite = TRUE)
  }else{
  subset_pressure_stack <- stack(subset_pressure_rasts)
raster::calc(subset_pressure_stack, fun=sum, na.rm=TRUE, 
filename = sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/figures/SI_pressure_maps/%s_%s_%s.tif", pressure, search1, search2), overwrite=TRUE)
cat(sprintf("checking: %s, %s, %s, N= %s \n", pressure, search1, search2, length(subset_pressure_rasts)))
}
}  
}




for(i in 1:dim(combined_cats)[1]){
  org <- combined_cats[i,1]
  product <- combined_cats[i,2]
## function
rescaled_animal_product(search1 = org, search2=product)
}


```

FOFM needs to be added up.

```{r}


all_rasters <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/_tif_equal_area_proj_per_cell_rescaled", pattern="fofm", full=TRUE)
stack_files <- stack(all_rasters)

raster::calc(stack_files, fun=sum, na.rm=TRUE, 
filename = "/home/shares/food-systems/Food_footprint/all_food_systems/analysis/figures/SI_pressure_maps/cpi_fofm_fisheries.tif", overwrite=TRUE)

pressure_list <- c("ghg", "nutrient", "disturbance", "water")

for(pressure in pressure_list){ # pressure = "ghg"

    subset_pressure_rasts <- grep(stressor, all_rasters, value=TRUE)
  subset_pressure_stack <- stack(subset_pressure_rasts)
raster::calc(subset_pressure_stack, fun=sum, na.rm=TRUE, 
filename = sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/figures/SI_pressure_maps/%s_fofm_fisheries.tif", pressure), overwrite=TRUE)
cat(sprintf("checking: %s, N= %s \n", pressure, length(subset_pressure_rasts)))
}


```


We have to go back to the beginning for crops:

```{r}

crop_files <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/datalayers", full=TRUE, pattern = "_crop_")

crop_farm_files <- grep("_x_", crop_files, value=TRUE, invert=TRUE)



for(file in crop_farm_files){ # 
  
  # file = crop_farm_files[1] 
  
  stressor_raster <- raster(file)
  
  name <- basename(file)
  name <- gsub(".tif", "", name)
  
  raster_area <- area(stressor_raster)
  #plot(raster_area)
  
if(grepl("disturbance", file)){
  
   stressor_area_adjust <-  stressor_raster/raster_area
   stressor_area_adjust[stressor_area_adjust >1] <- 1

   writeRaster(stressor_area_adjust,
               sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/figures/SI_pressure_maps/crop_manipulation/crop_tif_latlong_per_km2/%s_per_area.tif", name),
               overwrite=TRUE)

}else{
  
   stressor_area_adjust <- stressor_raster/raster_area
   
   writeRaster(stressor_area_adjust,
               sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/figures/SI_pressure_maps/crop_manipulation/crop_tif_latlong_per_km2/%s_per_area.tif", name),
               overwrite=TRUE)

  }
}

```


Next: project rasters to equal area gall peters so we can more easily scale data and such.
```{r}

gall_peters <- "+proj=cea +lon_0=0 +x_0=0 +y_0=0 +lat_ts=45 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
#raster template 
template_eq_area <- raster(res=0.083333333333333333)
extent(template_eq_area) <- c(-180, 180, -90, 90)
template_eq_area <- projectRaster(template_eq_area, crs=gall_peters, res=3000) 


```


```{r}

rasters_to_new_crs <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/figures/SI_pressure_maps/crop_manipulation/crop_tif_latlong_per_km2", pattern=".tif", full=TRUE)

for(rast in rasters_to_new_crs){ # rast <- rasters_to_new_crs[1]
  
rast_name <- basename(rast)
 getraster <- raster(rast)
 highres_raster <- disaggregate(getraster, fact=2)
 projectRaster(highres_raster, template_eq_area, method="ngb", over=TRUE, progress="text",
                       filename=sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/figures/SI_pressure_maps/crop_manipulation/crop_tif_equal_area_proj_per_km2/hi_res_gall_peter_%s", rast_name), overwrite=TRUE)

 gp_raster <- raster(sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/figures/SI_pressure_maps/crop_manipulation/crop_tif_equal_area_proj_per_km2/hi_res_gall_peter_%s", rast_name))
#plot(gp_raster)
 
 aggregate(gp_raster, fact=2, fun=mean, filename=sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/figures/SI_pressure_maps/crop_manipulation/crop_tif_equal_area_proj_per_km2/gall_peter_%s", rast_name), overwrite=TRUE)
}

```


Delete the hi_res version which is just a temp file.
```{r}
remove <- list.files(file.path("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/figures/SI_pressure_maps/crop_manipulation/crop_tif_equal_area_proj_per_km2"), full=TRUE, pattern="hi_res")
file.remove(remove)
```


convert to total per cell
```{r}
per_km_rasts <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/figures/SI_pressure_maps/crop_manipulation/crop_tif_equal_area_proj_per_km2", full=TRUE)

for(rast in per_km_rasts){ # rast = per_km_rasts[1]
 save_name <- basename(rast)
 save_name <- gsub("_per_area", "_per_cell", save_name)
    new_raster <- raster(rast)
  per_cell_rast <- new_raster*6000*6000/1000000
writeRaster(per_cell_rast, sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/figures/SI_pressure_maps/crop_manipulation/crop_tif_equal_area_proj_per_cell/%s", save_name), overwrite=TRUE)
#check <- raster(sprintf("/home/shares/food-systems/Food_footprint/final_data/_tif_equal_area_proj_per_cell/%s", save_name))
#cellStats(check, "sum", na.rm=TRUE)
}

```

Rescale

```{r}

rescaling_data <- read_csv(here("_analysis/rescale_values.csv"))

pressure_list <- c("ghg", "nutrient", "disturbance", "water")

for(stressor in pressure_list){ # stressor = "nutrient"
  
  resc_num <- rescaling_data$global_total[rescaling_data$pressure==stressor] 
  
stressor_system_files <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/figures/SI_pressure_maps/crop_manipulation/crop_tif_equal_area_proj_per_cell", pattern=stressor, full=TRUE)
  

  for(stressor_raster_filename in stressor_system_files){ #stressor_raster_filename= stressor_system_files[1]
file_name <- basename(stressor_raster_filename)
file_name <- gsub("_per_cell.tif", "", file_name)
file_name <- paste0(file_name, "_rescaled.tif")

stressor_raster <- raster(stressor_raster_filename)
rescaled_raster = raster::calc(stressor_raster, fun=function(x){ifelse(x>0, 
                                      (x/resc_num) * 1000000, 
                                       0)})
writeRaster(rescaled_raster, sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/figures/SI_pressure_maps/crop_manipulation/crop_tif_equal_area_proj_per_cell_rescaled/%s", file_name),
              overwrite=TRUE)
}
}

```

apply function from above to get the final rasters
```{r}

cats <- read_csv(here("_analysis/food_system_categories.csv"))

crops <- filter(cats, system == "crop") %>%
  select(organism, product = system) %>%
  unique()


## function
rescaled_animal_product <- function(search1 = "bana", search2="crop"){

crop_rasters <- list.files("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/figures/SI_pressure_maps/crop_manipulation/crop_tif_equal_area_proj_per_cell_rescaled", pattern=search1, full=TRUE)

stack_files <- stack(crop_rasters)

raster::calc(stack_files, fun=sum, na.rm=TRUE, 
filename = sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/figures/SI_pressure_maps/cpi_%s_%s.tif", search1, search2), overwrite=TRUE)

pressure_list <- c("ghg", "nutrient", "disturbance", "water")

for(pressure in pressure_list){ # pressure = "ghg"

    subset_pressure_rasts <- grep(pressure, crop_rasters, value=TRUE)
  
 file.copy(from = subset_pressure_rasts, 
           to = sprintf("/home/shares/food-systems/Food_footprint/all_food_systems/analysis/figures/SI_pressure_maps/%s_%s_%s.tif", pressure, search1, search2), 
           overwrite = TRUE)
}
}



for(i in 1:dim(crops)[1]){ #i=1
  org <- pull(crops[i,1])
  product <- pull(crops[i,2])
## function
rescaled_animal_product(search1 = org, search2=product)
}

```